
```{r}
library(tidyverse)
library(compositions)
library(stats)
library(car)
library(broom)
library(nnet)
library(GGally)
library(emmeans)

load(here::here('Data/integrated_dataset_2008-2022.Rdata'))
```


# Goal

For three classifications (HAB, BAB, No Bloom), what are their relations with different, non-independent algal groups?

Use relative biovolume as algal group variable (% of sample bv that is that algal group)

Can't put values in directly (correlation issues); use log ratios

Notes:
- for rn, no true data exploration phase
- for rn, didn't deal with station or temporal correlation

# Read in data

```{r}
# df_cen <- dat %>%
#   filter(Region == 'Central')
df_cen <- dat

df_cen <- df_cen %>%
  select(Month, Year, Station, AlgalGroup, tot_bio, relative_biovolume, harmful_class) %>%
  filter(!is.na(relative_biovolume)) %>%
  filter(!is.na(harmful_class))

focal_groups <- c('Cyanobacteria', 'Centric Diatoms', 'Pennate Diatoms',
                  'Cryptophytes', 'Green Algae')

df_cen <- df_cen %>%
    mutate(group = if_else(AlgalGroup %in% focal_groups, AlgalGroup, 'Other'))
```

# Multinomial 

```{r}
df_wide <- df_cen %>%
  group_by(Year, Month, Station, tot_bio, harmful_class, group) %>%
  summarise(biovolume = sum(relative_biovolume, na.rm = TRUE), .groups = 'drop') %>%
  pivot_wider(names_from = group, values_from = biovolume, values_fill = 0)

focal_groups <- c('Cyanobacteria', 'CentricDiatoms', 'PennateDiatoms',
                  'Cryptophytes', 'GreenAlgae')

df_wide <- df_wide %>%
  rename(CentricDiatoms = `Centric Diatoms`,
         PennateDiatoms = `Pennate Diatoms`,
         GreenAlgae = `Green Algae`)

comp_cols <- c(focal_groups, 'Other')

sbp <- matrix(c(
   1,  0,  0,  0, # cyano; cyano vs all
  -1,  1,  1,  0, # centric; diatoms vs crypt/green
  -1,  1, -1,  0, # pennate; centric vs pennate
  -1, -1,  0,  1, # crypto; crypto vs green
  -1, -1,  0, -1  # green
), nrow = 5, ncol = 4, byrow = TRUE,
dimnames = list(
  c('Cyanobacteria', 'CentricDiatoms', 'PennateDiatoms',
    'Cryptophytes', 'GreenAlgae'),
  c('b1_cyano_vs_rest', 'b2_diatoms_vs_cg',
    'b3_centric_vs_pennate', 'b4_crypto_vs_green')
))

comp_closed <- acomp(df_wide[, rownames(sbp)])

V <- gsi.buildilrBase(sbp)
ilr_coords <- ilr(comp_closed, V = V)
colnames(ilr_coords) <- colnames(sbp)

df_model <- bind_cols(
  df_wide %>% select(Year, Month, Station, tot_bio, harmful_class),
  as.data.frame(ilr_coords),
  log_tot_bio = log(df_wide$tot_bio)
)

df_model <- df_model %>%
  mutate(log_tot_bio = log_tot_bio - mean(log_tot_bio, na.rm = TRUE))
```

```{r}
df_model$harmful_class <- relevel(factor(df_model$harmful_class), ref = 'No bloom')

fit_multi <- nnet::multinom(harmful_class ~ b1_cyano_vs_rest + b2_diatoms_vs_cg +
                                      b3_centric_vs_pennate + b4_crypto_vs_green +
                                      log_tot_bio,
                      data = df_model)

tidy(fit_multi) 
```

```{r}
df_model$harmful_class <- relevel(factor(df_model$harmful_class), ref = 'BAB')

fit_bab <- nnet::multinom(harmful_class ~ b1_cyano_vs_rest + b2_diatoms_vs_cg +
                                    b3_centric_vs_pennate + b4_crypto_vs_green +
                                    log_tot_bio,
                    data = df_model)
tidy(fit_bab)
```


```{r}
vars <- c('b1_cyano_vs_rest', 'b2_diatoms_vs_cg',
          'b3_centric_vs_pennate', 'b4_crypto_vs_green', 'log_tot_bio')

all_trends <- map_dfr(vars, function(v) {
  emtrends(fit_multi, ~ harmful_class, var = v, mode = 'prob') %>%
    summary(infer = TRUE) %>%
    as.data.frame() %>%
    rename(trend = 2) %>%
    mutate(variable = v)
})

all_trends
```


# Two binomials

```{r}
df_model <- df_model %>%
  mutate(
    bloom = factor(if_else(harmful_class == 'No bloom', 'No bloom', 'Bloom')),
    hab_bab = factor(if_else(harmful_class == 'HAB', 'HAB', 'BAB'))
  )

df_model$bloom <- relevel(df_model$bloom, ref = 'No bloom')

fit_pt1 <- glm(bloom ~ b1_cyano_vs_rest + b2_diatoms_vs_cg +
                           b3_centric_vs_pennate + b4_crypto_vs_green +
                           log_tot_bio,
                  data = df_model,
                  family = binomial(link = 'logit'))

summary(fit_pt1)

df_bloom <- df_model %>% filter(bloom == 'Bloom')

fit_pt2 <- glm(hab_bab ~ b1_cyano_vs_rest + b2_diatoms_vs_cg +
                                b3_centric_vs_pennate + b4_crypto_vs_green +
                                log_tot_bio,
                  data = df_bloom,
                  family = binomial(link = 'logit'))

summary(fit_pt2)
```

# Not ILR Example

```{r}
df_alt <- df_wide %>%
  mutate(
    bloom = factor(if_else(harmful_class == 'No bloom', 'No bloom', 'Bloom')),
    hab_bab = factor(if_else(harmful_class == 'HAB', 'HAB', 'BAB')),
    log_tot_bio = log(tot_bio) - mean(log(tot_bio), na.rm = TRUE))

fit_alt1 <- glm(bloom ~ Cyanobacteria + CentricDiatoms + Cryptophytes +
                  PennateDiatoms + GreenAlgae + log_tot_bio,
                  data = df_alt,
                  family = binomial(link = 'logit'))

summary(fit_alt1)

df_altbloom <- df_alt %>% filter(bloom == 'Bloom')

fit_alt2 <- glm(hab_bab ~ Cyanobacteria + CentricDiatoms + Cryptophytes +
                  PennateDiatoms + GreenAlgae + log_tot_bio,
                  data = df_altbloom,
                  family = binomial(link = 'logit'))

summary(fit_alt2)
```

```{r}
df_alt$harmful_class <- relevel(factor(df_model$harmful_class), ref = 'No bloom')

fit_altmulti <- multinom(harmful_class ~ Cyanobacteria + CentricDiatoms +
                        Cryptophytes + PennateDiatoms + GreenAlgae +
                        log_tot_bio,
                      data = df_alt)

tidy(fit_altmulti) 
```

```{r}
vif(fit_altmulti)
```



```{r}
vif(fit_multi)
```

# Exploratory Graphs

```{r}
table(df_model$harmful_class)
```

```{r}
df_alt %>%
  pivot_longer(c(Cyanobacteria, CentricDiatoms, PennateDiatoms,
                 Cryptophytes, GreenAlgae, log_tot_bio),
               names_to = 'variable', values_to = 'value') %>%
  ggplot(aes(value, fill = harmful_class, color = harmful_class)) +
  geom_density(alpha = 0.4) +
  facet_wrap(~ variable, scales = 'free') +
  theme_minimal()
```

# GAM Example

```{r}
library(mgcv)
df_gam <- df_model
df_gam$harmful_class <- relevel(factor(df_model$harmful_class), ref = 'No bloom')
df_gam$hc_int <- as.integer(df_model$harmful_class) - 1

levels(df_gam$harmful_class)

fit_gam <- gam(
  list(
    hc_int ~ s(b1_cyano_vs_rest) + s(b2_diatoms_vs_cg) +
             s(b3_centric_vs_pennate) + s(b4_crypto_vs_green) + s(log_tot_bio),
    ~ s(b1_cyano_vs_rest) + s(b2_diatoms_vs_cg) +
      s(b3_centric_vs_pennate) + s(b4_crypto_vs_green) + s(log_tot_bio)
  ),
  family = multinom(K = 2),
  data = df_gam
)

summary(fit_gam)
```

```{r}
plot(fit_gam)
```

# ILR/Raw Comparsion

```{r}
sbp2 <- matrix(c(
  -1, -1,  0, -1, # green; green vs all
  -1,  1,  1,  0, # centric; diatoms vs crypt/green
  -1,  1, -1,  0, # pennate; centric vs pennate
  -1, -1,  0,  1, # crypto; crypto vs green
   1,  0,  0,  0  # cyano
), nrow = 5, ncol = 4, byrow = TRUE,
dimnames = list(
  c('Cyanobacteria', 'CentricDiatoms', 'PennateDiatoms',
    'Cryptophytes', 'GreenAlgae'),
  c('b1_green_vs_rest', 'b2_diatoms_vs_cg',
    'b3_centric_vs_pennate', 'b4_crypto_vs_cyano')
))

comp_closed2 <- acomp(df_wide[, rownames(sbp2)])

V2 <- gsi.buildilrBase(sbp2)
ilr_coords2 <- ilr(comp_closed2, V = V2)
colnames(ilr_coords2) <- colnames(sbp2)

df_model2 <- bind_cols(
  df_wide %>% select(Year, Month, Station, tot_bio, harmful_class),
  as.data.frame(ilr_coords2),
  log_tot_bio = log(df_wide$tot_bio)
)

df_model2 <- df_model2 %>%
  mutate(log_tot_bio = log_tot_bio - mean(log_tot_bio, na.rm = TRUE))
```

```{r}
df_model2$harmful_class <- relevel(factor(df_model$harmful_class), ref = 'No bloom')

fit_multi2 <- nnet::multinom(harmful_class ~ b1_green_vs_rest + b2_diatoms_vs_cg +
                                      b3_centric_vs_pennate + b4_crypto_vs_cyano +
                                      log_tot_bio,
                      data = df_model2)

tidy(fit_multi2) 
```

# PCA Checks

```{r}
comp_clr <- clr(comp_closed)

pca <- prcomp(comp_clr, center = TRUE, scale. = FALSE)

summary(pca)

round(pca$rotation, 3)
```


```{r}
scores <- as.data.frame(pca$x)
scores$harmful_class <- df_model$harmful_class

ggplot(scores, aes(PC1, PC2, color = harmful_class)) +
  geom_point(alpha = 0.6) +
  labs(x = paste0('PC1 (', round(100*summary(pca)$importance[2,1],1), '%)'),
       y = paste0('PC2 (', round(100*summary(pca)$importance[2,2],1), '%)')) +
  theme_minimal()
```

```{r}
ggplot(scores, aes(PC1, PC2)) +
  geom_density_2d(aes(color = harmful_class), linewidth = 0.4) +
  labs(x = paste0('PC1 (', round(100*summary(pca)$importance[2,1],1), '%)'),
       y = paste0('PC2 (', round(100*summary(pca)$importance[2,2],1), '%)')) +
  theme_minimal()
```


