```{r}
library(tidyverse)
library(here)
library(nnet)
library(mgcv)
library(corrplot)

conflicted::conflict_prefer('multinom', 'mgcv')
conflicted::conflict_prefer('select', 'dplyr')

load(here::here('Short Term Analysis/st_data/04_imputedata.Rda'))
```
# TODO: prioritize global model, then do the regional models
  - RE of region


# Goal

Multinomial GAM on short-term dataset.

# Data

Data is from the '04_imputedata.Rda' file, which contains one run of imputed data via `mice`'s multiple imputation algorithm. Characteristics:

- data currently at the Station level
   - Flow variables: system-wide
   - Clams: region-wide

- Algal Group data is currently taxon data aggregated per Station ($G_{s,g} = \sum{x_{s,t}}$)

Response variable: `harmful_class`: final classification of **HAB, BAB, or no bloom**. Categories were determined via \<method\> in \<filename\>.

- Only data from (date) onwards (bc...) is used.

- Classification determined by MVI scores.

## Regions/Flow Variables

Flow Variables:

- SAC (input)

- SJR (input)

- Exports (output) export to facilities in South Delta

- OUT (Net Delta Outflow Index; NDOI) where $$OUT = Inflow - Exports - Used$$

  - or is it NDO? (uses stations)

  - Inflow includes SAC and SJR

# Clean Data

## Remove AbsCellDensity and TotPhos cols

```{r}
df_mvi <- df_mvi %>%
  select(!c(contains('AbsCellDensity'),'TotPhos'))
```

## log variables

```{r}
log_vars <- c(
  'Chlorophyll',
  'Pheophytin',
  'Conductivity',
  'Turbidity',
  'DissAmmonia',
  'DissNitrateNitrite',
  'DissOrthophos',
  'DON',
  'TKN',
  'DissChloride',
  'TDS',
  'SAC',
  'SJR',
  'OUT',
  'EXPORTS',
  'monreg_biomass',
  'monreg_graze_rate',
  'monreg_filt_rate'
)

df_mvi <- df_mvi %>%
  mutate(
    across(
      all_of(log_vars),
      log,
      .names = 'log_{.col}'
    )
  ) %>%
  select(-all_of(log_vars))
```

## z-score (center & scale)

```{r}
# exclude rel cel density for now
df_mvi <- df_mvi %>%
  mutate(
    across(
      where(is.numeric) &
        !any_of(c(
          'Year', 'Month', 'Station', 'Survey', 'Season',
          'Region', 'mvi_class', 'modeled_class'
        )) &
        !contains('RelCellDensity'),
      ~ as.numeric(scale(.x, center = TRUE, scale = TRUE))
    )
  )
```

## refactor classification

```{r}
df_mvi <- df_mvi %>%
  mutate(mvi_class = relevel(factor(mvi_class), ref = 'No bloom'))
```


# Model Selection

Assumptions:

- biogeochemically, no predictor's effect should depend on region

- *however*, we interested in difference between regions (ie. subsets of predictor ranges), so we'll run separate models per region

```{r}
unique(df_mvi$Region)
```

# Pick Variables

## Flow

Four flow variables: SAC, SJR, EXPORTS, OUT
  - inputs: SAC/SJR/EXPORTS
  
  - output: OUT (modeled)

Shouldn't use all 4, collinear.

Solution:

 - run one model with inputs
 
 - run one model with output


```{r}
df_mvi %>%
  select(log_SAC, log_SJR, log_EXPORTS) %>%
  cor(use = 'pairwise') %>% round(2)
```

All can go in for now...but ideally choose between SAC/SJR.

```{r}
unique(df_mvi$Region)
```

## Clams

Clams graze phytoplankton, so graze rate is likely most relevant.

## Nutrients

DissAmmonia, DissOrthophos, DissNitrateNitrite make most sense.

```{r}
nut_vars <- c('log_DissAmmonia', 'log_DissNitrateNitrite', 'log_DissOrthophos',
              'log_DON', 'log_TKN', 'DissSilica', 'log_TDS', 'log_DissChloride', 'log_Chlorophyll')
```

```{r}
cm <- df_mvi %>%
  select(all_of(nut_vars)) %>%
  cor(use = 'pairwise.complete.obs')

corrplot(cm,
         method = 'color',
         type   = 'upper',
         addCoef.col = 'black',
         number.cex  = 0.7,
         tl.col = 'black',
         tl.srt = 45,
         diag   = FALSE,
         order  = 'hclust')
```

```{r}
fin_nuts <- c('log_DissAmmonia','log_DissOrthophos','log_DissNitrateNitrite')

cm <- df_mvi %>%
  select(all_of(fin_nuts)) %>%
  cor(use = 'pairwise.complete.obs')

cm
```

Not too correlated with each other, NitrateNitrite one to watch out for the most.

## Water Related

Chloride/TDS and Calcium aren't related to blooms/are related to Conductance.

## Solids

Usually caused by blooms I believe? Check literature.

## Initial model

```{r}
df_cen <- df_cen %>%
  mutate(mvi_num = as.integer(factor(mvi_class,
                                     levels = c('No bloom', 'BAB', 'HAB'))) - 1)

gam_ini <- gam(
  list(
    mvi_num ~ s(Temperature) + s(log_Conductivity) +,
              ~ s(Temperature) + s(log_Conductivity)
  ),
  family = multinom(K = 2),
  select = TRUE, # ridge theoretically, lasso behaviorally (check)
  data = df_cen
)

summary(gam_ini)
plot(gam_ini, pages = 1, scale = 0)
```


