---
title: "Seasonal Analysis"
author: "Alex Berk"
date: "2025-08-21"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Read in packages
```{r Read in Packages, include=FALSE}
library(tidyverse)
library(mgcv)
library(MuMIn)
library(gratia)
library(sf)

theme_set(theme_classic())
```

Read in Region Shapefile
```{r Read in Region Shapefile, include=FALSE}
#read in region data
regions <- st_read("Regions_shp/Rosies_regions_edited.shp")

#change the crs to UTM zone 10N (EPSG 32610)
regions<- st_transform(regions, crs =32610)
```

Read in Alex's csv
```{r Read in Alexs csv, include=FALSE}
#read in region data
wq <- read.csv("Data/edi_df_integrate_monthly.csv") %>%
  filter(ymd(year_month) >= ymd(19790101))
```

Create seasonal variable
```{r Create seasonal variable, include=FALSE}
#convert dates to months and seasons
wq$season <- replicate(nrow(wq),NA)
wq$season[wq$Month %in% c(1,2,3)] <- "Winter"
wq$season[wq$Month %in% c(4,5,6)] <- "Spring"
wq$season[wq$Month %in% c(7,8,9)] <- "Summer"
wq$season[wq$Month %in% c(10,11,12)] <- "Fall"

wq_season <- wq %>% group_by(Month) %>%
  summarize(season=unique(season)) 
```

Attach Regions and Clean up Data
```{r Attach Regions and Clean up Data, include=FALSE}

wq.sf <- wq %>%
  filter(!is.na(Longitude) & !is.na(Latitude)) %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) %>%
  #convert to UTMs so it's in the same coordinate reference system as the Delta shapefile
  st_transform(crs = 32610)

wq <- st_join(wq.sf, regions["Regions"], join = st_intersects, left= TRUE) %>%  
  filter(!is.na(Regions)) %>%
  st_drop_geometry() 

#Calculate monthly average values for each region
wq_r_sum <- wq %>% 
  group_by(Regions,year_month) %>%
  summarize_if(is.numeric,mean,na.rm=TRUE) %>%
  merge(wq_season,by="Month",all.x=T) %>%
  mutate(year_month = ymd(year_month)) %>%
  select(c("year_month", "Month", "season", "Regions", 
           "Chlorophyll",
           "DissAmmonia",
           "Secchi",
           "TotPhos",
           "DissNitrateNitrite",
           "Temperature", 
           "TurbidityNTU",
           "Conductivity",
           "SAC",
           "Index"))


#add 1-month lag
wq_r_sum_lagmonth1 <- wq_r_sum %>%
  mutate(lagmonth = year_month %m+% months(1)) %>%
  select(lagmonth,Regions,DissAmmonia:SAC) %>%
  rename_with(~ paste0("lag",.x)) %>%
  rename(lagmonth = "laglagmonth",Regions = "lagRegions")

#add 2-month lag
wq_r_sum_lagmonth2 <- wq_r_sum %>%
  mutate(lagmonth = year_month %m+% months(2)) %>%
  select(lagmonth,Regions,DissAmmonia:SAC) %>%
  rename_with(~ paste0("lag",.x)) %>%
  rename(lagmonth = "laglagmonth",Regions = "lagRegions")

#add 3-month lag
wq_r_sum_lagmonth3 <- wq_r_sum %>%
  mutate(lagmonth = year_month %m+% months(3)) %>%
  select(lagmonth,Regions,DissAmmonia:SAC) %>%
  rename_with(~ paste0("lag",.x)) %>%
  rename(lagmonth = "laglagmonth",Regions = "lagRegions")

wq_r_sum_lagseason <- wq_r_sum %>%
  mutate(year = year(year_month)) %>%
  group_by(Regions,year,season) %>%
  summarize(across(DissAmmonia:SAC,\(x) mean(x, na.rm = TRUE))) %>%
  rename_with(~paste0("lag",.x)) %>%
  rename(year = "lagyear",season = "lagseason",Regions="lagRegions") %>%
  mutate(lagseason = case_when(
    season == "Spring" ~ "Summer",
    season == "Summer" ~ "Fall",
    season == "Fall" ~ "Winter",
    season == "Winter" ~ "Spring" 
  )) %>%
  mutate(lagseasonyear = case_when(
    season == "Fall" ~ year+1,
    !(season == "Fall") ~ year
  )) %>%
  mutate(lagseasonyear = paste0(lagseason,lagseasonyear)) %>%
  ungroup %>%
  select(c("lagseasonyear","Regions","lagDissAmmonia","lagSecchi","lagTotPhos",
           "lagDissNitrateNitrite","lagTemperature","lagTurbidityNTU","lagConductivity",
           "lagSAC"))


wq_r_sum_lagyear <- wq_r_sum %>%
  mutate(lagyear = year_month %m+% months(12)) %>%
  select(lagyear,Regions,Index) %>%
  rename(lagyearIndex = "Index")

wq_r_sum1 <- wq_r_sum %>%
  merge(wq_r_sum_lagmonth1,by.x=c("year_month","Regions"),by.y=c("lagmonth","Regions"),all.x=T) %>%
  merge(wq_r_sum_lagyear,by.x=c("year_month","Regions"),by.y=c("lagyear","Regions"),all.x=T)

wq_r_sum2 <- wq_r_sum %>%
  merge(wq_r_sum_lagmonth2,by.x=c("year_month","Regions"),by.y=c("lagmonth","Regions"),all.x=T) %>%
  merge(wq_r_sum_lagyear,by.x=c("year_month","Regions"),by.y=c("lagyear","Regions"),all.x=T)

wq_r_sum3 <- wq_r_sum %>%
  merge(wq_r_sum_lagmonth3,by.x=c("year_month","Regions"),by.y=c("lagmonth","Regions"),all.x=T) %>%
  merge(wq_r_sum_lagyear,by.x=c("year_month","Regions"),by.y=c("lagyear","Regions"),all.x=T)

wq_r_sum_season <- wq_r_sum %>% 
  mutate(seasonyear = paste0(season,year(year_month)))%>%
  merge(wq_r_sum_lagseason,by.x=c("Regions","seasonyear"),by.y=c("Regions","lagseasonyear"),all.x=T) %>%
  merge(wq_r_sum_lagyear,by.x=c("year_month","Regions"),by.y=c("lagyear","Regions"),all.x=T)




#restructure data for facet wrap
wq_long1 <- wq_r_sum1 %>%
  pivot_longer(cols=DissAmmonia:lagyearIndex,names_to="Covariate_name",values_to="Covariate_value") %>%
  mutate(logvalue = log(Covariate_value)) %>%
  filter(!is.infinite(logvalue))

wq_long2 <- wq_r_sum2 %>%
  pivot_longer(cols=DissAmmonia:lagyearIndex,names_to="Covariate_name",values_to="Covariate_value") %>%
  mutate(logvalue = log(Covariate_value)) %>%
  filter(!is.infinite(logvalue))

wq_long3 <- wq_r_sum3 %>%
  pivot_longer(cols=DissAmmonia:lagyearIndex,names_to="Covariate_name",values_to="Covariate_value") %>%
  mutate(logvalue = log(Covariate_value)) %>%
  filter(!is.infinite(logvalue))

wq_longsn <- wq_r_sum_season %>%
  pivot_longer(cols=DissAmmonia:lagyearIndex,names_to="Covariate_name",values_to="Covariate_value") %>%
  mutate(logvalue = log(Covariate_value)) %>%
  filter(!is.infinite(logvalue))
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables with 1 month lag
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables}
plot <- wq_long1 %>% 
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, ncol = 4, scales = "free")+
  theme_bw()+
  ggtitle("Delta wide - 1-month lag")
plot
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables with 1 month lag (log transform) 
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables (log transform)}
plot <- wq_long1 %>% 
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, ncol = 4, scales = "free")+
  theme_bw()+
  ggtitle("Log Delta wide - 1 month lag")
plot
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables with 2-month lag
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables}
plot <- wq_long2 %>% 
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, ncol = 4, scales = "free")+
  theme_bw()+
  ggtitle("Delta wide - 2 month lag")
plot
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables with 2-month lag (log transform) 
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables (log transform)}
plot <- wq_long2 %>% 
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, ncol = 4, scales = "free")+
  theme_bw()+
  ggtitle("log Delta wide - 2 month lag")
plot
```


Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables with 3-month lag
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables}
plot <- wq_long3 %>% 
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, ncol = 4, scales = "free")+
  theme_bw()+
  ggtitle("Delta wide - 3-month lag")
plot
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables with 3-month lag (log transform) 
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables (log transform)}
plot <- wq_long3 %>% 
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, ncol = 4, scales = "free")+
  theme_bw()+
  ggtitle("log Delta wide - 3-month lag")
plot
```


Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables with seasonal lag
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables}
plot <- wq_longsn %>% 
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, ncol = 4, scales = "free")+
  theme_bw()+
  ggtitle("Delta wide - seasonal lag")
plot
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables with seasonal lag (log transform) 
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables (log transform)}
plot <- wq_longsn %>% 
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, ncol = 4, scales = "free")+
  theme_bw()+
  ggtitle("log Delta wide - seasonal lag")
plot
```






Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season}
plot_winter <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Winter") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Winter")
plot_winter

plot_spring <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Spring") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Spring")
plot_spring

plot_summer <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Summer") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Summer")
plot_summer


plot_fall <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Fall") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Fall")
plot_fall
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season (log transform)
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season (log transform)}
plot_winter2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Winter") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Winter")
plot_winter2

plot_spring2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Spring") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Spring")
plot_spring2

plot_summer2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Summer") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Summer")
plot_summer2


plot_fall2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Fall") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Fall")
plot_fall2
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Suisun Marsh
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Suisun Marsh}
plot_sm <- wq_long %>% filter(!is.na(Chlorophyll)) %>%
  filter(Regions == "Suisun Marsh") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Marsh")
plot_sm

plot_winter_sm <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Winter") %>%
  filter(Regions == "Suisun Marsh") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Marsh - Winter")
plot_winter_sm

plot_spring_sm <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Spring") %>%
  filter(Regions == "Suisun Marsh") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Marsh - Spring")
plot_spring_sm

plot_summer_sm <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Summer") %>%
  filter(Regions == "Suisun Marsh") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Marsh - Summer")
plot_summer_sm


plot_fall_sm <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Fall") %>%
  filter(Regions == "Suisun Marsh") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Marsh - Fall")
plot_fall_sm
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Suisun Marsh (log transform)
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Suisun Marsh (log transform)}
plot_sm2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>%
  filter(Regions == "Suisun Marsh") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Suisun Marsh")
plot_sm2

plot_winter_sm2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Winter") %>%
  filter(Regions == "Suisun Marsh") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Suisun Marsh - Winter")
plot_winter_sm2

plot_spring_sm2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Spring") %>%
  filter(Regions == "Suisun Marsh") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Suisun Marsh - Spring")
plot_spring_sm2

plot_summer_sm2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Summer") %>%
  filter(Regions == "Suisun Marsh") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Suisun Marsh - Summer")
plot_summer_sm2


plot_fall_sm2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Fall") %>%
  filter(Regions == "Suisun Marsh") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Suisun Marsh - Fall")
plot_fall_sm2
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Suisun Bay
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Suisun Bay}
plot_sb <- wq_long %>% filter(!is.na(Chlorophyll)) %>%
  filter(Regions == "Suisun Bay") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Bay")
plot_sb

plot_winter_sb <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Winter") %>%
  filter(Regions == "Suisun Bay") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Bay - Winter")
plot_winter_sb

plot_spring_sb <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Spring") %>%
  filter(Regions == "Suisun Bay") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Bay - Spring")
plot_spring_sb

plot_summer_sb <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Summer") %>%
  filter(Regions == "Suisun Bay") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Bay - Summer")
plot_summer_sb


plot_fall_sb <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Fall") %>%
  filter(Regions == "Suisun Bay") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Bay - Fall")
plot_fall_sb
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Suisun Bay (log transform)
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Suisun Marsh (log transform)}
plot_sb2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>%
  filter(Regions == "Suisun Bay") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Suisun Bay")
plot_sb2

plot_winter_sb2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Winter") %>%
  filter(Regions == "Suisun Bay") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Suisun Bay - Winter")
plot_winter_sb2

plot_spring_sb2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Spring") %>%
  filter(Regions == "Suisun Bay") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Suisun Bay - Spring")
plot_spring_sb2

plot_summer_sb2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Summer") %>%
  filter(Regions == "Suisun Bay") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Suisun Bay - Summer")
plot_summer_sb2


plot_fall_sb2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Fall") %>%
  filter(Regions == "Suisun Bay") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Suisun Bay - Fall")
plot_fall_sb2
```


Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: North Delta
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: North Delta (log transform)}

plot_nd <- wq_long %>% filter(!is.na(Chlorophyll)) %>%
  filter(Regions == "North Delta") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("North Delta")
plot_nd

plot_winter_nd <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Winter") %>%
  filter(Regions == "North Delta") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("North Delta - Winter")
plot_winter_nd

plot_spring_nd <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Spring") %>%
  filter(Regions == "North Delta") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("North Delta - Spring")
plot_spring_nd

plot_summer_nd <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Summer") %>%
  filter(Regions == "North Delta") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("North Delta - Summer")
plot_summer_nd


plot_fall_nd <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Fall") %>%
  filter(Regions == "North Delta") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("North Delta - Fall")
plot_fall_nd
```


Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: North Delta (log transform)
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: North Delta (log transform)}
plot_nd2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>%
  filter(Regions == "North Delta") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log North Delta")
plot_nd2

plot_winter_nd2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Winter") %>%
  filter(Regions == "North Delta") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log North Delta - Winter")
plot_winter_nd2

plot_spring_nd2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Spring") %>%
  filter(Regions == "North Delta") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log North Delta - Spring")
plot_spring_nd2

plot_summer_nd2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Summer") %>%
  filter(Regions == "North Delta") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log North Delta - Summer")
plot_summer_nd2


plot_fall_nd2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Fall") %>%
  filter(Regions == "North Delta") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log North Delta - Fall")
plot_fall_nd2
```
Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Confluence
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Confluence}
plot_cf <- wq_long %>% filter(!is.na(Chlorophyll)) %>%
  filter(Regions == "Confluence") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Confluence")
plot_cf

plot_winter_cf <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Winter") %>%
  filter(Regions == "Confluence") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Confluence - Winter")
plot_winter_cf

plot_spring_cf <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Spring") %>%
  filter(Regions == "Confluence") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Confluence - Spring")
plot_spring_cf

plot_summer_cf <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Summer") %>%
  filter(Regions == "Confluence") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Confluence - Summer")
plot_summer_cf


plot_fall_cf <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Fall") %>%
  filter(Regions == "Confluence") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Confluence - Fall")
plot_fall_cf
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Confluence (log transform)
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Confluence (log transform)}
plot_cf2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>%
  filter(Regions == "Confluence") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Confluence")
plot_cf2

plot_winter_cf2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Winter") %>%
  filter(Regions == "Confluence") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Confluence - Winter")
plot_winter_cf2

plot_spring_cf2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Spring") %>%
  filter(Regions == "Confluence") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Confluence - Spring")
plot_spring_cf2

plot_summer_cf2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Summer") %>%
  filter(Regions == "Confluence") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Confluence - Summer")
plot_summer_cf2


plot_fall_cf2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Fall") %>%
  filter(Regions == "Confluence") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Confluence - Fall")
plot_fall_cf2
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Central
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Central}
plot_cn <- wq_long %>% filter(!is.na(Chlorophyll)) %>%
  filter(Regions == "Central") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Central")
plot_cn

plot_winter_cn <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Winter") %>%
  filter(Regions == "Central") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Central - Winter")
plot_winter_cn

plot_spring_cn <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Spring") %>%
  filter(Regions == "Central") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Central - Spring")
plot_spring_cn

plot_summer_cn <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Summer") %>%
  filter(Regions == "Central") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Central - Summer")
plot_summer_cn


plot_fall_cn <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Fall") %>%
  filter(Regions == "Central") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Central - Fall")
plot_fall_cn
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Central (log transform)
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Central (log transform)}
plot_cn2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>%
  filter(Regions == "Central") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Central")
plot_cn2

plot_winter_cn2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Winter") %>%
  filter(Regions == "Central") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Central - Winter")
plot_winter_cn2

plot_spring_cn2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Spring") %>%
  filter(Regions == "Central") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Central - Spring")
plot_spring_cn2

plot_summer_cn2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Summer") %>%
  filter(Regions == "Central") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Central - Summer")
plot_summer_cn2


plot_fall_cn2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Fall") %>%
  filter(Regions == "Central") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log Central - Fall")
plot_fall_cn2
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: South
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: South}
plot_sd <- wq_long %>% filter(!is.na(Chlorophyll)) %>%
  filter(Regions == "South") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("South")
plot_sd

plot_winter_sd <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Winter") %>%
  filter(Regions == "South") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("South - Winter")
plot_winter_sd

plot_spring_sd <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Spring") %>%
  filter(Regions == "South") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("South - Spring")
plot_spring_sd

plot_summer_sd <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Summer") %>%
  filter(Regions == "South") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("South - Summer")
plot_summer_sd


plot_fall_sd <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Fall") %>%
  filter(Regions == "South") %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("South - Fall")
plot_fall_sd
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: South (no log transform)
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: South (no log transform)}
plot_sd2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>%
  filter(Regions == "South") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log South")
plot_sd2

plot_winter_sd2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Winter") %>%
  filter(Regions == "South") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log South - Winter")
plot_winter_sd2

plot_spring_sd2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Spring") %>%
  filter(Regions == "South") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log South - Spring")
plot_spring_sd2

plot_summer_sd2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Summer") %>%
  filter(Regions == "South") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log South - Summer")
plot_summer_sd2


plot_fall_sd2 <- wq_long %>% filter(!is.na(Chlorophyll)) %>% filter(season == "Fall") %>%
  filter(Regions == "South") %>%
ggplot(aes(x=logvalue, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Log South - Fall")
plot_fall_sd2
```

