---
title:
  "Binomial Regression"
author: "Alex Berk"
output: 
  html_document:
        toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

```{r, include=FALSE}

library(tidyverse)
library(lubridate)
library(mgcv)
library(marginaleffects)
library(MuMIn)
library(car)
library(corrplot)
library(stats)
library(tseries)
library(DHARMa)
library(mice)

conflicted::conflict_prefer("select","dplyr")
```

# Prepare data

```{r}

options(na.action = "na.exclude")

# variables to include:
c = c("bloom","Regions", "Date","Year", "Month",  "season","SAC","lagSAC",
       "EXPORT","logEXPORT","OUT","lagOUT", "SJR", "lagSJR", "Index","lagIndex","DissAmmonia",
       "lagDissAmmonia", "TotPhos","lagTotPhos","DissNitrateNitrite", 
       "lagDissNitrateNitrite","Temperature","lagTemperature","Secchi", 
       "lagSecchi","Conductivity",  "lagConductivity","season_num", "DissOrthophos", "lagDissOrthophos", "clam_step",
       "logDissNitrateNitrite","logDissAmmonia","logDissOrthophos","logConductivity","logSAC","logOUT")


df <- read.csv("Data/regional_integrated_dataset3.csv" )

df$Date = as.Date(df$Date)

# 1979 is the start of the study period.
df <- subset(df,df$Date >= "1979-01-01")

# Changing categorical variables to factors
df$Regions = as.factor(df$Regions)
df$season = as.factor(df$season)
df$season_num = as.numeric(df$season)


#Replacing clam variables with a step function (pre/post potamocorbula introduction to the Bay in 1987)
df <- df%>%mutate(clam_step=if_else(Year>1987, "post clam", "pre clam")) 
df$clam_step <- as.factor(df$clam_step)

# log transforming some variables to better visualize
df$logDissNitrateNitrite <- log(df$DissNitrateNitrite)
df$logDissAmmonia <- log(df$DissAmmonia)
df$logDissOrthophos <- log(df$DissOrthophos)
df$logConductivity <- log(df$Conductivity)
df$logSAC <- log(df$SAC)
df$logOUT <- log(df$OUT)
df$logEXPORT <- log(df$EXPORT)

#Binary transformation of bloom absence/presence
bloom_thresholds <- df %>%
  group_by(Regions) %>%
  summarise(mean = mean(Chlorophyll,na.rm=T),sd=sd(Chlorophyll,na.rm=T)) %>%
  mutate(bloom_thr = mean + sd) %>%
  select(-mean,-sd)

#mean/d group by region
#HABs vs BABs coming from different dataset
  
df <- df %>%
  merge(bloom_thresholds,by="Regions",all.x=T) %>%
  mutate(bloom = case_when(Chlorophyll > bloom_thr ~ 1,
                           Chlorophyll < bloom_thr ~ 0))

df <- subset(df, select = c) 

paste0("Number of blooms = ", sum(df$bloom))

# z-score standardizing the dataset for more logical comparison of predictors. This facilitates comparison of effect sizes between variables.
df <- df %>% mutate(across(!c(bloom,Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

df
```


# Impute Dataset
Taken from RH's Markdown file "Global_GAM_RH.html" for detailed checks. Only difference is the "bloom" value in the place of the "logChlorophyll" value.

Question: Should the bloom binary be calculated AFTER the imputation? Should Chlorophyll be imputed instead of the bloom factor? 

```{r}
ini <- mice(df, maxit=0)
ini$nmis
```
## Impute

```{r}
imp <- mice(df,seed = 123)
class (imp)
```

## Make imputed datasets

```{r}
df1<- complete(imp, action=1, include=FALSE)
df2<- complete(imp, action=2, include=FALSE)
df3<- complete(imp, action=3, include=FALSE)
df4<- complete(imp, action=4, include=FALSE)
df5<- complete(imp, action=5, include=FALSE)
```

## Add lag terms to the imputed dataset

```{r}
df_list = list(df1,df2,df3,df4,df5)

df_list_lag <- lapply(df_list, function(y) {
#Step 1: summarize by season 
wq_r_sum_lagseason <- df1 %>%
  group_by(Regions,Year,season) %>%
  summarize(across(where(is.numeric),\(x) mean(x, na.rm = TRUE)))

#Step 2: Add a column for the following season.
#This seems counter-intuitive because the lag should be the prior season,
#but this is a merge column: it will merge back with the prior season
#in the original table. 
wq_r_sum_lagseason <- wq_r_sum_lagseason %>%
  select(!c(lagOUT,lagSAC,lagSJR,lagTemperature,lagIndex,lagDissAmmonia,lagDissNitrateNitrite, lagTotPhos,lagConductivity,lagDissOrthophos))%>%
  rename_with(~paste0("lag",.x)) %>%
  rename(Year = "lagYear",season = "lagseason",Regions="lagRegions") %>% 
  mutate(lagseason = case_when(
    season == "Spring" ~ "Summer",
    season == "Summer" ~ "Autumn",
    season == "Autumn" ~ "Winter",
    season == "Winter" ~ "Spring" 
  )) %>%
  mutate(lagseasonyear = case_when(
    season == "Autumn" ~ Year+1,
    !(season == "Autumn") ~ Year
  )) %>%
  mutate(lagseasonyear = paste0(lagseason,lagseasonyear)) %>%
  ungroup %>% 
  select(lagseasonyear, Regions, laglogDissAmmonia,laglogDissNitrateNitrite,laglogConductivity,
         laglogSAC,laglogOUT, laglogDissOrthophos, laglogEXPORT)


y <- y %>% 
  mutate(seasonyear = paste0(season,year(Date))) %>%
  merge(wq_r_sum_lagseason,by.x=c("Regions","seasonyear"),by.y=c("Regions","lagseasonyear"),all.x=T) %>%
  select(-seasonyear)%>%
  relocate(Date)})

df_1 <- as.data.frame(df_list_lag[1])
df_2 <- as.data.frame(df_list_lag[2])
df_3 <- as.data.frame(df_list_lag[3])
df_4 <- as.data.frame(df_list_lag[4])
df_5 <- as.data.frame(df_list_lag[5])
```


# Fit models with imputed datasets
Use Ryan's top models from the Global GAM result to run a binomial GAM. 

## bm4.po4
Orthophosphos with SAC  (no EXPORT)
```{r}
bm4.po4.imp1 <- gam(bloom ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite+
               DissOrthophos +
               lagDissOrthophos,
             family=binomial(link = "logit"),data=df_1,select=T)

bm4.po4.imp2 <- gam(bloom ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite+
               DissOrthophos +
               lagDissOrthophos,
             family=binomial(link = "logit"),data=df_2,select=T)

bm4.po4.imp3 <- gam(bloom ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite+
               DissOrthophos +
               lagDissOrthophos,
             family=binomial(link = "logit"),data=df_3,select=T)

bm4.po4.imp4 <- gam(bloom ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite+
               DissOrthophos +
               lagDissOrthophos,
             family=binomial(link = "logit"),data=df_4,select=T)

bm4.po4.imp5 <- gam(bloom ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite+
               DissOrthophos +
               lagDissOrthophos,
             family=binomial(link = "logit"),data=df_5,select=T)

model_list_bm4 = list(bm4.po4.imp1,bm4.po4.imp2,bm4.po4.imp3,bm4.po4.imp4,bm4.po4.imp5)

```

## bm6.po4
Orthophosphos with SAC+ EXPORT

```{r}
bm6.po4.imp1 <- gam(bloom ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               EXPORT +
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite+
               DissOrthophos +
               lagDissOrthophos,
             family=binomial(link = "logit"),data=df_1,select=T)

bm6.po4.imp2 <- gam(bloom ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               EXPORT +
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite+
               DissOrthophos +
               lagDissOrthophos,
             family=binomial(link = "logit"),data=df_2,select=T)

bm6.po4.imp3 <- gam(bloom ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               EXPORT +
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite+
               DissOrthophos +
               lagDissOrthophos,
             family=binomial(link = "logit"),data=df_3,select=T)

bm6.po4.imp4 <- gam(bloom ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               EXPORT +
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite+
               DissOrthophos +
               lagDissOrthophos,
             family=binomial(link = "logit"),data=df_4,select=T)

bm6.po4.imp5 <- gam(bloom ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               EXPORT +
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite+
               DissOrthophos +
               lagDissOrthophos,
             family=binomial(link = "logit"),data=df_5,select=T)

model_list_bm6 <- list(bm6.po4.imp1,bm6.po4.imp2,bm6.po4.imp3,bm6.po4.imp4,bm6.po4.imp5)

```

# AIC, BIC (???)
Do I compare each imputation or do I do this after I pool them? Do I keep my pre-imp process?

## Compare with AIC

```{r}
AIC(bm4.po4.imp1,bm4.po4.imp2,bm4.po4.imp3,bm4.po4.imp4,bm4.po4.imp5,bm6.po4.imp1,bm6.po4.imp2,bm6.po4.imp3,bm6.po4.imp4,bm6.po4.imp5)
```

## Compare with BIC

```{r}
BIC(bm4.po4.imp1,bm4.po4.imp2,bm4.po4.imp3,bm4.po4.imp4,bm4.po4.imp5,bm6.po4.imp1,bm6.po4.imp2,bm6.po4.imp3,bm6.po4.imp4,bm6.po4.imp5)
```


# Shrink Models (???)
Do I shrink each imputation or do I pool them? Or do I do this pre-imputation? 
## bm.po4

```{r}
bm6.po4.imp1.shrink <- gam(bloom ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity) +
               s(SAC) + 
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               s(lagSAC) +
               s(lagDissAmmonia) +
               s(lagDissNitrateNitrite)+
               s(DissOrthophos) +
               s(lagDissOrthophos),
             family=binomial(link = "logit"),data=df_1,select=T)
summary(bm6.po4.imp1.shrink)
```


## bm.po6


```{r}
bm6.po4.imp1.shrink <- gam(bloom ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity) +
               s(SAC) +
               s(EXPORT) +   
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               s(lagSAC) +
               s(lagDissAmmonia) +
               s(lagDissNitrateNitrite)+
               s(DissOrthophos) +
               s(lagDissOrthophos),
             family=binomial(link = "logit"),data=df_1,select=T)
summary(bm6.po4.imp1.shrink)
```


