---
title:
  "Binomial Regression"
author: "Alex Berk"
output: 
  html_document:
        toc: TRUE
        toc_float: TRUE
        code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```


```{r, include=FALSE}

library(tidyverse)
library(lubridate)
library(mgcv)
library(marginaleffects)
library(MuMIn)
library(car)
library(corrplot)
library(stats)
library(tseries)
library(DHARMa)
library(mice)
library(lme4)

conflicted::conflict_prefer("select","dplyr")
conflicted::conflict_prefer("filter","dplyr")
```

# Prepare data

```{r}

# variables to include:
c = c("Chlorophyll", "Regions", "Date","Year", "Month",  "season","SAC","lagSAC",
       "EXPORT","logEXPORT","OUT","lagOUT", "DissAmmonia",
       "lagDissAmmonia", "TotPhos","lagTotPhos","DissNitrateNitrite", 
       "lagDissNitrateNitrite","Temperature","lagTemperature","Secchi", 
       "lagSecchi","Conductivity","lagConductivity","season_num", "DissOrthophos", "lagDissOrthophos", "clam_step","logDissNitrateNitrite","logDissAmmonia","logDissOrthophos","logConductivity","logSAC","logOUT")#, "Biomass","Grazing_rate","Filtration_rate", "SJR","lagSJR", "Index","lagIndex")

df <- read.csv("Data/regional_integrated_dataset3.csv" )

df$Date = as.Date(df$Date)

# 1979 is the start of the study period.
df <- subset(df,df$Date >= "1979-01-01" & df$Date < "2023-01-01")

# Changing categorical variables to factors
df$Regions = as.factor(df$Regions)
df$season = as.factor(df$season)
df$season_num = as.numeric(df$season)


#Replacing clam variables with a step function (pre/post potamocorbula introduction to the Bay in 1987)
df <- df%>%mutate(clam_step=if_else(Year>1987, "post clam", "pre clam")) 
df$clam_step <- as.factor(df$clam_step)


# log transforming some variables to better visualize
df$logDissNitrateNitrite <- log(df$DissNitrateNitrite)
df$logDissAmmonia <- log(df$DissAmmonia)
df$logDissOrthophos <- log(df$DissOrthophos)
df$logConductivity <- log(df$Conductivity)
df$logSAC <- log(df$SAC)
df$logOUT <- log(df$OUT)
df$logEXPORT <- log(df$EXPORT)

# Dropping NAs from the dataset, only for dredging.

 df <- subset(df, select = c) #%>%
    #drop_na()

```

## Separate Dataset Regionally
May need to separate by flow as well, check from Ryan

```{r}

df_flow <- df[,c("Date","Regions","SAC", "OUT",  "EXPORT", "logEXPORT","logOUT","lagOUT","logSAC","lagSAC")]

df_no_flow <- df[,!(names(df) %in% c("SAC", "logSAC", "OUT", "logOUT", "EXPORT", "logEXPORT","lagOUT", "lagSAC"))]

r <- unique(df_no_flow$Regions)
region_list <- list()
for (region in r){
  df_regional <- df_no_flow %>% filter(Regions == region)
  region <- gsub(" ","",region)
  region_list[[region]] <- df_regional
}

```

## Impute

```{r}
regional_imp <- list()
for (region_name in as.character(names(region_list))){
  current_df <- region_list[[region_name]]
  predictors <- quickpred(current_df)
  imp <- mice(current_df,seed = 123, print=F, predictorMatrix = predictors)
  new_name <- paste0(region_name, "_imp")
  regional_imp[[new_name]]<- imp
}
```

## Make combined imputed datasets
```{r}
regional_imp_dfs <- list()
action_seq <- as.list(seq(1,5))
for (region_imp in as.character(names(regional_imp))){
      current_imp <- regional_imp[[region_imp]]
      for (i in action_seq){
        df_new <- complete(current_imp,action = i, include=FALSE)
        df_new <- left_join( df_new,df_flow, by=c("Date","Regions"))
        new_name <- paste0(region_imp, "_df", i)
        regional_imp_dfs[[new_name]]<- df_new}}
```


## Add Lag terms to imputed datasets

```{r}
# This code is taken from clamdata_cleanup_AB_261125.R

df_list_lag <- lapply(regional_imp_dfs, function(y) {
#Step 1: summarize by season 
wq_r_sum_lagseason <- y %>%
  group_by(Regions,Year,season) %>%
  summarize(across(where(is.numeric),\(x) mean(x, na.rm = TRUE)), .groups="drop_last")

#Step 2: Add a column for the following season.
#This seems counter-intuitive because the lag should be the prior season,
#but this is a merge column: it will merge back with the prior season
#in the original table. 
wq_r_sum_lagseason <- wq_r_sum_lagseason %>%
  select(!c(lagOUT,lagSAC,lagTemperature,lagDissAmmonia,lagDissNitrateNitrite, lagTotPhos,lagConductivity,lagDissOrthophos))%>%
  rename_with(~paste0("lag",.x)) %>%
  rename(Year = "lagYear",season = "lagseason",Regions="lagRegions") %>% 
  mutate(lagseason = case_when(
    season == "Spring" ~ "Summer",
    season == "Summer" ~ "Autumn",
    season == "Autumn" ~ "Winter",
    season == "Winter" ~ "Spring" 
  )) %>%
  mutate(lagseasonyear = case_when(
    season == "Autumn" ~ Year+1,
    !(season == "Autumn") ~ Year
  )) %>%
  mutate(lagseasonyear = paste0(lagseason,lagseasonyear)) %>%
  ungroup %>% 
  select(lagseasonyear, Regions, laglogDissAmmonia,laglogDissNitrateNitrite,laglogConductivity,
         laglogSAC,laglogOUT, laglogDissOrthophos, laglogEXPORT)


y <- y %>% 
  mutate(seasonyear = paste0(season,year(Date))) %>%
  merge(wq_r_sum_lagseason,by.x=c("Regions","seasonyear"),by.y=c("Regions","lagseasonyear"),all.x=T) %>%
  select(-seasonyear)%>%
  relocate(Date)
})
```


## Combine regional imputed datsets
```{r}
### df1 (save natural scale as dataset for later calculations)
df1_nat <- distinct(rbind(df_list_lag$Central_imp_df1, df_list_lag$Confluence_imp_df1,
             df_list_lag$NorthDelta_imp_df1, df_list_lag$South_imp_df1,df_list_lag$SuisunBay_imp_df1)) 

```


## Add bloom threshold to imputation 1

```{r}
#Binary transformation of bloom absence/presence
bloom_thresholds <- df1_nat %>%
  group_by(Regions) %>%
  summarise(mean = mean(Chlorophyll,na.rm=T),sd=sd(Chlorophyll,na.rm=T)) %>%
  mutate(bloom_thr = mean + sd) %>%
  select(-mean,-sd)

#Imputation 1 
df1_nat <- df1_nat %>%
  merge(bloom_thresholds,by="Regions",all.x=T) %>%
  mutate(bloom = case_when(Chlorophyll > bloom_thr ~ 1,
                           Chlorophyll < bloom_thr ~ 0))

base::paste0("number of blooms: ",toString(sum(df1_nat$bloom))) 
```


## Z-score

```{r}
# z-score standardizing the dataset for easier comparison of predictors. This facilitates comparison of effect sizes between variables.

df1 <- df1_nat %>% mutate(across(!c(bloom,Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

```


# Fit models with imputed datasets

Use Ryan's top models from the Global GAM result to run a binomial GAM: Models 4 and 6 with PO4.

## bm3.po4
```{r}
bm3.po4.imp1 <- gam(bloom ~ DissAmmonia + 
                      DissNitrateNitrite + 
                      lagDissNitrateNitrite + 
                      Temperature + 
                      Conductivity + 
                      s(Regions, bs='re') + 
                      season + 
                      (OUT*season) + 
                      clam_step + 
                      OUT + 
                      lagOUT + 
                      lagDissAmmonia + 
                      DissOrthophos + 
                      lagDissOrthophos,
             family=binomial(link = "logit"),data=df1,select=T)
```

log transformations
```{r}
bm3.po4.imp1.log <- gam(bloom ~ logDissAmmonia + 
                      logDissNitrateNitrite + 
                      laglogDissNitrateNitrite + 
                      Temperature + 
                      logConductivity + 
                      s(Regions, bs='re') + 
                      season + 
                      (logOUT*season) + 
                      clam_step + 
                      logOUT + 
                      laglogOUT + 
                      laglogDissAmmonia + 
                      logDissOrthophos + 
                      laglogDissOrthophos,
             family=binomial(link = "logit"),data=df1,select=T)
```

glm
```{r}
bm3.po4.imp1.glm <- glmer(bloom ~ logDissAmmonia + 
                      logDissNitrateNitrite + 
                      laglogDissNitrateNitrite + 
                      Temperature + 
                      logConductivity + 
                      (1|Regions) + 
                      season + 
                      (logOUT*season) + 
                      clam_step + 
                      logOUT + 
                      laglogOUT + 
                      laglogDissAmmonia + 
                      logDissOrthophos + 
                      laglogDissOrthophos,
             family=binomial(link = "logit"),data=df1)
```


## bm4.po4

Orthophosphos with SAC (no EXPORT)

```{r}
bm4.po4.imp1 <- gam(bloom ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite+
               DissOrthophos +
               lagDissOrthophos,
             family=binomial(link = "logit"),data=df1,select=T)

```

With log transformations:

```{r}
bm4.po4.imp1.log <- gam(bloom ~ logDissAmmonia +
               logDissNitrateNitrite +
               Temperature +
               logConductivity +
               logSAC + 
               s(Regions, bs='re') +
               season +
               (logSAC*season) +
               clam_step +
               laglogSAC +
               laglogDissAmmonia +
               laglogDissNitrateNitrite+
               logDissOrthophos +
               laglogDissOrthophos,
             family=binomial(link = "logit"),data=df1,select=T)
```


glm:
```{r}
bm4.po4.imp1.glm <- glmer(bloom ~ logDissAmmonia +
               logDissNitrateNitrite +
               Temperature +
               logConductivity +
               logSAC + 
               (1|Regions) +
               season +
               (logSAC*season) +
               clam_step +
               laglogSAC +
               laglogDissAmmonia +
               laglogDissNitrateNitrite+
               logDissOrthophos +
               laglogDissOrthophos,
             family=binomial(link = "logit"),data=df1)
```



## bm6.po4

Orthophosphos with SAC+ EXPORT

```{r}
bm6.po4.imp1 <- gam(bloom ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               EXPORT +
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite+
               DissOrthophos +
               lagDissOrthophos,
             family=binomial(link = "logit"),data=df1,select=T)
```

Log transform:
```{r}
bm6.po4.imp1.log <- gam(bloom ~ logDissAmmonia +
               logDissNitrateNitrite +
               Temperature +
               logConductivity +
               logSAC +
               EXPORT +
               s(Regions, bs='re') +
               season +
               (logSAC*season) +
               clam_step +
               laglogSAC +
               laglogDissAmmonia +
               laglogDissNitrateNitrite+
               logDissOrthophos +
               laglogDissOrthophos,
             family=binomial(link = "logit"),data=df1,select=T)
```

glm:
```{r}
bm6.po4.imp1.glm <- glmer(bloom ~ logDissAmmonia +
               logDissNitrateNitrite +
               Temperature +
               logConductivity +
               logSAC +
               EXPORT +
               (1|Regions) +
               season +
               (logSAC*season) +
               clam_step +
               laglogSAC +
               laglogDissAmmonia +
               laglogDissNitrateNitrite+
               logDissOrthophos +
               laglogDissOrthophos,
             family=binomial(link = "logit"),data=df1)
```


# Shrink Models {.tabset}

## bm3.po4
```{r}
bm3.po4.imp1.shrink <- gam(bloom ~ s(DissAmmonia) + 
                             s(DissNitrateNitrite) + 
                             s(lagDissNitrateNitrite) + 
                             s(Temperature) + 
                             s(Conductivity) + 
                             s(Regions, bs='re') + 
                             season + 
                             s(OUT,by=season) + 
                             clam_step + 
                             s(OUT) + 
                             s(lagOUT) + 
                             s(lagDissAmmonia)+ 
                             s(DissOrthophos)+ 
                             s(lagDissOrthophos),
             family=binomial(link = "logit"),data=df1,select=T)
summary(bm3.po4.imp1.shrink)
```

##bm3.po4.log
```{r}
bm3.po4.imp1.log.shrink <- gam(bloom ~ s(logDissAmmonia) + 
                             s(logDissNitrateNitrite) + 
                             s(laglogDissNitrateNitrite) + 
                             s(Temperature) + 
                             s(logConductivity) + 
                             s(Regions, bs='re') + 
                             season + 
                             s(logOUT,by=season) + 
                             clam_step + 
                             s(logOUT) + 
                             s(laglogOUT) + 
                             s(laglogDissAmmonia)+ 
                             s(logDissOrthophos)+ 
                             s(laglogDissOrthophos),
             family=binomial(link = "logit"),data=df1,select=T)
summary(bm3.po4.imp1.log.shrink)
```


## bm4.po4

```{r}
bm4.po4.imp1.shrink <- gam(bloom ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity) +
               s(SAC) + 
               s(Regions, bs='re') +
               season +
               s(OUT,by=season) +
               clam_step +
               s(lagSAC) +
               s(lagDissAmmonia) +
               s(lagDissNitrateNitrite)+
               s(DissOrthophos) +
               s(lagDissOrthophos),
             family=binomial(link = "logit"),data=df1,select=T)
summary(bm4.po4.imp1.shrink)
```

## bm4.po4.log
```{r}
bm4.po4.imp1.log.shrink <- gam(bloom ~ s(logDissAmmonia) +
               s(logDissNitrateNitrite) +
               s(Temperature) +
               s(logConductivity) +
               s(logSAC) + 
               s(Regions, bs='re') +
               season +
               s(logSAC,by=season) +
               clam_step +
               s(laglogSAC) +
               s(laglogDissAmmonia) +
               s(laglogDissNitrateNitrite)+
               s(logDissOrthophos) +
               s(laglogDissOrthophos),
             family=binomial(link = "logit"),data=df1,select=T)
summary(bm4.po4.imp1.log.shrink)
```

## bm6.po4

```{r}
bm6.po4.imp1.shrink <- gam(bloom ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity) +
               s(SAC) +
               s(EXPORT) +   
               s(Regions, bs='re') +
               season +
               s(SAC,by=season) +
               clam_step +
               s(lagSAC) +
               s(lagDissAmmonia) +
               s(lagDissNitrateNitrite)+
               s(DissOrthophos) +
               s(lagDissOrthophos),
             family=binomial(link = "logit"),data=df1,select=T)
summary(bm6.po4.imp1.shrink)
```

## bm6.po4.log
```{r}
bm6.po4.imp1.log.shrink <- gam(bloom ~ s(logDissAmmonia) +
               s(logDissNitrateNitrite) +
               s(Temperature) +
               s(logConductivity) +
               s(logSAC) +
               s(logEXPORT) +   
               s(Regions, bs='re') +
               season +
               s(logSAC,by=season) +
               clam_step +
               s(laglogSAC) +
               s(laglogDissAmmonia) +
               s(laglogDissNitrateNitrite)+
               s(logDissOrthophos) +
               s(laglogDissOrthophos),
             family=binomial(link = "logit"),data=df1,select=T)
summary(bm6.po4.imp1.log.shrink)
```


# Diagnostics

## AIC

```{r}
AIC(bm3.po4.imp1,bm4.po4.imp1, bm6.po4.imp1,bm3.po4.imp1.log,bm4.po4.imp1.log, bm6.po4.imp1.log,bm3.po4.imp1.shrink, bm4.po4.imp1.shrink, bm6.po4.imp1.shrink,bm3.po4.imp1.log.shrink,bm4.po4.imp1.log.shrink, bm6.po4.imp1.log.shrink,
    bm3.po4.imp1.glm,bm4.po4.imp1.glm,bm6.po4.imp1.glm)
```


## BIC

```{r}
BIC(bm3.po4.imp1,bm4.po4.imp1, bm6.po4.imp1,bm3.po4.imp1.log,bm4.po4.imp1.log, bm6.po4.imp1.log,bm3.po4.imp1.shrink, bm4.po4.imp1.shrink, bm6.po4.imp1.shrink,bm3.po4.imp1.log.shrink,bm4.po4.imp1.log.shrink, bm6.po4.imp1.log.shrink,
    bm3.po4.imp1.glm,bm4.po4.imp1.glm,bm6.po4.imp1.glm)
```

## Residuals {.tabset}

### bm3 {.tabset}

#### bm3.po4.imp1
```{r}
testDispersion(bm3.po4.imp1)
plot(simulateResiduals(bm3.po4.imp1))
```

#### bm3.po4.imp1.log
```{r}
testDispersion(bm3.po4.imp1.log)
plot(simulateResiduals(bm3.po4.imp1.log))
```

#### bm3.po4.imp1.shrink
```{r}
gam.check(bm3.po4.imp1.shrink)
```

#### bm3.po4.imp1.log.shrink
```{r}
gam.check(bm3.po4.imp1.log.shrink)
```

### bm4 {.tabset}

#### bm4.po4.imp1
```{r}
testDispersion(bm4.po4.imp1)
plot(simulateResiduals(bm4.po4.imp1))
```

#### bm4.po4.imp1.log
```{r}
testDispersion(bm4.po4.imp1.log)
plot(simulateResiduals(bm4.po4.imp1.log))
```

#### bm4.po4.imp1.shrink
```{r}
gam.check(bm4.po4.imp1.shrink)
```

#### bm4.po4.imp1.log.shrink
```{r}
gam.check(bm4.po4.imp1.log.shrink)
```

### bm6 {.tabset}

#### bm6.po4.imp1
```{r}
testDispersion(bm6.po4.imp1)
plot(simulateResiduals(bm6.po4.imp1))
```

#### bm6.po4.imp1.log
```{r}
testDispersion(bm6.po4.imp1.log)
plot(simulateResiduals(bm6.po4.imp1.log))
```

#### bm6.po4.imp1.shrink
```{r}
gam.check(bm6.po4.imp1.shrink)
```

#### bm6.po4.imp1.log.shrink
```{r}
gam.check(bm6.po4.imp1.log.shrink)
```

### NEW - GLMs {.tabset}

#### bm3.po4.imp1.glm
```{r}
testDispersion(bm3.po4.imp1.glm)
plot(simulateResiduals(bm3.po4.imp1.glm))
```

#### bm4.po4.imp1.glm
```{r}
testDispersion(bm4.po4.imp1.glm)
plot(simulateResiduals(bm4.po4.imp1.glm))
```

#### bm6.po4.imp1.glm
```{r}
testDispersion(bm6.po4.imp1.glm)
plot(simulateResiduals(bm6.po4.imp1.glm))
```


# LD10
https://www.ccohs.ca/oshanswers/chemicals/ld50.html

have to back calc from z score
https://stackoverflow.com/questions/21594916/convert-normalized-variables-back-to-non-normalized-values 
x = std*z + m

## Vary temp and hold others constant

Will probably make this into a function to apply to other variables, but want to test it first!
```{r}

# #dataframe with the range of a parameter value and holding the others constant
# df_pred <- data.frame(
#   Temperature = seq(min(df1$Temperature,na.rm=T), max(df1$Temperature,na.rm=T), length.out = 1000),
#   clam_step = as.factor("post clam"),
#   season = as.factor("Summer"),
#   Regions = as.factor("Central")
#   )

# df_pred$Regions = fct_expand(df_pred$Regions,c("Central","Confluence","North Delta", "South", "Suisun Bay"))
# df_pred$season = fct_expand(df_pred$season,c("Autumn","Spring","Summer","Winter"))
# df_pred$clam_step = fct_expand(df_pred$clam_step,c("post clam","pre clam"))
# 
# df_mean <- df1 %>%
#   select(-clam_step,-Temperature) %>%
#   group_by() %>%
#   summarize(across(DissAmmonia:laglogEXPORT, ~ median(.x, na.rm = TRUE)))
# 
# df_pred <- bind_cols(df_pred,df_mean)  
# ```
# 
# ## Predict on the RESPONSE scale
# ```{r}
# 
# df_pred$predicted <- predict(bm4.po4.imp1.glm, newdata = df_pred, type = "response")
# ```
# 
# ## Linear interpolation to find where predicted probability crosses 0.10
# ```{r}
# 
# df1_stats <- df1_nat %>%
#   select(Date,DissAmmonia,DissNitrateNitrite,Temperature,lagDissAmmonia,
#          lagDissNitrateNitrite,DissOrthophos,lagDissOrthophos,
#          logDissAmmonia,logDissNitrateNitrite,laglogDissAmmonia,
#          laglogDissNitrateNitrite,logDissOrthophos,laglogDissOrthophos) %>%
#   pivot_longer(cols = !"Date",names_to = "Parameter") %>%
#   group_by(Parameter) %>%
#   summarize(mean = mean(value,na.rm=T),sd = sd(value,na.rm=T))
# 
# ld10 <- approx(x = df_pred$predicted, y = df_pred$Temperature, xout = 0.10)$y
# ld10_nat<-ld10*df1_stats[4,3]+ df1_stats[4,2]
# ld10_nat

```

```{r}
df1_stats2 <- df1_nat %>%
  select(colnames(model.frame(bm4.po4.imp1.glm)),-bloom) %>%
  pivot_longer(cols = is.numeric,names_to = "Parameter") %>%
  group_by(Parameter) %>%
  summarize(mean = mean(value,na.rm=T),sd = sd(value,na.rm=T))

ld10tab <- as.data.frame(matrix(data = NA,ncol=nrow(df1_stats2),nrow=1))
colnames(ld10tab) <- df1_stats2$Parameter 

for(i in colnames(ld10tab)) {
  #dataframe with the range of a parameter value and holding the others constant
  
  df_pred2 <- data.frame(
    parameter = seq(min(df1[[i]],na.rm=T), max(df1[[i]],na.rm=T), length.out = 1000),
    clam_step = as.factor("post clam"),
    season = as.factor("Summer"),
    Regions = as.factor("Central")) %>%
    rename(!!i := parameter)
  
  #expand factors
  df_pred2$Regions = fct_expand(df_pred2$Regions,c("Central","Confluence","North Delta",
                                                 "South", "Suisun Bay"))
  df_pred2$season = fct_expand(df_pred2$season,c("Autumn","Spring","Summer","Winter"))
  df_pred2$clam_step = fct_expand(df_pred2$clam_step,c("post clam","pre clam"))
  
  #find median of all other factors
  df_mean <- df1 %>%
    group_by() %>%
    summarize_if(is.numeric, ~ median(.x, na.rm = TRUE)) %>%
    select(c(colnames(ld10tab)),-c(i),-bloom)
  
  df_pred2 <- bind_cols(df_pred2,df_mean)
  
  #predict response scale
  df_pred2$predicted <- predict(bm4.po4.imp1.glm, newdata = df_pred, type = "response")
  
  ld10 <- approx(x = df_pred2$predicted, y = df_pred2[[i]], xout = 0.10)$y
  ld10_nat<-ld10*df1_stats2[df1_stats2$Parameter == i,3]+df1_stats2[df1_stats2$Parameter == i,2]
  ld10tab[[i]] <- as.numeric(ld10_nat)
  }
  

```
