---
title:
  "Binomial Regression"
author: "Alex Berk"
output: 
  html_document:
        toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

# CURRENT QUESTIONS 
1. When do I shrink now? All imputations? Before imputing? 
2. Similarly with AIC vs BIC - do I run this check for all imputations? Do I run models on the on the pre-imputed dataset first and use those AIC and BIC values? 
3. Should I calculate the bloom factor (0 or 1) after imputing so it reflects the imputed chlorophyll? (asked in light of Ryan re-calculating all the lag terms) 
4. Do I need to do log transformations as well with the binomial GAM or do you want to see results first?

```{r, include=FALSE}

library(tidyverse)
library(lubridate)
library(mgcv)
library(marginaleffects)
library(MuMIn)
library(car)
library(corrplot)
library(stats)
library(tseries)
library(DHARMa)
library(mice)

conflicted::conflict_prefer("select","dplyr")
conflicted::conflict_prefer("filter","dplyr")
```

# Prepare data

```{r}

# variables to include:
c = c("Chlorophyll", "Regions", "Date","Year", "Month",  "season","SAC","lagSAC",
       "EXPORT","logEXPORT","OUT","lagOUT", "DissAmmonia",
       "lagDissAmmonia", "TotPhos","lagTotPhos","DissNitrateNitrite", 
       "lagDissNitrateNitrite","Temperature","lagTemperature","Secchi", 
       "lagSecchi","Conductivity","lagConductivity","season_num", "DissOrthophos", "lagDissOrthophos", "clam_step","logDissNitrateNitrite","logDissAmmonia","logDissOrthophos","logConductivity","logSAC","logOUT")#, "Biomass","Grazing_rate","Filtration_rate", "SJR","lagSJR", "Index","lagIndex")

df <- read.csv("Data/regional_integrated_dataset3.csv" )

df$Date = as.Date(df$Date)

# 1979 is the start of the study period.
df <- subset(df,df$Date >= "1979-01-01")

# Changing categorical variables to factors
df$Regions = as.factor(df$Regions)
df$season = as.factor(df$season)
df$season_num = as.numeric(df$season)


#Replacing clam variables with a step function (pre/post potamocorbula introduction to the Bay in 1987)
df <- df%>%mutate(clam_step=if_else(Year>1987, "post clam", "pre clam")) 
df$clam_step <- as.factor(df$clam_step)


# log transforming some variables to better visualize
df$logDissNitrateNitrite <- log(df$DissNitrateNitrite)
df$logDissAmmonia <- log(df$DissAmmonia)
df$logDissOrthophos <- log(df$DissOrthophos)
df$logConductivity <- log(df$Conductivity)
df$logSAC <- log(df$SAC)
df$logOUT <- log(df$OUT)
df$logEXPORT <- log(df$EXPORT)

# Dropping NAs from the dataset, only for dredging.

 df <- subset(df, select = c) #%>%
    #drop_na()

```

## Separate Dataset Regionally

```{r}
region_list <- list()
for (region in unique(df$Regions)){
  df_regional <- df %>% filter(Regions == region)
  region <- gsub(" ","",region)
  region_list[[region]] <- df_regional
}
```

## Impute

```{r}
regional_imp <- list()
for (region_name in as.character(names(region_list))){
  current_df <- region_list[[region_name]]
  imp <- mice(current_df,seed = 123, print=F)
  new_name <- paste0(region_name, "_imp")
  regional_imp[[new_name]]<- imp
}
```

## Make combined imputed datasets
```{r}
regional_imp_dfs <- list()
action_seq <- as.list(seq(1,5))
for (region_imp in as.character(names(regional_imp))){
      current_imp <- regional_imp[[region_imp]]
      for (i in action_seq){
        df_new <- complete(current_imp,action = i, include=FALSE)
        new_name <- paste0(region_imp, "_df", i)
        regional_imp_dfs[[new_name]]<- df_new}}
```


## Add Lag terms to imputed datasets

```{r}
df_list_lag <- lapply(regional_imp_dfs, function(y) {
#Step 1: summarize by season 
wq_r_sum_lagseason <- y %>%
  group_by(Regions,Year,season) %>%
  summarize(across(where(is.numeric),\(x) mean(x, na.rm = TRUE)))

#Step 2: Add a column for the following season.
#This seems counter-intuitive because the lag should be the prior season,
#but this is a merge column: it will merge back with the prior season
#in the original table. 
wq_r_sum_lagseason <- wq_r_sum_lagseason %>%
  select(!c(lagOUT,lagSAC,lagTemperature,lagDissAmmonia,lagDissNitrateNitrite, lagTotPhos,lagConductivity,lagDissOrthophos))%>%
  rename_with(~paste0("lag",.x)) %>%
  rename(Year = "lagYear",season = "lagseason",Regions="lagRegions") %>% 
  mutate(lagseason = case_when(
    season == "Spring" ~ "Summer",
    season == "Summer" ~ "Autumn",
    season == "Autumn" ~ "Winter",
    season == "Winter" ~ "Spring" 
  )) %>%
  mutate(lagseasonyear = case_when(
    season == "Autumn" ~ Year+1,
    !(season == "Autumn") ~ Year
  )) %>%
  mutate(lagseasonyear = paste0(lagseason,lagseasonyear)) %>%
  ungroup %>% 
  select(lagseasonyear, Regions, laglogDissAmmonia,laglogDissNitrateNitrite,laglogConductivity,
         laglogSAC,laglogOUT, laglogDissOrthophos, laglogEXPORT)


y <- y %>% 
  mutate(seasonyear = paste0(season,year(Date))) %>%
  merge(wq_r_sum_lagseason,by.x=c("Regions","seasonyear"),by.y=c("Regions","lagseasonyear"),all.x=T) %>%
  select(-seasonyear)%>%
  relocate(Date)})
```


## Combine regional imputed datsets
```{r}
df_1 <- rbind(df_list_lag$Central_imp_df1, df_list_lag$Confluence_imp_df1,
             df_list_lag$NorthDelta_imp_df1, df_list_lag$South_imp_df1,df_list_lag$SuisunBay_imp_df1)

df_2 <- rbind(df_list_lag$Central_imp_df2, df_list_lag$Confluence_imp_df2,
             df_list_lag$NorthDelta_imp_df2, df_list_lag$South_imp_df2,df_list_lag$SuisunBay_imp_df2)

df_3 <- rbind(df_list_lag$Central_imp_df3, df_list_lag$Confluence_imp_df3,
             df_list_lag$NorthDelta_imp_df3, df_list_lag$South_imp_df3,df_list_lag$SuisunBay_imp_df3)

df_4 <- rbind(df_list_lag$Central_imp_df4, df_list_lag$Confluence_imp_df4,
             df_list_lag$NorthDelta_imp_df4, df_list_lag$South_imp_df4,df_list_lag$SuisunBay_imp_df4)

df_5 <- rbind(df_list_lag$Central_imp_df5, df_list_lag$Confluence_imp_df5,
             df_list_lag$NorthDelta_imp_df5, df_list_lag$South_imp_df5,df_list_lag$SuisunBay_imp_df5)
```


## Add bloom threshold to imputation 1

```{r}
#Binary transformation of bloom absence/presence
bloom_thresholds <- df_1 %>%
  group_by(Regions) %>%
  summarise(mean = mean(Chlorophyll,na.rm=T),sd=sd(Chlorophyll,na.rm=T)) %>%
  mutate(bloom_thr = mean + sd) %>%
  select(-mean,-sd)

#Imputation 1 
df_1 <- df_1 %>%
  merge(bloom_thresholds,by="Regions",all.x=T) %>%
  mutate(bloom = case_when(Chlorophyll > bloom_thr ~ 1,
                           Chlorophyll < bloom_thr ~ 0))

base::paste0("number of blooms: ",toString(sum(df_1$bloom))) 
```


## Z-score

```{r}
# z-score standardizing the dataset for easier comparison of predictors. This facilitates comparison of effect sizes between variables.

df1 <- df_1 %>% mutate(across(!c(bloom,Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

```


# Fit models with imputed datasets

Use Ryan's top models from the Global GAM result to run a binomial GAM: Models 4 and 6 with PO4.

## bm3.po4
```{r}
bm3.po4.imp1 <- gam(bloom ~ DissAmmonia + 
                      DissNitrateNitrite + 
                      lagDissNitrateNitrite + 
                      Temperature + 
                      Conductivity + 
                      s(Regions, bs='re') + 
                      season + 
                      (OUT*season) + 
                      clam_step + 
                      OUT + 
                      lagOUT + 
                      lagDissAmmonia + 
                      DissOrthophos + 
                      lagDissOrthophos,
             family=binomial(link = "logit"),data=df1,select=T)
```

log transformations
```{r}
bm3.po4.imp1.log <- gam(bloom ~ logDissAmmonia + 
                      logDissNitrateNitrite + 
                      laglogDissNitrateNitrite + 
                      Temperature + 
                      logConductivity + 
                      s(Regions, bs='re') + 
                      season + 
                      (OUT*season) + 
                      clam_step + 
                      logOUT + 
                      laglogOUT + 
                      laglogDissAmmonia + 
                      logDissOrthophos + 
                      laglogDissOrthophos,
             family=binomial(link = "logit"),data=df1,select=T)
```


## bm4.po4

Orthophosphos with SAC (no EXPORT)

```{r}
bm4.po4.imp1 <- gam(bloom ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite+
               DissOrthophos +
               lagDissOrthophos,
             family=binomial(link = "logit"),data=df1,select=T)

```

With log transformations:

```{r}
bm4.po4.imp1.log <- gam(bloom ~ logDissAmmonia +
               logDissNitrateNitrite +
               Temperature +
               logConductivity +
               SAC + 
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               laglogSAC +
               laglogDissAmmonia +
               laglogDissNitrateNitrite+
               logDissOrthophos +
               laglogDissOrthophos,
             family=binomial(link = "logit"),data=df1,select=T)
```

## bm6.po4

Orthophosphos with SAC+ EXPORT

```{r}
bm6.po4.imp1 <- gam(bloom ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               EXPORT +
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite+
               DissOrthophos +
               lagDissOrthophos,
             family=binomial(link = "logit"),data=df1,select=T)
```

Log transform:
```{r}
bm6.po4.imp1.log <- gam(bloom ~ logDissAmmonia +
               logDissNitrateNitrite +
               Temperature +
               logConductivity +
               SAC +
               EXPORT +
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               laglogSAC +
               laglogDissAmmonia +
               laglogDissNitrateNitrite+
               logDissOrthophos +
               laglogDissOrthophos,
             family=binomial(link = "logit"),data=df1,select=T)
```

# AIC, BIC

Included log models. 

## Compare with AIC

```{r}
AIC(bm3.po4.imp1,bm4.po4.imp1, bm6.po4.imp1,bm3.po4.imp1.log,bm4.po4.imp1.log, bm6.po4.imp1.log)
```

## Compare with BIC

```{r}
BIC(bm3.po4.imp1, bm4.po4.imp1, bm6.po4.imp1,bm3.po4.imp1.log,bm4.po4.imp1.log, bm6.po4.imp1.log)
```

# Shrink Models

## bm3.po4
```{r}
bm3.po4.imp1.shrink <- gam(bloom ~ s(DissAmmonia) + 
                             s(DissNitrateNitrite) + 
                             s(lagDissNitrateNitrite) + 
                             s(Temperature) + 
                             s(Conductivity) + 
                             s(Regions, bs='re') + 
                             season + 
                             (OUT*season) + 
                             clam_step + 
                             s(OUT) + 
                             s(lagOUT) + 
                             s(lagDissAmmonia)+ 
                             s(DissOrthophos)+ 
                             s(lagDissOrthophos),
             family=binomial(link = "logit"),data=df1,select=T)
summary(bm3.po4.imp1.shrink)
```

log transformed
```{r}
bm3.po4.imp1.log.shrink <- gam(bloom ~ s(logDissAmmonia) + 
                             s(logDissNitrateNitrite) + 
                             s(laglogDissNitrateNitrite) + 
                             s(Temperature) + 
                             s(logConductivity) + 
                             s(Regions, bs='re') + 
                             season + 
                             (OUT*season) + 
                             clam_step + 
                             s(logOUT) + 
                             s(laglogOUT) + 
                             s(laglogDissAmmonia)+ 
                             s(logDissOrthophos)+ 
                             s(laglogDissOrthophos),
             family=binomial(link = "logit"),data=df1,select=T)
summary(bm3.po4.imp1.log.shrink)
```


## bm4.po4

```{r}
bm4.po4.imp1.shrink <- gam(bloom ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity) +
               s(SAC) + 
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               s(lagSAC) +
               s(lagDissAmmonia) +
               s(lagDissNitrateNitrite)+
               s(DissOrthophos) +
               s(lagDissOrthophos),
             family=binomial(link = "logit"),data=df1,select=T)
summary(bm4.po4.imp1.shrink)
```

log transformed
```{r}
bm4.po4.imp1.log.shrink <- gam(bloom ~ s(logDissAmmonia) +
               s(logDissNitrateNitrite) +
               s(Temperature) +
               s(logConductivity) +
               s(logSAC) + 
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               s(laglogSAC) +
               s(laglogDissAmmonia) +
               s(laglogDissNitrateNitrite)+
               s(logDissOrthophos) +
               s(laglogDissOrthophos),
             family=binomial(link = "logit"),data=df1,select=T)
summary(bm4.po4.imp1.log.shrink)
```

## bm6.po4

```{r}
bm6.po4.imp1.shrink <- gam(bloom ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity) +
               s(SAC) +
               s(EXPORT) +   
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               s(lagSAC) +
               s(lagDissAmmonia) +
               s(lagDissNitrateNitrite)+
               s(DissOrthophos) +
               s(lagDissOrthophos),
             family=binomial(link = "logit"),data=df1,select=T)
summary(bm6.po4.imp1.shrink)
```

log transformed
```{r}
bm6.po4.imp1.log.shrink <- gam(bloom ~ s(logDissAmmonia) +
               s(logDissNitrateNitrite) +
               s(Temperature) +
               s(logConductivity) +
               s(logSAC) +
               s(logEXPORT) +   
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               s(laglogSAC) +
               s(laglogDissAmmonia) +
               s(laglogDissNitrateNitrite)+
               s(logDissOrthophos) +
               s(laglogDissOrthophos),
             family=binomial(link = "logit"),data=df1,select=T)
summary(bm6.po4.imp1.log.shrink)
```

# LD50
https://www.ccohs.ca/oshanswers/chemicals/ld50.html

have to back calc from z score
https://stackoverflow.com/questions/21594916/convert-normalized-variables-back-to-non-normalized-values 
x = std*z + m

## Vary temp and hold others constant

Will probably make this into a function to apply to other variables, but want to test it first!
```{r}
df_1.2 <- data.frame(
  Temperature = seq(min(df1$Temperature), max(df1$Temperature), length.out = 1000),
  clam_step = as.factor("post clam"),
  season = as.factor("Summer"),
  Regions = as.factor("Central")
  )

df_1.2$Regions = fct_expand(df_1.2$Regions,c("Central","Confluence","North Delta", "South", "Suisun Bay"))
df_1.2$season = fct_expand(df_1.2$season,c("Autumn","Spring","Summer","Winter"))
df_1.2$clam_step = fct_expand(df_1.2$clam_step,c("post clam","pre clam"))

df_mean <- df1 %>%
  select(-clam_step,-Temperature) %>%
  group_by() %>%
  summarize(across(SAC:laglogEXPORT, ~ median(.x, na.rm = TRUE)))

df_1.2 <- bind_cols(df_1.2,df_mean) 
```

## Predict on the RESPONSE scale
```{r}

df_1.2$predicted <- predict(bm4.po4.imp1, newdata = df_1.2, type = "response")
plot(df_1.2$Temperature,df_1.2$predicted)
```

## Linear interpolation to find where predicted probability crosses 0.50
```{r}
ld50 <- approx(x = df_1.2$predicted, y = df_1.2$Temperature, xout = 0.50)$y
ld50

```


## Old wrong way

```{r}
df1_stats <- df_1 %>% 
  select(Date,DissAmmonia,DissNitrateNitrite,Temperature,lagDissAmmonia,
         lagDissNitrateNitrite,DissOrthophos,lagDissOrthophos,
         logDissAmmonia,logDissNitrateNitrite,laglogDissAmmonia,
         laglogDissNitrateNitrite,logDissOrthophos,laglogDissOrthophos) %>%
  pivot_longer(cols = !"Date",names_to = "Parameter") %>%
  group_by(Parameter) %>%
  summarize(mean = mean(value,na.rm=T),sd = sd(value,na.rm=T)) 

#LD50 <- -bm4.po4.imp1$coefficients[1]/bm4.po4.imp1$coefficients["Temperature"] 

# to back calc from z score: x = std*z + m
LD50_table <- df1_stats %>%
  mutate(intercept_bm4 = bm4.po4.imp1$coefficients[1]) %>%
  mutate(coeff_bm4 = bm4.po4.imp1$coefficients[df1_stats$Parameter]) %>%
  mutate(ld50_bm4_z = -intercept_bm4/coeff_bm4) %>%
  mutate(ld50_bm4_nat = ld50_bm4_z*sd+mean) #%>%
  # mutate(intercept_bm4log = bm4.po4.imp1.log$coefficients[1]) %>%
  # mutate(coeff_bm4log = bm4.po4.imp1.log$coefficients[df1_stats$Parameter])%>%
  # mutate(ld50_bm4log_z = -intercept_bm4log/coeff_bm4log) %>%
  # mutate(ld50_bm4log_nat = ld50_bm4log_z*sd+mean)

LD50_table

```

