---
title: "Phytoplankton Community Ordination Plots"
author: "Schuyler Nardelli"
date: "2026-09-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data
```{r, warning=FALSE, message=FALSE}

# Clear environment and plots
rm(list = ls()); graphics.off()

# Load libraries
library(tidyverse)
library(vegan)
library(RColorBrewer)
library(plotly)
library(sf)
library(ggpubr)
library(ggrepel)
library(patchwork)

# Load integrated phytoplankton + environmental dataset
load("Data/integrated_dataset_2008-2024_imp.Rdata") 

#Reload PESP dataset with Genus and AlgalGroup
pesp <- read.csv("Data/PESP_enumeration.csv") %>%
  mutate(Date = as.Date(Date)) %>%
  arrange(Date, Time)

sum_na <- function(x) ifelse(all(is.na(x)), NA_real_, sum(x, na.rm = TRUE))

pesp1 <- pesp %>%
  group_by(Survey,Date,Station,Latitude,Longitude,SampleMethod,SampleDepth,DepthType,Lab,
           CountMethodSample,CountMethodTaxa,Magnification,AlgalGroup,Genus) %>%
  summarize(Cells_per_mL = sum_na(Cells_per_mL),
            Units_per_mL = sum_na(Units_per_mL),
            Biovolume_per_mL = sum_na(Biovolume_per_mL))

# Calculate relative abundance/biovolume for each sample
tot <- pesp1 %>% group_by(Date,Station) %>% summarise(tot_cells = sum_na(Cells_per_mL),
                                                      tot_units = sum_na(Units_per_mL),
                                                      tot_bio = sum_na(Biovolume_per_mL))
pesp1 <- pesp1 %>% left_join(tot,by=c("Date","Station"))
pesp1$relative_cell_density <- (pesp1$Cells_per_mL/pesp1$tot_cells)*100;
pesp1$relative_unit_density <- (pesp1$Units_per_mL/pesp1$tot_units)*100;
pesp1$relative_biovolume <- (pesp1$Biovolume_per_mL/pesp1$tot_bio)*100;

rm(sum_na,tot)

# Merge PESP with discrete datasets
# Merge wq and df
x <- df[,c("Date","SAC","SJR","OUT","EXPORTS","Index")]
wq_df <- left_join(wq, x, by = "Date")
wq_df <- subset(wq_df, select = -c(Latitude, Longitude))

# Merge with PESP data
dat <- left_join(pesp1, wq_df, by = c("Survey","Date","Station"))

rm(x,wq_df,wq_df)

# Add Month, Year, and Season
dat$Month <- month(dat$Date)
dat$Year <- year(dat$Date)

dat$Season<- NA
I = which(dat$Month %in% c(7:9)); dat$Season[I]<-"Summer"
I = which(dat$Month %in% c(10:12)); dat$Season[I]<-"Autumn"
I = which(dat$Month %in% c(1:3)); dat$Season[I]<-"Winter"
I = which(dat$Month %in% c(4:6)); dat$Season[I]<-"Spring"

rm(I)

# Add in Region
points_df <- dat[,c("Station","Survey","Latitude","Longitude")]
points_df <- unique(points_df)
points_df <- points_df %>% filter(!is.na(Latitude))
points_df$id <- 1:length(points_df$Survey)

Region <- st_read('Data/Regions_shp/Rosies_Regions_edited.shp')

points_sf = points_df %>% st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326)

points_with_region <- st_join(points_sf, Region, join = st_within) %>%  
  st_drop_geometry() 

stat <- left_join(points_sf,points_with_region,by=c("Station","Survey","id"))
coord <- data.frame(st_coordinates(stat$geometry))
stat$Latitude <- coord$Y; stat$Longitude <- coord$X; rm(coord)
stat <- stat %>% st_drop_geometry() %>% select(-"id")
dat <- left_join(dat,stat,by=c("Station","Survey","Latitude","Longitude")) %>%
  rename(Region = Regions)

rm(stat,points_df,points_sf,points_with_region,Region)

# Add monthly grazing rate
# Add in region
points_df <- clam[,c("station_name","Latitude","Longitude")]
points_df <- unique(points_df)
points_df <- points_df %>% filter(!is.na(Latitude))
points_df$id <- 1:length(points_df$station_name)

Region <- st_read('Data/Regions_shp/Rosies_Regions_edited.shp')

points_sf = points_df %>% st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326)

points_with_region <- st_join(points_sf, Region, join = st_within) %>%  
  st_drop_geometry() 

stat <- left_join(points_sf,points_with_region,by=c("station_name"))
coord <- data.frame(st_coordinates(stat$geometry))
stat$Latitude <- coord$Y; stat$Longitude <- coord$X; rm(coord)
stat <- stat %>% st_drop_geometry() 

clam <- left_join(clam,stat,by=c("station_name","Latitude","Longitude"))

rm(stat,points_df,points_sf,points_with_region,Region)

monthly_clam <- clam %>%
  filter(!is.na(Region)) %>%
  group_by(Year,Month,Region) %>%
  summarize(monthly_biomass = mean(Biomass,na.rm=TRUE),
            monthly_grazing_rate = mean(Grazing_rate,na.rm=TRUE),
            monthly_filtration_rate = mean(Filtration_rate,na.rm=TRUE)) 

dat <- left_join(dat, monthly_clam, by = c("Year","Month","Region"))

# Filter dataset
dat1 <- dat %>% filter(Region %in% c("Central", "Confluence", "Suisun Bay") &   
                        Date > "2007-12-31" &    
                        Survey %in% c("DWR-EMP","USGS-BP"))   

# Perry: rn, from "Datetime" to "VSS"
dat1 <- dat1 %>% ungroup()%>% filter(!if_all(24:67, is.na)) # Screens out PESP samples without corresponding WQ data
dat1 <- dat1 %>% arrange(Date)

```

## Flag "bloom" or "no bloom" for each sample using chl threshold
Considered "bloom" if Chla is 1 standard deviation above the mean value for the entire time series
```{r, warning=FALSE, message=FALSE}

dat1$bloom <- NA

# Suisun Bay
test <- dat1 %>% filter(Region=="Suisun Bay") %>% group_by(Date) %>% summarize(Chl=mean(Chlorophyll,na.rm=TRUE))
x = mean(test$Chl,na.rm=TRUE) + sd(test$Chl,na.rm=TRUE)
I <- which(dat1$Region == "Suisun Bay" & dat1$Chlorophyll > x); dat1$bloom[I] <- "Yes"; rm(I)
I <- which(dat1$Region == "Suisun Bay" & dat1$Chlorophyll <= x); dat1$bloom[I] <- "No"; rm(I,x)

p1<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Suisun Bay"),aes(x=Date,y=Chlorophyll),color="black")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay" & bloom=="Yes"), aes(x = Date, y = Chlorophyll),color="blue")+ggtitle("Suisun Bay")

# Confluence
test <- dat1 %>% filter(Region=="Confluence") %>% group_by(Date) %>% summarize(Chl=mean(Chlorophyll,na.rm=TRUE))
x = mean(test$Chl,na.rm=TRUE) + sd(test$Chl,na.rm=TRUE)
I <- which(dat1$Region == "Confluence" & dat1$Chlorophyll > x); dat1$bloom[I] <- "Yes"; rm(I)
I <- which(dat1$Region == "Confluence" & dat1$Chlorophyll <= x); dat1$bloom[I] <- "No"; rm(I,test,x)

p2<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Confluence"),aes(x=Date,y=Chlorophyll),color="black")+
  geom_point(data = dat1 %>% filter(Region=="Confluence" & bloom=="Yes"), aes(x = Date, y = Chlorophyll),color="blue")+ggtitle("Confluence")

# Central Delta
test <- dat1 %>% filter(Region=="Central") %>% group_by(Date) %>% summarize(Chl=mean(Chlorophyll,na.rm=TRUE))
x = mean(test$Chl,na.rm=TRUE) + sd(test$Chl,na.rm=TRUE)
I <- which(dat1$Region == "Central" & dat1$Chlorophyll > x); dat1$bloom[I] <- "Yes"; rm(I)
I <- which(dat1$Region == "Central" & dat1$Chlorophyll <= x); dat1$bloom[I] <- "No"; rm(I,test,x)

p3<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Central"),aes(x=Date,y=Chlorophyll),color="black")+
  geom_point(data = dat1 %>% filter(Region=="Central" & bloom=="Yes"), aes(x = Date, y = Chlorophyll),color="blue")+ggtitle("Central")

ggarrange(p1,p2,p3,ncol=1)

rm(p1,p2,p3)

```

## Calculate HAB/BAB/No bloom using MVI
"HAB" if MVI = 3-5 AND occurs from June to October. 
"BAB" if MVI = 1-2 from June to October OR if the sample was collected during a bloom between November and May.
"No bloom" if sample was collected between November and May and chlorophyll is below the bloom threshold.
```{r, warning=FALSE, message=FALSE}

dat1 <- dat1 %>%
  group_by(Region) %>%
  mutate(chl_threshold = mean(Chlorophyll, na.rm = TRUE) + sd(Chlorophyll, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(
    mvi_class = case_when(
      is.na(Microcystis) ~ NA_character_,
      Microcystis >= 3 & Month %in% 6:10 ~ "HAB",
      Microcystis <= 2 & Month %in% 6:10 ~ "BAB",
      Month %in% c(11, 12, 1:5) ~ "BAB",
      TRUE ~ NA_character_
    ),
    mvi_class = case_when(
      mvi_class == "BAB" & bloom == "Yes" ~ "BAB",
      mvi_class == "BAB" & bloom == "No" ~ "No bloom",
      TRUE ~ mvi_class
    )
  )

dat <- dat1[, c("Survey","Year","Season","Month","Date","Region","Station","Latitude","Longitude",
                 "SampleMethod","SampleDepth","DepthType","Lab","CountMethodSample","CountMethodTaxa",
                 "Magnification","AlgalGroup","Genus","Cells_per_mL","Units_per_mL","Biovolume_per_mL",
                 "tot_cells","tot_units","tot_bio","relative_cell_density","relative_unit_density",
                 "relative_biovolume","chl_threshold","bloom","mvi_class",
                 "Sample_depth_surface","Sample_depth_nutr_surface","Sample_depth_bottom","Microcystis",
                 "Chlorophyll","Pheophytin","Temperature","Temperature_bottom","Conductivity",
                 "Conductivity_bottom","Salinity","Salinity_bottom","Secchi","TurbidityNTU",
                 "TurbidityNTU_bottom","TurbidityFNU","TurbidityFNU_bottom","DissolvedOxygen",
                 "DissolvedOxygen_bottom","DissolvedOxygenPercent","DissolvedOxygenPercent_bottom","pH",
                 "DissAmmonia","DissNitrateNitrite","TotPhos","DissOrthophos","TON","DON","TKN",
                 "DissBromide","DissCalcium","TotChloride","DissChloride","DissSilica","TOC","DOC","TDS",
                 "TSS","VSS","SAC","SJR","OUT","EXPORTS","Index","monthly_biomass","monthly_grazing_rate","monthly_filtration_rate")]

```


## NMDS - GENERA - full year
```{r, warning=FALSE, message=FALSE}

# Format data
phyt <- dat %>% 
  filter(Year >= 2014, !is.na(Biovolume_per_mL), Survey == "DWR-EMP") %>% #, Month %in% 6:10
  group_by(Year,Season,Month,Date,Region,Station,mvi_class,Temperature,Conductivity,Secchi,TurbidityFNU,DissolvedOxygen,pH,
         DissAmmonia,DissNitrateNitrite,TotPhos,DissOrthophos,DissSilica,OUT,monthly_grazing_rate,Genus) %>%
  summarise(biovol = sum(Biovolume_per_mL, na.rm = TRUE),.groups = "drop") 

com.df <- phyt %>% pivot_wider(names_from = Genus, values_from = biovol, values_fill = 0)
env.df <- com.df[,1:20]
com.df <- com.df[,21:199]
sites <- env.df %>% select(Date, Station, Year, Month)

# Run NMDS
com.trans <- sqrt(com.df)               # or log1p(com.df)
com.trans <- wisconsin(com.trans)
set.seed(20)
com.MDS <- metaMDS(com.trans, distance = "bray", k = 3, trymax = 100, autotransform = FALSE)

###NMDS output
# points: sample scores
# dims: number of NMS axes
# stress: stress of final solutions
# data: what was ordinated and any transformations
# distance: distance metric used
# converged: whether or not solution converged
# tries: number of configurations tried
# species: scores of variables
# call: how function was called

#check the stress
com.MDS$stress # stress = 0.19
##ABOUT STRESS: A good rule of thumb- 
#stress < 0.05 provides an excellent representation in reduced dimensions, 
# < 0.1 is great,
# < 0.2 is good/ok, 
# > 0.2 provides a poor representation

envfit(com.MDS, env.df[,c(3,5,7:20)],na.rm=TRUE)

# ============================================================
# BUILD THE RDA MODEL  (this was missing)
# ============================================================
# Predictor set — the five variables meaningful at genus level
vars <- c("Temperature","Conductivity","TurbidityFNU","DissAmmonia","monthly_grazing_rate")

# Hellinger transform for RDA (appropriate for community data + Euclidean-based RDA)
com.hell <- decostand(com.df, method = "hellinger")

# Drop rows with missing environmental data — from BOTH matrices, same index
keep <- complete.cases(env.df[, vars])
com.hell.sub <- com.hell[keep, ]
env.sub      <- env.df[keep, ]

rda.mod <- rda(com.hell.sub ~ Temperature + Conductivity + TurbidityFNU +
                              DissAmmonia + monthly_grazing_rate,
               data = env.sub)

# ============================================================
# DIAGNOSTICS
# ============================================================

# ------------------------------------------------------------
# PART 1 — RDA diagnostics (constrained model: testable)
# ------------------------------------------------------------
cat("############### RDA DIAGNOSTICS ###############\n")
cat("\nRows used:", nrow(env.sub), "of", nrow(env.df),
    "(", nrow(env.df) - nrow(env.sub), "dropped for missingness )\n")

cat("\n=== VIF (want < 5) ===\n");                print(vif.cca(rda.mod))
cat("\n=== Overall significance ===\n");          print(anova(rda.mod, permutations = 999))
cat("\n=== Marginal (per-variable) ===\n");        print(anova(rda.mod, by = "margin", permutations = 999))
cat("\n=== Axis significance ===\n");              print(anova(rda.mod, by = "axis", permutations = 999))
cat("\n=== Variance explained (adj. R2) ===\n");   print(RsquareAdj(rda.mod))

# axis identity from biplot scores
bp <- scores(rda.mod, display = "bp", choices = 1:3)
cat("\n=== Biplot scores (variable loadings per axis) ===\n"); print(round(bp, 3))
rda_axis_lab <- function(bp, axis) {
  col <- bp[, axis]; top <- names(which.max(abs(col)))
  sprintf("%s: %s (%s)", axis, top, ifelse(col[top] > 0, "+", "-"))
}
cat("\n=== Suggested RDA axis identities ===\n")
for (a in colnames(bp)) cat(rda_axis_lab(bp, a), "\n")

# ------------------------------------------------------------
# PART 2 — NMDS diagnostics (unconstrained: fit + group tests)
# ------------------------------------------------------------
cat("\n\n############### NMDS DIAGNOSTICS ###############\n")

# (a) stress & convergence
cat("\n=== Stress ===\n"); print(com.MDS$stress)
cat("Rule of thumb: <0.1 great, <0.2 usable, >0.2 poor\n")
cat("\n=== Convergence ===\n")
cat("Converged:", com.MDS$converged, " | tries:", com.MDS$tries, "\n")
# note: if converged = FALSE, consider raising trymax or k

# (b) Shepard plot: non-metric & linear fit R2 (how well distances are preserved)
cat("\n=== Shepard / stress plot (opens a plot) ===\n")
stressplot(com.MDS)   # visual; reports non-metric & linear R2 on the plot

# (c) scree: stress vs number of dimensions (is k adequate?)
cat("\n=== Stress by dimensionality (k = 1..4) — this refits, may be slow ===\n")
set.seed(20)
k_stress <- sapply(1:4, function(k) {
  m <- metaMDS(com.trans, distance = "bray", k = k, trymax = 20,
               autotransform = FALSE, trace = 0)
  m$stress
})
print(data.frame(k = 1:4, stress = round(k_stress, 3)))
cat("Look for the 'elbow' — where added dimensions stop reducing stress much.\n")

# (d) envfit: post-hoc variable correlations (NOT a model test — descriptive)
cat("\n=== envfit (post-hoc variable correlations) ===\n")
ef <- envfit(com.MDS, env.df[, vars], na.rm = TRUE, permutations = 999, choices = c(1,2))
print(ef)

# NMDS axis identity from envfit vectors, weighted by sqrt(r2)
ev <- as.data.frame(scores(ef, display = "vectors", choices = c(1,2)))
ev$r2 <- ef$vectors$r
nmds_axis_lab <- function(ev, axis) {
  w <- abs(ev[[axis]]) * sqrt(ev$r2); top <- rownames(ev)[which.max(w)]
  sprintf("%s: %s (%s)", axis, top, ifelse(ev[top, axis] > 0, "+", "-"))
}
cat("\n=== Suggested NMDS axis identities ===\n")
cat(nmds_axis_lab(ev, "NMDS1"), "\n"); cat(nmds_axis_lab(ev, "NMDS2"), "\n")

# (e) PERMANOVA — the actual TEST of group differences on the dissimilarity matrix
#     This is the NMDS-world equivalent of asking "do regions/seasons differ?"
cat("\n=== PERMANOVA (adonis2) on Bray-Curtis ===\n")
# align community + grouping, dropping NA rows in the grouping vars
grp <- env.df[, c("Region","Season")]
ok  <- complete.cases(grp)
d.bray <- vegdist(com.trans[ok, ], method = "bray")
set.seed(20)
perm <- adonis2(d.bray ~ Region + Season, data = grp[ok, ],
                permutations = 999, by = "margin")
print(perm)

# (f) betadisper — ESSENTIAL companion: is a significant PERMANOVA due to
#     different group *centroids* (real separation) or different *dispersion* (spread)?
cat("\n=== betadisper (homogeneity of dispersion) ===\n")
cat("\n-- Region --\n")
bd.region <- betadisper(d.bray, grp$Region[ok])
print(anova(bd.region))     # significant => groups differ in spread, not just location
cat("\n-- Season --\n")
bd.season <- betadisper(d.bray, grp$Season[ok])
print(anova(bd.season))

# ============================================================
# Assumes all-months genus objects in scope: com.df, env.df, com.MDS, rda.mod
# (rda.mod built on all-months data as in the diagnostic block)
# ============================================================

clean_labs <- function(v) {
  v <- gsub("TurbidityFNU",         "Turbidity",    v)
  v <- gsub("monthly_grazing_rate", "Grazing rate", v)
  v <- gsub("DissAmmonia",          "Ammonia",      v)
  v
}

# ============================================================
# AXIS INTERPRETATION CHECKS (both methods, same logic)
# ============================================================

# --- RDA: biplot scores tell you the axis identity directly ---
bp <- scores(rda.mod, display = "bp", choices = 1:2)
rda_axis_lab <- function(bp, axis) {
  col <- bp[, axis]; top <- names(which.max(abs(col)))
  sprintf("%s (%s, %s)", axis, clean_labs(top), ifelse(col[top] > 0, "+", "-"))
}
cat("RDA1:", rda_axis_lab(bp, "RDA1"), "\n")
cat("RDA2:", rda_axis_lab(bp, "RDA2"), "\n")

# --- NMDS: no biplot scores; use envfit vectors weighted by r2 ---
ef <- envfit(com.MDS, env.df[, vars], na.rm = TRUE, permutations = 999, choices = c(1,2))
ev <- as.data.frame(scores(ef, display = "vectors", choices = c(1,2)))
ev$r2 <- ef$vectors$r     # r2 for each vector
nmds_axis_lab <- function(ev, axis) {
  w <- abs(ev[[axis]]) * sqrt(ev$r2)          # weight direction by strength
  top <- rownames(ev)[which.max(w)]
  sprintf("%s (%s, %s)", axis, clean_labs(top), ifelse(ev[top, axis] > 0, "+", "-"))
}
cat("NMDS1:", nmds_axis_lab(ev, "NMDS1"), "\n")
cat("NMDS2:", nmds_axis_lab(ev, "NMDS2"), "\n")
# Both should report turbidity (axis 1) and temperature (axis 2)

# ============================================================
# FIGURE 1: GENUS RDA  (RDA2 relabeled to temperature)
# ============================================================
site.rda <- as.data.frame(scores(rda.mod, display = "sites", choices = 1:2))
site.rda$Region <- env.sub$Region
site.rda$Season <- env.sub$Season          # for the faceted version

sp.rda <- as.data.frame(scores(rda.mod, display = "species", choices = 1:2))
sp.rda$genus <- rownames(sp.rda)
sp.rda$d <- sqrt(sp.rda$RDA1^2 + sp.rda$RDA2^2)
top.rda <- sp.rda[sp.rda$genus != "Unknown", ]
top.rda <- top.rda[order(-top.rda$d), ][1:12, ]

bp.rda <- as.data.frame(scores(rda.mod, display = "bp", choices = 1:2))
bp.rda$var <- clean_labs(rownames(bp.rda))

arr <- 0.6; labm <- 1.12
xlab_rda <- "RDA1 (turbidity gradient)"
ylab_rda <- "RDA2 (temperature gradient)"     # <-- corrected label

rda.env <- ggplot(site.rda, aes(RDA1, RDA2)) +
  geom_point(aes(color = Region), alpha = 0.4, size = 1.6) +
  stat_ellipse(aes(color = Region), level = 0.68) +
  geom_segment(data = bp.rda, aes(x=0, y=0, xend=RDA1*arr, yend=RDA2*arr),
               arrow = arrow(length = unit(0.2,"cm")),
               color = "grey15", linewidth = 0.6, inherit.aes = FALSE) +
    geom_text_repel(data = bp.rda,aes(x = RDA1*arr*labm, y = RDA2*arr*labm, label = var),
            color = "grey15", size = 3.2, fontface = "bold",
            segment.color = "grey60", segment.size = 0.3, box.padding = 0.5, 
            max.overlaps = Inf, seed = 1, inherit.aes = FALSE) +
  coord_equal() +
  scale_x_continuous(expand = expansion(mult = 0.18)) +
  scale_y_continuous(expand = expansion(mult = 0.18)) +
  labs(x = xlab_rda, y = ylab_rda, color = "Region") +
  theme_minimal(base_size = 12)

lim.rda <- max(abs(c(top.rda$RDA1, top.rda$RDA2))) * 1.5
rda.sp <- ggplot(top.rda, aes(RDA1, RDA2)) +
  geom_point(color = "steelblue4", size = 2) +
  geom_text_repel(aes(label = genus), color = "steelblue4",
                  size = 3, fontface = "italic", max.overlaps = Inf, seed = 1) +
  geom_segment(data = bp.rda, aes(x=0, y=0, xend=RDA1*arr, yend=RDA2*arr),
               arrow = arrow(length = unit(0.2,"cm")),
               color = "grey15", linewidth = 0.6, inherit.aes = FALSE) +
  geom_text_repel(data = bp.rda,aes(x = RDA1*arr*labm, y = RDA2*arr*labm, label = var),
            color = "grey15", size = 3.2, fontface = "bold",
            segment.color = "grey60", segment.size = 0.3, box.padding = 0.5, 
            max.overlaps = Inf, seed = 1, inherit.aes = FALSE) +
  coord_equal(xlim = c(-lim.rda, lim.rda), ylim = c(-lim.rda, lim.rda)) +
  labs(x = xlab_rda, y = ylab_rda) +
  theme_minimal(base_size = 12)

fig.rda <- (rda.env + rda.sp) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")",
                  title = "Genus-level RDA (all months)") &
  theme(legend.position = "bottom", plot.tag = element_text(face = "bold"))

# ============================================================
# FIGURE 2: GENUS NMDS  (interpretation now references temperature)
# ============================================================
site.mds <- as.data.frame(scores(com.MDS, display = "sites", choices = c(1,2)))
site.mds$Region <- env.df$Region
site.mds$Season <- env.df$Season

sp.mds <- as.data.frame(scores(com.MDS, display = "species", choices = c(1,2)))
sp.mds$genus <- rownames(sp.mds)
top.mds <- sp.mds[sp.mds$genus %in% top.rda$genus, ]

vec.mds <- as.data.frame(scores(ef, display = "vectors", choices = c(1,2)))
vec.mds$var <- clean_labs(rownames(vec.mds))

vrange   <- max(abs(c(vec.mds$NMDS1, vec.mds$NMDS2)))
arr.site <- 0.85 * max(abs(c(site.mds$NMDS1, site.mds$NMDS2))) / vrange
arr.sp   <- 0.85 * max(abs(c(top.mds$NMDS1,  top.mds$NMDS2)))  / vrange

mds.env <- ggplot(site.mds, aes(NMDS1, NMDS2)) +
  geom_point(aes(color = Region), alpha = 0.4, size = 1.6) +
  stat_ellipse(aes(color = Region), level = 0.68) +
  geom_segment(data = vec.mds, aes(x=0, y=0, xend=NMDS1*arr.site, yend=NMDS2*arr.site),
               arrow = arrow(length = unit(0.2,"cm")),
               color = "grey15", linewidth = 0.6, inherit.aes = FALSE) +
  geom_text(data = vec.mds, aes(x=NMDS1*arr.site*labm, y=NMDS2*arr.site*labm, label=var),
            color = "grey15", size = 3.2, fontface = "bold", inherit.aes = FALSE) +
  coord_equal() +
  scale_x_continuous(expand = expansion(mult = 0.18)) +
  scale_y_continuous(expand = expansion(mult = 0.18)) +
  annotate("text", x = -Inf, y = -Inf, hjust = -0.1, vjust = -0.8,
           label = paste0("Stress = ", round(com.MDS$stress, 2)), size = 3) +
  labs(x = "NMDS1", y = "NMDS2", color = "Region") +
  theme_minimal(base_size = 12)

lim.mds <- max(abs(c(top.mds$NMDS1, top.mds$NMDS2))) * 1.5
mds.sp <- ggplot(top.mds, aes(NMDS1, NMDS2)) +
  geom_point(color = "steelblue4", size = 2) +
  geom_text_repel(aes(label = genus), color = "steelblue4",
                  size = 3, fontface = "italic", max.overlaps = Inf, seed = 1) +
  geom_segment(data = vec.mds, aes(x=0, y=0, xend=NMDS1*arr.sp, yend=NMDS2*arr.sp),
               arrow = arrow(length = unit(0.2,"cm")),
               color = "grey15", linewidth = 0.6, inherit.aes = FALSE) +
  geom_text(data = vec.mds, aes(x=NMDS1*arr.sp*labm, y=NMDS2*arr.sp*labm, label=var),
            color = "grey15", size = 3.2, fontface = "bold", inherit.aes = FALSE) +
  coord_equal(xlim = c(-lim.mds, lim.mds), ylim = c(-lim.mds, lim.mds)) +
  labs(x = "NMDS1", y = "NMDS2") +
  theme_minimal(base_size = 12)

fig.mds <- (mds.env + mds.sp) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")",
                  title = "Genus-level NMDS (all months)") &
  theme(legend.position = "bottom", plot.tag = element_text(face = "bold"))

# ============================================================
# FIGURE 3: SEASON-FACETED NMDS site ordination
# Shows seasonal turnover directly (points colored by Region, faceted by Season)
# ============================================================
# ensure Season is an ordered factor so facets read Winter->Autumn
season_levels <- c("Winter","Spring","Summer","Autumn")
site.mds$Season <- factor(site.mds$Season,
                          levels = season_levels[season_levels %in% unique(site.mds$Season)])

fig.season <- ggplot(site.mds, aes(NMDS1, NMDS2)) +
  geom_point(aes(color = Region), alpha = 0.4, size = 1.4) +
  stat_ellipse(aes(color = Region), level = 0.68) +
  geom_segment(data = vec.mds, aes(x=0, y=0, xend=NMDS1*arr.site, yend=NMDS2*arr.site),
               arrow = arrow(length = unit(0.15,"cm")),
               color = "grey15", linewidth = 0.5, inherit.aes = FALSE) +
  geom_text(data = vec.mds, aes(x=NMDS1*arr.site*labm, y=NMDS2*arr.site*labm, label=var),
            color = "grey15", size = 2.6, fontface = "bold", inherit.aes = FALSE) +
  coord_equal() +
  facet_wrap(~ Season) +
  labs(x = "NMDS1", y = "NMDS2", color = "Region",
       title = "Seasonal turnover in phytoplankton community (NMDS)") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

# ---- Display / save ----
fig.rda
fig.mds
fig.season
# ggsave("genus_RDA_allmonths.png",  fig.rda,    width = 10, height = 5.5, dpi = 300)
# ggsave("genus_NMDS_allmonths.png", fig.mds,    width = 10, height = 5.5, dpi = 300)
# ggsave("genus_NMDS_byseason.png",  fig.season, width = 9,  height = 8,   dpi = 300)


```

## NMDS - GENERA - summer
```{r, warning=FALSE, message=FALSE}

# Format data
phyt <- dat %>% 
  filter(Year >= 2014, !is.na(Biovolume_per_mL), Survey == "DWR-EMP", Month %in% 6:10) %>% 
  group_by(Year,Season,Month,Date,Region,Station,mvi_class,Temperature,Conductivity,Secchi,TurbidityFNU,DissolvedOxygen,pH,
         DissAmmonia,DissNitrateNitrite,TotPhos,DissOrthophos,DissSilica,OUT,monthly_grazing_rate,Genus) %>%
  summarise(biovol = sum(Biovolume_per_mL, na.rm = TRUE),.groups = "drop") 

com.df <- phyt %>% pivot_wider(names_from = Genus, values_from = biovol, values_fill = 0)
env.df <- com.df[,1:20]
com.df <- com.df[,21:199]
sites <- env.df %>% select(Date, Station, Year, Month)

# Run NMDS
com.trans <- sqrt(com.df)               # or log1p(com.df)
com.trans <- wisconsin(com.trans)
set.seed(20)
com.MDS <- metaMDS(com.trans, distance = "bray", k = 3, trymax = 100, autotransform = FALSE)

###NMDS output
# points: sample scores
# dims: number of NMS axes
# stress: stress of final solutions
# data: what was ordinated and any transformations
# distance: distance metric used
# converged: whether or not solution converged
# tries: number of configurations tried
# species: scores of variables
# call: how function was called

#check the stress
com.MDS$stress # stress = 0.19
##ABOUT STRESS: A good rule of thumb- 
#stress < 0.05 provides an excellent representation in reduced dimensions, 
# < 0.1 is great,
# < 0.2 is good/ok, 
# > 0.2 provides a poor representation

envfit(com.MDS, env.df[,c(3,5,7:20)],na.rm=TRUE)

library(vegan)
library(ggplot2)
library(ggrepel)
library(patchwork)

# ============================================================
# ASSUMES the GENUS-level objects are in scope:
#   com.df      = genus community matrix (896 x 178)
#   env.df      = environmental data (aligned rows)
#   com.MDS     = genus NMDS (k=3, stress 0.19)
# If com.hell doesn't exist yet, this rebuilds it from genus com.df.
# ============================================================

# ---- Genus RDA ----
com.hell <- decostand(com.df, method = "hellinger")
vars <- c("Temperature","Conductivity","TurbidityFNU","DissAmmonia","monthly_grazing_rate")
keep <- complete.cases(env.df[, vars])
com.hell.sub <- com.hell[keep, ]
env.sub      <- env.df[keep, ]

rda.mod <- rda(com.hell.sub ~ Temperature + Conductivity + TurbidityFNU +
                              DissAmmonia + monthly_grazing_rate, data = env.sub)

# ---- Shared label cleaner ----
clean_labs <- function(v) {
  v <- gsub("TurbidityFNU",         "Turbidity",    v)
  v <- gsub("monthly_grazing_rate", "Grazing rate", v)
  v <- gsub("DissAmmonia",          "Ammonia",      v)
  v
}

# ============================================================
# FIGURE 1: GENUS RDA
# ============================================================
site.rda <- as.data.frame(scores(rda.mod, display = "sites", choices = 1:2))
site.rda$Region <- env.sub$Region

sp.rda <- as.data.frame(scores(rda.mod, display = "species", choices = 1:2))
sp.rda$genus <- rownames(sp.rda)
sp.rda$d <- sqrt(sp.rda$RDA1^2 + sp.rda$RDA2^2)
top.rda <- sp.rda[sp.rda$genus != "Unknown", ]
top.rda <- top.rda[order(-top.rda$d), ][1:12, ]

bp.rda <- as.data.frame(scores(rda.mod, display = "bp", choices = 1:2))
bp.rda$var <- clean_labs(rownames(bp.rda))

arr <- 0.6; labm <- 1.12

rda.env <- ggplot(site.rda, aes(RDA1, RDA2)) +
  geom_point(aes(color = Region), alpha = 0.4, size = 1.6) +
  stat_ellipse(aes(color = Region), level = 0.68) +
  geom_segment(data = bp.rda, aes(x=0, y=0, xend=RDA1*arr, yend=RDA2*arr),
               arrow = arrow(length = unit(0.2,"cm")),
               color = "grey15", linewidth = 0.6, inherit.aes = FALSE) +
  geom_text(data = bp.rda, aes(x=RDA1*arr*labm, y=RDA2*arr*labm, label=var),
            color = "grey15", size = 3.2, fontface = "bold", inherit.aes = FALSE) +
  coord_equal() +
  scale_x_continuous(expand = expansion(mult = 0.18)) +
  scale_y_continuous(expand = expansion(mult = 0.18)) +
  labs(x = "RDA1 (turbidity gradient)", y = "RDA2 (grazing gradient)", color = "Region") +
  theme_minimal(base_size = 12)

lim.rda <- max(abs(c(top.rda$RDA1, top.rda$RDA2))) * 1.5
rda.sp <- ggplot(top.rda, aes(RDA1, RDA2)) +
  geom_point(color = "steelblue4", size = 2) +
  geom_text_repel(aes(label = genus), color = "steelblue4",
                  size = 3, fontface = "italic", max.overlaps = Inf, seed = 1) +
  geom_segment(data = bp.rda, aes(x=0, y=0, xend=RDA1*arr, yend=RDA2*arr),
               arrow = arrow(length = unit(0.2,"cm")),
               color = "grey15", linewidth = 0.6, inherit.aes = FALSE) +
  geom_text(data = bp.rda, aes(x=RDA1*arr*labm, y=RDA2*arr*labm, label=var),
            color = "grey15", size = 3.2, fontface = "bold", inherit.aes = FALSE) +
  coord_equal(xlim = c(-lim.rda, lim.rda), ylim = c(-lim.rda, lim.rda)) +
  labs(x = "RDA1 (turbidity gradient)", y = "RDA2 (grazing gradient)") +
  theme_minimal(base_size = 12)

fig.rda <- (rda.env + rda.sp) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")",
                  title = "Genus-level RDA") &
  theme(legend.position = "bottom", plot.tag = element_text(face = "bold"))

# ============================================================
# FIGURE 2: GENUS NMDS  (same 12 genera as the RDA)
# ============================================================
ef <- envfit(com.MDS, env.df[, vars], na.rm = TRUE, permutations = 999, choices = c(1,2))

site.mds <- as.data.frame(scores(com.MDS, display = "sites", choices = c(1,2)))
site.mds$Region <- env.df$Region

sp.mds <- as.data.frame(scores(com.MDS, display = "species", choices = c(1,2)))
sp.mds$genus <- rownames(sp.mds)
top.mds <- sp.mds[sp.mds$genus %in% top.rda$genus, ]   # same taxa as RDA panel

vec.mds <- as.data.frame(scores(ef, display = "vectors", choices = c(1,2)))
vec.mds$var <- clean_labs(rownames(vec.mds))

# arrows scaled separately to each panel's cloud (NMDS site & species scales differ)
vrange   <- max(abs(c(vec.mds$NMDS1, vec.mds$NMDS2)))
arr.site <- 0.85 * max(abs(c(site.mds$NMDS1, site.mds$NMDS2))) / vrange
arr.sp   <- 0.85 * max(abs(c(top.mds$NMDS1,  top.mds$NMDS2)))  / vrange

mds.env <- ggplot(site.mds, aes(NMDS1, NMDS2)) +
  geom_point(aes(color = Region), alpha = 0.4, size = 1.6) +
  stat_ellipse(aes(color = Region), level = 0.68) +
  geom_segment(data = vec.mds, aes(x=0, y=0, xend=NMDS1*arr.site, yend=NMDS2*arr.site),
               arrow = arrow(length = unit(0.2,"cm")),
               color = "grey15", linewidth = 0.6, inherit.aes = FALSE) +
  geom_text(data = vec.mds, aes(x=NMDS1*arr.site*labm, y=NMDS2*arr.site*labm, label=var),
            color = "grey15", size = 3.2, fontface = "bold", inherit.aes = FALSE) +
  coord_equal() +
  scale_x_continuous(expand = expansion(mult = 0.18)) +
  scale_y_continuous(expand = expansion(mult = 0.18)) +
  annotate("text", x = -Inf, y = -Inf, hjust = -0.1, vjust = -0.8,
           label = "Stress = 0.19", size = 3) +
  labs(x = "NMDS1", y = "NMDS2", color = "Region") +
  theme_minimal(base_size = 12)

lim.mds <- max(abs(c(top.mds$NMDS1, top.mds$NMDS2))) * 1.5
mds.sp <- ggplot(top.mds, aes(NMDS1, NMDS2)) +
  geom_point(color = "steelblue4", size = 2) +
  geom_text_repel(aes(label = genus), color = "steelblue4",
                  size = 3, fontface = "italic", max.overlaps = Inf, seed = 1) +
  geom_segment(data = vec.mds, aes(x=0, y=0, xend=NMDS1*arr.sp, yend=NMDS2*arr.sp),
               arrow = arrow(length = unit(0.2,"cm")),
               color = "grey15", linewidth = 0.6, inherit.aes = FALSE) +
  geom_text(data = vec.mds, aes(x=NMDS1*arr.sp*labm, y=NMDS2*arr.sp*labm, label=var),
            color = "grey15", size = 3.2, fontface = "bold", inherit.aes = FALSE) +
  coord_equal(xlim = c(-lim.mds, lim.mds), ylim = c(-lim.mds, lim.mds)) +
  labs(x = "NMDS1", y = "NMDS2") +
  theme_minimal(base_size = 12)

fig.mds <- (mds.env + mds.sp) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")",
                  title = "Genus-level NMDS") &
  theme(legend.position = "bottom", plot.tag = element_text(face = "bold"))

# ---- Display / save ----
fig.rda
fig.mds
# ggsave("genus_RDA.png", fig.rda, width = 10, height = 5.5, dpi = 300)
# ggsave("genus_NMDS.png", fig.mds, width = 10, height = 5.5, dpi = 300)


# --- assumes the all-months objects are in scope ---
#   com.df (genus matrix, all months), env.df (aligned), rda.mod already built
# If rda.mod isn't rebuilt on the all-months data yet, rebuild it:
com.hell <- decostand(com.df, method = "hellinger")
vars <- c("Temperature","Conductivity","TurbidityFNU","DissAmmonia","monthly_grazing_rate")
keep <- complete.cases(env.df[, vars])
com.hell.sub <- com.hell[keep, ]
env.sub      <- env.df[keep, ]
rda.mod <- rda(com.hell.sub ~ Temperature + Conductivity + TurbidityFNU +
                              DissAmmonia + monthly_grazing_rate, data = env.sub)

# --- how many rows actually went in (after NA drop)? ---
cat("Rows used:", nrow(env.sub), "of", nrow(env.df),
    "(", nrow(env.df) - nrow(env.sub), "dropped for missingness )\n\n")

# --- core diagnostics ---
cat("=== VIF ===\n");             print(vif.cca(rda.mod))
cat("\n=== Overall significance ===\n"); print(anova(rda.mod, permutations = 999))
cat("\n=== Marginal (per-variable) ===\n"); print(anova(rda.mod, by = "margin", permutations = 999))
cat("\n=== Axis significance ===\n");       print(anova(rda.mod, by = "axis", permutations = 999))
cat("\n=== Variance explained ===\n");      print(RsquareAdj(rda.mod))

# --- biplot scores: which variable defines each axis ---
bp <- scores(rda.mod, display = "bp", choices = 1:3)
cat("\n=== Biplot scores (env variable loadings on each axis) ===\n")
print(round(bp, 3))

# --- auto-identify the dominant variable on RDA1 and RDA2 ---
lab_for_axis <- function(bp, axis) {
  col <- bp[, axis]
  top <- names(which.max(abs(col)))
  sign <- ifelse(col[top] > 0, "+", "-")
  sprintf("%s (dominated by %s, %s)", axis, top, sign)
}
cat("\n=== Suggested axis identities ===\n")
cat(lab_for_axis(bp, "RDA1"), "\n")
cat(lab_for_axis(bp, "RDA2"), "\n")

# --- genus scores along each axis (top/bottom responders) ---
sp1 <- sort(scores(rda.mod, display = "species", choices = 1)[,1])
sp2 <- sort(scores(rda.mod, display = "species", choices = 2)[,1])
cat("\n=== RDA1: 6 most negative / 6 most positive genera ===\n")
print(round(c(head(sp1,6), tail(sp1,6)), 3))
cat("\n=== RDA2: 6 most negative / 6 most positive genera ===\n")
print(round(c(head(sp2,6), tail(sp2,6)), 3))

## OTHER############################################
#First we will extract the coordinates
coords <- com.MDS$points
#add site and year
coords <- cbind(sites, coords)

#Since our env data was taken supplementary to the survey we can use the function envfit()
envft <- envfit(com.MDS, env.df[,c(3,5,7:20)],na.rm=TRUE,choices=c(1,2))
envft

ggplot(data = coords, aes(x = MDS1, y = MDS2, fill = as.factor(env.df$mvi_class))) +
  geom_hline(yintercept=0) + geom_vline(xintercept=0) +
  geom_point(data = coords, size = 2.5, alpha = 0.9,pch=21) +
  scale_fill_manual(values = colorRampPalette(brewer.pal(3, "RdYlBu"))(3),name="")+
  theme(axis.title = element_text(size = 10, colour = "black"),
        panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "black"),
        axis.text = element_text(size = 12, colour = "black"), legend.key = element_blank(),
        legend.title = element_text(size = 12, colour = "black"),
        legend.text = element_text(size = 9, colour = "black"), legend.position = "right",
        axis.title.x = element_text(size = 12),
        plot.title = element_text(size = 12, hjust = 0.5)) +
  labs(x = "NMDS 1", y = "NMDS 2") 

ggplot(data = coords, aes(x = MDS1, y = MDS2, fill = as.factor(env.df$Month))) +
  geom_hline(yintercept=0) + geom_vline(xintercept=0) +
  geom_point(data = coords, size = 2.5, alpha = 0.9,pch=21) +
  scale_fill_manual(values = colorRampPalette(brewer.pal(5, "RdYlBu"))(5),name="Month")+
  theme(axis.title = element_text(size = 10, colour = "black"),
        panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "black"),
        axis.text = element_text(size = 12, colour = "black"), legend.key = element_blank(),
        legend.title = element_text(size = 12, colour = "black"),
        legend.text = element_text(size = 9, colour = "black"), legend.position = "right",
        axis.title.x = element_text(size = 12),
        plot.title = element_text(size = 12, hjust = 0.5)) +
  labs(x = "NMDS 1", y = "NMDS 2") 

ggplot(data = coords, aes(x = MDS1, y = MDS2, fill = env.df$TurbidityFNU)) +
  geom_hline(yintercept=0) + geom_vline(xintercept=0) +
  geom_point(data = coords, size = 2.5, alpha = 0.9,pch=21) +
  scale_fill_distiller(palette = "RdYlBu", name = "Turbidity (FNU)") +
  theme(axis.title = element_text(size = 10, colour = "black"),
        panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "black"),
        axis.text = element_text(size = 12, colour = "black"), legend.key = element_blank(),
        legend.title = element_text(size = 12, colour = "black"),
        legend.text = element_text(size = 9, colour = "black"), legend.position = "right",
        axis.title.x = element_text(size = 12),
        plot.title = element_text(size = 12, hjust = 0.5)) +
  labs(x = "NMDS 1", y = "NMDS 2") 

ggplot(data = coords, aes(x = MDS1, y = MDS2, fill = env.df$TurbidityFNU)) +
  geom_hline(yintercept=0) + geom_vline(xintercept=0) +
  geom_point(data = coords, size = 2.5, alpha = 0.9,pch=21) +
  scale_fill_distiller(palette = "RdYlBu", name = "Turbidity (FNU)") +
  theme(axis.title = element_text(size = 10, colour = "black"),
        panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "black"),
        axis.text = element_text(size = 12, colour = "black"), legend.key = element_blank(),
        legend.title = element_text(size = 12, colour = "black"),
        legend.text = element_text(size = 9, colour = "black"), legend.position = "right",
        axis.title.x = element_text(size = 12),
        plot.title = element_text(size = 12, hjust = 0.5)) +
  labs(x = "NMDS 1", y = "NMDS 2") 
```


## NMDS - ALGAL GROUP - summer
```{r, warning=FALSE, message=FALSE}

# Format data
phyt <- dat %>% 
  filter(Year >= 2014, !is.na(Biovolume_per_mL), Survey == "DWR-EMP", Month %in% 6:10) %>%
  group_by(Year,Season,Month,Date,Region,Station,mvi_class,Temperature,Conductivity,Secchi,TurbidityFNU,DissolvedOxygen,pH,
         DissAmmonia,DissNitrateNitrite,TotPhos,DissOrthophos,DissSilica,OUT,monthly_grazing_rate,AlgalGroup) %>%
  summarise(biovol = sum(Biovolume_per_mL, na.rm = TRUE),.groups = "drop") 

com.df <- phyt %>% pivot_wider(names_from = AlgalGroup, values_from = biovol, values_fill = 0)
env.df <- com.df[,1:20]
com.df <- com.df[,21:32]
sites <- env.df %>% select(Date, Station, Year, Month)

# Run NMDS
com.trans <- sqrt(com.df)               # or log1p(com.df)
com.trans <- wisconsin(com.trans)
set.seed(20)
com.MDS <- metaMDS(com.trans, distance = "bray", k = 3, trymax = 100, autotransform = FALSE)

###NMDS output
# points: sample scores
# dims: number of NMS axes
# stress: stress of final solutions
# data: what was ordinated and any transformations
# distance: distance metric used
# converged: whether or not solution converged
# tries: number of configurations tried
# species: scores of variables
# call: how function was called

#check the stress
com.MDS$stress # stress = 0.146
##ABOUT STRESS: A good rule of thumb- 
#stress < 0.05 provides an excellent representation in reduced dimensions, 
# < 0.1 is great,
# < 0.2 is good/ok, 
# > 0.2 provides a poor representation

envfit(com.MDS, env.df[,c(3,5,7:20)],na.rm=TRUE)

library(ggplot2)
library(ggrepel)
library(patchwork)

ef <- envfit(com.MDS, env.df[, c("TurbidityFNU","Secchi","monthly_grazing_rate")],
             na.rm = TRUE, permutations = 999, choices = c(1,2))

site.mds <- as.data.frame(scores(com.MDS, display = "sites", choices = c(1,2)))
site.mds$Region <- env.df$Region

sp.mds <- as.data.frame(scores(com.MDS, display = "species", choices = c(1,2)))
sp.mds$AlgalGroup <- rownames(sp.mds)

vec.mds <- as.data.frame(scores(ef, display = "vectors", choices = c(1,2)))
vec.mds$var <- rownames(vec.mds)
vec.mds$var <- gsub("TurbidityFNU",         "Turbidity",    vec.mds$var)
vec.mds$var <- gsub("monthly_grazing_rate", "Grazing rate", vec.mds$var)

# --- Scale each panel's arrows to that panel's point cloud ---
sp_range   <- max(abs(c(sp.mds$NMDS1,   sp.mds$NMDS2)))
site_range <- max(abs(c(site.mds$NMDS1, site.mds$NMDS2)))
vec_range  <- max(abs(c(vec.mds$NMDS1,  vec.mds$NMDS2)))

arr_mult      <- 0.85 * sp_range   / vec_range
arr_mult_site <- 0.85 * site_range / vec_range
lab_mult <- 1.10

# --- Panel A: sites ---
mds.env.panel <- ggplot(site.mds, aes(NMDS1, NMDS2)) +
  geom_point(aes(color = Region), alpha = 0.4, size = 1.6) +
  stat_ellipse(aes(color = Region), level = 0.68) +
  geom_segment(data = vec.mds,
               aes(x = 0, y = 0, xend = NMDS1*arr_mult_site, yend = NMDS2*arr_mult_site),
               arrow = arrow(length = unit(0.2,"cm")),
               color = "grey15", linewidth = 0.6, inherit.aes = FALSE) +
  geom_text(data = vec.mds,
            aes(x = NMDS1*arr_mult_site*lab_mult, y = NMDS2*arr_mult_site*lab_mult, label = var),
            color = "grey15", size = 3.2, fontface = "bold", inherit.aes = FALSE) +
  coord_equal() +
  annotate("text", x = -Inf, y = -Inf, hjust = -0.1, vjust = -0.8,
           label = "Stress = 0.15", size = 3) +
  labs(x = "NMDS1", y = "NMDS2", color = "Region") +
  theme_minimal(base_size = 12)

# --- Panel B: groups — points FIRST, arrows scaled to fit, explicit limits ---
lim <- sp_range * 1.4    # give the labels room

mds.sp.panel <- ggplot(sp.mds, aes(NMDS1, NMDS2)) +
  geom_point(color = "steelblue4", size = 2.5) +
  geom_text_repel(aes(label = AlgalGroup), color = "steelblue4",
                  size = 3, max.overlaps = Inf, seed = 1) +
  geom_segment(data = vec.mds,
               aes(x = 0, y = 0, xend = NMDS1*arr_mult, yend = NMDS2*arr_mult),
               arrow = arrow(length = unit(0.2,"cm")),
               color = "grey15", linewidth = 0.6, inherit.aes = FALSE) +
  geom_text(data = vec.mds,
            aes(x = NMDS1*arr_mult*lab_mult, y = NMDS2*arr_mult*lab_mult, label = var),
            color = "grey15", size = 3.2, fontface = "bold", inherit.aes = FALSE) +
  coord_equal(xlim = c(-lim, lim), ylim = c(-lim, lim)) +
  labs(x = "NMDS1", y = "NMDS2") +
  theme_minimal(base_size = 12)

(mds.env.panel + mds.sp.panel) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") &
  theme(legend.position = "bottom",
        plot.tag = element_text(face = "bold"))


#####RDA model################################################

library(vegan)
library(ggplot2)
library(ggrepel)
library(patchwork)

# --- IMPORTANT: build com.hell from the AlgalGroup community matrix ---
# (com.df here must be the 12-column AlgalGroup matrix, not the genus one)
com.hell <- decostand(com.df, method = "hellinger")

# --- Model ---
vars <- c("TurbidityFNU","monthly_grazing_rate")
keep <- complete.cases(env.df[, vars])
com.hell.sub <- com.hell[keep, ]
env.sub      <- env.df[keep, ]

rda.mod <- rda(com.hell.sub ~ TurbidityFNU + monthly_grazing_rate,
               data = env.sub)

# --- Diagnostics ---
vif.cca(rda.mod)
anova(rda.mod, permutations = 999)
anova(rda.mod, by = "margin", permutations = 999)
anova(rda.mod, by = "axis", permutations = 999)
RsquareAdj(rda.mod)
scores(rda.mod, display = "bp", choices = 1:3)

# --- Extract scores ---
site.rda <- as.data.frame(scores(rda.mod, display = "sites", choices = 1:2))
site.rda$Region <- env.sub$Region

# 12 groups — keep them all, no distance filtering, no "Unknown" exclusion
sp.rda <- as.data.frame(scores(rda.mod, display = "species", choices = 1:2))
sp.rda$AlgalGroup <- rownames(sp.rda)

# --- Environmental arrows ---
bp <- as.data.frame(scores(rda.mod, display = "bp", choices = 1:2))
bp$var <- rownames(bp)
bp$var <- gsub("TurbidityFNU",         "Turbidity",    bp$var)
bp$var <- gsub("monthly_grazing_rate", "Grazing rate", bp$var)
bp$var <- gsub("DissAmmonia",          "Ammonia",      bp$var)

arr_mult <- 0.6
lab_mult <- 1.12

# --- Panel A: sites + ellipses + arrows ---
env.panel <- ggplot(site.rda, aes(RDA1, RDA2)) +
  geom_point(aes(color = Region), alpha = 0.4, size = 1.6) +
  stat_ellipse(aes(color = Region), level = 0.68) +
  geom_segment(data = bp,
               aes(x = 0, y = 0, xend = RDA1*arr_mult, yend = RDA2*arr_mult),
               arrow = arrow(length = unit(0.2,"cm")),
               color = "grey15", linewidth = 0.6, inherit.aes = FALSE) +
  geom_text(data = bp,
            aes(x = RDA1*arr_mult*lab_mult, y = RDA2*arr_mult*lab_mult, label = var),
            color = "grey15", size = 3.2, fontface = "bold", inherit.aes = FALSE) +
  coord_equal() +
  scale_x_continuous(expand = expansion(mult = 0.18)) +
  scale_y_continuous(expand = expansion(mult = 0.18)) +
  labs(x = "RDA1", y = "RDA2", color = "Region") +
  theme_minimal(base_size = 12)

# --- Panel B: groups + arrows (points first, explicit limits) ---
lim <- max(abs(c(sp.rda$RDA1, sp.rda$RDA2))) * 1.5

sp.panel <- ggplot(sp.rda, aes(RDA1, RDA2)) +
  geom_point(color = "steelblue4", size = 2.5) +
  geom_text_repel(aes(label = AlgalGroup), color = "steelblue4",
                  size = 3, max.overlaps = Inf, seed = 1) +
  geom_segment(data = bp,
               aes(x = 0, y = 0, xend = RDA1*arr_mult, yend = RDA2*arr_mult),
               arrow = arrow(length = unit(0.2,"cm")),
               color = "grey15", linewidth = 0.6, inherit.aes = FALSE) +
  geom_text(data = bp,
            aes(x = RDA1*arr_mult*lab_mult, y = RDA2*arr_mult*lab_mult, label = var),
            color = "grey15", size = 3.2, fontface = "bold", inherit.aes = FALSE) +
  coord_equal(xlim = c(-lim, lim), ylim = c(-lim, lim)) +
  labs(x = "RDA1", y = "RDA2") +
  theme_minimal(base_size = 12)

# --- Combine ---
(env.panel + sp.panel) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") &
  theme(legend.position = "bottom",
        plot.tag = element_text(face = "bold"))

## OTHER############################################
#First we will extract the coordinates
coords <- com.MDS$points
#add site and year
coords <- cbind(sites, coords)

#Since our env data was taken supplementary to the survey we can use the function envfit()
envft <- envfit(com.MDS, env.df[,c(3,5,7:20)],na.rm=TRUE,choices=c(1,2))
envft

ggplot(data = coords, aes(x = MDS1, y = MDS2, fill = as.factor(env.df$mvi_class))) +
  geom_hline(yintercept=0) + geom_vline(xintercept=0) +
  geom_point(data = coords, size = 2.5, alpha = 0.9,pch=21) +
  scale_fill_manual(values = colorRampPalette(brewer.pal(3, "RdYlBu"))(3),name="")+
  theme(axis.title = element_text(size = 10, colour = "black"),
        panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "black"),
        axis.text = element_text(size = 12, colour = "black"), legend.key = element_blank(),
        legend.title = element_text(size = 12, colour = "black"),
        legend.text = element_text(size = 9, colour = "black"), legend.position = "right",
        axis.title.x = element_text(size = 12),
        plot.title = element_text(size = 12, hjust = 0.5)) +
  labs(x = "NMDS 1", y = "NMDS 2") 

ggplot(data = coords, aes(x = MDS1, y = MDS2, fill = as.factor(env.df$Month))) +
  geom_hline(yintercept=0) + geom_vline(xintercept=0) +
  geom_point(data = coords, size = 2.5, alpha = 0.9,pch=21) +
  scale_fill_manual(values = colorRampPalette(brewer.pal(5, "RdYlBu"))(5),name="Month")+
  theme(axis.title = element_text(size = 10, colour = "black"),
        panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "black"),
        axis.text = element_text(size = 12, colour = "black"), legend.key = element_blank(),
        legend.title = element_text(size = 12, colour = "black"),
        legend.text = element_text(size = 9, colour = "black"), legend.position = "right",
        axis.title.x = element_text(size = 12),
        plot.title = element_text(size = 12, hjust = 0.5)) +
  labs(x = "NMDS 1", y = "NMDS 2") 

ggplot(data = coords, aes(x = MDS1, y = MDS2, fill = env.df$TurbidityFNU)) +
  geom_hline(yintercept=0) + geom_vline(xintercept=0) +
  geom_point(data = coords, size = 2.5, alpha = 0.9,pch=21) +
  scale_fill_distiller(palette = "RdYlBu", name = "Turbidity (FNU)") +
  theme(axis.title = element_text(size = 10, colour = "black"),
        panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "black"),
        axis.text = element_text(size = 12, colour = "black"), legend.key = element_blank(),
        legend.title = element_text(size = 12, colour = "black"),
        legend.text = element_text(size = 9, colour = "black"), legend.position = "right",
        axis.title.x = element_text(size = 12),
        plot.title = element_text(size = 12, hjust = 0.5)) +
  labs(x = "NMDS 1", y = "NMDS 2") 

ggplot(data = coords, aes(x = MDS1, y = MDS2, fill = env.df$TurbidityFNU)) +
  geom_hline(yintercept=0) + geom_vline(xintercept=0) +
  geom_point(data = coords, size = 2.5, alpha = 0.9,pch=21) +
  scale_fill_distiller(palette = "RdYlBu", name = "Turbidity (FNU)") +
  theme(axis.title = element_text(size = 10, colour = "black"),
        panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "black"),
        axis.text = element_text(size = 12, colour = "black"), legend.key = element_blank(),
        legend.title = element_text(size = 12, colour = "black"),
        legend.text = element_text(size = 9, colour = "black"), legend.position = "right",
        axis.title.x = element_text(size = 12),
        plot.title = element_text(size = 12, hjust = 0.5)) +
  labs(x = "NMDS 1", y = "NMDS 2") 

```

## Cluster analysis
```{r, warning=FALSE, message=FALSE}

# Format data
phyt <- dat %>% 
  filter(Year >= 2014, !is.na(Biovolume_per_mL), Survey == "DWR-EMP", Month %in% 6:10) %>%
  group_by(Year,Season,Month,Date,Region,Station,mvi_class,Temperature,Conductivity,Secchi,TurbidityFNU,DissolvedOxygen,pH,
         DissAmmonia,DissNitrateNitrite,TotPhos,DissOrthophos,DissSilica,OUT,monthly_grazing_rate,AlgalGroup,Genus) %>%
  summarise(biovol = sum(Biovolume_per_mL, na.rm = TRUE),.groups = "drop") %>%
  mutate(sqbiovol= sqrt(biovol))

com.df <- phyt %>% pivot_wider(names_from = Genus, values_from = biovol, values_fill = 0)
env.df <- com.df[,1:21]
com.df <- com.df[,22:200]
sites <- env.df %>% select(Date, Station, Year, Month)
com.t <- decostand(com.df, method = "total")   # or "total", or log

# Cluster analysis
com.df[is.na(com.df)] <- 0
  dis <- vegdist (com.t, method = 'bray') 
  hc <- hclust(dis,method="ward.D2")

plot(hc,main = ' ',ylab="Bray-Curtis Distance")
rect.hclust(hc, k = 4, border = 2:5) 

# Assess number of clusters
library(cluster)
library(factoextra)

gap <- clusGap(as.matrix(dis), FUN = hcut, K.max = 10, B = 100)
fviz_gap_stat(gap)

fviz_nbclust(as.matrix(dis), FUN = hcut, method = "wss",
             hc_method = "ward.D2", diss = dis)

fviz_nbclust(as.matrix(dis), FUN = hcut, method = "silhouette",
             hc_method = "ward.D2", diss = dis)

# Final number of clusters to use
grp <- cutree(hc, k = 4)

# Cluster figures
env.df$group <- factor(grp[rownames(env.df)])
ggplot(data=env.df)+geom_boxplot(aes(x=group,y=Temperature,group=group))
ggplot(data=env.df)+geom_boxplot(aes(x=group,y=log10(Conductivity),group=group))
ggplot(data=env.df)+geom_boxplot(aes(x=group,y=TurbidityFNU,group=group))
ggplot(data=env.df)+geom_boxplot(aes(x=group,y=log10(DissAmmonia),group=group))
ggplot(data=env.df)+geom_boxplot(aes(x=group,y=log10(DissNitrateNitrite),group=group))
ggplot(data=env.df)+geom_boxplot(aes(x=group,y=log10(DissOrthophos),group=group))
ggplot(data=env.df)+geom_boxplot(aes(x=group,y=log10(DissSilica),group=group))
ggplot(data=env.df)+geom_boxplot(aes(x=group,y=log10(OUT),group=group))
ggplot(data=env.df)+geom_boxplot(aes(x=group,y=monthly_grazing_rate,group=group))

ggplot(env.df, aes(x=group, fill = Region)) +geom_bar(position="fill") 
ggplot(env.df, aes(x=group, fill = as.factor(Month))) +geom_bar(position="fill") 
ggplot(env.df, aes(x=group, fill = mvi_class)) +geom_bar(position="fill") 

# Look at species composition in each cluster

# 3. attach cluster membership (make sure names align!)
biovol <- phyt %>% 
  group_by(Date, Station, AlgalGroup) %>%
  summarise(biovol = mean(biovol), .groups = "drop") %>%   # or first()/max() — see note
  pivot_wider(names_from = AlgalGroup, values_from = biovol, values_fill = 0)

biovol_df <- as.data.frame(biovol) %>%
  rownames_to_column("sample") %>%
  mutate(cluster = factor(grp[sample]))

# 4. sum biovolume per genus within each cluster, then convert to % of cluster total
cluster_comp <- biovol_df %>%
  pivot_longer(-c(Date,Station,sample, cluster), names_to = "genus", values_to = "biovol") %>%
  group_by(cluster, genus) %>%
  summarise(total_biovol = sum(biovol,na.rm=TRUE), .groups = "drop") %>%
  group_by(cluster) %>%
  mutate(pct = total_biovol / sum(total_biovol,na.rm=TRUE) * 100) %>%   # % of pooled cluster biovolume
  ungroup()

library(forcats)
top_genera <- cluster_comp %>%
  group_by(genus) %>% summarise(m = max(pct,na.rm=TRUE)) %>%
  slice_max(m, n = 20) %>% pull(genus)

plot_df <- cluster_comp %>%
  mutate(genus2 = fct_other(genus, keep = top_genera)) %>%
  group_by(cluster, genus2) %>%
  summarise(pct = sum(pct,na.rm=TRUE), .groups = "drop")

p<-ggplot(plot_df, aes(cluster, pct, fill = genus2)) +
  geom_col() +
  scale_fill_viridis_d() +
  labs(x = "Cluster", y = "% of pooled biovolume", fill = "Genus") +
  theme_minimal()
ggplotly(p)

```

## MVI divisions
```{r, warning=FALSE, message=FALSE}

# Format data
phyt <- dat %>% 
  filter(Year >= 2014, !is.na(Biovolume_per_mL), Survey == "DWR-EMP", !is.na(mvi_class)) %>%
  group_by(Year,Season,Month,Date,Region,Station,mvi_class,Temperature,Conductivity,Secchi,TurbidityFNU,DissolvedOxygen,pH,
         DissAmmonia,DissNitrateNitrite,TotPhos,DissOrthophos,DissSilica,OUT,monthly_grazing_rate,AlgalGroup,Genus) %>%
  summarise(biovol = sum(Biovolume_per_mL, na.rm = TRUE),.groups = "drop") 

# Figures
ggplot(data=phyt)+geom_boxplot(aes(x=mvi_class,y=Temperature,group=mvi_class))
ggplot(data=phyt)+geom_boxplot(aes(x=mvi_class,y=log10(Conductivity),group=mvi_class))
ggplot(data=phyt)+geom_boxplot(aes(x=mvi_class,y=TurbidityFNU,group=mvi_class))
ggplot(data=phyt)+geom_boxplot(aes(x=mvi_class,y=log10(DissAmmonia),group=mvi_class))
ggplot(data=phyt)+geom_boxplot(aes(x=mvi_class,y=log10(DissNitrateNitrite),group=mvi_class))
ggplot(data=phyt)+geom_boxplot(aes(x=mvi_class,y=log10(DissOrthophos),group=mvi_class))
ggplot(data=phyt)+geom_boxplot(aes(x=mvi_class,y=log10(DissSilica),group=mvi_class))
ggplot(data=phyt)+geom_boxplot(aes(x=mvi_class,y=log10(OUT),group=mvi_class))
ggplot(data=phyt)+geom_boxplot(aes(x=mvi_class,y=monthly_grazing_rate,group=mvi_class))

ggplot(phyt, aes(x=mvi_class, fill = Region)) +geom_bar(position="fill") 
ggplot(phyt, aes(x=mvi_class, fill = as.factor(Month))) +geom_bar(position="fill") 

# Look at GENUS composition in each cluster

# 3. attach cluster membership (make sure names align!)
biovol <- phyt %>% 
  group_by(Date, Region, Station, mvi_class,Genus) %>%
  summarise(biovol = mean(biovol), .groups = "drop") %>%   # or first()/max() — see note
  pivot_wider(names_from = Genus, values_from = biovol, values_fill = 0)

biovol_df <- as.data.frame(biovol) %>%
  rownames_to_column("sample")

# 4. sum biovolume per genus within each cluster, then convert to % of cluster total
cluster_comp <- biovol_df %>%
  pivot_longer(-c(Date,Region,Station,sample, mvi_class), names_to = "genus", values_to = "biovol") %>%
  group_by(Region,mvi_class, genus) %>%
  summarise(total_biovol = sum(biovol,na.rm=TRUE), .groups = "drop") %>%
  group_by(Region,mvi_class) %>%
  mutate(pct = total_biovol / sum(total_biovol,na.rm=TRUE) * 100) %>%   # % of pooled cluster biovolume
  ungroup()

library(forcats)
top_genera <- cluster_comp %>%
  group_by(genus) %>% summarise(m = max(pct,na.rm=TRUE)) %>%
  slice_max(m, n = 20) %>% pull(genus)

plot_df <- cluster_comp %>%
  mutate(genus2 = fct_other(genus, keep = top_genera)) %>%
  group_by(Region,mvi_class, genus2) %>%
  summarise(pct = sum(pct,na.rm=TRUE), .groups = "drop")

p<-ggplot(plot_df, aes(mvi_class, pct, fill = genus2)) +
  geom_col() +
  scale_fill_viridis_d() +
  labs(x = "Cluster", y = "% of pooled biovolume", fill = "Genus") +
  facet_wrap(~Region)+
  theme_minimal()
ggplotly(p)

# Look at ALGAL GROUP composition in each cluster

# 3. attach cluster membership (make sure names align!)
biovol <- phyt %>% 
  group_by(Date, Region, Station, mvi_class,AlgalGroup) %>%
  summarise(biovol = mean(biovol), .groups = "drop") %>%   # or first()/max() — see note
  pivot_wider(names_from = AlgalGroup, values_from = biovol, values_fill = 0)

biovol_df <- as.data.frame(biovol) %>%
  rownames_to_column("sample")

# 4. sum biovolume per genus within each cluster, then convert to % of cluster total
cluster_comp <- biovol_df %>%
  pivot_longer(-c(Date,Region,Station,sample, mvi_class), names_to = "algalgroup", values_to = "biovol") %>%
  group_by(Region,mvi_class, algalgroup) %>%
  summarise(total_biovol = sum(biovol,na.rm=TRUE), .groups = "drop") %>%
  group_by(Region,mvi_class) %>%
  mutate(pct = total_biovol / sum(total_biovol,na.rm=TRUE) * 100) %>%   # % of pooled cluster biovolume
  ungroup()

plot_df <- cluster_comp %>%
  group_by(Region,mvi_class, algalgroup) %>%
  summarise(pct = sum(pct,na.rm=TRUE), .groups = "drop")

p<-ggplot(plot_df, aes(mvi_class, pct, fill = algalgroup)) +
  geom_col() +
  scale_fill_manual(values = colorRampPalette(brewer.pal(13, "BrBG"))(13),name="Algal Group")+
  labs(x = "Cluster", y = "% of pooled biovolume", fill = "Genus") +
  facet_wrap(~Region)+
  theme_minimal()
ggplotly(p)





library(dplyr)

# Your desired group order (top-to-bottom in legend)
group_levels <- c("Centric Diatoms", "Chrysophytes", "Cryptophytes",
                  "Cyanobacteria", "Euglenoids", "Green Algae",
                  "Pennate Diatoms", "Ciliates", "Dinoflagellates")

# Named color vector: one color per group, in your order
group_colors <- setNames(
  colorRampPalette(brewer.pal(9, "BrBG"))(length(group_levels)),
  group_levels
)

# Groups that reach >= 1% in at least one bar segment
keep_groups <- plot_df %>%
  filter(pct >= 1) %>%
  pull(algalgroup) %>%
  unique()

# Keep desired ordering, restricted to survivors
keep_ordered <- group_levels[group_levels %in% keep_groups]

plot_df2 <- plot_df %>%
  mutate(algalgroup = ifelse(algalgroup %in% keep_groups,
                             as.character(algalgroup), NA),
         algalgroup = factor(algalgroup, levels = group_levels))

ggplot(plot_df2, aes(mvi_class, pct, fill = algalgroup)) +
  geom_col(color="black") +
  scale_fill_manual(
    values = group_colors,     # named -> color is locked to the group
    name = "Algal Group",
    na.value = "grey80",
    breaks = keep_ordered,      # legend order + which entries show
    limits = group_levels       # keeps full mapping consistent
  ) +
  labs(x = "", y = "% of pooled biovolume") +
  facet_wrap(~Region) +
  theme_minimal()
```
