---
title:
  "Global GAM"
author: "Ryan Hankins"
output: 
  html_document:
        toc: TRUE
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```


```{r, include=FALSE}

library(mgcv)
library(tidyverse)
library(marginaleffects)
library(MuMIn)
library(car)
library(corrplot)
library(stats)
```


# GAM:
Chla~ NH4 + (NO2+NO3) + temperature  + conductivity +  SAC + 1|Region + Season + Pre/post-clam  + seasonal lag(SAC) + seasonal lag(NH4)  + seasonal lag (NO2+NO3) 

```{r}

c = c( "logchla", "Date","Year", "Month",  "season","Regions","SAC","lagSAC",
       "EXPORT","OUT","lagOUT", "Index","lagIndex","DissAmmonia",
       "lagDissAmmonia", "TotPhos","lagTotPhos","DissNitrateNitrite", 
       "lagDissNitrateNitrite","Temperature","lagTemperature","Secchi", 
       "lagSecchi","Conductivity",  "lagConductivity","Biomass","Grazing_rate",
      "Filtration_rate","season_num")

df <- read.csv("Data/regional_integrated_dataset3.csv" )

df$Date = as.Date(df$Date)

#change Regions to factor for GAM to work as a random effect
df$Regions = as.factor(df$Regions)

#change season to numeric value for cyclic cubic spline?
df$season_num = as.numeric(as.factor(df$season))

df$logchla <- log(df$Chlorophyll)

df <- df %>%
  select(all_of(c)) %>%
  drop_na(all_of(c))
head(df)
```


# Choose the model with the best combination of flow variables

```{r}
options(na.action = "na.fail")
# OUT (no EXPORT, no step)
m1 <- gam((logchla) ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               OUT + 
               s(Regions, bs='re') +
               season +
               (OUT*season) +
               Grazing_rate +
               lagOUT +
               lagDissAmmonia +
               lagDissNitrateNitrite,
             data = df,
             method = 'ML'
             )

# SAC (no EXPORT, no step)
m2 <- gam((logchla) ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               Grazing_rate +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite,
             data = df,
             method = 'ML'
             )

# Create dataframe with pre/post clam dummy variable
df=df%>%mutate(clam_step=if_else(Year>1987, "post clam", "pre clam")) #create dummy variable for sampling period


# OUT + Step (no EXPORT)
m3 <- gam((logchla) ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               OUT + 
               s(Regions, bs='re') +
               season +
               (OUT*season) +
            #replace grazing rate with calm step function
               clam_step +
               lagOUT +
               lagDissAmmonia +
               lagDissNitrateNitrite,
             data = df,
             method = 'ML'
             )

# SAC + Step (no EXPORT)
m4 <- gam((logchla) ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               s(Regions, bs='re') +
               season +
               (SAC*season) +
            #replace grazing rate with clam step function
               clam_step +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite,
             data = df,
             method = 'ML'
             )

# OUT + EXPORT + Step
m5 <- gam((logchla) ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               OUT + 
               EXPORT +
               s(Regions, bs='re') +
               season +
               (OUT*season) +
            #replace grazing rate with clam step function
               clam_step +
               lagOUT +
               lagDissAmmonia +
               lagDissNitrateNitrite,
             data = df,
             method = 'ML'
             )

#SAC+ EXPORT + Step (best performing combo)
m6 <- gam((logchla) ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               EXPORT +
               s(Regions, bs='re') +
               season +
               (SAC*season) +
            #replace grazing rate with clam step function
               clam_step +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite,
             data = df,
             method = 'ML'
             )

# m6 but with Season as a numeric smooth term with bs=cc (cyclic cubic spline)

m7 <- gam((logchla) ~ DissAmmonia +
               DissNitrateNitrite +
               Temperature +
               Conductivity +
               SAC + 
               EXPORT +
               s(Regions, bs='re') +
               s(season_num, bs='cc', k=4) +
               (SAC*season) +
            #replace grazing rate with clam step function
               clam_step +
               lagSAC +
               lagDissAmmonia +
               lagDissNitrateNitrite,
             data = df,
             method = 'ML'
             )

```
## Compare with AIC
```{r}
AIC(m1,m2,m3, m4, m5, m6, m7)
```

```{r}
BIC(m1,m2,m3, m4, m5, m6, m7)
```

## Decisions/observations:
- Best combo of flow variables is SAC + EXPORT
- Not including NDOI (OUT) because it is a delta-wide flow measure. 
- remove month random effect
- keep season, try cyclic cubic spline bs="cc", set k=4 (m7)
- add region as a random effect
- models performed better with clam data as a step function

## to do later:
-explore marginal effects of variables 
  
# Dredge top models:
```{r}
# OUT (no EXPORT, no step)
#m.1.dredge <- dredge(m1)
#m.1.dredge

```

```{r}
# SAC (no EXPORT, no step)
m.2.dredge <- dredge(m2)
head(m.2.dredge, 3)
```

```{r}
# OUT + Step (no EXPORT)
#m.3.dredge <- dredge(m3)
#m.3.dredge
```

```{r}
# SAC + Step (no EXPORT)
m.4.dredge <- dredge(m4)
head(m.4.dredge, 3)
```

```{r}
# OUT + EXPORT + Step
#m.5.dredge <- dredge(m5)
#m.5.dredge
```

```{r}

#SAC+ EXPORT + Step (best performing combo)
m.6.dredge <- dredge(m6)
head(m.6.dredge, 3)
```

```{r}
# m6 but with as a numeric smooth term with bs=cc (cyclic cubic spline)
m.7.dredge <- dredge(m7)
head(m.7.dredge, 3)
```

```{r}
#top.m1 <- get.models(m.1.dredge, 1)[[1]]
top.m2 <- get.models(m.2.dredge, 1)[[1]]
#top.m3 <- get.models(m.3.dredge, 1)[[1]]
top.m4 <- get.models(m.4.dredge, 1)[[1]]
#top.m5 <- get.models(m.5.dredge, 1)[[1]]
top.m6 <- get.models(m.6.dredge, 1)[[1]]
top.m7 <- get.models(m.7.dredge, 1)[[1]]

```

```{r}
AIC(top.m2, top.m4, top.m6, top.m7)
```

```{r}
BIC(top.m2, top.m4, top.m6, top.m7)
```

# Shrink top model:
- EDF close to 0 it is effectively removed from the model
- linear terms have EDF close to 1

## identify linear terms & shrink predictors

```{r}
# setting select=TRUE 
m.shrink <- gam((logchla) ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity) +
               s(SAC) + 
               s(EXPORT) +
               s(Regions, bs='re') +
               s(season_num, bs='cc', k=4) +
               (SAC*season) +
               clam_step +
               s(lagSAC) +
               s(lagDissAmmonia) +
               s(lagDissNitrateNitrite),
             data = df,
             method = 'ML', 
              select=TRUE
             )


summary(m.shrink)
```
## Shrink observations
- EXPORT ~ 1 (linear)
- SAC, season_num ~0 (effectively removed from model)

# Look at concurvity of smooth terms

<0.3 - low, >0.5 - some concern, >0.8 - very high

```{r}
concurvity(m.shrink)
```
```{r}
concurvity(m.shrink, full =FALSE)
```
## concurvity notes:
- temp x dissamm = 0.4
- season_num x temperature = 0.7
- lagAmmonia x LagDissNitrateNitrite = 0.4
- lagDissNitrateNitrite x Regions =0.4

# Look at variance inflation factor of linear terms
### GIF^(1/(2DF)) >2 - moderate collinearity, GIF>3.16 = high collinearity
https://019b2da8-edfb-a262-61be-7973c056d9ae.share.connect.posit.cloud/mlr-collinearity.html

```{r}

m.l<- lm(formula = logchla ~ SAC+ EXPORT + clam_step,
            data=df)
vif(m.l)
```

# Plot

```{r}
# Plot data with model fit
c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "SAC",
          "EXPORT",
          "season_num",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia",
          "lagDissNitrateNitrite")

lapply(c, function(x) {
  plot_predictions(m.shrink, x , points = .5) +
    theme_classic(base_size = 12)
})

```
```{r}
gam.check(m.shrink)
```


# Data exploration                                                            
- outputs can be found in presentations folder, logchla exploration

# Determine distribution of Chla-a
```{r}
# hist(df$Chlorophyll)
# sqrt_chla <- sqrt(df$Chlorophyll)
# hist(sqrt_chla)
# df$log_chla <- log(df$Chlorophyll)
# hist(df$logchla)
# log10_chla <- log10(df$Chlorophyll)
# hist(log10_chla)
# 
# 
# log seems to have the best fit.
# 
# # biplots
# 
# #scatter{r}
# lapply(c, function(x) {
#   ggplot(df, mapping =aes(x = .data[[x]], y = logchla, color= Regions)) +
#     ylab("log(Chlorphyll a)")+
#     geom_point()+
#     facet_wrap(~season, ncol=1)+
#     theme_classic()})
# 
# # linear regression
# lapply(c, function(x) {
#   ggplot(df, mapping =aes(x = .data[[x]], y = logchla, color= Regions)) +
#     ylab("log(Chlorphyll a)")+
#     geom_smooth(method="lm", se=FALSE)+
#     facet_wrap(~season, ncol=1)+
#     theme_classic()})
# 
# 
# ggplot(df, mapping =aes(x = Date, y=log_chla, color=Regions))+
#   ylab("log(Chlorphyll a)")+
#   geom_line()+
#   scale_x_date(date_breaks= "10 years")+
#   facet_wrap(~season, ncol=1)+
#   theme_classic()
# 
# ggplot(df, mapping =aes(x = Date, y=log_chla))+
#   ylab("log(Chlorphyll a)")+
#   geom_line()+
#   scale_x_date(date_breaks= "10 years")+
#   facet_wrap(~season, ncol=1)+
#   theme_classic()
# 
# # linear regression
# lapply(c, function(x) {
#   ggplot(df, mapping =aes(x = .data[[x]], y = logchla, color=season)) +
#     ylab("log(Chlorphyll a)")+
#     geom_smooth(method="lm", se=FALSE)+
#     geom_point()+
#     theme_classic()})
# 
# #Corr plot
# df_numeric = df %>% select(where(is.numeric))
# m = cor(df_numeric)
# 
# corrplot(m,
#          method = "color",
#          type = "upper",
#          addCoef.col = "black",
#          number.cex = 0.7,
#          tl.cex = 0.8,
#          tl.col = "black",
#          col = colorRampPalette(c("blue", "white", "red"))(200)
# )
# 
# #######################################
# # Min Max normalized data exploration #
# #######################################
# 
# 
# df_minmax <- df %>% mutate(across(!c(Regions,season), ~as.vector(scale(.,center=min(.), scale = max(.)-min(.)))))
# 
# 
# # linear regression
# lapply(c, function(x) {
#   ggplot(df_minmax, mapping =aes(x = .data[[x]], y = logchla, color=season)) +
#     ylab("log(Chlorphyll a)")+
#     geom_smooth(method="lm", se=FALSE)+
#     geom_point()+
#     theme_classic()})
# 
# #######################################
# # z-score standardized data exploration #
# #######################################
# 
# 
# df_std <- df %>% mutate(across(!c(Regions,season), ~as.vector(scale(.))))
# 
# 
# # linear regression
# lapply(c, function(x) {
#   ggplot(df_std, mapping =aes(x = .data[[x]], y = logchla, color=season)) +
#     ylab("log(Chlorphyll a)")+
#     geom_smooth(method="lm", se=FALSE)+
#     geom_point()+
#     theme_classic()})
# lapply(c, function(x) {
#   ggplot(df_std, mapping =aes(x = .data[[x]], y = logchla)) +
#     ylab("log(Chlorphyll a)")+
#     geom_smooth(method="lm", se=TRUE)+
#     geom_point()+
#     theme_classic()})
# 
# #############################
# # Box Plot for Step Change  #
# #############################
# 
# 
# ggplot(df, mapping =aes(x = year_factor, y = logchla)) +
#   ylab("log(Chlorphyll a)")+
#   geom_boxplot()+
#   theme_classic()

```
## Resources
https://gsp.humboldt.edu/olm/R/05_03_GAM.html

https://www.geeksforgeeks.org/r-language/generalized-additive-models-using-r/





