---
title: "Classifying HABs, BABs, and no blooms in 2008-2022 dataset"
author: "Schuyler Nardelli"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data
Load in PESP, WQ, dayflow, and clam data. For WQ and clam data, make sure there is only one sample associated with each date and station.
```{r, warning=FALSE, message=FALSE}

# Clear environment and plots
rm(list = ls()); graphics.off()

# Load libraries
library(tidyverse)
library(readxl)
library(sf)
library(ggmap)
library(ggpubr)
library(EDIutils)

#--------------------------------------------------
# Helper function for loading EDI entities
#--------------------------------------------------

read_edi_entity <- function(package_id, entity_name) {
  entity_id <- read_data_entity_names(package_id) %>%
    filter(entityName == entity_name) %>%
    pull(entityId)
  raw_file <- read_data_entity(package_id, entity_id)
  readr::read_csv(raw_file, show_col_types = FALSE)
}

#--------------------------------------------------
# Load PESP data
#--------------------------------------------------

pesp <- read_edi_entity(package_id = "edi.2209.2",entity_name = "PESP_enumeration.csv") %>%
  mutate(Date = as.Date(Date)) %>%
  arrange(Date, Time)

#--------------------------------------------------
# Load discrete WQ data
#--------------------------------------------------

wq <- read_edi_entity(package_id = "edi.731.7",entity_name = "Delta_Integrated_WQ.csv") %>%
  mutate(Date = as.Date(Date)) %>%
  separate(Station,into = c("Survey", "Station"),sep = " ") %>%
  select(Station,Date,Sample_depth_surface,Sample_depth_nutr_surface,Sample_depth_bottom,
         Temperature,Temperature_bottom,Conductivity,Conductivity_bottom,Salinity,Salinity_bottom,Secchi,
         TurbidityNTU,TurbidityNTU_bottom,TurbidityFNU,TurbidityFNU_bottom,DissolvedOxygen,
         DissolvedOxygen_bottom,DissolvedOxygenPercent,DissolvedOxygenPercent_bottom,pH,pH_bottom,
         Microcystis,Chlorophyll,Pheophytin,TotAmmonia,DissAmmonia,DissNitrateNitrite,TotPhos,
         DissOrthophos,TON,DON,TKN,TotAlkalinity,DissBromide,DissCalcium,TotChloride,DissChloride,
         DissSilica,TOC,DOC,TDS,TSS,VSS) %>%
  group_by(Date,Station)%>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))
wq[is.na(wq)] <- NA

#--------------------------------------------------
# Load EMP benthic data
#--------------------------------------------------

clam <- read_csv("Data/combined_clams_dataset2.csv",show_col_types = FALSE) %>%
  mutate(station_name = str_remove(station_name, "^EMP "),Date = as.Date(Date))%>%
  group_by(Date,station_name)%>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))
clam[is.na(clam)] <- NA

#--------------------------------------------------
# Load Dayflow data
#--------------------------------------------------

df <- read_csv("Data/Dayflow/dayflow-results-1997-2023.csv",show_col_types = FALSE) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"),WY = ifelse(Month >= 10, Year + 1, Year)) %>%
  left_join(read_csv("Data/Dayflow/wy_class_ind.csv",show_col_types = FALSE),by = "WY")

rm(read_edi_entity)

```


## Calculate monthly clam values
```{r, warning=FALSE, message=FALSE}

# Add in region
points_df <- clam[,c("station_name","Latitude","Longitude")]
points_df <- unique(points_df)
points_df <- points_df %>% filter(!is.na(Latitude))
points_df$id <- 1:length(points_df$station_name)

Region <- st_read('Data/Regions_shp/Rosies_Regions_edited.shp')

points_sf = points_df %>% st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326)

points_with_region <- st_join(points_sf, Region, join = st_within) %>%  
  st_drop_geometry() 

stat <- left_join(points_sf,points_with_region,by=c("station_name"))
coord <- data.frame(st_coordinates(stat$geometry))
stat$Latitude <- coord$Y; stat$Longitude <- coord$X; rm(coord)
stat <- stat %>% st_drop_geometry() 

clam <- left_join(clam,stat,by=c("station_name","Latitude","Longitude")) %>%
  rename(Region = Regions)

rm(stat,points_df,points_sf,points_with_region,Region)

monthly_clam <- clam %>%
  filter(!is.na(Region)) %>%
  group_by(Year,Month,Region) %>%
  summarize(monthly_biomass = mean(Biomass,na.rm=TRUE),
            monthly_grazing_rate = mean(Grazing_rate,na.rm=TRUE),
            monthly_filtration_rate = mean(Filtration_rate,na.rm=TRUE)) 


```


## Calculate HAB/BAB using ptox metric
0 = absence of potentially toxigenic species, 1 = presence of potentially toxigenic species.
```{r, warning=FALSE, message=FALSE}

pesp <- pesp %>%
  group_by(Date,Station) %>%
  mutate(ptox = ifelse(any((Genus %in%
                  c("Microcoleus","Cuspidothrix","Aphanizomenon","Microcystis","Planktothrix") &
                  Taxon != "Aphanizomenon flos-aquae") |
        Taxon %in% c("Dolichospermum circinale","Microseira wollei")),1,0)) %>%
  ungroup()

```


## Aggregate PESP data by Algal Group
```{r, warning=FALSE, message=FALSE}

sum_na <- function(x) ifelse(all(is.na(x)), NA_real_, sum(x, na.rm = TRUE))

pesp1 <- pesp %>%
  group_by(Survey,Date,Station,Latitude,Longitude,SampleMethod,SampleDepth,DepthType,Lab,
           CountMethodSample,CountMethodTaxa,Magnification,ptox,AlgalGroup) %>%
  summarize(Cells_per_mL = sum_na(Cells_per_mL),
            Units_per_mL = sum_na(Units_per_mL),
            Biovolume_per_mL = sum_na(Biovolume_per_mL))

# Calculate relative abundance/biovolume for each sample
tot <- pesp1 %>% group_by(Date,Station) %>% summarise(tot_cells = sum_na(Cells_per_mL),
                                                      tot_units = sum_na(Units_per_mL),
                                                      tot_bio = sum_na(Biovolume_per_mL))
pesp1 <- pesp1 %>% left_join(tot,by=c("Date","Station"))
pesp1$relative_cell_density <- (pesp1$Cells_per_mL/pesp1$tot_cells)*100;
pesp1$relative_unit_density <- (pesp1$Units_per_mL/pesp1$tot_units)*100;
pesp1$relative_biovolume <- (pesp1$Biovolume_per_mL/pesp1$tot_bio)*100;

rm(sum_na,tot)

```


## Merge PESP and discrete datasets
```{r, warning=FALSE, message=FALSE}

# Merge wq and df
x <- df[,c("Date","SAC","SJR","OUT","EXPORTS","Index")]
wq_df <- left_join(wq, x, by = "Date")

# Merge with clam data
x <- clam[,c("Date","station_name","Biomass","Grazing_rate","Filtration_rate")]
colnames(x)[colnames(x) == "station_name"] <- "Station"
wq_df_clam <- left_join(wq_df, x, by = c("Date","Station"))

# Merge with PESP data
dat <- left_join(pesp1, wq_df_clam, by = c("Date","Station"))

rm(x,wq_df_clam,wq_df)

```


## Add in Year, Month, and Season
```{r, warning=FALSE, message=FALSE}

dat$Month <- month(dat$Date)
dat$Year <- year(dat$Date)

dat$Season<- NA
I = which(dat$Month %in% c(7:9)); dat$Season[I]<-"Summer"
I = which(dat$Month %in% c(10:12)); dat$Season[I]<-"Autumn"
I = which(dat$Month %in% c(1:3)); dat$Season[I]<-"Winter"
I = which(dat$Month %in% c(4:6)); dat$Season[I]<-"Spring"

rm(I)

```

## Calculate HAB/BAB using MVI
HAB (value of 1) if MVI = 3-5 AND occurs from June to October. 
BAB (value of 0) if MVI = 1-2 from June to October OR if the sample was collected between November and May.
```{r, warning=FALSE, message=FALSE}

dat <- dat %>%
  group_by(Date,Station) %>%
  mutate(mvi = ifelse(is.na(Microcystis),NA,ifelse(Microcystis %in% 3:5 & Month %in% 6:10, 1, 0))) %>%
  ungroup()

```


## Add in regional designation
```{r, warning=FALSE, message=FALSE}

points_df <- dat[,c("Station","Survey","Latitude","Longitude")]
points_df <- unique(points_df)
points_df <- points_df %>% filter(!is.na(Latitude))
points_df$id <- 1:length(points_df$Survey)

Region <- st_read('Data/Regions_shp/Rosies_Regions_edited.shp')

points_sf = points_df %>% st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326)

points_with_region <- st_join(points_sf, Region, join = st_within) %>%  
  st_drop_geometry() 

stat <- left_join(points_sf,points_with_region,by=c("Station","Survey","id"))
coord <- data.frame(st_coordinates(stat$geometry))
stat$Latitude <- coord$Y; stat$Longitude <- coord$X; rm(coord)
stat <- stat %>% st_drop_geometry() %>% select(-"id")
dat <- left_join(dat,stat,by=c("Station","Survey","Latitude","Longitude")) %>%
  rename(Region = Regions)

rm(stat,points_df,points_sf,points_with_region,Region)

```


## Filter dataset
```{r, warning=FALSE, message=FALSE}

dat1 <- dat %>% filter(Region %in% c("Central", "Confluence", "Suisun Bay") &   
                        Date > "2007-12-31" &    
                        Survey %in% c("DWR-EMP","USGS-BP"))   

dat1 <- dat1 %>% filter(!if_all(24:62, is.na)) # Screens out PESP samples without corresponding WQ data
dat1 <- dat1 %>% arrange(Date)

```


## Flag "bloom" or "no bloom" for each sample using chl threshold
Considered "bloom" if Chla is 1 standard deviation above the mean value for the entire time series
```{r, warning=FALSE, message=FALSE}

dat1$bloom <- NA

# Suisun Bay
test <- dat1 %>% filter(Region=="Suisun Bay") %>% group_by(Date) %>% summarize(Chl=mean(Chlorophyll,na.rm=TRUE))
x = mean(test$Chl,na.rm=TRUE) + sd(test$Chl,na.rm=TRUE)
I <- which(dat1$Region == "Suisun Bay" & dat1$Chlorophyll > x); dat1$bloom[I] <- "Yes"; rm(I)
I <- which(dat1$Region == "Suisun Bay" & dat1$Chlorophyll <= x); dat1$bloom[I] <- "No"; rm(I,x)

p1<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Suisun Bay"),aes(x=Date,y=Chlorophyll),color="black")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay" & bloom=="Yes"), aes(x = Date, y = Chlorophyll),color="blue")+ggtitle("Suisun Bay")

# Confluence
test <- dat1 %>% filter(Region=="Confluence") %>% group_by(Date) %>% summarize(Chl=mean(Chlorophyll,na.rm=TRUE))
x = mean(test$Chl,na.rm=TRUE) + sd(test$Chl,na.rm=TRUE)
I <- which(dat1$Region == "Confluence" & dat1$Chlorophyll > x); dat1$bloom[I] <- "Yes"; rm(I)
I <- which(dat1$Region == "Confluence" & dat1$Chlorophyll <= x); dat1$bloom[I] <- "No"; rm(I,test,x)

p2<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Confluence"),aes(x=Date,y=Chlorophyll),color="black")+
  geom_point(data = dat1 %>% filter(Region=="Confluence" & bloom=="Yes"), aes(x = Date, y = Chlorophyll),color="blue")+ggtitle("Confluence")

# Central Delta
test <- dat1 %>% filter(Region=="Central") %>% group_by(Date) %>% summarize(Chl=mean(Chlorophyll,na.rm=TRUE))
x = mean(test$Chl,na.rm=TRUE) + sd(test$Chl,na.rm=TRUE)
I <- which(dat1$Region == "Central" & dat1$Chlorophyll > x); dat1$bloom[I] <- "Yes"; rm(I)
I <- which(dat1$Region == "Central" & dat1$Chlorophyll <= x); dat1$bloom[I] <- "No"; rm(I,test,x)

p3<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Central"),aes(x=Date,y=Chlorophyll),color="black")+
  geom_point(data = dat1 %>% filter(Region=="Central" & bloom=="Yes"), aes(x = Date, y = Chlorophyll),color="blue")+ggtitle("Central")

ggarrange(p1,p2,p3,ncol=1)

rm(p1,p2,p3)

```


## Add in monthly clam data
```{r, warning=FALSE, message=FALSE}

dat1 <- left_join(dat1, monthly_clam, by = c("Year","Month","Region"))

```

## Data coverage
```{r, warning=FALSE, message=FALSE}

ggplot(data=dat1,aes(x=Date,y=Biovolume_per_mL))+geom_point()+facet_wrap(~Region)+
  ylab("Biovolume (ug/mL)")
ggplot(data=dat1,aes(x=Date,y=Cells_per_mL))+geom_point()+facet_wrap(~Region)


# Region
deltabox <- make_bbox(lat = Latitude, lon = Longitude, data = dat1, f = c(.2, .2)) #sets bounding box based on station coordinates, f value adds a buffer around the perimeter
register_stadiamaps(key = "b4a4b8da-f8e2-4ba1-bcdc-91e9d5c83d00")
stamenmap <- get_stadiamap(deltabox, zoom = 10, maptype = "alidade_smooth", force = TRUE, crop = TRUE)

ggmap(stamenmap) +
  geom_point(data = dat1, 
             mapping = aes(x = Longitude, y = Latitude,fill=Region),size=3,pch=21) +
  #geom_text(data = dat, aes(x = Longitude, y = Latitude,label = Station),fontface = #"bold",cex=3,vjust=0.5,hjust=1.3)+
  labs(title = "", x = "Longitude", y = "Latitude") +
  theme_light(base_size = 12) +
  theme(panel.grid = element_blank(),
        strip.background = element_rect(fill = "black"),
        strip.text = element_text(size = 10),axis.text.x = element_text(angle = 90))+
  xlab("")+ylab("")+theme_light(base_size = 16) 

# Number of samples per region per season over time
dat1 %>% group_by(Region, Year, Season)%>%summarize(n=n()) %>%
  ggplot(aes(x=Year,y=Region))+geom_tile(aes(fill=n))+
  scale_fill_gradient(low="white",high="red")+
  geom_text(aes(label=round(n,1)),size=4)+
  xlab("")+ylab("")+
  facet_grid(Season~.)+
  theme_bw(base_size=14)+
  theme(legend.position = "none")

# Number of samples per region per month over time
dat1 %>% group_by(Region, Year, Month)%>%summarize(n=n()) %>%
  ggplot(aes(x=Year,y=Region))+geom_tile(aes(fill=n))+
  scale_fill_gradient(low="white",high="red")+
  geom_text(aes(label=round(n,1)),size=4)+
  xlab("")+ylab("")+
  facet_grid(Month~.)+
  theme_bw(base_size=14)+
  theme(legend.position = "none")

# Number of samples per station and season over time
dat1 %>% group_by(Station, Region,Season,Year)%>%summarize(n=n()) %>%
  ggplot(aes(x=Year,y=as.factor(Station)))+geom_tile(aes(fill=n))+
  scale_fill_gradient(low="white",high="red")+
  geom_text(aes(label=round(n,1)),size=4)+
  xlab("")+ylab("")+facet_grid(Region~Season,scales = 'free_y',space='free')+
  theme_bw(base_size=14)+
  theme(legend.position = "none")+
  theme(axis.text.x=element_text(angle = -90, vjust = 0.3))

rm(deltabox,stamenmap)
```


## Backcalculate HAB/BAB using MVI and ptox
```{r, warning=FALSE, message=FALSE}

library(lme4)

# Prepare data
dat2 <- dat1 %>% filter(Month %in% (6:10))

train_dat <- dat2 %>% filter(!is.na(ptox) & !is.na(mvi) & !is.na(Chlorophyll)) %>%
  select(Month,Date,Region,Station,Chlorophyll,mvi,ptox) %>%
  group_by(Month,Date,Region,Station,Chlorophyll,mvi,ptox) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))

missing_dat <- dat2 %>% filter(is.na(mvi) & !is.na(ptox) & !is.na(Chlorophyll)) %>%
  select(Month,Date,Region,Station,Chlorophyll,mvi,ptox) %>%
  group_by(Month,Date,Region,Station,Chlorophyll,mvi,ptox) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))

# Fit logistic model
model <- glmer(mvi ~ ptox + Region * Chlorophyll + (1 | Month), family = binomial, data = train_dat)

# Use model to predict values on training data
train_dat$prob_harmful <- predict(model,newdata = train_dat,type = "response")

train_dat <- train_dat %>% 
  mutate(pred_class = case_when(prob_harmful >= 0.4 ~ "HAB", prob_harmful < 0.4 ~ "BAB")) %>%
  mutate(HAB_BAB_class = case_when(mvi == 1 ~ "HAB", mvi == 0 ~ "BAB"))

# Use model to predict values on missing data
missing_dat$prob_harmful <- predict(model,newdata = missing_dat,type = "response")

missing_dat <- missing_dat %>% 
  mutate(pred_class = case_when(prob_harmful >= 0.4 ~ "HAB", prob_harmful < 0.4 ~ "BAB"))
missing_dat$HAB_BAB_class <- missing_dat$pred_class;

# Add modeled data to main data frame
dat1$prob_harmful <- NA; dat1$pred_class <- NA; dat1$HAB_BAB_class <- NA; dat1$modeled <- NA
dat1 <- dat1 %>%
  mutate(
    prob_harmful  = ifelse(Month %in% c(11, 12, 1:5), NA_real_, prob_harmful),
    pred_class = ifelse(Month %in% c(11, 12, 1:5), NA_character_, pred_class),
    HAB_BAB_class = ifelse(Month %in% c(11, 12, 1:5), "BAB", HAB_BAB_class),
    modeled       = ifelse(Month %in% c(11, 12, 1:5), "No", modeled)
  )

train_dat$modeled <- "No"
missing_dat$modeled <- "Yes"

dat1 <- dat1 %>%
  rows_update(missing_dat %>% select(Month, Date, Station, prob_harmful, pred_class, HAB_BAB_class, modeled),
              by = c("Month", "Date", "Station")) %>%
  rows_update(train_dat %>% select(Month, Date, Station, prob_harmful, pred_class, HAB_BAB_class, modeled),
              by = c("Month", "Date", "Station"))

# Add in third designation - HAB, BAB and no bloom
dat1 <- dat1 %>%
  mutate(harmful_class = case_when(
    HAB_BAB_class == "HAB" ~ "HAB",
    HAB_BAB_class == "BAB" & bloom == "Yes" ~ "BAB",
    HAB_BAB_class == "BAB" & bloom == "No" ~ "No bloom",
    TRUE ~ NA_character_
  ))

rm(dat2,model)

```


## Reorder and save final dataset
```{r, warning=FALSE, message=FALSE}

dat <- dat1[, c("Survey","Year","Season","Month","Date","Region","Station","Latitude","Longitude",
                 "SampleMethod","SampleDepth","DepthType","Lab","CountMethodSample","CountMethodTaxa",
                 "Magnification","AlgalGroup","Cells_per_mL","Units_per_mL","Biovolume_per_mL","tot_cells",
                 "tot_units","tot_bio","relative_cell_density","relative_unit_density","relative_biovolume",
                 "bloom","ptox","mvi","prob_harmful","pred_class","HAB_BAB_class","modeled","harmful_class",
                 "Sample_depth_surface","Sample_depth_nutr_surface","Sample_depth_bottom","Microcystis",
                 "Chlorophyll","Pheophytin","Temperature","Temperature_bottom","Conductivity",
                 "Conductivity_bottom","Salinity","Salinity_bottom","Secchi","TurbidityNTU",
                 "TurbidityNTU_bottom","TurbidityFNU","TurbidityFNU_bottom","DissolvedOxygen",
                 "DissolvedOxygen_bottom","DissolvedOxygenPercent","DissolvedOxygenPercent_bottom","pH",
                 "DissAmmonia","DissNitrateNitrite","TotPhos","DissOrthophos","TON","DON","TKN",
                 "DissBromide","DissCalcium","TotChloride","DissChloride","DissSilica","TOC","DOC","TDS",
                 "TSS","VSS","SAC","SJR","OUT","EXPORTS","Index","Biomass","Grazing_rate","Filtration_rate",
                 "monthly_biomass","monthly_grazing_rate","monthly_filtration_rate")]

save(dat, clam, df, pesp, wq, file = "Data/integrated_dataset_2008-2022.RData")

```


## HAB/BAB modeled vs observed figures
```{r, warning=FALSE, message=FALSE}

# Plot final dataset
p<- ggplot()+
  geom_path(data = dat1 %>% filter(Region=="Confluence" & !is.na(Chlorophyll)),
            aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Confluence")+
  geom_point(data = dat1 %>% filter(Region=="Confluence" & !is.na(Chlorophyll)),
             aes(x=Date,y=Chlorophyll,fill=as.factor(harmful_class),shape=as.factor(modeled)),size=1)+
  scale_fill_manual(values = c("blue", "red","grey"))+scale_shape_manual(values = c(21, 24)) + 
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+labs(fill="HAB vs. BAB",shape="Modeled?",x="",y="")

p1<- ggplot()+
  geom_path(data = dat1 %>% filter(Region=="Suisun Bay" & !is.na(Chlorophyll)),
            aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Bay")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay" & !is.na(Chlorophyll)),
             aes(x=Date,y=Chlorophyll,fill=as.factor(harmful_class),shape=as.factor(modeled)),size=1)+
  scale_fill_manual(values = c("blue", "red","grey"))+scale_shape_manual(values = c(21, 24)) + 
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+labs(fill="HAB vs. BAB",shape="Modeled?",x="",y="")


p2 <- ggplot()+
  geom_path(data = dat1 %>% filter(Region=="Central" & !is.na(Chlorophyll)),
            aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Central")+
  geom_point(data = dat1 %>% filter(Region=="Central" & !is.na(Chlorophyll)),
             aes(x=Date,y=Chlorophyll,fill=as.factor(harmful_class),shape=as.factor(modeled)),size=1)+
  scale_fill_manual(values = c("blue", "red","grey"))+scale_shape_manual(values = c(21, 24)) + 
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+labs(fill="HAB vs. BAB",shape="Modeled?",x="",y="")

tiff("../phyto_synthesis/Figures/HAB v BAB/final_hab_bab_dataset.tiff", units="in", width=12, height=8, res=300)
ggarrange(p2,p,p1,ncol=1,common.legend=TRUE)
dev.off()


# Compare real MVI to predicted MVI in training data
train_dat <- train_dat %>%
  mutate(color = case_when(
    pred_class == "BAB" & HAB_BAB_class == "BAB" ~ "blue",
    pred_class == "HAB" & HAB_BAB_class == "HAB" ~ "red",
    pred_class == "BAB" & HAB_BAB_class == "HAB" ~ "orange",
    pred_class == "HAB" & HAB_BAB_class == "BAB" ~ "green",
    TRUE ~ NA_character_
  ))
train_dat <- train_dat %>% arrange(Date)

p <- ggplot()+
  geom_point(data = train_dat %>% filter(Region=="Central"),
             aes(x=Date,y=Chlorophyll,fill=as.factor(color)),pch=21,size=2)+
  ggtitle("Central Delta, training")+
  scale_fill_manual(values = c("blue", "green","orange","red"),
                    labels=c("Mod BAB, Obs BAB","Mod HAB, Obs BAB","Mod BAB, Obs HAB","Mod HAB, Obs HAB"))+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+
  labs(fill="Modeled vs. Observed",x="",y="Chl-a (ug/L)")

p1 <- ggplot()+
  geom_point(data = train_dat %>% filter(Region=="Confluence"),
             aes(x=Date,y=Chlorophyll,fill=as.factor(color)),pch=21,size=2)+
  ggtitle("Confluence, training")+
  scale_fill_manual(values = c("blue", "green","orange","red"),
                    labels=c("Mod BAB, Obs BAB","Mod HAB, Obs BAB","Mod BAB, Obs HAB","Mod HAB, Obs HAB"))+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+
  labs(fill="Modeled vs. Observed",x="",y="Chl-a (ug/L)")

p2 <- ggplot()+
  geom_point(data = train_dat %>% filter(Region=="Suisun Bay"),
             aes(x=Date,y=Chlorophyll,fill=as.factor(color)),pch=21,size=2)+
  ggtitle("Suisun Bay, training")+
  scale_fill_manual(values = c("blue", "green","orange","red"),
                    labels=c("Mod BAB, Obs BAB","Mod HAB, Obs BAB","Mod BAB, Obs HAB","Mod HAB, Obs HAB"))+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+
  labs(fill="Modeled vs. Observed",x="",y="Chl-a (ug/L)")

tiff("../phyto_synthesis/Figures/HAB v BAB/real_v_modeled.tiff", units="in", width=12, height=8, res=300)
ggarrange(p,p1,p2,ncol=1,common.legend=TRUE,legend="bottom")
dev.off()

```


## HAB/BAB model validation
```{r, warning=FALSE, message=FALSE}

library(pROC)      # AUC
library(caret) 
library(lme4)

# Prepare data
dat2 <- dat1 %>% filter(Month %in% (6:10))

train_dat <- dat2 %>% filter(!is.na(ptox) & !is.na(mvi) & !is.na(Chlorophyll)) %>%
  select(Month,Date,Region,Station,Chlorophyll,mvi,ptox) %>%
  group_by(Month,Date,Region,Station,Chlorophyll,mvi,ptox) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))

missing_dat <- dat2 %>% filter(is.na(mvi) & !is.na(ptox) & !is.na(Chlorophyll)) %>%
  select(Month,Date,Region,Station,Chlorophyll,mvi,ptox) %>%
  group_by(Month,Date,Region,Station,Chlorophyll,mvi,ptox) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))

# Fit logistic model
model <- glmer(mvi ~ ptox + Region * Chlorophyll + (1 | Month), family = binomial, data = train_dat)
summary(model)

# Cross validation
set.seed(123)
ctrl <- trainControl(method = "cv",number = 5,classProbs = TRUE,summaryFunction = twoClassSummary)
train_dat$harmful_factor <- factor(ifelse(train_dat$mvi == 1, "yes", "no"))

cv_model <- train(harmful_factor ~ ptox + Region * Chlorophyll + (1 | Month),
  data = train_dat,method = "glm",family = "binomial",trControl = ctrl,metric = "PR")
mean(cv_model$resample$ROC)
sd(cv_model$resample$ROC)

# Check model output results
train_dat$prob_harmful <- predict(model, type = "response")
ggplot(data=train_dat)+geom_point(aes(y=prob_harmful,x=Chlorophyll))
ggplot(data=train_dat)+geom_point(aes(y=prob_harmful,x=ptox))

# Validate model

# AUC
roc_obj <- roc(train_dat$mvi, train_dat$prob_harmful)
auc_val <- auc(roc_obj)

print(auc_val)
# AUC ≥ 0.75 → strong enough to justify back-calculation
# AUC 0.65–0.75 → usable, but report uncertainty carefully
# AUC < 0.65 → ptox alone may not be reliable predictor

# AUC-PR

# Get predicted probabilities (not class predictions)
probs <- predict(model,newdata = train_dat,type = "response")
actual <- train_dat$mvi

# Precision-Recall curve
pr <- PRROC::pr.curve(scores.class0 = probs, weights.class0 = actual, curve = TRUE)
plot(pr)
pr$auc.integral  # Your AUC-PR value

# Confusion matrices
train_dat$prob_harmful <- predict(model,newdata = train_dat,type = "response")

train_dat <- train_dat %>% 
  mutate(pred_class = case_when(prob_harmful >= 0.4 ~ "HAB", prob_harmful < 0.4 ~ "BAB")) %>%
  mutate(HAB_BAB_class = case_when(mvi == 1 ~ "HAB", mvi == 0 ~ "BAB"))

confusionMatrix(factor(train_dat$pred_class),factor(train_dat$HAB_BAB_class), positive = "HAB")

```



## (no longer use) Method 1: Define HAB v BAB using MVI
Considered a BAB if MVI 1-2 and a HAB if MVI 3-5
```{r, warning=FALSE, message=FALSE}

# Define HAB as MVI 3-5
dat1$HAB_MVI <- NA
I <- which(dat1$Microcystis %in% c(3:5) & dat1$Month %in% c(6:10)); dat1$HAB_MVI[I] <- "HAB"; rm(I)
I <- which(dat1$Microcystis %in% c(3:5) & dat1$Month %in% c(1:5,11:12)); dat1$HAB_MVI[I] <- "BAB"; rm(I)
I <- which(dat1$Microcystis %in% c(1:2)); dat1$HAB_MVI[I] <- "BAB"; rm(I)
dat1 <- dat1 %>% filter(Date > "2008-01-01")

# Plot
p<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Suisun Marsh"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Marsh")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Marsh"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),pch=21,color="gray")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Marsh"&bloom=="Yes"&HAB_MVI=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="green")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Marsh"&bloom=="Yes"&HAB_MVI=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p11<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Suisun Bay"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Bay, MVI")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="gray")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay"&bloom=="Yes"&HAB_MVI=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay"&bloom=="Yes"&HAB_MVI=="HAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+ylab("")+xlab("")
p9<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Confluence"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Confluence, MVI")+
  geom_point(data = dat1 %>% filter(Region=="Confluence"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="gray")+
  geom_point(data = dat1 %>% filter(Region=="Confluence"&bloom=="Yes"&HAB_MVI=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Region=="Confluence"&bloom=="Yes"&HAB_MVI=="HAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+ylab("Chlorophyll (µg/L)")+xlab("")
p3<-ggplot()+geom_path(data = dat1 %>% filter(Region=="North Delta"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("North Delta")+
  geom_point(data = dat1 %>% filter(Region=="North Delta"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Region=="North Delta"&bloom=="Yes"&HAB_MVI=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="North Delta"&bloom=="Yes"&HAB_MVI=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p10<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Central"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Central, MVI")+
  geom_point(data = dat1 %>% filter(Region=="Central"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="gray")+
  geom_point(data = dat1 %>% filter(Region=="Central"&bloom=="Yes"&HAB_MVI=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Region=="Central"&bloom=="Yes"&HAB_MVI=="HAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+ylab("")+xlab("")
p5<-ggplot()+geom_path(data = dat1 %>% filter(Region=="South"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("South")+
  geom_point(data = dat1 %>% filter(Region=="South"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Region=="South"&bloom=="Yes"&HAB_MVI=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="South"&bloom=="Yes"&HAB_MVI=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
#ggarrange(p,p1,p2,p3,p4,p5,ncol=2,nrow=3)

tiff("../phyto_synthesis/Figures/HABvBAB_MVI.jpeg", units="in", width=5, height=6, res=300)
ggarrange(p11,p9,p10,ncol=1)
dev.off()

rm(p,p1,p2,p3,p4,p5)

```

## (no longer use) Method 2a: Define HAB v BAB using cyanobacterial density from WHO (2006)
Density thresholds from World Health Organization guidance levels, found in Kudela et al. (2023).
Could classify "HAB" as medium and high risk, or just high risk.
```{r, warning=FALSE, message=FALSE}

# Aggregate by group
grp <- dat1 %>% group_by(Date,Survey,Region,Station,AlgalGroup) %>%
  summarize(Density = sum(Cells_per_mL),Biovolume = sum(Biovolume_per_mL))
tot <- dat1 %>% group_by(Date,Survey,Region,Station) %>% summarise(Tot_Density = sum(Cells_per_mL),Tot_Biovolume = sum(Biovolume_per_mL))
grp <- grp %>% left_join(tot,by=c("Date","Survey","Region","Station"))

c <- grp %>% filter(AlgalGroup == "Cyanobacteria")

# Define high, medium, and low risk WHO density thresholds
c$HAB_cyano <- NA
I <- which(c$Density <20000); c$HAB_cyano[I] <- "Low"; rm(I)
I <- which(c$Density >=20000 & c$Density <100000); c$HAB_cyano[I] <- "Med"; rm(I)
I <- which(c$Density >=100000); c$HAB_cyano[I] <- "High"; rm(I)
c <- c[,c(1:4,6,10)]
c <- c %>% rename(Density_cyano = Density)

dat1 <- left_join(dat1, c, by = c("Date","Survey","Region","Station"))

# Plot
p<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Suisun Marsh"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Marsh")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Marsh"&bloom=="Yes"&HAB_cyano=="Low"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Marsh"&bloom=="Yes"&HAB_cyano=="Med"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Marsh"&bloom=="Yes"&HAB_cyano=="High"),aes(x=Date,y=Chlorophyll),color="red")
p1<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Suisun Bay"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Bay")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay"&bloom=="Yes"&HAB_cyano=="Low"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay"&bloom=="Yes"&HAB_cyano=="Med"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay"&bloom=="Yes"&HAB_cyano=="High"),aes(x=Date,y=Chlorophyll),color="red")
p2<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Confluence"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Confluence")+
  geom_point(data = dat1 %>% filter(Region=="Confluence"&bloom=="Yes"&HAB_cyano=="Low"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="Confluence"&bloom=="Yes"&HAB_cyano=="Med"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Region=="Confluence"&bloom=="Yes"&HAB_cyano=="High"),aes(x=Date,y=Chlorophyll),color="red")
p3<-ggplot()+geom_path(data = dat1 %>% filter(Region=="North Delta"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("North Delta")+
  geom_point(data = dat1 %>% filter(Region=="North Delta"&bloom=="Yes"&HAB_cyano=="Low"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="North Delta"&bloom=="Yes"&HAB_cyano=="Med"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Region=="North Delta"&bloom=="Yes"&HAB_cyano=="High"),aes(x=Date,y=Chlorophyll),color="red")
p4<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Central"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Central")+
  geom_point(data = dat1 %>% filter(Region=="Central"&bloom=="Yes"&HAB_cyano=="Low"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="Central"&bloom=="Yes"&HAB_cyano=="Med"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Region=="Central"&bloom=="Yes"&HAB_cyano=="High"),aes(x=Date,y=Chlorophyll),color="red")
p5<-ggplot()+geom_path(data = dat1 %>% filter(Region=="South"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("South")+
  geom_point(data = dat1 %>% filter(Region=="South"&bloom=="Yes"&HAB_cyano=="Low"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="South"&bloom=="Yes"&HAB_cyano=="Med"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Region=="South"&bloom=="Yes"&HAB_cyano=="High"),aes(x=Date,y=Chlorophyll),color="red")
#ggarrange(p,p1,p2,p3,p4,p5,ncol=2,nrow=3)
ggarrange(p1,p2,p4,ncol=1)

ggplot() + geom_histogram(data=c, aes(x=Density_cyano))

rm(p,p1,p2,p3,p4,p5,c,grp,tot)

```

## (no longer use) Method 2b: Define HAB v BAB using cyanobacterial biovolume from WHO (2021)
Biovolume thresholds from World Health Organization guidance levels, 2021.
Could classify "HAB" as alert levels 1 and 2, or just alert level 2.
```{r, warning=FALSE, message=FALSE}

# Aggregate by group
grp <- dat1 %>% group_by(Date,Survey,Region,Station,AlgalGroup) %>%
  summarize(Density = sum(Cells_per_mL),Biovolume = sum(Biovolume_per_mL))
tot <- dat1 %>% group_by(Date,Survey,Region,Station) %>% summarise(Tot_Density = sum(Cells_per_mL),Tot_Biovolume = sum(Biovolume_per_mL))
grp <- grp %>% left_join(tot,by=c("Date","Survey","Region","Station"))

c <- grp %>% filter(AlgalGroup == "Cyanobacteria")

# Define alert level 1 and 2 WHO biovolume thresholds
c$HAB_cyano2 <- NA
I <- which(c$Biovolume <4e+6); c$HAB_cyano2[I] <- "BAB"; rm(I)
I <- which(c$Biovolume >=4e+6 & c$Biovolume <8e+6); c$HAB_cyano2[I] <- "AlertL1"; rm(I)
I <- which(c$Biovolume >=8e+6); c$HAB_cyano2[I] <- "AlertL2"; rm(I)
c <- c[,c(1:4,7,10)]
c <- c %>% rename(Biovolume_cyano = Biovolume)

dat1 <- left_join(dat1, c, by = c("Date","Survey","Region","Station"))

# Plot
p<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Suisun Marsh"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Marsh")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Marsh"&bloom=="Yes"&HAB_cyano2=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Marsh"&bloom=="Yes"&HAB_cyano2=="AlertL1"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Marsh"&bloom=="Yes"&HAB_cyano2=="AlertL2"),aes(x=Date,y=Chlorophyll),color="red")
p1<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Suisun Bay"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Bay")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="gray")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay"&bloom=="Yes"&HAB_cyano2=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay"&bloom=="Yes"&HAB_cyano2=="AlertL1"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="orange")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay"&bloom=="Yes"&HAB_cyano2=="AlertL2"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+ylab("")+xlab("")
p2<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Confluence"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Confluence")+
  geom_point(data = dat1 %>% filter(Region=="Confluence"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="gray")+
  geom_point(data = dat1 %>% filter(Region=="Confluence"&bloom=="Yes"&HAB_cyano2=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Region=="Confluence"&bloom=="Yes"&HAB_cyano2=="AlertL1"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="orange")+
  geom_point(data = dat1 %>% filter(Region=="Confluence"&bloom=="Yes"&HAB_cyano2=="AlertL2"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+ylab("Chlorophyll (µg/L)")+xlab("")
p3<-ggplot()+geom_path(data = dat1 %>% filter(Region=="North Delta"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("North Delta")+
  geom_point(data = dat1 %>% filter(Region=="North Delta"&bloom=="Yes"&HAB_cyano2=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="North Delta"&bloom=="Yes"&HAB_cyano2=="AlertL1"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Region=="North Delta"&bloom=="Yes"&HAB_cyano2=="AlertL2"),aes(x=Date,y=Chlorophyll),color="red")
p4<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Central"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Central")+
  geom_point(data = dat1 %>% filter(Region=="Central"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="gray")+
  geom_point(data = dat1 %>% filter(Region=="Central"&bloom=="Yes"&HAB_cyano2=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Region=="Central"&bloom=="Yes"&HAB_cyano2=="AlertL1"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="orange")+
  geom_point(data = dat1 %>% filter(Region=="Central"&bloom=="Yes"&HAB_cyano2=="AlertL2"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+ylab("")+xlab("")
p5<-ggplot()+geom_path(data = dat1 %>% filter(Region=="South"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("South")+
  geom_point(data = dat1 %>% filter(Region=="South"&bloom=="Yes"&HAB_cyano2=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="South"&bloom=="Yes"&HAB_cyano2=="AlertL1"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Region=="South"&bloom=="Yes"&HAB_cyano2=="AlertL2"),aes(x=Date,y=Chlorophyll),color="red")
#ggarrange(p,p1,p2,p3,p4,p5,ncol=2,nrow=3)

tiff("../phyto_synthesis/Figures/HABvBAB_WHO.jpeg", units="in", width=5, height=6, res=300)
ggarrange(p1,p2,p4,ncol=1)
dev.off()



rm(p,p1,p2,p3,p4,p5,c,grp,tot)

```

## (no longer use) Method 3: Define HAB v BAB using cyanobacterial relative abundance
Define HAB as a cyanobacterial relative biovolume > 50%
```{r, warning=FALSE, message=FALSE}

# Aggregate by group
grp <- dat1 %>% group_by(Date,Survey,Region,Station,AlgalGroup) %>%
  summarize(Density = sum(Cells_per_mL),Biovolume = sum(Biovolume_per_mL))
tot <- dat1 %>% group_by(Date,Survey,Region,Station) %>% summarise(Tot_Density = sum(Cells_per_mL),Tot_Biovolume = sum(Biovolume_per_mL))
grp <- grp %>% left_join(tot,by=c("Date","Survey","Region","Station"))
grp$Rel_Biovolume <- (grp$Biovolume/grp$Tot_Biovolume)*100;
grp$Rel_Density <- (grp$Density/grp$Tot_Density)*100;

c <- grp %>% filter(AlgalGroup == "Cyanobacteria")

# Define HAB as a cyanobacterial relative biovolume > 50%
x = mean(c$Rel_Biovolume,na.rm=TRUE) + sd(c$Rel_Biovolume,na.rm=TRUE)
c$HAB_cyano_ra <- NA
I <- which(c$Rel_Biovolume <x); c$HAB_cyano_ra[I] <- "BAB"; rm(I)
I <- which(c$Rel_Biovolume >=x); c$HAB_cyano_ra[I] <- "HAB"; rm(I)
c <- c[,c(1:4,10,12)]
c <- c %>% rename(Rel_Biovolume_cyano = Rel_Biovolume)

dat1 <- left_join(dat1, c, by = c("Date","Survey","Region","Station"))

# Plot
p<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Suisun Marsh"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Marsh")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Marsh"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Marsh"&bloom=="Yes"&HAB_cyano_ra=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Marsh"&bloom=="Yes"&HAB_cyano_ra=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p1<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Suisun Bay"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Bay")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay"&bloom=="Yes"&HAB_cyano_ra=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay"&bloom=="Yes"&HAB_cyano_ra=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p2<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Confluence"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Confluence")+
  geom_point(data = dat1 %>% filter(Region=="Confluence"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Region=="Confluence"&bloom=="Yes"&HAB_cyano_ra=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="Confluence"&bloom=="Yes"&HAB_cyano_ra=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p3<-ggplot()+geom_path(data = dat1 %>% filter(Region=="North Delta"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("North Delta")+
  geom_point(data = dat1 %>% filter(Region=="North Delta"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Region=="North Delta"&bloom=="Yes"&HAB_cyano_ra=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="North Delta"&bloom=="Yes"&HAB_cyano_ra=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p4<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Central"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Central")+
  geom_point(data = dat1 %>% filter(Region=="Central"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Region=="Central"&bloom=="Yes"&HAB_cyano_ra=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="Central"&bloom=="Yes"&HAB_cyano_ra=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p5<-ggplot()+geom_path(data = dat1 %>% filter(Region=="South"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("South")+
  geom_point(data = dat1 %>% filter(Region=="South"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Region=="South"&bloom=="Yes"&HAB_cyano_ra=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="South"&bloom=="Yes"&HAB_cyano_ra=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
#ggarrange(p,p1,p2,p3,p4,p5,ncol=2,nrow=3)
ggarrange(p1,p2,p4,ncol=1)

ggplot() + geom_histogram(data=c, aes(x=Rel_Biovolume_cyano))

rm(c,grp,tot,p,p1,p2,p3,p4,p5,x)

```

## (no longer use) Method 4: Define HAB v BAB using presence of ptox species
```{r, warning=FALSE, message=FALSE}

# Flag ptox species
dat1$ptox_flag <- NA
I <- which(dat1$Genus %in% c("Microcoleus","Cuspidothrix","Aphanizomenon",
                             "Microcystis","Planktothrix") &
             !dat1$Taxon == "Aphanizomenon flos-aquae" | 
             dat1$Taxon %in% c("Dolichospermum circinale","Microseira wollei"))
dat1$ptox_flag[I]<- "Yes"; 
dat1$ptox_flag[is.na(dat1$ptox_flag)] <- "No"


# Aggregate by ptox presence/absence
#dat1$ptox_flag <- factor(dat1$ptox_flag)
grp <- dat1 %>% group_by(Date,Month,Survey,Region,Station,ptox_flag) %>% #,.drop = FALSE) %>%
  summarize(Density = sum(Cells_per_mL),Biovolume = sum(Biovolume_per_mL))
tot <- dat1 %>% group_by(Date,Month,Survey,Region,Station) %>% summarise(Tot_Density = sum(Cells_per_mL),Tot_Biovolume = sum(Biovolume_per_mL))
grp <- grp %>% left_join(tot,by=c("Date","Month","Survey","Region","Station"))
grp$Rel_Biovolume <- (grp$Biovolume/grp$Tot_Biovolume)*100;
grp$Rel_Density <- (grp$Density/grp$Tot_Density)*100;

c <- grp %>% filter(ptox_flag == "Yes")

# Define HAB when ptox > 0
c$HAB_ptox <- NA
I <- which(c$Month %in% c(6:10)); c$HAB_ptox[I] <- "HAB"; rm(I)
I <- which(c$Month %in% c(1:5,11:12)); c$HAB_ptox[I] <- "BAB"; rm(I)
c <- c[,c(1:5,11,13)]
c <- c %>% rename(Rel_Biovolume_ptox = Rel_Biovolume)

dat1 <- left_join(dat1, c, by = c("Date","Month","Survey","Region","Station"))
dat1$HAB_ptox[is.na(dat1$HAB_ptox)] <- "BAB"
I = which(!is.na(dat1$Biovolume_per_mL) & is.na(dat1$Rel_Biovolume_ptox)) 
dat1$Rel_Biovolume_ptox[I]<- 0


# Plot
p2<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Suisun Marsh"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Marsh")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Marsh"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Marsh"&bloom=="Yes"&HAB_ptox=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Marsh"&bloom=="Yes"&HAB_ptox=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p3<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Suisun Bay"& Date > "2008-01-01"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Bay,ptox presence/absence")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay"&bloom=="Yes"& Date > "2008-01-01"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay"&bloom=="Yes"&HAB_ptox=="BAB"& Date > "2008-01-01"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay"&bloom=="Yes"&HAB_ptox=="HAB"& Date > "2008-01-01"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+xlab("")+ylab("")
p4<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Confluence" & Date > "2008-01-01"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Confluence,ptox presence/absence")+
  geom_point(data = dat1 %>% filter(Region=="Confluence"&bloom=="Yes"& Date > "2008-01-01"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Region=="Confluence"&bloom=="Yes"&HAB_ptox=="BAB"& Date > "2008-01-01"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Region=="Confluence"&bloom=="Yes"&HAB_ptox=="HAB"& Date > "2008-01-01"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+xlab("")+ylab("Chlorophyll (µg/L)")
p5<-ggplot()+geom_path(data = dat1 %>% filter(Region=="North Delta"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("North Delta")+
  geom_point(data = dat1 %>% filter(Region=="North Delta"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Region=="North Delta"&bloom=="Yes"&HAB_ptox=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="North Delta"&bloom=="Yes"&HAB_ptox=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p6<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Central"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Central,ptox presence/absence")+
  geom_point(data = dat1 %>% filter(Region=="Central"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Region=="Central"&bloom=="Yes"&HAB_ptox=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Region=="Central"&bloom=="Yes"&HAB_ptox=="HAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+xlab("")+ylab("")
p7<-ggplot()+geom_path(data = dat1 %>% filter(Region=="South"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("South")+
  geom_point(data = dat1 %>% filter(Region=="South"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Region=="South"&bloom=="Yes"&HAB_ptox=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Region=="South"&bloom=="Yes"&HAB_ptox=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
#ggarrange(p,p1,p2,p3,p4,p5,ncol=2,nrow=3)

tiff("../phyto_synthesis/Figures/HABvBAB_ptox.jpeg", units="in", width=5, height=6, res=300)
ggarrange(p3,p4,p6,ncol=1)
dev.off()

```


## (no longer use) Compare ptox to MVI
```{r, warning=FALSE, message=FALSE}

ggplot(data=dat1 %>% filter(bloom == "Yes" & Microcystis > 1))+geom_point(aes(x=Microcystis,y=Rel_Biovolume_ptox))

test <- dat1 %>% select(Date,Region,Station,bloom,Microcystis,Rel_Biovolume_ptox,HAB_MVI,HAB_ptox)
test<-unique(test)
test <- test %>% filter(!is.na(HAB_MVI) & !is.na(HAB_ptox) &bloom == "Yes")
test$compare <- ifelse(test$HAB_MVI == test$HAB_ptox, "same", "different")
# 88 blooms with MVI and ptox

test1 <- test %>% filter(compare == "same") 
# 75 blooms that classify differently, 13 blooms that classify differently

test1 <- test %>% filter(compare == "different" & HAB_MVI == "HAB")
# Of those 13 differences, 
  # 2 instances where MVI classifies BAB and ptox classifies HAB (MVI = 2 for both)
  # 11 instances where MVI classifies HAB and ptox classifies BAB (MVI = 3 for 6 of those instances and 4 for 5 of those instances)

test <- dat1 %>% select(Date,Region,Station,bloom,Microcystis,Rel_Biovolume_ptox,HAB_MVI,HAB_ptox)
test<-unique(test)
test1 <- test %>% filter(bloom == "Yes" & HAB_ptox == "HAB")

ggplot(data=test1)+geom_histogram(aes(Rel_Biovolume_ptox))


```


## (no longer use) Compare agreement between methods
```{r, warning=FALSE, message=FALSE}

# Method 1: MVI
test <- dat1 %>% group_by(Date,Station,bloom,Microcystis,HAB_MVI) %>%
  summarize(Chl = mean(Chlorophyll)) %>%
  filter(bloom == "Yes" & Microcystis > 0 & HAB_MVI == "BAB")

#185 blooms after 2007
#88 blooms with corresponding MVI data
#19 HABs, 69 BABs

# Method 2: Cyanobacterial density thresholds
test <- dat1 %>% group_by(Date,Station,bloom,Density_cyano,HAB_cyano) %>%
  summarize(Chl = mean(Chlorophyll)) %>%
  filter(bloom == "Yes" & Density_cyano > 0 & HAB_cyano %in% c("Med","High"))

#185 blooms after 2007
#162 blooms with corresponding density data when cyanobacteria were present
#31 blooms where thresholds were "Low" risk
#131 blooms where thresholds were "Medium" or "High" risk
#88 blooms where thresholds were "High" risk

# Method 3: Cyanobacterial relative biovolume > 50%
test <- dat1 %>% group_by(Date,Station,bloom,Rel_Biovolume_cyano,HAB_cyano_ra) %>%
  summarize(Chl = mean(Chlorophyll)) %>%
  filter(bloom == "Yes" & Rel_Biovolume_cyano > 0 & HAB_cyano_ra == "BAB")

#185 blooms after 2007
#122 blooms with corresponding biovolume data when cyanobacteria were present
#18 HABs, 104 BABs


# Plots
test <-data.frame(MVI = dat1$HAB_MVI, Density = dat1$HAB_cyano, RelBio = dat1$HAB_cyano_ra)
test[is.na(test)] <- "No data"

test$MVI <- factor(test$MVI,levels=c("No data","BAB","HAB"))
test$Density <- factor(test$Density, levels=c("No data","Low","Med","High"))
test$RelBio <- factor(test$RelBio, levels=c("No data","BAB","HAB"))

test %>% group_by(MVI, Density)%>%summarize(n=n()) %>%
  ggplot(aes(x=MVI,y=Density))+geom_tile(aes(fill=n))+
  scale_fill_gradient(low="white",high="red")+
  geom_text(aes(label=round(n,1)),size=4)+
  xlab("Method 1: MVI")+ylab("Method 2: Density Threshold")+
  theme_bw(base_size=14)+
  theme(legend.position = "none")

test %>% group_by(MVI, RelBio)%>%summarize(n=n()) %>%
  ggplot(aes(x=MVI,y=RelBio))+geom_tile(aes(fill=n))+
  scale_fill_gradient(low="white",high="red")+
  geom_text(aes(label=round(n,1)),size=4)+
  xlab("Method 1: MVI")+ylab("Method 3: Relative Biovolume > 50%")+
  theme_bw(base_size=14)+
  theme(legend.position = "none")

test %>% group_by(Density, RelBio)%>%summarize(n=n()) %>%
  ggplot(aes(x=Density,y=RelBio))+geom_tile(aes(fill=n))+
  scale_fill_gradient(low="white",high="red")+
  geom_text(aes(label=round(n,1)),size=4)+
  xlab("Method 2: Density Threshold")+ylab("Method 3: Relative Biovolume > 50%")+
  theme_bw(base_size=14)+
  theme(legend.position = "none")

```

## (no longer use) Test models
```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(pROC)      # AUC
library(caret) 
library(lme4)

# Prepare data
dat1$HAB_ptox2 <- ifelse(dat1$HAB_ptox == "HAB", 1, 0)
dat1$HAB_MVI2 <- ifelse(dat1$HAB_MVI == "HAB", 1, 0)
dat1 <- dat1 %>% filter(Region %in% c("Central","Suisun Bay","Confluence"))

#test
dat1$mvi<-NA; I = which(dat1$Microcystis %in% c(1:2)); dat1$mvi[I]<- 0
I = which(dat1$Microcystis %in% c(3:5)); dat1$mvi[I]<- 1
dat1$ptox <- ifelse(dat1$ptox_flag == "Yes", 1, 0)
train_dat <- dat1 %>% filter(!is.na(ptox) & !is.na(mvi) & Month %in% c(6:10)) %>%
  select(Month,Date,Region,Station,Chlorophyll,Temperature,Microcystis,ptox_flag,mvi,ptox) %>%
  group_by(Month,Date,Region,Station,Chlorophyll,Temperature,Microcystis,ptox_flag,mvi,ptox) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))
train_dat2 <- dat1 %>% filter(!is.na(ptox) & !is.na(mvi) & Month %in% c(6:10)&bloom=="Yes") %>%
  select(Month,Date,Region,Station,Chlorophyll,Temperature,Microcystis,ptox_flag,mvi,ptox) %>%
  group_by(Month,Date,Region,Station,Chlorophyll,Temperature,Microcystis,ptox_flag,mvi,ptox) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))

# Preliminary figures
ggplot(data=train_dat)+
  geom_point(aes(x=Chlorophyll,y=Microcystis,color=as.factor(ptox)))+
  ggtitle("All sampling dates")


counts1 <- train_dat %>%group_by(Microcystis) %>%
  summarise(n = sum(!is.na(Chlorophyll)), .groups = "drop")
counts2 <- train_dat2 %>%group_by(Microcystis) %>%
  summarise(n = sum(!is.na(Chlorophyll)), .groups = "drop")
p<-ggplot(data=train_dat)+
  geom_boxplot(aes(x=Microcystis,y=Chlorophyll,group=Microcystis))+scale_y_log10()+
  geom_text(data = counts1,aes(x = Microcystis, y = max(dat1$Chlorophyll, na.rm = TRUE), 
      label = paste0("n=", n)),vjust = -0.5) +
  ggtitle("All sampling dates")
p1<-ggplot(data=train_dat2)+
  geom_boxplot(aes(x=Microcystis,y=Chlorophyll,group=Microcystis))+scale_y_log10()+
  geom_text(data = counts2,aes(x = Microcystis, y = max(dat1$Chlorophyll, na.rm = TRUE), 
        label = paste0("n=", n)),vjust = -0.5) +
  ggtitle("Only blooms")
ggarrange(p,p1,ncol=1)

counts1 <- train_dat %>%group_by(ptox_flag) %>%
  summarise(n = sum(!is.na(Chlorophyll)), .groups = "drop")
counts2 <- train_dat2 %>%group_by(ptox_flag) %>%
  summarise(n = sum(!is.na(Chlorophyll)), .groups = "drop")
p<-ggplot(data=train_dat)+
  geom_boxplot(aes(x=ptox_flag,y=Chlorophyll,group=ptox_flag))+scale_y_log10()+
  geom_text(data = counts1,aes(x = ptox_flag, y = max(dat1$Chlorophyll, na.rm = TRUE), 
      label = paste0("n=", n)),vjust = -0.5) +
  ggtitle("All sampling dates")
p1<-ggplot(data=train_dat2)+
  geom_boxplot(aes(x=ptox_flag,y=Chlorophyll,group=ptox_flag))+scale_y_log10()+
  geom_text(data = counts2,aes(x = ptox_flag, y = max(dat1$Chlorophyll, na.rm = TRUE), 
        label = paste0("n=", n)),vjust = -0.5) +
  ggtitle("Only blooms")
ggarrange(p,p1,ncol=1)

counts_all <- train_dat %>%group_by(Month, Microcystis) %>%summarise(n = n(), .groups = "drop")
counts_bloom <- train_dat2 %>%group_by(Month, Microcystis) %>%summarise(n = n(), .groups = "drop")
p <- ggplot(counts_all) +
  geom_point(aes(x = as.factor(Month),y = Microcystis,color = n),size=9)+
  geom_text(aes(x = as.factor(Month),y = Microcystis,label = n),size = 4,fontface = "bold") +
  scale_color_distiller(palette = "YlOrRd", direction = 1) +
  ggtitle("All sampling dates")
p1 <- ggplot(counts_bloom) +
  geom_point(aes(x = as.factor(Month),y = Microcystis,color = n),size=9)+
  geom_text(aes(x = as.factor(Month),y = Microcystis,label = n),size = 4,fontface = "bold") +
  scale_color_distiller(palette = "YlOrRd", direction = 1) +ggtitle("Only blooms")
ggarrange(p,p1,ncol=1)

counts_all <- train_dat %>%group_by(Month, ptox_flag) %>%summarise(n = n(), .groups = "drop")
counts_bloom <- train_dat2 %>%group_by(Month, ptox_flag) %>%summarise(n = n(), .groups = "drop")
p <- ggplot(counts_all) +
  geom_point(aes(x = as.factor(Month),y = ptox_flag,color = n),size=9)+
  geom_text(aes(x = as.factor(Month),y = ptox_flag,label = n),size = 4,fontface = "bold") +
  scale_color_distiller(palette = "YlOrRd", direction = 1) +
  ggtitle("All sampling dates")
p1 <- ggplot(counts_bloom) +
  geom_point(aes(x = as.factor(Month),y = ptox_flag,color = n),size=9)+
  geom_text(aes(x = as.factor(Month),y = ptox_flag,label = n),size = 4,fontface = "bold") +
  scale_color_distiller(palette = "YlOrRd", direction = 1) +ggtitle("Only blooms")
ggarrange(p,p1,ncol=1)

# Collinearity between temp and ptox
cor(train_dat$Temperature, train_dat$HAB_ptox2, use = "complete.obs")
ggplot(data=train_dat)+geom_point(aes(x=HAB_ptox2,y=Temperature))
# 0.1621026

# Collinearity between chl and ptox
cor(train_dat$Chlorophyll, train_dat$HAB_ptox2, use = "complete.obs")
# 0.0146

# Cross validation
set.seed(123)
ctrl <- trainControl(method = "cv",number = 5,classProbs = TRUE,summaryFunction = twoClassSummary)
train_dat$harmful_factor <- factor(ifelse(train_dat$mvi == 1, "yes", "no"))
train_dat2$harmful_factor <- factor(ifelse(train_dat2$mvi == 1, "yes", "no"))

# Model 1a: 
model <- glmer(mvi ~ ptox * Chlorophyll + Region * Chlorophyll + (1 | Month), 
      family = binomial, data = train_dat2)
summary(model)

cv_model <- train(harmful_factor ~ ptox * Chlorophyll + Region * Chlorophyll + (1 | Month),
  data = train_dat2,method = "glm",family = "binomial",trControl = ctrl,metric = "PR")
mean(cv_model$resample$ROC)
sd(cv_model$resample$ROC)

# Model 2a: 
model <- glmer(mvi ~ ptox + Region * Chlorophyll + (1 | Month), 
      family = binomial, data = train_dat2)
summary(model)

cv_model <- train(harmful_factor ~ ptox + Region * Chlorophyll + (1 | Month),
  data = train_dat2,method = "glm",family = "binomial",trControl = ctrl,metric = "PR")
mean(cv_model$resample$ROC)
sd(cv_model$resample$ROC)

# Model 3a:
model <- glm(mvi ~ ptox + Chlorophyll,data = train_dat2,family = binomial)
summary(model)

cv_model <- train(harmful_factor ~ ptox + Chlorophyll,
  data = train_dat2,method = "glm",family = "binomial",trControl = ctrl,metric = "PR")
mean(cv_model$resample$ROC)
sd(cv_model$resample$ROC)

# Model 4a: 
model <- glmer(mvi ~ ptox + Chlorophyll + (1 | Month), 
      family = binomial, data = train_dat2)
summary(model)

cv_model <- train(harmful_factor ~ ptox + Chlorophyll + (1 | Month),
  data = train_dat2,method = "glm",family = "binomial",trControl = ctrl,metric = "PR")
mean(cv_model$resample$ROC)
sd(cv_model$resample$ROC)

# Model 1b: 
model <- glmer(mvi ~ ptox * Chlorophyll + Region * Chlorophyll + (1 | Month), 
      family = binomial, data = train_dat)
summary(model)

cv_model <- train(harmful_factor ~ ptox * Chlorophyll + Region * Chlorophyll + (1 | Month),
  data = train_dat,method = "glm",family = "binomial",trControl = ctrl,metric = "PR")
mean(cv_model$resample$ROC)
sd(cv_model$resample$ROC)

library(DHARMa)
sim_res <- simulateResiduals(model)
plot(sim_res)

# Model 2b: 
model <- glmer(mvi ~ ptox + Region * Chlorophyll + (1 | Month), 
      family = binomial, data = train_dat)
summary(model)

cv_model <- train(harmful_factor ~ ptox + Region * Chlorophyll + (1 | Month),
  data = train_dat,method = "glm",family = "binomial",trControl = ctrl,metric = "PR")
mean(cv_model$resample$ROC)
sd(cv_model$resample$ROC)

sim_res <- simulateResiduals(model)
plot(sim_res)

# Model 3b:
model <- glm(mvi ~ ptox + Chlorophyll,data = train_dat,family = binomial)
summary(model)

cv_model <- train(harmful_factor ~ ptox + Chlorophyll,
  data = train_dat,method = "glm",family = "binomial",trControl = ctrl,metric = "PR")
mean(cv_model$resample$ROC)
sd(cv_model$resample$ROC)

# Model 4b: 
model <- glmer(mvi ~ ptox + Chlorophyll + (1 | Month), 
      family = binomial, data = train_dat)
summary(model)

cv_model <- train(harmful_factor ~ ptox + Chlorophyll + (1 | Month),
  data = train_dat,method = "glm",family = "binomial",trControl = ctrl,metric = "PR")
mean(cv_model$resample$ROC)
sd(cv_model$resample$ROC)

# Model 5b: 
model <- glmer(mvi ~ ptox + (1 | Month), 
      family = binomial, data = train_dat)
summary(model)

cv_model <- train(harmful_factor ~ ptox + Chlorophyll + (1 | Month),
  data = train_dat,method = "glm",family = "binomial",trControl = ctrl,metric = "PR")
mean(cv_model$resample$ROC)
sd(cv_model$resample$ROC)

# Model 5a: 
model <- glmer(mvi ~ ptox + (1 | Month), 
      family = binomial, data = train_dat2)
summary(model)

cv_model <- train(harmful_factor ~ ptox + Chlorophyll + (1 | Month),
  data = train_dat,method = "glm",family = "binomial",trControl = ctrl,metric = "PR")
mean(cv_model$resample$ROC)
sd(cv_model$resample$ROC)
```

