---
title: "HABs vs BABs"
author: "Schuyler Nardelli"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prep Workspace
```{r, warning=FALSE, message=FALSE}
# Clear environment, close any open plots
rm(list = ls()); graphics.off()
  
# Load in libraries
library(tidyverse)
library(readxl)
library(sf)
library("ggmap")
library(ggpubr)
library(EDIutils)

# Load PESP data
packageId <-"edi.2209.2"
read_data_package_archive(packageId, path = tempdir()) 
res <- read_data_entity_names(packageId)

entityName <- "PESP_enumeration.csv"
entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
pesp <- readr::read_csv(file = raw)

pesp$Date <- as.Date(pesp$Date)
pesp <- pesp %>% arrange(Date, Time)
rm(res,entityId,entityName,packageId,raw)


# Load Discrete WQ data package (with MVI)
packageId <-"edi.731.7"
read_data_package_archive(packageId, path = tempdir()) 
res <- read_data_entity_names(packageId)

entityName <- "Delta_Integrated_WQ.csv"
entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
wq <- readr::read_csv(file = raw)
rm(entityId,entityName,raw)

entityName <- "Delta_Integrated_WQ_metadata.csv"
entityId <- res$entityId[res$entityName == entityName]
raw <- read_data_entity(packageId, entityId)
wq_meta <- readr::read_csv(file = raw)
rm(res,entityId,entityName,packageId,raw)

wq$Date <- as.Date(wq$Date)
wq <- wq %>% separate(Station, into = c("Survey", "Station"), sep = " ")
wq <- wq[, -c(2,4:5)]
wq <- wq[,c("Source","Station","Date","Datetime","Sample_depth_surface","Tide",
           "Temperature","Conductivity","Salinity","Secchi","TurbidityNTU","TurbidityFNU",
           "DissolvedOxygen","DissolvedOxygenPercent","pH","Microcystis","Chlorophyll","Pheophytin",
           "TotAmmonia","DissAmmonia","DissNitrateNitrite","TotPhos","DissOrthophos","TON","DON",
           "TKN","TotAlkalinity","DissBromide","DissCalcium","TotChloride","DissChloride","DissSilica",
           "TOC","DOC","TDS","TSS","VSS")]

# Load EMP benthic data
clam <- read.csv("Data/combined_clams_dataset.csv")
clam$station_name <- gsub("EMP ", "", clam$station_name)
clam$Date <- as.Date(clam$Date)

# Load Dayflow data
df <- read.csv("Data/Dayflow/dayflow-results-1997-2023.csv")
df$Date <- as.Date(df$Date,format = "%m/%d/%Y")
ind <- read.csv("Data/Dayflow/wy_class_ind.csv")
df$WY <- ifelse(df$Month >= 10, df$Year + 1, df$Year)
df <- merge(df,ind,by="WY",all.x=TRUE)

```


## Merge PESP and discrete datasets
Only include samples associated with PESP samples. Will want fields for date, station, region, chl, other discrete properties.
```{r, warning=FALSE, message=FALSE}

# Merge wq and df
x <- df[,c("Date","SAC","OUT","Index")]
wq_df <- merge(wq, x, by = "Date", all.x = TRUE)

# Merge with benthic data
x <- clam[,c("Date","station_name","Corbicula_biomass","Potamocorbula_biomass","Corbicula_GR","Potamocorbula_GR","Corbicula_FR","Potamocorbula_FR")]
colnames(x)[colnames(x) == "station_name"] <- "Station"
wq_df_clam <- left_join(wq_df, x, by = c("Date","Station"))

# Merge with PESP data
dat <- left_join(pesp, wq_df_clam, by = c("Date","Station"))

# Add season info
dat$Month <- month(dat$Date)
dat$Year <- year(dat$Date)
dat$Season<- NA
I = which(dat$Month %in% c(7:9)); dat$Season[I]<-"Summer"
I = which(dat$Month %in% c(10:12)); dat$Season[I]<-"Fall"
I = which(dat$Month %in% c(1:3)); dat$Season[I]<-"Winter"
I = which(dat$Month %in% c(4:6)); dat$Season[I]<-"Spring"

rm(x,wq_df_clam,wq_df, clam, df, ind,wq,pesp,wq_meta, I)

```


## Add in regional designation
```{r, warning=FALSE, message=FALSE}

points_df <- dat[,c("Station","Survey","Latitude","Longitude")]
points_df <- unique(points_df)
points_df <- points_df %>% filter(!is.na(Latitude))
points_df$id <- 1:length(points_df$Survey)

regions <- st_read('Data/Regions_shp/Rosies_regions_edited.shp')

points_sf = points_df %>% st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326)

points_with_region <- st_join(points_sf, regions, join = st_within) %>%  
  st_drop_geometry() 

stat <- left_join(points_sf,points_with_region,by=c("Station","Survey","id"))
coord <- data.frame(st_coordinates(stat$geometry))
stat$Latitude <- coord$Y; stat$Longitude <- coord$X; rm(coord)
stat <- stat %>% st_drop_geometry() %>% select(-"id")
dat <- left_join(dat,stat,by=c("Station","Survey","Latitude","Longitude"))

rm(stat,points_df,points_sf,points_with_region,regions)

```


## Filter dataset
```{r, warning=FALSE, message=FALSE}

dat1 <- dat %>% filter(!is.na(Regions) &   # Exclude regions outside Suisun Bay and Delta
                        #Date > "2007-12-31" &    # Exclude data before 2008
                        Survey %in% c("DWR-EMP","USGS-BP"))   # Exclude surveys besides EMP and BP

dat1 <- dat1[!is.na(dat1$Source), ] # Screens out PESP samples without corresponding WQ data
dat1 <- dat1 %>% arrange(Date, Time)

rm(dat)

```


## Data coverage
```{r, warning=FALSE, message=FALSE}

ggplot(data=dat1,aes(x=Date,y=Biovolume_per_mL))+geom_point()+facet_wrap(~Regions)+
  ylab("Biovolume (ug/mL)")
ggplot(data=dat1,aes(x=Date,y=Cells_per_mL))+geom_point()+facet_wrap(~Regions)


# Regions
deltabox <- make_bbox(lat = Latitude, lon = Longitude, data = dat1, f = c(.2, .2)) #sets bounding box based on station coordinates, f value adds a buffer around the perimeter
register_stadiamaps(key = "b4a4b8da-f8e2-4ba1-bcdc-91e9d5c83d00")
stamenmap <- get_stadiamap(deltabox, zoom = 10, maptype = "alidade_smooth", force = TRUE, crop = TRUE)

ggmap(stamenmap) +
  geom_point(data = dat1, 
             mapping = aes(x = Longitude, y = Latitude,fill=Regions),size=3,pch=21) +
  #geom_text(data = dat, aes(x = Longitude, y = Latitude,label = Station),fontface = #"bold",cex=3,vjust=0.5,hjust=1.3)+
  labs(title = "", x = "Longitude", y = "Latitude") +
  theme_light(base_size = 12) +
  theme(panel.grid = element_blank(),
        strip.background = element_rect(fill = "black"),
        strip.text = element_text(size = 10),axis.text.x = element_text(angle = 90))+
  xlab("")+ylab("")+theme_light(base_size = 16) 

# Number of samples per region per season over time
dat1 %>% group_by(Regions, Year, Season)%>%summarize(n=n()) %>%
  ggplot(aes(x=Year,y=Regions))+geom_tile(aes(fill=n))+
  scale_fill_gradient(low="white",high="red")+
  geom_text(aes(label=round(n,1)),size=4)+
  xlab("")+ylab("")+
  facet_grid(Season~.)+
  theme_bw(base_size=14)+
  theme(legend.position = "none")

# Number of samples per region per month over time
dat1 %>% group_by(Regions, Year, Month)%>%summarize(n=n()) %>%
  ggplot(aes(x=Year,y=Regions))+geom_tile(aes(fill=n))+
  scale_fill_gradient(low="white",high="red")+
  geom_text(aes(label=round(n,1)),size=4)+
  xlab("")+ylab("")+
  facet_grid(Month~.)+
  theme_bw(base_size=14)+
  theme(legend.position = "none")

# Number of samples per station and season over time
dat1 %>% group_by(Station, Regions,Season,Year)%>%summarize(n=n()) %>%
  ggplot(aes(x=Year,y=as.factor(Station)))+geom_tile(aes(fill=n))+
  scale_fill_gradient(low="white",high="red")+
  geom_text(aes(label=round(n,1)),size=4)+
  xlab("")+ylab("")+facet_grid(Regions~Season,scales = 'free_y',space='free')+
  theme_bw(base_size=14)+
  theme(legend.position = "none")+
  theme(axis.text.x=element_text(angle = -90, vjust = 0.3))

rm(deltabox,stamenmap)
```


## Flag "bloom" or "no bloom" for each sample using chl threshold
Considered "bloom" if Chla is 1 standard deviation above the mean value for the entire time series
```{r, warning=FALSE, message=FALSE}

dat1$bloom <- NA

# Suisun Marsh
test <- dat1 %>% filter(Regions=="Suisun Marsh") %>% group_by(Date) %>% summarize(Chl=mean(Chlorophyll,na.rm=TRUE))
x = mean(test$Chl,na.rm=TRUE) + sd(test$Chl,na.rm=TRUE)
I <- which(dat1$Regions == "Suisun Marsh" & dat1$Chlorophyll > x); dat1$bloom[I] <- "Yes"; rm(I)
I <- which(dat1$Regions == "Suisun Marsh" & dat1$Chlorophyll <= x); dat1$bloom[I] <- "No"; rm(I,test,x)

p<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Suisun Marsh"),aes(x=Date,y=Chlorophyll),color="black")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Marsh" & bloom=="Yes"), aes(x = Date, y = Chlorophyll),color="blue")+ggtitle("Suisun Marsh")

# Suisun Bay
test <- dat1 %>% filter(Regions=="Suisun Bay") %>% group_by(Date) %>% summarize(Chl=mean(Chlorophyll,na.rm=TRUE))
x = mean(test$Chl,na.rm=TRUE) + sd(test$Chl,na.rm=TRUE)
I <- which(dat1$Regions == "Suisun Bay" & dat1$Chlorophyll > x); dat1$bloom[I] <- "Yes"; rm(I)
I <- which(dat1$Regions == "Suisun Bay" & dat1$Chlorophyll <= x); dat1$bloom[I] <- "No"; rm(I,x)

p1<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Suisun Bay"),aes(x=Date,y=Chlorophyll),color="black")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Bay" & bloom=="Yes"), aes(x = Date, y = Chlorophyll),color="blue")+ggtitle("Suisun Bay")

# Confluence
test <- dat1 %>% filter(Regions=="Confluence") %>% group_by(Date) %>% summarize(Chl=mean(Chlorophyll,na.rm=TRUE))
x = mean(test$Chl,na.rm=TRUE) + sd(test$Chl,na.rm=TRUE)
I <- which(dat1$Regions == "Confluence" & dat1$Chlorophyll > x); dat1$bloom[I] <- "Yes"; rm(I)
I <- which(dat1$Regions == "Confluence" & dat1$Chlorophyll <= x); dat1$bloom[I] <- "No"; rm(I,test,x)

p2<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Confluence"),aes(x=Date,y=Chlorophyll),color="black")+
  geom_point(data = dat1 %>% filter(Regions=="Confluence" & bloom=="Yes"), aes(x = Date, y = Chlorophyll),color="blue")+ggtitle("Confluence")

# North Delta
test <- dat1 %>% filter(Regions=="North Delta") %>% group_by(Date) %>% summarize(Chl=mean(Chlorophyll,na.rm=TRUE))
x = mean(test$Chl,na.rm=TRUE) + sd(test$Chl,na.rm=TRUE)
I <- which(dat1$Regions == "North Delta" & dat1$Chlorophyll > x); dat1$bloom[I] <- "Yes"; rm(I)
I <- which(dat1$Regions == "North Delta" & dat1$Chlorophyll <= x); dat1$bloom[I] <- "No"; rm(I,test,x)

p3<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="North Delta"),aes(x=Date,y=Chlorophyll),color="black")+
  geom_point(data = dat1 %>% filter(Regions=="North Delta" & bloom=="Yes"), aes(x = Date, y = Chlorophyll),color="blue")+ggtitle("North Delta")

# Central Delta
test <- dat1 %>% filter(Regions=="Central") %>% group_by(Date) %>% summarize(Chl=mean(Chlorophyll,na.rm=TRUE))
x = mean(test$Chl,na.rm=TRUE) + sd(test$Chl,na.rm=TRUE)
I <- which(dat1$Regions == "Central" & dat1$Chlorophyll > x); dat1$bloom[I] <- "Yes"; rm(I)
I <- which(dat1$Regions == "Central" & dat1$Chlorophyll <= x); dat1$bloom[I] <- "No"; rm(I,test,x)

p4<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Central"),aes(x=Date,y=Chlorophyll),color="black")+
  geom_point(data = dat1 %>% filter(Regions=="Central" & bloom=="Yes"), aes(x = Date, y = Chlorophyll),color="blue")+ggtitle("Central")

# South Delta
test <- dat1 %>% filter(Regions=="South") %>% group_by(Date) %>% summarize(Chl=mean(Chlorophyll,na.rm=TRUE))
x = mean(test$Chl,na.rm=TRUE) + sd(test$Chl,na.rm=TRUE)
I <- which(dat1$Regions == "South" & dat1$Chlorophyll > x); dat1$bloom[I] <- "Yes"; rm(I)
I <- which(dat1$Regions == "South" & dat1$Chlorophyll <= x); dat1$bloom[I] <- "No"; rm(I,test,x)

p5<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="South"),aes(x=Date,y=Chlorophyll),color="black")+
  geom_point(data = dat1 %>% filter(Regions=="South" & bloom=="Yes"), aes(x = Date, y = Chlorophyll),color="blue")+  ggtitle("South")

#jpeg("Figures/Bloom detection/habbab_dens.jpg", units="in", width=16, height=9, res=300)
ggarrange(p,p1,p2,p3,p4,p5,ncol=2,nrow=3)
#dev.off()

rm(p,p1,p2,p3,p4,p5)

```

## Method 1: Define HAB v BAB using MVI
Considered a BAB if MVI 1-2 and a HAB if MVI 3-5
```{r, warning=FALSE, message=FALSE}

# Define HAB as MVI 3-5
dat1$HAB_MVI <- NA
I <- which(dat1$Microcystis %in% c(3:5) & dat1$Month %in% c(6:10)); dat1$HAB_MVI[I] <- "HAB"; rm(I)
I <- which(dat1$Microcystis %in% c(3:5) & dat1$Month %in% c(1:5,11:12)); dat1$HAB_MVI[I] <- "BAB"; rm(I)
I <- which(dat1$Microcystis %in% c(1:2)); dat1$HAB_MVI[I] <- "BAB"; rm(I)

# Plot
p<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Suisun Marsh"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Marsh")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Marsh"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),pch=21,color="gray")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Marsh"&bloom=="Yes"&HAB_MVI=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="green")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Marsh"&bloom=="Yes"&HAB_MVI=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p1<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Suisun Bay"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Bay")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Bay"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="gray")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Bay"&bloom=="Yes"&HAB_MVI=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Bay"&bloom=="Yes"&HAB_MVI=="HAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+ylab("")+xlab("")
p2<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Confluence"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Confluence")+
  geom_point(data = dat1 %>% filter(Regions=="Confluence"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="gray")+
  geom_point(data = dat1 %>% filter(Regions=="Confluence"&bloom=="Yes"&HAB_MVI=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Regions=="Confluence"&bloom=="Yes"&HAB_MVI=="HAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+ylab("Chlorophyll (µg/L)")+xlab("")
p3<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="North Delta"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("North Delta")+
  geom_point(data = dat1 %>% filter(Regions=="North Delta"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Regions=="North Delta"&bloom=="Yes"&HAB_MVI=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="North Delta"&bloom=="Yes"&HAB_MVI=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p4<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Central"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Central")+
  geom_point(data = dat1 %>% filter(Regions=="Central"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="gray")+
  geom_point(data = dat1 %>% filter(Regions=="Central"&bloom=="Yes"&HAB_MVI=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Regions=="Central"&bloom=="Yes"&HAB_MVI=="HAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+ylab("")+xlab("")
p5<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="South"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("South")+
  geom_point(data = dat1 %>% filter(Regions=="South"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Regions=="South"&bloom=="Yes"&HAB_MVI=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="South"&bloom=="Yes"&HAB_MVI=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
#ggarrange(p,p1,p2,p3,p4,p5,ncol=2,nrow=3)

tiff("../phyto_synthesis/Figures/HABvBAB_MVI.jpeg", units="in", width=5, height=6, res=300)
ggarrange(p1,p2,p4,ncol=1)
dev.off()

rm(p,p1,p2,p3,p4,p5)

```

## Method 2a: Define HAB v BAB using cyanobacterial density from WHO (2006)
Density thresholds from World Health Organization guidance levels, found in Kudela et al. (2023).
Could classify "HAB" as medium and high risk, or just high risk.
```{r, warning=FALSE, message=FALSE}

# Aggregate by group
grp <- dat1 %>% group_by(Date,Survey,Regions,Station,AlgalGroup) %>%
  summarize(Density = sum(Cells_per_mL),Biovolume = sum(Biovolume_per_mL))
tot <- dat1 %>% group_by(Date,Survey,Regions,Station) %>% summarise(Tot_Density = sum(Cells_per_mL),Tot_Biovolume = sum(Biovolume_per_mL))
grp <- grp %>% left_join(tot,by=c("Date","Survey","Regions","Station"))

c <- grp %>% filter(AlgalGroup == "Cyanobacteria")

# Define high, medium, and low risk WHO density thresholds
c$HAB_cyano <- NA
I <- which(c$Density <20000); c$HAB_cyano[I] <- "Low"; rm(I)
I <- which(c$Density >=20000 & c$Density <100000); c$HAB_cyano[I] <- "Med"; rm(I)
I <- which(c$Density >=100000); c$HAB_cyano[I] <- "High"; rm(I)
c <- c[,c(1:4,6,10)]
c <- c %>% rename(Density_cyano = Density)

dat1 <- left_join(dat1, c, by = c("Date","Survey","Regions","Station"))

# Plot
p<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Suisun Marsh"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Marsh")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Marsh"&bloom=="Yes"&HAB_cyano=="Low"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Marsh"&bloom=="Yes"&HAB_cyano=="Med"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Marsh"&bloom=="Yes"&HAB_cyano=="High"),aes(x=Date,y=Chlorophyll),color="red")
p1<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Suisun Bay"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Bay")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Bay"&bloom=="Yes"&HAB_cyano=="Low"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Bay"&bloom=="Yes"&HAB_cyano=="Med"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Bay"&bloom=="Yes"&HAB_cyano=="High"),aes(x=Date,y=Chlorophyll),color="red")
p2<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Confluence"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Confluence")+
  geom_point(data = dat1 %>% filter(Regions=="Confluence"&bloom=="Yes"&HAB_cyano=="Low"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="Confluence"&bloom=="Yes"&HAB_cyano=="Med"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Regions=="Confluence"&bloom=="Yes"&HAB_cyano=="High"),aes(x=Date,y=Chlorophyll),color="red")
p3<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="North Delta"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("North Delta")+
  geom_point(data = dat1 %>% filter(Regions=="North Delta"&bloom=="Yes"&HAB_cyano=="Low"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="North Delta"&bloom=="Yes"&HAB_cyano=="Med"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Regions=="North Delta"&bloom=="Yes"&HAB_cyano=="High"),aes(x=Date,y=Chlorophyll),color="red")
p4<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Central"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Central")+
  geom_point(data = dat1 %>% filter(Regions=="Central"&bloom=="Yes"&HAB_cyano=="Low"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="Central"&bloom=="Yes"&HAB_cyano=="Med"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Regions=="Central"&bloom=="Yes"&HAB_cyano=="High"),aes(x=Date,y=Chlorophyll),color="red")
p5<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="South"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("South")+
  geom_point(data = dat1 %>% filter(Regions=="South"&bloom=="Yes"&HAB_cyano=="Low"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="South"&bloom=="Yes"&HAB_cyano=="Med"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Regions=="South"&bloom=="Yes"&HAB_cyano=="High"),aes(x=Date,y=Chlorophyll),color="red")
#ggarrange(p,p1,p2,p3,p4,p5,ncol=2,nrow=3)
ggarrange(p1,p2,p4,ncol=1)

ggplot() + geom_histogram(data=c, aes(x=Density_cyano))

rm(p,p1,p2,p3,p4,p5,c,grp,tot)

```

## Method 2b: Define HAB v BAB using cyanobacterial biovolume from WHO (2021)
Biovolume thresholds from World Health Organization guidance levels, 2021.
Could classify "HAB" as alert levels 1 and 2, or just alert level 2.
```{r, warning=FALSE, message=FALSE}

# Aggregate by group
grp <- dat1 %>% group_by(Date,Survey,Regions,Station,AlgalGroup) %>%
  summarize(Density = sum(Cells_per_mL),Biovolume = sum(Biovolume_per_mL))
tot <- dat1 %>% group_by(Date,Survey,Regions,Station) %>% summarise(Tot_Density = sum(Cells_per_mL),Tot_Biovolume = sum(Biovolume_per_mL))
grp <- grp %>% left_join(tot,by=c("Date","Survey","Regions","Station"))

c <- grp %>% filter(AlgalGroup == "Cyanobacteria")

# Define alert level 1 and 2 WHO biovolume thresholds
c$HAB_cyano2 <- NA
I <- which(c$Biovolume <4e+6); c$HAB_cyano2[I] <- "BAB"; rm(I)
I <- which(c$Biovolume >=4e+6 & c$Biovolume <8e+6); c$HAB_cyano2[I] <- "AlertL1"; rm(I)
I <- which(c$Biovolume >=8e+6); c$HAB_cyano2[I] <- "AlertL2"; rm(I)
c <- c[,c(1:4,7,10)]
c <- c %>% rename(Biovolume_cyano = Biovolume)

dat1 <- left_join(dat1, c, by = c("Date","Survey","Regions","Station"))

# Plot
p<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Suisun Marsh"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Marsh")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Marsh"&bloom=="Yes"&HAB_cyano2=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Marsh"&bloom=="Yes"&HAB_cyano2=="AlertL1"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Marsh"&bloom=="Yes"&HAB_cyano2=="AlertL2"),aes(x=Date,y=Chlorophyll),color="red")
p1<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Suisun Bay"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Bay")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Bay"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="gray")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Bay"&bloom=="Yes"&HAB_cyano2=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Bay"&bloom=="Yes"&HAB_cyano2=="AlertL1"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="orange")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Bay"&bloom=="Yes"&HAB_cyano2=="AlertL2"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+ylab("")+xlab("")
p2<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Confluence"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Confluence")+
  geom_point(data = dat1 %>% filter(Regions=="Confluence"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="gray")+
  geom_point(data = dat1 %>% filter(Regions=="Confluence"&bloom=="Yes"&HAB_cyano2=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Regions=="Confluence"&bloom=="Yes"&HAB_cyano2=="AlertL1"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="orange")+
  geom_point(data = dat1 %>% filter(Regions=="Confluence"&bloom=="Yes"&HAB_cyano2=="AlertL2"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+ylab("Chlorophyll (µg/L)")+xlab("")
p3<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="North Delta"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("North Delta")+
  geom_point(data = dat1 %>% filter(Regions=="North Delta"&bloom=="Yes"&HAB_cyano2=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="North Delta"&bloom=="Yes"&HAB_cyano2=="AlertL1"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Regions=="North Delta"&bloom=="Yes"&HAB_cyano2=="AlertL2"),aes(x=Date,y=Chlorophyll),color="red")
p4<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Central"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Central")+
  geom_point(data = dat1 %>% filter(Regions=="Central"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="gray")+
  geom_point(data = dat1 %>% filter(Regions=="Central"&bloom=="Yes"&HAB_cyano2=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Regions=="Central"&bloom=="Yes"&HAB_cyano2=="AlertL1"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="orange")+
  geom_point(data = dat1 %>% filter(Regions=="Central"&bloom=="Yes"&HAB_cyano2=="AlertL2"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+ylab("")+xlab("")
p5<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="South"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("South")+
  geom_point(data = dat1 %>% filter(Regions=="South"&bloom=="Yes"&HAB_cyano2=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="South"&bloom=="Yes"&HAB_cyano2=="AlertL1"),aes(x=Date,y=Chlorophyll),color="orange")+
  geom_point(data = dat1 %>% filter(Regions=="South"&bloom=="Yes"&HAB_cyano2=="AlertL2"),aes(x=Date,y=Chlorophyll),color="red")
#ggarrange(p,p1,p2,p3,p4,p5,ncol=2,nrow=3)

tiff("../phyto_synthesis/Figures/HABvBAB_WHO.jpeg", units="in", width=5, height=6, res=300)
ggarrange(p1,p2,p4,ncol=1)
dev.off()



rm(p,p1,p2,p3,p4,p5,c,grp,tot)

```

## Method 3: Define HAB v BAB using cyanobacterial relative abundance
Define HAB as a cyanobacterial relative biovolume > 50%
```{r, warning=FALSE, message=FALSE}

# Aggregate by group
grp <- dat1 %>% group_by(Date,Survey,Regions,Station,AlgalGroup) %>%
  summarize(Density = sum(Cells_per_mL),Biovolume = sum(Biovolume_per_mL))
tot <- dat1 %>% group_by(Date,Survey,Regions,Station) %>% summarise(Tot_Density = sum(Cells_per_mL),Tot_Biovolume = sum(Biovolume_per_mL))
grp <- grp %>% left_join(tot,by=c("Date","Survey","Regions","Station"))
grp$Rel_Biovolume <- (grp$Biovolume/grp$Tot_Biovolume)*100;
grp$Rel_Density <- (grp$Density/grp$Tot_Density)*100;

c <- grp %>% filter(AlgalGroup == "Cyanobacteria")

# Define HAB as a cyanobacterial relative biovolume > 50%
x = mean(c$Rel_Biovolume,na.rm=TRUE) + sd(c$Rel_Biovolume,na.rm=TRUE)
c$HAB_cyano_ra <- NA
I <- which(c$Rel_Biovolume <x); c$HAB_cyano_ra[I] <- "BAB"; rm(I)
I <- which(c$Rel_Biovolume >=x); c$HAB_cyano_ra[I] <- "HAB"; rm(I)
c <- c[,c(1:4,10,12)]
c <- c %>% rename(Rel_Biovolume_cyano = Rel_Biovolume)

dat1 <- left_join(dat1, c, by = c("Date","Survey","Regions","Station"))

# Plot
p<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Suisun Marsh"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Marsh")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Marsh"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Marsh"&bloom=="Yes"&HAB_cyano_ra=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Marsh"&bloom=="Yes"&HAB_cyano_ra=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p1<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Suisun Bay"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Bay")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Bay"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Bay"&bloom=="Yes"&HAB_cyano_ra=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Bay"&bloom=="Yes"&HAB_cyano_ra=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p2<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Confluence"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Confluence")+
  geom_point(data = dat1 %>% filter(Regions=="Confluence"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Regions=="Confluence"&bloom=="Yes"&HAB_cyano_ra=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="Confluence"&bloom=="Yes"&HAB_cyano_ra=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p3<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="North Delta"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("North Delta")+
  geom_point(data = dat1 %>% filter(Regions=="North Delta"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Regions=="North Delta"&bloom=="Yes"&HAB_cyano_ra=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="North Delta"&bloom=="Yes"&HAB_cyano_ra=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p4<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Central"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Central")+
  geom_point(data = dat1 %>% filter(Regions=="Central"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Regions=="Central"&bloom=="Yes"&HAB_cyano_ra=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="Central"&bloom=="Yes"&HAB_cyano_ra=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p5<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="South"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("South")+
  geom_point(data = dat1 %>% filter(Regions=="South"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Regions=="South"&bloom=="Yes"&HAB_cyano_ra=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="South"&bloom=="Yes"&HAB_cyano_ra=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
#ggarrange(p,p1,p2,p3,p4,p5,ncol=2,nrow=3)
ggarrange(p1,p2,p4,ncol=1)

ggplot() + geom_histogram(data=c, aes(x=Rel_Biovolume_cyano))

rm(c,grp,tot,p,p1,p2,p3,p4,p5,x)

```

## Method 4: Define HAB v BAB using presence of ptox species
```{r, warning=FALSE, message=FALSE}

# Flag ptox species
dat1$ptox_flag <- NA
I <- which(dat1$Genus %in% c("Microcoleus","Cuspidothrix","Aphanizomenon",
                             "Microcystis","Planktothrix") &
             !dat1$Taxon == "Aphanizomenon flos-aquae" | 
             dat1$Taxon %in% c("Dolichospermum circinale","Microseira wollei"))
dat1$ptox_flag[I]<- "Yes"; 
dat1$ptox_flag[is.na(dat1$ptox_flag)] <- "No"


# Aggregate by ptox presence/absence
#dat1$ptox_flag <- factor(dat1$ptox_flag)
grp <- dat1 %>% group_by(Date,Month,Survey,Regions,Station,ptox_flag) %>% #,.drop = FALSE) %>%
  summarize(Density = sum(Cells_per_mL),Biovolume = sum(Biovolume_per_mL))
tot <- dat1 %>% group_by(Date,Month,Survey,Regions,Station) %>% summarise(Tot_Density = sum(Cells_per_mL),Tot_Biovolume = sum(Biovolume_per_mL))
grp <- grp %>% left_join(tot,by=c("Date","Month","Survey","Regions","Station"))
grp$Rel_Biovolume <- (grp$Biovolume/grp$Tot_Biovolume)*100;
grp$Rel_Density <- (grp$Density/grp$Tot_Density)*100;

c <- grp %>% filter(ptox_flag == "Yes")

# Define HAB when ptox > 0
c$HAB_ptox <- NA
I <- which(c$Month %in% c(6:10)); c$HAB_ptox[I] <- "HAB"; rm(I)
I <- which(c$Month %in% c(1:5,11:12)); c$HAB_ptox[I] <- "BAB"; rm(I)
c <- c[,c(1:5,11,13)]
c <- c %>% rename(Rel_Biovolume_ptox = Rel_Biovolume)

dat1 <- left_join(dat1, c, by = c("Date","Month","Survey","Regions","Station"))
dat1$HAB_ptox[is.na(dat1$HAB_ptox)] <- "BAB"
I = which(!is.na(dat1$Biovolume_per_mL) & is.na(dat1$Rel_Biovolume_ptox)) 
dat1$Rel_Biovolume_ptox[I]<- 0


# Plot
p<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Suisun Marsh"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Marsh")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Marsh"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Marsh"&bloom=="Yes"&HAB_ptox=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Marsh"&bloom=="Yes"&HAB_ptox=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p1<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Suisun Bay"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Bay")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Bay"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Bay"&bloom=="Yes"&HAB_ptox=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Regions=="Suisun Bay"&bloom=="Yes"&HAB_ptox=="HAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+xlab("")+ylab("")
p2<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Confluence"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Confluence")+
  geom_point(data = dat1 %>% filter(Regions=="Confluence"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Regions=="Confluence"&bloom=="Yes"&HAB_ptox=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Regions=="Confluence"&bloom=="Yes"&HAB_ptox=="HAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+xlab("")+ylab("Chlorophyll (µg/L)")
p3<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="North Delta"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("North Delta")+
  geom_point(data = dat1 %>% filter(Regions=="North Delta"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Regions=="North Delta"&bloom=="Yes"&HAB_ptox=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="North Delta"&bloom=="Yes"&HAB_ptox=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
p4<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="Central"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Central")+
  geom_point(data = dat1 %>% filter(Regions=="Central"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Regions=="Central"&bloom=="Yes"&HAB_ptox=="BAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="blue")+
  geom_point(data = dat1 %>% filter(Regions=="Central"&bloom=="Yes"&HAB_ptox=="HAB"),aes(x=Date,y=Chlorophyll),size=2,pch=21,fill="red")+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+xlab("")+ylab("")
p5<-ggplot()+geom_path(data = dat1 %>% filter(Regions=="South"),aes(x=Date,y=Chlorophyll),color="black")+ggtitle("South")+
  geom_point(data = dat1 %>% filter(Regions=="South"&bloom=="Yes"),aes(x=Date,y=Chlorophyll),color="gray")+
  geom_point(data = dat1 %>% filter(Regions=="South"&bloom=="Yes"&HAB_ptox=="BAB"),aes(x=Date,y=Chlorophyll),color="green")+
  geom_point(data = dat1 %>% filter(Regions=="South"&bloom=="Yes"&HAB_ptox=="HAB"),aes(x=Date,y=Chlorophyll),color="red")
#ggarrange(p,p1,p2,p3,p4,p5,ncol=2,nrow=3)

tiff("../phyto_synthesis/Figures/HABvBAB_ptox.jpeg", units="in", width=5, height=6, res=300)
ggarrange(p1,p2,p4,ncol=1)
dev.off()

```


## Compare ptox to MVI
```{r, warning=FALSE, message=FALSE}

ggplot(data=dat1 %>% filter(bloom == "Yes" & Microcystis > 1))+geom_point(aes(x=Microcystis,y=Rel_Biovolume_ptox))

test <- dat1 %>% select(Date,Regions,Station,bloom,Microcystis,Rel_Biovolume_ptox,HAB_MVI,HAB_ptox)
test<-unique(test)
test <- test %>% filter(!is.na(HAB_MVI) & bloom == "Yes")
test$compare <- ifelse(test$HAB_MVI == test$HAB_ptox, "same", "different")
# 88 blooms with MVI and ptox

test1 <- test %>% filter(compare == "same") 
# 75 blooms that classify differently, 13 blooms that classify differently

test1 <- test %>% filter(compare == "different" & HAB_MVI == "HAB")
# Of those 13 differences, 
  # 2 instances where MVI classifies BAB and ptox classifies HAB (MVI = 2 for both)
  # 11 instances where MVI classifies HAB and ptox classifies BAB (MVI = 3 for 6 of those instances and 4 for 5 of those instances)

test <- dat1 %>% select(Date,Regions,Station,bloom,Microcystis,Rel_Biovolume_ptox,HAB_MVI,HAB_ptox)
test<-unique(test)
test1 <- test %>% filter(bloom == "Yes" & HAB_ptox == "HAB")

ggplot(data=test1)+geom_histogram(aes(Rel_Biovolume_ptox))


```


## Compare agreement between methods
```{r, warning=FALSE, message=FALSE}

# Method 1: MVI
test <- dat1 %>% group_by(Date,Station,bloom,Microcystis,HAB_MVI) %>%
  summarize(Chl = mean(Chlorophyll)) %>%
  filter(bloom == "Yes" & Microcystis > 0 & HAB_MVI == "BAB")

#185 blooms after 2007
#88 blooms with corresponding MVI data
#19 HABs, 69 BABs

# Method 2: Cyanobacterial density thresholds
test <- dat1 %>% group_by(Date,Station,bloom,Density_cyano,HAB_cyano) %>%
  summarize(Chl = mean(Chlorophyll)) %>%
  filter(bloom == "Yes" & Density_cyano > 0 & HAB_cyano %in% c("Med","High"))

#185 blooms after 2007
#162 blooms with corresponding density data when cyanobacteria were present
#31 blooms where thresholds were "Low" risk
#131 blooms where thresholds were "Medium" or "High" risk
#88 blooms where thresholds were "High" risk

# Method 3: Cyanobacterial relative biovolume > 50%
test <- dat1 %>% group_by(Date,Station,bloom,Rel_Biovolume_cyano,HAB_cyano_ra) %>%
  summarize(Chl = mean(Chlorophyll)) %>%
  filter(bloom == "Yes" & Rel_Biovolume_cyano > 0 & HAB_cyano_ra == "BAB")

#185 blooms after 2007
#122 blooms with corresponding biovolume data when cyanobacteria were present
#18 HABs, 104 BABs


# Plots
test <-data.frame(MVI = dat1$HAB_MVI, Density = dat1$HAB_cyano, RelBio = dat1$HAB_cyano_ra)
test[is.na(test)] <- "No data"

test$MVI <- factor(test$MVI,levels=c("No data","BAB","HAB"))
test$Density <- factor(test$Density, levels=c("No data","Low","Med","High"))
test$RelBio <- factor(test$RelBio, levels=c("No data","BAB","HAB"))

test %>% group_by(MVI, Density)%>%summarize(n=n()) %>%
  ggplot(aes(x=MVI,y=Density))+geom_tile(aes(fill=n))+
  scale_fill_gradient(low="white",high="red")+
  geom_text(aes(label=round(n,1)),size=4)+
  xlab("Method 1: MVI")+ylab("Method 2: Density Threshold")+
  theme_bw(base_size=14)+
  theme(legend.position = "none")

test %>% group_by(MVI, RelBio)%>%summarize(n=n()) %>%
  ggplot(aes(x=MVI,y=RelBio))+geom_tile(aes(fill=n))+
  scale_fill_gradient(low="white",high="red")+
  geom_text(aes(label=round(n,1)),size=4)+
  xlab("Method 1: MVI")+ylab("Method 3: Relative Biovolume > 50%")+
  theme_bw(base_size=14)+
  theme(legend.position = "none")

test %>% group_by(Density, RelBio)%>%summarize(n=n()) %>%
  ggplot(aes(x=Density,y=RelBio))+geom_tile(aes(fill=n))+
  scale_fill_gradient(low="white",high="red")+
  geom_text(aes(label=round(n,1)),size=4)+
  xlab("Method 2: Density Threshold")+ylab("Method 3: Relative Biovolume > 50%")+
  theme_bw(base_size=14)+
  theme(legend.position = "none")

```
