---
title: "Classifying HABs, BABs, and no blooms in 2008-2024 dataset - EMP AND PETERSON DATA"
author: "Schuyler Nardelli"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data
Load in PESP, WQ, dayflow, and clam data. For WQ and clam data, make sure there is only one sample associated with each date and station.
```{r, warning=FALSE, message=FALSE}

# Clear environment and plots
rm(list = ls()); graphics.off()

# Load libraries
library(tidyverse)
library(readxl)
library(sf)
library(DBI)
library(odbc)
library(ggmap)
library(ggpubr)
library(EDIutils)

#--------------------------------------------------
# Helper function for loading EDI entities
#--------------------------------------------------

read_edi_entity <- function(package_id, entity_name) {
  entity_id <- read_data_entity_names(package_id) %>%
    filter(entityName == entity_name) %>%
    pull(entityId)
  raw_file <- read_data_entity(package_id, entity_id)
  readr::read_csv(raw_file, show_col_types = FALSE)
}

#--------------------------------------------------
# Load PESP data
#--------------------------------------------------

pesp <- read_edi_entity(package_id = "edi.2209.2",entity_name = "PESP_enumeration.csv") %>%
  mutate(Date = as.Date(Date)) %>%
  arrange(Date, Time)

#--------------------------------------------------
# Load EMP discrete WQ data
#--------------------------------------------------

emp <- read_edi_entity(package_id = "edi.458.14",entity_name = "EMP_DWQ_1975-2024") 
emp$analyte_unit <- paste(emp$Analyte, emp$Result_Unit, sep = "-") 

id_cols <- c("Date", "Time", "Station")

values_wide <- emp %>%
  group_by(across(all_of(id_cols)), analyte_unit) %>%
  summarise(Result_Value = first(na.omit(Result_Value)), .groups = "drop") %>%
  pivot_wider(
    id_cols   = all_of(id_cols),
    names_from = analyte_unit,
    values_from = Result_Value
  )

depths_wide <- emp %>%
  filter(!is.na(Sampling_Depth)) %>%
  mutate(depth_type = if_else(Sampling_Depth <= 3, "surface", "bottom")) %>%
  group_by(across(all_of(id_cols))) %>%
  summarise(
    Sample_depth_surface = first(Sampling_Depth[depth_type == "surface"]),
    Sample_depth_bottom  = first(Sampling_Depth[depth_type == "bottom"]),
    .groups = "drop"
  )

wide <- left_join(depths_wide, values_wide,by = id_cols)
wide <- wide %>% mutate(across(c(4:30,32:40,44:54), as.numeric))

dat1 <- data.frame(Survey="DWR-EMP",
                   Station=wide$Station,
                   Latitude=wide$`Latitude-decimal degrees`,
                   Longitude=wide$`Longitude-decimal degrees`,
                   Date=wide$Date,
                   Datetime=as.POSIXct(paste(wide$Date, wide$Time), format="%Y-%m-%d %H:%M:%S"),
                   Sample_depth_surface=wide$Sample_depth_surface,
                   Sample_depth_nutr_surface=wide$Sample_depth_surface,
                   Sample_depth_bottom=wide$Sample_depth_bottom,
                   Temperature=wide$`Water Temperature-degC`,
                   Temperature_bottom=wide$`(Bottom) Water Temperature-degC`,
                   Conductivity=wide$`Specific Conductance-uS/cm`,
                   Conductivity_bottom=wide$`(Bottom) Specific Conductance-uS/cm`,
                   Salinity=NA,
                   Salinity_bottom=NA,
                   Secchi=wide$`Secchi Depth-cm`,
                   TurbidityNTU=wide$`Turbidity-NTU`,
                   TurbidityNTU_bottom=NA,
                   TurbidityFNU=wide$`Turbidity-FNU`,
                   TurbidityFNU_bottom=wide$`(Bottom) Turbidity-FNU`,
                   DissolvedOxygen=wide$`Dissolved Oxygen mg/L-mg/L`,
                   DissolvedOxygen_bottom=wide$`(Bottom) Dissolved Oxygen mg/L-mg/L`,
                   DissolvedOxygenPercent=wide$`Dissolved Oxygen % Saturation-% Saturation`,
                   DissolvedOxygenPercent_bottom=wide$`(Bottom) Dissolved Oxygen % Saturation-% Saturation`,
                   pH=wide$`pH-unitless`,
                   pH_bottom=wide$`(Bottom) pH-unitless`,
                   Microcystis=wide$`MVI-NA`,
                   Chlorophyll=wide$`Chlorophyll a-ug/L`,
                   Chlorophyll_bottom=wide$`(Bottom) Chlorophyll Fluorescence ug/L-ug/L`,
                   Pheophytin=wide$`Pheophytin a-ug/L`,
                   TotAmmonia=wide$`Total Ammonia-mg/L`,
                   DissAmmonia=wide$`Dissolved Ammonia-mg/L`,
                   DissNitrateNitrite=wide$`Dissolved Nitrate + Nitrite-mg/L`,
                   TotPhos=wide$`Total Phosphorus-mg/L`,
                   DissOrthophos=wide$`Dissolved Orthophosphate-mg/L`,
                   TON=wide$`Total Organic Nitrogen-mg/L`,
                   DON=wide$`Dissolved Organic Nitrogen-mg/L`,
                   TKN=wide$`Total Kjeldahl Nitrogen-mg/L`,
                   TotAlkalinity=wide$`Total Alkalinity-mg/L`,
                   DissBromide=wide$`Dissolved Bromide-mg/L`,
                   DissCalcium=wide$`Dissolved Calcium-mg/L`,
                   TotChloride=wide$`Total Chloride-mg/L`,
                   DissChloride=wide$`Dissolved Chloride-mg/L`,
                   DissSilica=wide$`Dissolved Silica-mg/L`,
                   TOC=wide$`Total Organic Carbon-mg/L`,
                   DOC=wide$`Dissolved Organic Carbon-mg/L`,
                   TDS=wide$`Total Dissolved Solids-mg/L`,
                   TSS=wide$`Total Suspended Solids-mg/L`,
                   VSS=wide$`Volatile Suspended Solids-mg/L`)

#--------------------------------------------------
# Load Peterson data
#--------------------------------------------------

con <- dbConnect(odbc(),
                 Driver = "MySQL ODBC 9.7 Unicode Driver",
                 Server = "mysql8.natweb.usgs.gov",
                 Database = "sfbay",
                 UID = "sfbay",
                 PWD = "TnMNAXA9AypR", #Ask Schuyler for PW if you need it!
                 Port = 3306,
                 SSLMODE = "REQUIRED")

# Read in cruise database
cruiseall <- dbReadTable(con, "cruiseall")

# Read in nutrient database
nutrients <- dbReadTable(con, "nutrients")

# Disconnect database
dbDisconnect(con)
rm(con)

id_cols <- c("date", "time", "stat")

binned1 <-  cruiseall %>% select(date,time,stat,depth,dscrchl,dscroxy,oxy,oxysat,temp,salin) %>% 
  filter(!is.na(dscrchl) | !is.na(dscroxy)) %>%
  mutate(depth_bin = if_else(depth <= 3, "surface", "deep")) %>%
  group_by(across(all_of(id_cols)), depth_bin) %>%
  summarise(depth   = mean(depth,   na.rm = TRUE),
    dscrchl = mean(dscrchl, na.rm = TRUE),
    dscroxy = mean(dscroxy, na.rm = TRUE),
    oxy     = mean(oxy,     na.rm = TRUE),
    oxysat  = mean(oxysat,  na.rm = TRUE),
    temp  = mean(temp,  na.rm = TRUE),
    salin  = mean(salin,  na.rm = TRUE),
        .groups = "drop") %>%
  mutate(across(where(is.numeric), ~ ifelse(is.nan(.x), NA, .x)))

binned2 <-  nutrients %>%
  mutate(depth_bin = if_else(depth <= 3, "surface", "deep")) %>%
  group_by(across(all_of(id_cols)), depth_bin) %>%
  summarise(depth   = mean(depth,   na.rm = TRUE),
    no2 = mean(no2, na.rm = TRUE),
    NO32 = mean(NO32, na.rm = TRUE),
    nh4     = mean(nh4,     na.rm = TRUE),
    po4  = mean(po4,  na.rm = TRUE),
    si  = mean(si,  na.rm = TRUE),
    .groups = "drop") %>%
  mutate(across(where(is.numeric), ~ ifelse(is.nan(.x), NA, .x)))


wide1 <- binned1 %>%
  pivot_wider(id_cols     = c(date, time, stat),
    names_from  = depth_bin,
    values_from = c(depth, dscrchl, dscroxy, oxy, oxysat, temp, salin),
    names_glue  = "{.value}_{depth_bin}")

wide2 <- binned2 %>%
  pivot_wider(id_cols     = c(date, time, stat),
    names_from  = depth_bin,
    values_from = c(depth, no2, NO32, nh4, po4, si),
    names_glue  = "{.value}_{depth_bin}") %>%
  rename(nut_depth_surface=depth_surface,nut_depth_deep=depth_deep)

joined <- full_join(wide1, wide2, by = c("date", "stat", "time"))
joined <- joined %>%
  mutate(oxy_surface_combined = coalesce(dscroxy_surface, oxy_surface),
         oxy_deep_combined = coalesce(dscroxy_deep, oxy_deep))

stat <- read_excel("Data/Peterson_Stations.xlsx")
stat <- stat %>% rename(stat=Station)
stat$stat <- as.numeric(stat$stat)
joined <- left_join(joined,stat,by="stat")

dat2 <- data.frame(Survey="USGS-BP",
                   Station=joined$stat,
                   Latitude=joined$`Latitude (degrees)`,
                   Longitude=joined$`Longitude (degrees)`,
                   Date=joined$date,
                   Datetime=as.POSIXct(paste(joined$date, joined$time), format="%Y-%m-%d %H:%M:%S"),
                   Sample_depth_surface=joined$depth_surface,
                   Sample_depth_nutr_surface=joined$nut_depth_surface,
                   Sample_depth_bottom=joined$depth_deep,
                   Temperature=joined$temp_surface,
                   Temperature_bottom=joined$temp_deep,
                   Conductivity=NA,
                   Conductivity_bottom=NA,
                   Salinity=joined$salin_surface,
                   Salinity_bottom=joined$salin_deep,
                   Secchi=NA,
                   TurbidityNTU=NA,
                   TurbidityNTU_bottom=NA,
                   TurbidityFNU=NA,
                   TurbidityFNU_bottom=NA,
                   DissolvedOxygen=joined$oxy_surface_combined,
                   DissolvedOxygen_bottom=joined$oxy_deep_combined,
                   DissolvedOxygenPercent=joined$oxysat_surface,
                   DissolvedOxygenPercent_bottom=joined$oxysat_deep,
                   pH=NA,
                   pH_bottom=NA,
                   Microcystis=NA,
                   Chlorophyll=joined$dscrchl_surface,
                   Chlorophyll_bottom=joined$dscrchl_deep,
                   Pheophytin=NA,
                   TotAmmonia=NA,
                   DissAmmonia=joined$nh4_surface,
                   DissNitrateNitrite=joined$NO32_surface,
                   TotPhos=NA,
                   DissOrthophos=joined$po4_surface,
                   TON=NA,
                   DON=NA,
                   TKN=NA,
                   TotAlkalinity=NA,
                   DissBromide=NA,
                   DissCalcium=NA,
                   TotChloride=NA,
                   DissChloride=NA,
                   DissSilica=joined$si_surface,
                   TOC=NA,
                   DOC=NA,
                   TDS=NA,
                   TSS=NA,
                   VSS=NA)

  
#--------------------------------------------------
# Merge WQ datasets
#--------------------------------------------------

wq <- rbind(dat1,dat2)
wq$Longitude <- as.numeric(wq$Longitude)
rm(binned1,binned2,cruiseall,dat1,dat2,depths_wide,emp,joined,nutrients,stat,values_wide,wide,wide1,wide2,x,id_cols,read_edi_entity)

#--------------------------------------------------
# Load EMP benthic data
#--------------------------------------------------

clam <- read_csv("Data/combined_clams_dataset2.csv",show_col_types = FALSE) %>%
  mutate(station_name = str_remove(station_name, "^EMP "),Date = as.Date(Date))%>%
  group_by(Date,station_name)%>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))
clam[is.na(clam)] <- NA

#--------------------------------------------------
# Load Dayflow data
#--------------------------------------------------

df <- read_csv("Data/Dayflow/dayflow-results-1997-2023.csv",show_col_types = FALSE) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"),WY = ifelse(Month >= 10, Year + 1, Year)) %>%
  left_join(read_csv("Data/Dayflow/wy_class_ind.csv",show_col_types = FALSE),by = "WY")

```


## Calculate presence/absence of potential toxigenic species
0 = absence of potentially toxigenic species, 1 = presence of potentially toxigenic species.
```{r, warning=FALSE, message=FALSE}

pesp <- pesp %>%
  group_by(Date,Station) %>%
  mutate(ptox = ifelse(any((Genus %in%
                  c("Microcoleus","Cuspidothrix","Aphanizomenon","Microcystis","Planktothrix") &
                  Taxon != "Aphanizomenon flos-aquae") |
        Taxon %in% c("Dolichospermum circinale","Microseira wollei")),1,0)) %>%
  ungroup()

```


## Aggregate PESP data by Algal Group
```{r, warning=FALSE, message=FALSE}

sum_na <- function(x) ifelse(all(is.na(x)), NA_real_, sum(x, na.rm = TRUE))

pesp1 <- pesp %>%
  group_by(Survey,Date,Station,Latitude,Longitude,SampleMethod,SampleDepth,DepthType,Lab,
           CountMethodSample,CountMethodTaxa,Magnification,ptox,AlgalGroup) %>%
  summarize(Cells_per_mL = sum_na(Cells_per_mL),
            Units_per_mL = sum_na(Units_per_mL),
            Biovolume_per_mL = sum_na(Biovolume_per_mL))

# Calculate relative abundance/biovolume for each sample
tot <- pesp1 %>% group_by(Date,Station) %>% summarise(tot_cells = sum_na(Cells_per_mL),
                                                      tot_units = sum_na(Units_per_mL),
                                                      tot_bio = sum_na(Biovolume_per_mL))
pesp1 <- pesp1 %>% left_join(tot,by=c("Date","Station"))
pesp1$relative_cell_density <- (pesp1$Cells_per_mL/pesp1$tot_cells)*100;
pesp1$relative_unit_density <- (pesp1$Units_per_mL/pesp1$tot_units)*100;
pesp1$relative_biovolume <- (pesp1$Biovolume_per_mL/pesp1$tot_bio)*100;

rm(sum_na,tot)

```


## Merge PESP and discrete datasets
```{r, warning=FALSE, message=FALSE}

# Merge wq and df
x <- df[,c("Date","SAC","SJR","OUT","EXPORTS","Index")]
wq_df <- left_join(wq, x, by = "Date")
wq_df <- subset(wq_df, select = -c(Latitude, Longitude))

# Merge with PESP data
dat <- left_join(pesp1, wq_df, by = c("Survey","Date","Station"))

rm(x,wq_df,wq_df)

```

## Add in Year, Month, and Season
```{r, warning=FALSE, message=FALSE}

dat$Month <- month(dat$Date)
dat$Year <- year(dat$Date)

dat$Season<- NA
I = which(dat$Month %in% c(7:9)); dat$Season[I]<-"Summer"
I = which(dat$Month %in% c(10:12)); dat$Season[I]<-"Autumn"
I = which(dat$Month %in% c(1:3)); dat$Season[I]<-"Winter"
I = which(dat$Month %in% c(4:6)); dat$Season[I]<-"Spring"

rm(I)

```

## Add in regional designation
```{r, warning=FALSE, message=FALSE}

points_df <- dat[,c("Station","Survey","Latitude","Longitude")]
points_df <- unique(points_df)
points_df <- points_df %>% filter(!is.na(Latitude))
points_df$id <- 1:length(points_df$Survey)

Region <- st_read('Data/Regions_shp/Rosies_Regions_edited.shp')

points_sf = points_df %>% st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326)

points_with_region <- st_join(points_sf, Region, join = st_within) %>%  
  st_drop_geometry() 

stat <- left_join(points_sf,points_with_region,by=c("Station","Survey","id"))
coord <- data.frame(st_coordinates(stat$geometry))
stat$Latitude <- coord$Y; stat$Longitude <- coord$X; rm(coord)
stat <- stat %>% st_drop_geometry() %>% select(-"id")
dat <- left_join(dat,stat,by=c("Station","Survey","Latitude","Longitude")) %>%
  rename(Region = Regions)

rm(stat,points_df,points_sf,points_with_region,Region)

```


## Add in monthly clam data
```{r, warning=FALSE, message=FALSE}

# Add in region
points_df <- clam[,c("station_name","Latitude","Longitude")]
points_df <- unique(points_df)
points_df <- points_df %>% filter(!is.na(Latitude))
points_df$id <- 1:length(points_df$station_name)

Region <- st_read('Data/Regions_shp/Rosies_Regions_edited.shp')

points_sf = points_df %>% st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326)

points_with_region <- st_join(points_sf, Region, join = st_within) %>%  
  st_drop_geometry() 

stat <- left_join(points_sf,points_with_region,by=c("station_name"))
coord <- data.frame(st_coordinates(stat$geometry))
stat$Latitude <- coord$Y; stat$Longitude <- coord$X; rm(coord)
stat <- stat %>% st_drop_geometry() 

clam <- left_join(clam,stat,by=c("station_name","Latitude","Longitude")) %>%
  rename(Region = Regions)

rm(stat,points_df,points_sf,points_with_region,Region)

monthly_clam <- clam %>%
  filter(!is.na(Region)) %>%
  group_by(Year,Month,Region) %>%
  summarize(monthly_biomass = mean(Biomass,na.rm=TRUE),
            monthly_grazing_rate = mean(Grazing_rate,na.rm=TRUE),
            monthly_filtration_rate = mean(Filtration_rate,na.rm=TRUE)) 

dat <- left_join(dat, monthly_clam, by = c("Year","Month","Region"))

```


## Filter dataset
```{r, warning=FALSE, message=FALSE}

dat1 <- dat %>% filter(Region %in% c("Central", "Confluence", "Suisun Bay") &   
                        Date > "2007-12-31" &    
                        Survey %in% c("DWR-EMP","USGS-BP"))   

dat1 <- dat1 %>% ungroup()%>% filter(!if_all(24:67, is.na)) # Screens out PESP samples without corresponding WQ data
dat1 <- dat1 %>% arrange(Date)

```


## Flag "bloom" or "no bloom" for each sample using chl threshold
Considered "bloom" if Chla is 1 standard deviation above the mean value for the entire time series
```{r, warning=FALSE, message=FALSE}

dat1$bloom <- NA

# Suisun Bay
test <- dat1 %>% filter(Region=="Suisun Bay") %>% group_by(Date) %>% summarize(Chl=mean(Chlorophyll,na.rm=TRUE))
x = mean(test$Chl,na.rm=TRUE) + sd(test$Chl,na.rm=TRUE)
I <- which(dat1$Region == "Suisun Bay" & dat1$Chlorophyll > x); dat1$bloom[I] <- "Yes"; rm(I)
I <- which(dat1$Region == "Suisun Bay" & dat1$Chlorophyll <= x); dat1$bloom[I] <- "No"; rm(I,x)

p1<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Suisun Bay"),aes(x=Date,y=Chlorophyll),color="black")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay" & bloom=="Yes"), aes(x = Date, y = Chlorophyll),color="blue")+ggtitle("Suisun Bay")

# Confluence
test <- dat1 %>% filter(Region=="Confluence") %>% group_by(Date) %>% summarize(Chl=mean(Chlorophyll,na.rm=TRUE))
x = mean(test$Chl,na.rm=TRUE) + sd(test$Chl,na.rm=TRUE)
I <- which(dat1$Region == "Confluence" & dat1$Chlorophyll > x); dat1$bloom[I] <- "Yes"; rm(I)
I <- which(dat1$Region == "Confluence" & dat1$Chlorophyll <= x); dat1$bloom[I] <- "No"; rm(I,test,x)

p2<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Confluence"),aes(x=Date,y=Chlorophyll),color="black")+
  geom_point(data = dat1 %>% filter(Region=="Confluence" & bloom=="Yes"), aes(x = Date, y = Chlorophyll),color="blue")+ggtitle("Confluence")

# Central Delta
test <- dat1 %>% filter(Region=="Central") %>% group_by(Date) %>% summarize(Chl=mean(Chlorophyll,na.rm=TRUE))
x = mean(test$Chl,na.rm=TRUE) + sd(test$Chl,na.rm=TRUE)
I <- which(dat1$Region == "Central" & dat1$Chlorophyll > x); dat1$bloom[I] <- "Yes"; rm(I)
I <- which(dat1$Region == "Central" & dat1$Chlorophyll <= x); dat1$bloom[I] <- "No"; rm(I,test,x)

p3<-ggplot()+geom_path(data = dat1 %>% filter(Region=="Central"),aes(x=Date,y=Chlorophyll),color="black")+
  geom_point(data = dat1 %>% filter(Region=="Central" & bloom=="Yes"), aes(x = Date, y = Chlorophyll),color="blue")+ggtitle("Central")

ggarrange(p1,p2,p3,ncol=1)

rm(p1,p2,p3)

```

## Calculate HAB/BAB/No bloom using MVI
"HAB" if MVI = 3-5 AND occurs from June to October. 
"BAB" if MVI = 1-2 from June to October OR if the sample was collected during a bloom between November and May.
"No bloom" if sample was collected between November and May and chlorophyll is below the bloom threshold.
```{r, warning=FALSE, message=FALSE}

dat1 <- dat1 %>%
  group_by(Region) %>%
  mutate(chl_threshold = mean(Chlorophyll, na.rm = TRUE) + sd(Chlorophyll, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(
    mvi_class = case_when(
      is.na(Microcystis) ~ NA_character_,
      Microcystis >= 3 & Month %in% 6:10 ~ "HAB",
      Microcystis <= 2 & Month %in% 6:10 ~ "BAB",
      Month %in% c(11, 12, 1:5) ~ "BAB",
      TRUE ~ NA_character_
    ),
    mvi_class = case_when(
      mvi_class == "BAB" & bloom == "Yes" ~ "BAB",
      mvi_class == "BAB" & bloom == "No" ~ "No bloom",
      TRUE ~ mvi_class
    )
  )

```


## Data coverage
```{r, warning=FALSE, message=FALSE}

ggplot(data=dat1,aes(x=Date,y=Biovolume_per_mL))+geom_point()+facet_wrap(~Region)+
  ylab("Biovolume (ug/mL)")
ggplot(data=dat1,aes(x=Date,y=Cells_per_mL))+geom_point()+facet_wrap(~Region)


# Region
deltabox <- make_bbox(lat = Latitude, lon = Longitude, data = dat1, f = c(.2, .2)) #sets bounding box based on station coordinates, f value adds a buffer around the perimeter
register_stadiamaps(key = "b4a4b8da-f8e2-4ba1-bcdc-91e9d5c83d00")
stamenmap <- get_stadiamap(deltabox, zoom = 10, maptype = "alidade_smooth", force = TRUE, crop = TRUE)

ggmap(stamenmap) +
  geom_point(data = dat1, 
             mapping = aes(x = Longitude, y = Latitude,fill=Region),size=3,pch=21) +
  #geom_text(data = dat, aes(x = Longitude, y = Latitude,label = Station),fontface = #"bold",cex=3,vjust=0.5,hjust=1.3)+
  labs(title = "", x = "Longitude", y = "Latitude") +
  theme_light(base_size = 12) +
  theme(panel.grid = element_blank(),
        strip.background = element_rect(fill = "black"),
        strip.text = element_text(size = 10),axis.text.x = element_text(angle = 90))+
  xlab("")+ylab("")+theme_light(base_size = 16) 

# Number of samples per region per season over time
dat1 %>% group_by(Region, Year, Season)%>%summarize(n=n()) %>%
  ggplot(aes(x=Year,y=Region))+geom_tile(aes(fill=n))+
  scale_fill_gradient(low="white",high="red")+
  geom_text(aes(label=round(n,1)),size=4)+
  xlab("")+ylab("")+
  facet_grid(Season~.)+
  theme_bw(base_size=14)+
  theme(legend.position = "none")

# Number of samples per region per month over time
dat1 %>% group_by(Region, Year, Month)%>%summarize(n=n()) %>%
  ggplot(aes(x=Year,y=Region))+geom_tile(aes(fill=n))+
  scale_fill_gradient(low="white",high="red")+
  geom_text(aes(label=round(n,1)),size=4)+
  xlab("")+ylab("")+
  facet_grid(Month~.)+
  theme_bw(base_size=14)+
  theme(legend.position = "none")

# Number of samples per station and season over time
dat1 %>% group_by(Station, Region,Season,Year)%>%summarize(n=n()) %>%
  ggplot(aes(x=Year,y=as.factor(Station)))+geom_tile(aes(fill=n))+
  scale_fill_gradient(low="white",high="red")+
  geom_text(aes(label=round(n,1)),size=4)+
  xlab("")+ylab("")+facet_grid(Region~Season,scales = 'free_y',space='free')+
  theme_bw(base_size=14)+
  theme(legend.position = "none")+
  theme(axis.text.x=element_text(angle = -90, vjust = 0.3))

rm(deltabox,stamenmap)
```


## Backcalculate HAB/BAB using MVI and ptox
```{r, warning=FALSE, message=FALSE}

library(lme4)

# Prepare data
dat2 <- dat1 %>% filter(Month %in% (6:10))

train_dat <- dat2 %>% filter(!is.na(ptox) & !is.na(Microcystis) & !is.na(Chlorophyll)) %>%
  select(Month,Date,Region,Station,Chlorophyll,Microcystis,ptox) %>%
  group_by(Month,Date,Region,Station,Chlorophyll,Microcystis,ptox) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))%>%
  mutate(mvi = ifelse(is.na(Microcystis),NA,ifelse(Microcystis %in% 3:5 & Month %in% 6:10, 1, 0)))

missing_dat <- dat2 %>% filter(is.na(Microcystis) & !is.na(ptox) & !is.na(Chlorophyll)) %>%
  select(Month,Date,Region,Station,Chlorophyll,Microcystis,ptox) %>%
  group_by(Month,Date,Region,Station,Chlorophyll,Microcystis,ptox) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE)) %>%
  mutate(mvi = ifelse(is.na(Microcystis),NA,ifelse(Microcystis %in% 3:5 & Month %in% 6:10, 1, 0)))

all_dat <- dat2 %>% filter(!is.na(ptox) & !is.na(Chlorophyll)) %>%
  select(Month,Date,Region,Station,Chlorophyll,Microcystis,ptox) %>%
  group_by(Month,Date,Region,Station,Chlorophyll,Microcystis,ptox) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE)) %>%
  mutate(mvi = ifelse(is.na(Microcystis),NA,ifelse(Microcystis %in% 3:5 & Month %in% 6:10, 1, 0)))

# Fit logistic model
model <- glmer(mvi ~ ptox + Region * Chlorophyll + (1 | Month), family = binomial, data = train_dat)

# Use model to predict values on training data
train_dat$modeled_prob_harmful <- predict(model,newdata = train_dat,type = "response")

train_dat <- train_dat %>% 
  mutate(modeled_class = case_when(modeled_prob_harmful >= 0.4 ~ "HAB", modeled_prob_harmful < 0.4 ~ "BAB"))

# Use model to predict values on missing data
missing_dat$modeled_prob_harmful <- predict(model,newdata = missing_dat,type = "response")

missing_dat <- missing_dat %>% 
  mutate(modeled_class = case_when(modeled_prob_harmful >= 0.4 ~ "HAB", modeled_prob_harmful < 0.4 ~ "BAB"))

# Use model to predict values on entire dataset data
all_dat$modeled_prob_harmful <- predict(model,newdata = all_dat,type = "response")

all_dat <- all_dat %>% 
  mutate(modeled_class = case_when(modeled_prob_harmful >= 0.4 ~ "HAB", modeled_prob_harmful < 0.4 ~ "BAB"))

dat1 <- left_join(dat1,all_dat,by=c("Month","Date","Region","Station","Chlorophyll","Microcystis","ptox"))

dat1 <- dat1 %>%
  mutate(
    modeled_class = case_when(
      # existing (or filled) BAB that's below threshold -> demote to No bloom
      modeled_class == "BAB" & bloom == "No" ~ "No bloom",
      # existing BAB above threshold stays BAB
      modeled_class == "BAB" & bloom == "Yes" ~ "BAB",
      # keep any other existing non-NA value as-is
      !is.na(modeled_class) ~ modeled_class,
      # fill NAs with the month + chlorophyll logic
      Month %in% c(11, 12, 1:5) & bloom == "Yes" ~ "BAB",
      Month %in% c(11, 12, 1:5) & bloom == "No" ~ "No bloom",
      TRUE ~ NA_character_
    )
  )

rm(dat2,model)

```


## Reorder and save final dataset
```{r, warning=FALSE, message=FALSE}

dat <- dat1[, c("Survey","Year","Season","Month","Date","Region","Station","Latitude","Longitude",
                 "SampleMethod","SampleDepth","DepthType","Lab","CountMethodSample","CountMethodTaxa",
                 "Magnification","AlgalGroup","Cells_per_mL","Units_per_mL","Biovolume_per_mL","tot_cells",
                 "tot_units","tot_bio","relative_cell_density","relative_unit_density","relative_biovolume",
                 "chl_threshold","bloom","ptox","modeled_prob_harmful","modeled_class","mvi_class",
                 "Sample_depth_surface","Sample_depth_nutr_surface","Sample_depth_bottom","Microcystis",
                 "Chlorophyll","Pheophytin","Temperature","Temperature_bottom","Conductivity",
                 "Conductivity_bottom","Salinity","Salinity_bottom","Secchi","TurbidityNTU",
                 "TurbidityNTU_bottom","TurbidityFNU","TurbidityFNU_bottom","DissolvedOxygen",
                 "DissolvedOxygen_bottom","DissolvedOxygenPercent","DissolvedOxygenPercent_bottom","pH",
                 "DissAmmonia","DissNitrateNitrite","TotPhos","DissOrthophos","TON","DON","TKN",
                 "DissBromide","DissCalcium","TotChloride","DissChloride","DissSilica","TOC","DOC","TDS",
                 "TSS","VSS","SAC","SJR","OUT","EXPORTS","Index","monthly_biomass","monthly_grazing_rate","monthly_filtration_rate")]

save(dat, clam, df, pesp, wq, file = "Data/integrated_dataset_2008-2024.RData")

```


## (need to be updated) HAB/BAB modeled vs observed figures
```{r, warning=FALSE, message=FALSE}

# Plot final dataset
p<- ggplot()+
  geom_path(data = dat1 %>% filter(Region=="Confluence" & !is.na(Chlorophyll)),
            aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Confluence")+
  geom_point(data = dat1 %>% filter(Region=="Confluence" & !is.na(Chlorophyll)),
             aes(x=Date,y=Chlorophyll,fill=as.factor(harmful_class),shape=as.factor(modeled)),size=1)+
  scale_fill_manual(values = c("blue", "red","grey"))+scale_shape_manual(values = c(21, 24)) + 
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+labs(fill="HAB vs. BAB",shape="Modeled?",x="",y="")

p1<- ggplot()+
  geom_path(data = dat1 %>% filter(Region=="Suisun Bay" & !is.na(Chlorophyll)),
            aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Suisun Bay")+
  geom_point(data = dat1 %>% filter(Region=="Suisun Bay" & !is.na(Chlorophyll)),
             aes(x=Date,y=Chlorophyll,fill=as.factor(harmful_class),shape=as.factor(modeled)),size=1)+
  scale_fill_manual(values = c("blue", "red","grey"))+scale_shape_manual(values = c(21, 24)) + 
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+labs(fill="HAB vs. BAB",shape="Modeled?",x="",y="")


p2 <- ggplot()+
  geom_path(data = dat1 %>% filter(Region=="Central" & !is.na(Chlorophyll)),
            aes(x=Date,y=Chlorophyll),color="black")+ggtitle("Central")+
  geom_point(data = dat1 %>% filter(Region=="Central" & !is.na(Chlorophyll)),
             aes(x=Date,y=Chlorophyll,fill=as.factor(harmful_class),shape=as.factor(modeled)),size=1)+
  scale_fill_manual(values = c("blue", "red","grey"))+scale_shape_manual(values = c(21, 24)) + 
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+labs(fill="HAB vs. BAB",shape="Modeled?",x="",y="")

tiff("../phyto_synthesis/Figures/HAB v BAB/final_hab_bab_dataset.tiff", units="in", width=12, height=8, res=300)
ggarrange(p2,p,p1,ncol=1,common.legend=TRUE)
dev.off()


# Compare real MVI to predicted MVI in training data
train_dat <- train_dat %>%
  mutate(color = case_when(
    pred_class == "BAB" & HAB_BAB_class == "BAB" ~ "blue",
    pred_class == "HAB" & HAB_BAB_class == "HAB" ~ "red",
    pred_class == "BAB" & HAB_BAB_class == "HAB" ~ "orange",
    pred_class == "HAB" & HAB_BAB_class == "BAB" ~ "green",
    TRUE ~ NA_character_
  ))
train_dat <- train_dat %>% arrange(Date)

p <- ggplot()+
  geom_point(data = train_dat %>% filter(Region=="Central"),
             aes(x=Date,y=Chlorophyll,fill=as.factor(color)),pch=21,size=2)+
  ggtitle("Central Delta, training")+
  scale_fill_manual(values = c("blue", "green","orange","red"),
                    labels=c("Mod BAB, Obs BAB","Mod HAB, Obs BAB","Mod BAB, Obs HAB","Mod HAB, Obs HAB"))+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+
  labs(fill="Modeled vs. Observed",x="",y="Chl-a (ug/L)")

p1 <- ggplot()+
  geom_point(data = train_dat %>% filter(Region=="Confluence"),
             aes(x=Date,y=Chlorophyll,fill=as.factor(color)),pch=21,size=2)+
  ggtitle("Confluence, training")+
  scale_fill_manual(values = c("blue", "green","orange","red"),
                    labels=c("Mod BAB, Obs BAB","Mod HAB, Obs BAB","Mod BAB, Obs HAB","Mod HAB, Obs HAB"))+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+
  labs(fill="Modeled vs. Observed",x="",y="Chl-a (ug/L)")

p2 <- ggplot()+
  geom_point(data = train_dat %>% filter(Region=="Suisun Bay"),
             aes(x=Date,y=Chlorophyll,fill=as.factor(color)),pch=21,size=2)+
  ggtitle("Suisun Bay, training")+
  scale_fill_manual(values = c("blue", "green","orange","red"),
                    labels=c("Mod BAB, Obs BAB","Mod HAB, Obs BAB","Mod BAB, Obs HAB","Mod HAB, Obs HAB"))+
  theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+
  labs(fill="Modeled vs. Observed",x="",y="Chl-a (ug/L)")

tiff("../phyto_synthesis/Figures/HAB v BAB/real_v_modeled.tiff", units="in", width=12, height=8, res=300)
ggarrange(p,p1,p2,ncol=1,common.legend=TRUE,legend="bottom")
dev.off()

```


## (need to be updated) HAB/BAB model validation
```{r, warning=FALSE, message=FALSE}

library(pROC)      # AUC
library(caret) 
library(lme4)

# Prepare data
dat2 <- dat1 %>% filter(Month %in% (6:10))

train_dat <- dat2 %>% filter(!is.na(ptox) & !is.na(mvi) & !is.na(Chlorophyll)) %>%
  select(Month,Date,Region,Station,Chlorophyll,mvi,ptox) %>%
  group_by(Month,Date,Region,Station,Chlorophyll,mvi,ptox) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))

missing_dat <- dat2 %>% filter(is.na(mvi) & !is.na(ptox) & !is.na(Chlorophyll)) %>%
  select(Month,Date,Region,Station,Chlorophyll,mvi,ptox) %>%
  group_by(Month,Date,Region,Station,Chlorophyll,mvi,ptox) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))

# Fit logistic model
model <- glmer(mvi ~ ptox + Region * Chlorophyll + (1 | Month), family = binomial, data = train_dat)
summary(model)

# Cross validation
set.seed(123)
ctrl <- trainControl(method = "cv",number = 5,classProbs = TRUE,summaryFunction = twoClassSummary)
train_dat$harmful_factor <- factor(ifelse(train_dat$mvi == 1, "yes", "no"))

cv_model <- train(harmful_factor ~ ptox + Region * Chlorophyll + (1 | Month),
  data = train_dat,method = "glm",family = "binomial",trControl = ctrl,metric = "PR")
mean(cv_model$resample$ROC)
sd(cv_model$resample$ROC)

# Check model output results
train_dat$prob_harmful <- predict(model, type = "response")
ggplot(data=train_dat)+geom_point(aes(y=prob_harmful,x=Chlorophyll))
ggplot(data=train_dat)+geom_point(aes(y=prob_harmful,x=ptox))

# Validate model

# AUC
roc_obj <- roc(train_dat$mvi, train_dat$prob_harmful)
auc_val <- auc(roc_obj)

print(auc_val)
# AUC ≥ 0.75 → strong enough to justify back-calculation
# AUC 0.65–0.75 → usable, but report uncertainty carefully
# AUC < 0.65 → ptox alone may not be reliable predictor

# AUC-PR

# Get predicted probabilities (not class predictions)
probs <- predict(model,newdata = train_dat,type = "response")
actual <- train_dat$mvi

# Precision-Recall curve
pr <- PRROC::pr.curve(scores.class0 = probs, weights.class0 = actual, curve = TRUE)
plot(pr)
pr$auc.integral  # Your AUC-PR value

# Confusion matrices
train_dat$prob_harmful <- predict(model,newdata = train_dat,type = "response")

train_dat <- train_dat %>% 
  mutate(pred_class = case_when(prob_harmful >= 0.4 ~ "HAB", prob_harmful < 0.4 ~ "BAB")) %>%
  mutate(HAB_BAB_class = case_when(mvi == 1 ~ "HAB", mvi == 0 ~ "BAB"))

confusionMatrix(factor(train_dat$pred_class),factor(train_dat$HAB_BAB_class), positive = "HAB")

```
