---
title: "Seasonal Analysis"
author: "Shaela Noble"
date: "2025-06-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Read in packages
```{r Read in Packages, include=FALSE}
library(ggplot2)
library(plyr)
library(dplyr)
library(splines)
library(tidyverse)
library(caret)
library(ggrepel)
library(here)
library(FSA)
library(nlstools)
library(corrplot)
library(sf)
library(forcats)
theme_set(theme_classic())
```

Read in Region Shapefile
```{r Read in Region Shapefile, include=FALSE}
#read in region data
Rosies_regions <- st_read("clams_example/Rosies_regions_edited/Rosies_regions_edited.shp")

#change the crs to UTM zone 10N (EPSG 32610)
Rosies_regions<- st_transform(Rosies_regions, crs =32610)
```

Read in Alex's csv
```{r Read in Alexs csv, include=FALSE}
#read in region data
wq_flow <- read.csv("Data/edi_df_integrate_monthly.csv")
```

Create seasonal variable
```{r Create seasonal variable, include=FALSE}
#convert dates to months and seasons
wq_season<-wq_flow %>% mutate(
       Year=if_else(Month==12, Year+1, Year), # Move Decembers to the following year
       Season=case_when(Month%in%3:5 ~ "Spring", # Create seasonal variables
                        Month%in%6:8 ~ "Summer",
                        Month%in%9:11 ~ "Fall",
                        Month%in%c(12, 1, 2) ~ "Winter",
                        TRUE ~ NA_character_))
```

Attach Regions and Clean up Data
```{r Attach Regions and Clean up Data, include=FALSE}
#filter data frame to contain variables of interest 
wq_data<-wq_season %>% select(year_month, Season, Source, Station, Latitude, Longitude, Depth, Temperature, Conductivity, Conductivity_bottom, Secchi, Chlorophyll, DissAmmonia, DissNitrateNitrite, DissOrthophos, SAC, SJR, OUT, EXPORT)

#filter out NA coord
wq_data <- filter(wq_data, is.na(Latitude) == FALSE)

#convert file to simple features so we can map stations to subregions
wq_data_sf = wq_data %>% st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) %>%
  #convert to UTMs so it's in the same coordinate reference system as Rosie's regions shapefile
  st_transform(crs = 32610)

#Join regions to WQ data
wq_data_regions <- st_join(wq_data_sf, Rosies_regions["Regions"], join = st_intersects, left= TRUE) %>%  
  filter(!is.na(Regions)) %>%
  st_drop_geometry() 

#restructure data for facet wrap
wq_data_facet <- wq_data_regions %>% select(Source, Station, year_month, Regions, Season, Temperature, Conductivity, Secchi, Chlorophyll, DissAmmonia, DissNitrateNitrite, DissOrthophos, SAC, SJR, OUT, EXPORT) %>% mutate(log_Conductivity = log(Conductivity), log_Secchi = log(Secchi), log_DissAmmonia = log(DissAmmonia), log_DissNitrateNitrite = log(DissNitrateNitrite), log_DissOrthophos = log(DissOrthophos), log_SAC = log(SAC), log_SJR = log(SJR), log_OUT = log(OUT), log_EXPORT = log(EXPORT)) %>% pivot_longer(cols=c(-Source, -Station, -year_month, -Chlorophyll, -Season, - Regions), names_to="Covariate_name", values_to="Covariate_value")

wq_data_facet2 <- wq_data_regions %>% select(Source, Station, year_month, Regions, Season, Temperature, Conductivity, Secchi, Chlorophyll, DissAmmonia, DissNitrateNitrite, DissOrthophos, SAC, SJR, OUT, EXPORT) %>%  pivot_longer(cols=c(-Source, -Station, -year_month, -Chlorophyll, -Season, - Regions), names_to="Covariate_name", values_to="Covariate_value")
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables}
plot <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>% 
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, ncol = 4, scales = "free")+
  theme_bw()+
  ggtitle("Delta wide")
plot
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables (no log transform)
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables (no log transform)}
plot2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>% 
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, ncol = 4, scales = "free")+
  theme_bw()+
  ggtitle("Delta wide")
plot2
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season}
plot_winter <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Winter") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Winter")
plot_winter

plot_spring <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Spring") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Spring")
plot_spring

plot_summer <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Summer") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Summer")
plot_summer

plot_fall <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Fall") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Fall")
plot_fall
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season (no log transform)
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season (no log transform)}
plot_winter2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Winter") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Winter")
plot_winter2

plot_spring2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Spring") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Spring")
plot_spring2

plot_summer2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Summer") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Summer")
plot_summer2

plot_fall2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Fall") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Fall")
plot_fall2
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Suisun Marsh
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Suisun Marsh}
plot_winter_sm <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Winter") %>% filter(Regions == "Suisun Marsh") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Marsh - Winter")
plot_winter_sm

plot_spring_sm <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Spring") %>% filter(Regions == "Suisun Marsh") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Marsh - Spring")
plot_spring_sm

plot_summer_sm <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Summer") %>% filter(Regions == "Suisun Marsh") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT",  "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Marsh - Summer")
plot_summer_sm

plot_fall_sm <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Fall") %>% filter(Regions == "Suisun Marsh") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Marsh - Fall")
plot_fall_sm
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Suisun Marsh (no log transform)
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Suisun Marsh (no log transform)}
plot_winter_sm2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Winter") %>% filter(Regions == "Suisun Marsh") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Marsh - Winter")
plot_winter_sm2

plot_spring_sm2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Spring") %>% filter(Regions == "Suisun Marsh") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Marsh - Spring")
plot_spring_sm2

plot_summer_sm2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Summer") %>% filter(Regions == "Suisun Marsh") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Marsh - Summer")
plot_summer_sm2

plot_fall_sm2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Fall") %>% filter(Regions == "Suisun Marsh") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Marsh - Fall")
plot_fall_sm2
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Suisun Bay
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Suisun Bay}
plot_winter_sb <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Winter") %>% filter(Regions == "Suisun Bay") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Bay - Winter")
plot_winter_sb

plot_spring_sb <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Spring") %>% filter(Regions == "Suisun Bay") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Bay - Spring")
plot_spring_sb

plot_summer_sb <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Summer") %>% filter(Regions == "Suisun Bay") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Bay - Summer")
plot_summer_sb

plot_fall_sb <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Fall") %>% filter(Regions == "Suisun Bay") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Bay - Fall")
plot_fall_sb
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Suisun Bay (no log transform)
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Suisun Bay (no log transform)}
plot_winter_sb2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Winter") %>% filter(Regions == "Suisun Bay") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Bay - Winter")
plot_winter_sb2

plot_spring_sb2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Spring") %>% filter(Regions == "Suisun Bay") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Bay - Spring")
plot_spring_sb2

plot_summer_sb2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Summer") %>% filter(Regions == "Suisun Bay") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Bay - Summer")
plot_summer_sb2

plot_fall_sb2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Fall") %>% filter(Regions == "Suisun Bay") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Suisun Bay - Fall")
plot_fall_sb2
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: North Delta
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: North Delta}
plot_winter_nd <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Winter") %>% filter(Regions == "North Delta") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT","EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("North Delta - Winter")
plot_winter_nd

plot_spring_nd <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Spring") %>% filter(Regions == "North Delta") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("North Delta - Spring")
plot_spring_nd

plot_summer_nd <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Summer") %>% filter(Regions == "North Delta") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("North Delta - Summer")
plot_summer_nd

plot_fall_nd <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Fall") %>% filter(Regions == "North Delta") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("North Delta - Fall")
plot_fall_nd
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: North Delta (no log transform)
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: North Delta (no log transform)}
plot_winter_nd2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Winter") %>% filter(Regions == "North Delta") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("North Delta - Winter")
plot_winter_nd2

plot_spring_nd2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Spring") %>% filter(Regions == "North Delta") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("North Delta - Spring")
plot_spring_nd2

plot_summer_nd2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Summer") %>% filter(Regions == "North Delta") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("North Delta - Summer")
plot_summer_nd2

plot_fall_nd2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Fall") %>% filter(Regions == "North Delta") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("North Delta - Fall")
plot_fall_nd2
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Confluence
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Confluence}
plot_winter_cf <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Winter") %>% filter(Regions == "Confluence") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Confluence - Winter")
plot_winter_cf

plot_spring_cf <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Spring") %>% filter(Regions == "Confluence") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Confluence - Spring")
plot_spring_cf

plot_summer_cf <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Summer") %>% filter(Regions == "Confluence") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Confluence - Summer")
plot_summer_cf

plot_fall_cf <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Fall") %>% filter(Regions == "Confluence") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Confluence - Fall")
plot_fall_cf
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Confluence (no log transform)
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Confluence (no log transform)}
plot_winter_cf2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Winter") %>% filter(Regions == "Confluence") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Confluence - Winter")
plot_winter_cf2

plot_spring_cf2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Spring") %>% filter(Regions == "Confluence") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Confluence - Spring")
plot_spring_cf2

plot_summer_cf2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Summer") %>% filter(Regions == "Confluence") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Confluence - Summer")
plot_summer_cf2

plot_fall_cf2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Fall") %>% filter(Regions == "Confluence") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Confluence - Fall")
plot_fall_cf2
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Central
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Central}
plot_winter_cn <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Winter") %>% filter(Regions == "Central") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Central - Winter")
plot_winter_cn

plot_spring_cn <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Spring") %>% filter(Regions == "Central") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Central - Spring")
plot_spring_cn

plot_summer_cn <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Summer") %>% filter(Regions == "Central") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Central - Summer")
plot_summer_cn

plot_fall_cn <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Fall") %>% filter(Regions == "Central") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Central - Fall")
plot_fall_cn
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Central (no log transform)
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: Central (no log transform)}
plot_winter_cn2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Winter") %>% filter(Regions == "Central") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Central - Winter")
plot_winter_cn2

plot_spring_cn2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Spring") %>% filter(Regions == "Central") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Central - Spring")
plot_spring_cn2

plot_summer_cn2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Summer") %>% filter(Regions == "Central") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Central - Summer")
plot_summer_cn2

plot_fall_cn2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Fall") %>% filter(Regions == "Central") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("Central - Fall")
plot_fall_cn2
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: South
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: South}
plot_winter_sd <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Winter") %>% filter(Regions == "South") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("South - Winter")
plot_winter_sd

plot_spring_sd <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Spring") %>% filter(Regions == "South") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("South - Spring")
plot_spring_sd

plot_summer_sd <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Summer") %>% filter(Regions == "South") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("South - Summer")
plot_summer_sd

plot_fall_sd <- wq_data_facet %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Fall") %>% filter(Regions == "South") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("South - Fall")
plot_fall_sd
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: South (no log transform)
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables by Season and Across Regions: South (no log transform)}
plot_winter_sd2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Winter") %>% filter(Regions == "South") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("South - Winter")
plot_winter_sd2

plot_spring_sd2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Spring") %>% filter(Regions == "South") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("South - Spring")
plot_spring_sd2

plot_summer_sd2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Summer") %>% filter(Regions == "South") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("South - Summer")
plot_summer_sd2

plot_fall_sd2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll)) %>% filter(Season == "Fall") %>% filter(Regions == "South") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw() +
  ggtitle("South - Fall")
plot_fall_sd2
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables for Different Time Periods (Pre-Overbite Clam, Pre-POD, and Post-POD)
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables for Different Time Periods (Pre-Overbite Clam, Pre-POD, and Post-POD)}
plot_preclam <- wq_data_facet %>% filter(!is.na(Chlorophyll))%>% mutate(Date = as.Date(year_month, "%Y-%m-%d")) %>% filter(Date >= "1979-01-01" & Date < "1987-01-01") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>% 
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, ncol = 4, scales = "free")+
  theme_bw()+
  ggtitle("Pre-Overbite Clam (1979-1986)")
plot_preclam

plot_prePOD <- wq_data_facet %>% filter(!is.na(Chlorophyll))%>% mutate(Date = as.Date(year_month, "%Y-%m-%d")) %>% filter(Date >= "1987-01-01" & Date < "2002-01-01") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>% 
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, ncol = 4, scales = "free")+
  theme_bw()+
  ggtitle("Pre-POD (1987-2002)")
plot_prePOD

plot_postPOD <- wq_data_facet %>% filter(!is.na(Chlorophyll))%>% mutate(Date = as.Date(year_month, "%Y-%m-%d")) %>% filter(Date >= "2002-01-01") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Conductivity", "log_Conductivity", "Secchi", "log_Secchi", "DissAmmonia", "log_DissAmmonia", "DissNitrateNitrite", "log_DissNitrateNitrite", "DissOrthophos", "log_DissOrthophos", "SAC", "log_SAC", "SJR", "log_SJR", "OUT", "log_OUT", "EXPORT", "log_EXPORT", "Temperature")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, ncol = 4, scales = "free")+
  theme_bw()+
  ggtitle("Post-POD (2002-present)")
plot_postPOD
```

Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables for Different Time Periods (Pre-Overbite Clam, Pre-POD, and Post-POD) (no log transform)
```{r Explore Relationships Among Discrete Chlorophyll A and Other Predictor Variables for Different Time Periods (Pre-Overbite Clam, Pre-POD, and Post-POD) (no log transform)}
plot_preclam2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll))%>% mutate(Date = as.Date(year_month, "%Y-%m-%d")) %>% filter(Date >= "1979-01-01" & Date < "1987-01-01") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>% 
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, ncol = 4, scales = "free")+
  theme_bw()+
  ggtitle("Pre-Overbite Clam (1979-1986)")
plot_preclam2

plot_prePOD2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll))%>% mutate(Date = as.Date(year_month, "%Y-%m-%d")) %>% filter(Date >= "1987-01-01" & Date < "2002-01-01") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>% 
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, ncol = 4, scales = "free")+
  theme_bw()+
  ggtitle("Pre-POD (1987-2002)")
plot_prePOD2

plot_postPOD2 <- wq_data_facet2 %>% filter(!is.na(Chlorophyll))%>% mutate(Date = as.Date(year_month, "%Y-%m-%d")) %>% filter(Date >= "2002-01-01") %>% mutate(Covariate_name = forcats::fct_relevel(Covariate_name, "Temperature", "Conductivity", "Secchi",  "DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "SAC", "SJR", "OUT", "EXPORT")) %>%
ggplot(aes(x=Covariate_value, y=(Chlorophyll)))+
  geom_point()+
  facet_wrap(~Covariate_name, ncol = 4, scales = "free")+
  theme_bw()+
  ggtitle("Post-POD (2002-present)")
plot_postPOD2
```

Discrete Chl-a Plots
```{r Discrete Chl-a Plots}
chl_a <- wq_data_regions %>% filter(!is.na(Chlorophyll)) %>% 
  ggplot(aes(x=year_month, y=Chlorophyll)) +
  geom_line()
chl_a

chl_a_season <- wq_data_regions %>% filter(!is.na(Chlorophyll)) %>% mutate(Date = as.Date(year_month, "%Y-%m-%d")) %>% 
  ggplot(aes(x=Date, y=sqrt(Chlorophyll), color=Regions)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~Season, ncol = 1, scales = "free") +
  scale_x_date(breaks = date_breaks("10 years"),
  labels = date_format("%Y"))
chl_a_season
```