---
title:
  "Global GAM"
author:
  "Ryan Hankins"
output: 
  html_document:
        toc: TRUE
        toc_float: TRUE
        code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, tidy = "styler")
```
# Purpose

The purpose of these models is to understand the effects of environmental drivers and antecedent conditions on Delta-wide phytoplankton biomass (chl-a). 

# Prepare Data

```{r, include=FALSE}

library(mgcv)
library(tidyverse)
library(marginaleffects)
library(MuMIn)
library(car)
library(corrplot)
library(stats)
library(marginaleffects)
library(mice)
library(lattice)
library(ggplot2)
library(data.table)
```


```{r}

# variables to include in the analysis:
c = c("Chlorophyll","logchla", "Regions", "Date","Year", "Month",  "season","SAC","lagSAC",
       "EXPORT","logEXPORT","OUT","lagOUT", "DissAmmonia",
       "lagDissAmmonia", "TotPhos","lagTotPhos","DissNitrateNitrite", 
       "lagDissNitrateNitrite","Temperature","lagTemperature","Secchi", 
       "lagSecchi","Conductivity","lagConductivity","season_num", "DissOrthophos", "lagDissOrthophos", "clam_step","logDissNitrateNitrite","logDissAmmonia","logDissOrthophos","logConductivity","logSAC","logOUT")#, "Biomass","Grazing_rate","Filtration_rate", "SJR","lagSJR", "Index","lagIndex")

df <- read.csv("Data/regional_integrated_dataset3.csv" )

df$Date = as.Date(df$Date)

# 1979 is the start of the study period.
df <- subset(df,df$Date >= "1979-01-01")
# 2022 is the end of the study period
df <- subset(df,df$Date <= "2022-12-31")

# Changing categorical variables to factors
# need to change regions to factor farther down so dataframe splitting by region works properly
#df$Regions = as.factor(df$Regions)
df$season = as.factor(df$season)
df$season_num = as.numeric(df$season)


# Replacing clam variables with a step function (pre/post potamocorbula introduction to the Bay in 1987)
df <- df%>%mutate(clam_step=if_else(Year>1987, "post clam", "pre clam")) 
df$clam_step <- as.factor(df$clam_step)

# log transforming chl-a to achieve normal distribution of response variable.
df$logchla <- log(df$Chlorophyll)

# log transforming some variables to better visualize
df$logDissNitrateNitrite <- log(df$DissNitrateNitrite)
df$logDissAmmonia <- log(df$DissAmmonia)
df$logDissOrthophos <- log(df$DissOrthophos)
df$logConductivity <- log(df$Conductivity)
df$logSAC <- log(df$SAC)
df$logOUT <- log(df$OUT)
df$logEXPORT <- log(df$EXPORT)


 df <- subset(df, select = c)
 
```


# Impute

We are using Multiple Imputations by Chained Equations (MICE) to impute missing data.

- Transforming the data first and then imputing is the preferred strategy.
(https://019b2da8-edfb-a262-61be-7973c056d9ae.share.connect.posit.cloud/mi-fitting.html#ref-von_hippel_2009)

- See this link for a graphic of the imputation process: https://stackoverflow.com/questions/54075836/multiple-imputed-datasets-pooling-results

- Multiple imputed versions of the data set are created, then each imputed data set is analysed with the GAM, then GAM outputs from each dataset are compared.

- We are using Predictive Mean Matching (pmm) to impute.
    
## See missing data 

```{r}
ini <- mice(df, maxit=0)
ini$nmis
```


## Separate dataset regionally

```{r}
# We need to remove the flow variables so we don't have duplicated values for each region

df_flow <- df[,c("Date","Regions","SAC", "OUT",  "EXPORT", "logEXPORT","logOUT","lagOUT","logSAC","lagSAC")]

df_no_flow <- df[,!(names(df) %in% c("SAC", "logSAC", "OUT", "logOUT", "EXPORT", "logEXPORT","lagOUT", "lagSAC"))]

r <- unique(df_no_flow$Regions)
region_list <- list()
for (region in r){
  df_regional <- df_no_flow %>% filter(Regions == region)
  region <- gsub(" ","",region)
  region_list[[region]] <- df_regional
}
```

## Impute

Impute all of the regional datasets 5 times

```{r include=FALSE}

# This creates a list of mids objects

regional_imp <- list()
for (region_name in as.character(names(region_list))){
  current_df <- region_list[[region_name]]
  predictors <- quickpred(current_df)
  imp <- mice(current_df,seed = 123, print=F, predictorMatrix = predictors)
  new_name <- paste0(region_name, "_imp")
  regional_imp[[new_name]]<- imp
}
```

## Inspect convergence regionally {.tabset}

The plot shows the mean and the standard deviation of the imputed values only. Generally, we want to see the streams intermingle and be free of any trends at the later iterations. (https://www.gerkovink.com/miceVignettes/Convergence_pooling/Convergence_and_pooling.html)

### Central

```{r}
# Check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$Central_imp, maxit=35, print=F)
plot(imp40)
```


### Confluence

```{r}
#check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$Confluence_imp, maxit=35, print=F)
plot(imp40)
```


### North

```{r}
#check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$NorthDelta_imp, maxit=35, print=F)
plot(imp40)
```

### South

```{r}
#check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$South_imp, maxit=35, print=F)
plot(imp40)
```


### Suisun Bay

```{r}
#check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$SuisunBay_imp, maxit=35, print=F)
plot(imp40)
```


## Make combined imputed datasets

```{r}
# Create imputed datasets for each region
regional_imp_dfs <- list()
action_seq <- as.list(seq(1,5))
for (region_imp in as.character(names(regional_imp))){
      current_imp <- regional_imp[[region_imp]]
      for (i in action_seq){
        df_new <- complete(current_imp,action = i, include=FALSE)
        df_new <- left_join( df_new,df_flow, by=c("Date","Regions"))
        new_name <- paste0(region_imp, "_df", i)
        regional_imp_dfs[[new_name]]<- df_new}}

```


## Add lag terms to the imputed datasets

```{r}
# This code is taken from clamdata_cleanup_AB_261125.R

df_list_lag <- lapply(regional_imp_dfs, function(y) {
#Step 1: summarize by season 
wq_r_sum_lagseason <- y %>%
  group_by(Regions,Year,season) %>%
  summarize(across(where(is.numeric),\(x) mean(x, na.rm = TRUE)), .groups="drop_last")

#Step 2: Add a column for the following season.
#This seems counter-intuitive because the lag should be the prior season,
#but this is a merge column: it will merge back with the prior season
#in the original table. 
wq_r_sum_lagseason <- wq_r_sum_lagseason %>%
  select(!c(lagOUT,lagSAC,lagTemperature,lagDissAmmonia,lagDissNitrateNitrite, lagTotPhos,lagConductivity,lagDissOrthophos))%>%
  rename_with(~paste0("lag",.x)) %>%
  rename(Year = "lagYear",season = "lagseason",Regions="lagRegions") %>% 
  mutate(lagseason = case_when(
    season == "Spring" ~ "Summer",
    season == "Summer" ~ "Autumn",
    season == "Autumn" ~ "Winter",
    season == "Winter" ~ "Spring" 
  )) %>%
  mutate(lagseasonyear = case_when(
    season == "Autumn" ~ Year+1,
    !(season == "Autumn") ~ Year
  )) %>%
  mutate(lagseasonyear = paste0(lagseason,lagseasonyear)) %>%
  ungroup %>% 
  select(lagseasonyear, Regions, laglogDissAmmonia,laglogDissNitrateNitrite,laglogConductivity,
         laglogSAC,laglogOUT, laglogDissOrthophos, laglogEXPORT)


y <- y %>% 
  mutate(seasonyear = paste0(season,year(Date))) %>%
  merge(wq_r_sum_lagseason,by.x=c("Regions","seasonyear"),by.y=c("Regions","lagseasonyear"),all.x=T) %>%
  select(-seasonyear)%>%
  relocate(Date)
})

```

## Combine the imputed datasets

Now we have to combine the 25 regional datasets into 5 datasets that include all regions.This is also where we are z-scoring the datasets for better comparison of the predictors' relative effects.

```{r}
### df1
df1 <- distinct(rbind(df_list_lag$Central_imp_df1, df_list_lag$Confluence_imp_df1,
             df_list_lag$NorthDelta_imp_df1, df_list_lag$South_imp_df1,df_list_lag$SuisunBay_imp_df1)) %>%
  # z-score as the last step before the GAM
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

### df2
df2 <- distinct(rbind(df_list_lag$Central_imp_df2, df_list_lag$Confluence_imp_df2,
             df_list_lag$NorthDelta_imp_df2, df_list_lag$South_imp_df2,df_list_lag$SuisunBay_imp_df2))%>%
 # z-score 
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

### df3

df3 <- distinct(rbind(df_list_lag$Central_imp_df3, df_list_lag$Confluence_imp_df3,
             df_list_lag$NorthDelta_imp_df3, df_list_lag$South_imp_df3,df_list_lag$SuisunBay_imp_df3))%>%
  # z-score 
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

### df4
df4 <- distinct(rbind(df_list_lag$Central_imp_df4, df_list_lag$Confluence_imp_df4,
             df_list_lag$NorthDelta_imp_df4, df_list_lag$South_imp_df4,df_list_lag$SuisunBay_imp_df4))%>%
  # z-score 
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))


### df5

df5 <- distinct(rbind(df_list_lag$Central_imp_df5, df_list_lag$Confluence_imp_df5,
             df_list_lag$NorthDelta_imp_df5, df_list_lag$South_imp_df5,df_list_lag$SuisunBay_imp_df5))%>%
  # z- score
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

# Make regions a factor
df1$Regions <- as.factor(df1$Regions)
df2$Regions <- as.factor(df2$Regions)
df3$Regions <- as.factor(df3$Regions)
df4$Regions <- as.factor(df4$Regions)
df5$Regions <- as.factor(df5$Regions)

```


# Fit models with imputed datasets {.tabset}

## m3.po3

```{r}

m3.po4.imp1 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + s(OUT, by = season) + clam_step + s(OUT) + s(lagOUT) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df1, method = 'ML', select=TRUE)

m3.po4.imp2 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + s(OUT, by = season) + clam_step + s(OUT) + s(lagOUT) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df2, method = 'ML', select=TRUE)

m3.po4.imp3 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + s(OUT, by = season) + clam_step + s(OUT) + s(lagOUT) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df3, method = 'ML', select=TRUE)

m3.po4.imp4 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + s(OUT, by = season) + clam_step + s(OUT) + s(lagOUT) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df4, method = 'ML', select=TRUE)

m3.po4.imp5 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) +  s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + s(OUT, by = season) + clam_step + s(OUT) + s(lagOUT) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df5, method = 'ML', select=TRUE)
```

Fit again with log transformed terms

```{r}

m3.po4.imp1.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logOUT, by = season) + clam_step + s(logOUT) + s(laglogOUT) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df1, method = 'ML', select=TRUE)

m3.po4.imp2.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) +  s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logOUT, by = season) + clam_step + s(logOUT) + s(laglogOUT) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df2, method = 'ML', select=TRUE)

m3.po4.imp3.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logOUT, by = season) + clam_step + s(logOUT) + s(laglogOUT) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df3, method = 'ML', select=TRUE)

m3.po4.imp4.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite)  + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logOUT, by = season) + clam_step + s(logOUT) + s(laglogOUT) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df4, method = 'ML', select=TRUE)

m3.po4.imp5.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite)  + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logOUT, by = season) + clam_step + s(logOUT) + s(laglogOUT) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df5, method = 'ML', select=TRUE)
```

## m4.po4

```{r}
# fit m4.po4 with all 5 imputed datasets

m4.po4.imp1 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + s(SAC, by = season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df1, method = 'ML', select=TRUE)

m4.po4.imp2 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + s(SAC, by = season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df2, method = 'ML', select=TRUE)

m4.po4.imp3 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + s(SAC, by =season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df3, method = 'ML', select=TRUE)

m4.po4.imp4 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + s(SAC, by =season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df4, method = 'ML', select=TRUE)

m4.po4.imp5 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) +  s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + s(SAC, by =season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df5, method = 'ML', select=TRUE)
```

Fit again with log transformed terms

```{r}

m4.po4.imp1.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logSAC, by =season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df1, method = 'ML', select=TRUE)

m4.po4.imp2.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) +  s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logSAC, by =season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df2, method = 'ML', select=TRUE)

m4.po4.imp3.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logSAC, by =season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df3, method = 'ML', select=TRUE)

m4.po4.imp4.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite)  + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logSAC, by =season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df4, method = 'ML', select=TRUE)

m4.po4.imp5.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite)  + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logSAC, by =season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df5, method = 'ML', select=TRUE)
```


## m6.po4

```{r}
m6.po4.imp1 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + s(SAC, by =season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(EXPORT)+ s(lagDissOrthophos),data= df1, method = 'ML', select=TRUE)

m6.po4.imp2 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite)  + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + s(SAC, by =season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(EXPORT)+ s(lagDissOrthophos),data= df2, method = 'ML', select=TRUE)

m6.po4.imp3 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite)  + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + s(SAC, by =season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(EXPORT)+ s(lagDissOrthophos),data= df3, method = 'ML', select=TRUE)

m6.po4.imp4 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite)  + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + s(SAC, by =season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(EXPORT)+ s(lagDissOrthophos),data= df4, method = 'ML', select=TRUE)

m6.po4.imp5 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite)  + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + s(SAC, by =season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(EXPORT)+ s(lagDissOrthophos),data= df5, method = 'ML', select=TRUE)

```

Fit again with log transformations

```{r}
m6.po4.imp1.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logSAC, by =season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(EXPORT)+ s(laglogDissOrthophos),data= df1, method = 'ML', select=TRUE)

m6.po4.imp2.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logSAC, by =season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(EXPORT)+ s(laglogDissOrthophos),data= df2, method = 'ML', select=TRUE)

m6.po4.imp3.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite)+ s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logSAC, by =season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(EXPORT)+ s(laglogDissOrthophos),data= df3, method = 'ML', select=TRUE)

m6.po4.imp4.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logSAC, by =season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(EXPORT)+ s(laglogDissOrthophos),data= df4, method = 'ML', select=TRUE)

m6.po4.imp5.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logSAC, by = season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(EXPORT)+ s(laglogDissOrthophos),data= df5, method = 'ML', select=TRUE)

```

## Compare AIC/BIC

To see the comparison of AIC/BIC of all the models in the model selection process, see Model_Selection.Rmd

Here, I am comparing all of the imputed versions of the top 3 models with and without log transformed predictors.

```{r echo=FALSE}

aic_all<- AIC(m3.po4.imp1,m3.po4.imp2,m3.po4.imp3,m3.po4.imp4,m3.po4.imp5, m4.po4.imp1,m4.po4.imp2,m4.po4.imp3,m4.po4.imp4,m4.po4.imp5, m6.po4.imp1,m6.po4.imp2,m6.po4.imp3,m6.po4.imp4,m6.po4.imp5, m3.po4.imp1.log,m3.po4.imp2.log,m3.po4.imp3.log,m3.po4.imp4.log,m3.po4.imp5.log, m4.po4.imp1.log,m4.po4.imp2.log,m4.po4.imp3.log,m4.po4.imp4.log,m4.po4.imp5.log, m6.po4.imp1.log,m6.po4.imp2.log,m6.po4.imp3.log,m6.po4.imp4.log,m6.po4.imp5.log)

bic_all <- BIC(m3.po4.imp1,m3.po4.imp2,m3.po4.imp3,m3.po4.imp4,m3.po4.imp5, m4.po4.imp1,m4.po4.imp2,m4.po4.imp3,m4.po4.imp4,m4.po4.imp5, m6.po4.imp1,m6.po4.imp2,m6.po4.imp3,m6.po4.imp4,m6.po4.imp5, m3.po4.imp1.log,m3.po4.imp2.log,m3.po4.imp3.log,m3.po4.imp4.log,m3.po4.imp5.log, m4.po4.imp1.log,m4.po4.imp2.log,m4.po4.imp3.log,m4.po4.imp4.log,m4.po4.imp5.log, m6.po4.imp1.log,m6.po4.imp2.log,m6.po4.imp3.log,m6.po4.imp4.log,m6.po4.imp5.log)
```


This table ranks the models by their AIC and shows the difference in AIC/BIC values of each model from the lowest AIC/BIC of all models. The code is taken from here: https://interagencyecologicalprogram.github.io/ND-FASTR/rtm_chl_models_flow_weekly_avg.html#Purpose 


```{r}
# AIC values
aic_all_t <- aic_all %>%
  as_tibble(rownames = "Model") %>%
  mutate(Model_Number = 1:30, .before = Model)

# BIC values
bic_all_t <- bic_all %>%
  as_tibble(rownames = "Model")

# Combine AIC and BIC and calculate differences from lowest value
aic_bic_all <-
  left_join(aic_all_t, bic_all_t, by = join_by(Model, df)) %>%
  mutate(across(c(AIC, BIC), ~ .x - min(.x), .names = "{.col}_delta")) %>%
  select(starts_with("Model"), df, starts_with("AIC"), starts_with("BIC"))

# Sort by AIC
aic_bic_all %>% arrange(AIC)

```

# Plot predictions {.tabset}

## m3.po4 plotted predictions {.tabset}

```{r}
model_list_m3 = list(m3.po4.imp1,m3.po4.imp2,m3.po4.imp3,m3.po4.imp4,m3.po4.imp5)
```

### DissAmmonia

```{r}

lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "DissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### DissNitrateNitrite

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "DissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissNitrateNitrite

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "lagDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### Temperature

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "Temperature", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Conductivity

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "Conductivity", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### season

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "season", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Regions

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "Regions", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### clam_step

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "clam_step", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### OUT

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "OUT", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagOUT

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "lagOUT", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissAmmonia

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "lagDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### DissOrthophos

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "DissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissOrthophos

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "lagDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### OUT:season

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = c("OUT","season"), points = 0.5) +
      theme_classic(base_size = 12)
})
```


## Log m3.po4 plotted predictions {.tabset}

```{r}
model_list_m3_log = list(m3.po4.imp1.log,m3.po4.imp2.log,m3.po4.imp3.log,m3.po4.imp4.log,m3.po4.imp5.log)
```

### logDissAmmonia

```{r}

lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "logDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logDissNitrateNitrite

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "logDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogDissNitrateNitrite

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logConductivity

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "logConductivity", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logOUT

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "logOUT", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogOUT

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "laglogOUT", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogDissAmmonia

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logDissOrthophos

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "logDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogDissOrthophos

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Regions

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "Regions", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logOUT:season

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = c("logOUT","season"), points = 0.5) +
      theme_classic(base_size = 12)
})
```

## m4.po4 plotted predictions {.tabset}

```{r}
model_list_m4 = list(m4.po4.imp1,m4.po4.imp2,m4.po4.imp3,m4.po4.imp4,m4.po4.imp5)
```

### DissAmmonia

```{r}

lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "DissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### DissNitrateNitrite

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "DissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```
### lagDissNitrateNitrite

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "lagDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Temperature

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "Temperature", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Conductivity

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "Conductivity", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### season

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "season", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Regions

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "Regions", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### clam_step

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "clam_step", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### lagSAC

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "lagSAC", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissAmmonia

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "lagDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### DissOrthophos

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "DissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissOrthophos

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "lagDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### SAC:season

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = c("SAC","season"), points = 0.5) +
      theme_classic(base_size = 12)
})
```

## Log m4.po4 plotted predictions {.tabset}

```{r}
model_list_m4_log = list(m4.po4.imp1.log,m4.po4.imp2.log,m4.po4.imp3.log,m4.po4.imp4.log,m4.po4.imp5.log)
```

### logDissAmmonia

```{r}

lapply(model_list_m4_log, function(mod) {
    plot_predictions(mod, condition = "logDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logDissNitrateNitrite

```{r}
lapply(model_list_m4_log, function(mod) {
    plot_predictions(mod, condition = "logDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```
### laglogDissNitrateNitrite

```{r}
lapply(model_list_m4_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logConductivity

```{r}
lapply(model_list_m4_log, function(mod) {
    plot_predictions(mod, condition = "logConductivity", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogSAC

```{r}
lapply(model_list_m4_log, function(mod) {
    plot_predictions(mod, condition = "laglogSAC", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### loglagDissAmmonia

```{r}
lapply(model_list_m4_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logDissOrthophos

```{r}
lapply(model_list_m4_log, function(mod) {
    plot_predictions(mod, condition = "logDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogDissOrthophos

```{r}
lapply(model_list_m4_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### logSAC:season

```{r}
lapply(model_list_m4_log, function(mod) {
    plot_predictions(mod, condition = c("logSAC","season"), points = 0.5) +
      theme_classic(base_size = 12)
})
```

## m6.po4 plotted predicitons {.tabset}

```{r}
model_list_m6 = list(m6.po4.imp1,m6.po4.imp2,m6.po4.imp3,m6.po4.imp4,m6.po4.imp5)
```

### DissAmmonia

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "DissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### DissNitrateNitrite

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "DissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissNitrateNitrite

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "lagDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Temperature

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "Temperature", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Conductivity

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "Conductivity", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### season

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "season", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Regions

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "Regions", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### clam_step

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "clam_step", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagSAC

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "lagSAC", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### EXPORT

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "EXPORT", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissAmmonia

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "lagDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### DissOrthophos

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "DissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissOrthophos

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "lagDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### SAC:season

```{r}
lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = c("SAC","season"), points = 0.5) +
      theme_classic(base_size = 12)
})
```

## Log m6.po4 plotted predicitons {.tabset}

```{r}
model_list_m6_log = list(m6.po4.imp1.log,m6.po4.imp2.log,m6.po4.imp3.log,m6.po4.imp4.log,m6.po4.imp5.log)
```


### logDissAmmonia

```{r}

lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = "logDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

###logDissNitrateNitrite

```{r}

lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = "logDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogDissNitrateNitrite

```{r}

lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logConductivity

```{r}

lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = "logConductivity", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogSAC

```{r}

lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = "laglogSAC", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### EXPORT

```{r}

lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = "EXPORT", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogDissAmmonia

```{r}

lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logDissOrthophos

```{r}

lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = "logDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogDissOrthophos

```{r}

lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logSAC:season

```{r}
lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = c("logSAC","season"), points = 0.5) +
      theme_classic(base_size = 12)
})
```

# Plot Marginal Effects

Only plotting the marginal effects of the first model

## m3.po4 marginal effects

Plotting marginal effects for first imputed dataset

```{r}
c <- c("logDissAmmonia",
          "laglogDissAmmonia",
          "logDissNitrateNitrite",
          "laglogDissNitrateNitrite",
          "Temperature",
          "logConductivity",
          "logOUT",
          "laglogOUT")

lapply(c, function(x) {
  plot_slopes(m3.po4.imp1.log, variables = x, condition = x)
})

```

## m4.po4 marginal effects

Plotting marginal effects for first imputed dataset

```{r}
c <- c("logDissAmmonia",
          "laglogDissAmmonia",
          "logDissNitrateNitrite",
          "laglogDissNitrateNitrite",
          "Temperature",
          "logConductivity",
          "laglogSAC")

lapply(c, function(x) {
  plot_slopes(m4.po4.imp1.log, variables = x, condition = x)
})

```


## m6.po4 marginal effects

Plotting marginal effects for first imputed dataset

```{r}
c <- c("logDissAmmonia",
          "laglogDissAmmonia",
          "logDissNitrateNitrite",
          "laglogDissNitrateNitrite",
          "Temperature",
          "logConductivity",
          "laglogSAC",
           "EXPORT")

lapply(c, function(x) {
  plot_slopes(m6.po4.imp1.log, variables = x, condition = x)
})

```

# Results {.tabset}

## m3.po4  {.tabset}

### GAM check

```{r}
all_diagnostics <- sapply(model_list_m3_log, function(x) gam.check(x))
```

### Summary

```{r}
summary(m3.po4.imp1.log)
```

## m4.po4 {.tabset}

### GAM check

```{r}
all_diagnostics <- sapply(model_list_m4_log, function(x) gam.check(x))
```

### Summary

```{r}
summary(m4.po4.imp1.log)
```

## m6.po4 {.tabset}

### GAM check

```{r}
all_diagnostics <- sapply(model_list_m6_log, function(x) gam.check(x))
```

### Summary

```{r}
summary(m6.po4.imp1.log)
```

# Remove flow:season interaction 

```{r}
m3.po4.imp1.log.nofs <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + clam_step + s(logOUT) + s(laglogOUT) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df1, method = 'ML', select=TRUE)

m4.po4.imp1.log.nofs <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df1, method = 'ML', select=TRUE)

m6.po4.imp1.log.nofs <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(EXPORT)+ s(laglogDissOrthophos),data= df1, method = 'ML', select=TRUE)

m3.po4.imp1.nofs <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + clam_step + s(OUT) + s(lagOUT) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df1, method = 'ML', select=TRUE)

m4.po4.imp1.nofs <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df1, method = 'ML', select=TRUE)

m6.po4.imp1.nofs <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(EXPORT)+ s(lagDissOrthophos),data= df1, method = 'ML', select=TRUE)
```

## Compare AIC/BIC with original models 

We are interested in what effect removing the flow:season interaction has on the performance of the models. I replicated the top models without the flow:season interaction (denoted by ".nofs"), so we can compare them to our original models.


```{r echo= FALSE}

aic_all<- AIC(m3.po4.imp1.log.nofs , m4.po4.imp1.log.nofs, m6.po4.imp1.log.nofs,m3.po4.imp1.log, m4.po4.imp1.log, m6.po4.imp1.log, m3.po4.imp1.nofs , m4.po4.imp1.nofs, m6.po4.imp1.nofs, m3.po4.imp1, m4.po4.imp1,m6.po4.imp1)

bic_all <- BIC(m3.po4.imp1.log.nofs , m4.po4.imp1.log.nofs, m6.po4.imp1.log.nofs,m3.po4.imp1.log, m4.po4.imp1.log,m6.po4.imp1.log,m3.po4.imp1.nofs , m4.po4.imp1.nofs, m6.po4.imp1.nofs,m3.po4.imp1, m4.po4.imp1,m6.po4.imp1)
```


This table ranks the models by their AIC and shows the difference in AIC/BIC values of each model from the lowest AIC/BIC of all models. The code is taken from here: https://interagencyecologicalprogram.github.io/ND-FASTR/rtm_chl_models_flow_weekly_avg.html#Purpose 


```{r}
# AIC values
aic_all_t <- aic_all %>%
  as_tibble(rownames = "Model")

# BIC values
bic_all_t <- bic_all %>%
  as_tibble(rownames = "Model")

# Combine AIC and BIC and calculate differences from lowest value
aic_bic_all <-
  left_join(aic_all_t, bic_all_t, by = join_by(Model, df)) %>%
  mutate(across(c(AIC, BIC), ~ .x - min(.x), .names = "{.col}_delta")) %>%
  select(starts_with("Model"), df, starts_with("AIC"), starts_with("BIC"))

# Sort by AIC
aic_bic_all %>% arrange(AIC)

```

# Resources

https://gsp.humboldt.edu/olm/R/05_03_GAM.html

https://www.geeksforgeeks.org/r-language/generalized-additive-models-using-r/

https://019b2da8-edfb-a262-61be-7973c056d9ae.share.connect.posit.cloud/mlr-collinearity.html 

GAM interpretation:
https://ecogambler.netlify.app/blog/interpreting-gams/

NA exclude
https://stats.stackexchange.com/questions/492955/should-i-use-na-omit-or-na-exclude-in-a-linear-model-in-r 
