---
title:
  "Global GAM"
author:
  "Ryan Hankins"
output: 
  html_document:
        toc: TRUE
        toc_float: TRUE
        code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, tidy = "styler")
```

# Prepare Data

```{r, include=FALSE}

library(mgcv)
library(tidyverse)
library(marginaleffects)
library(MuMIn)
library(car)
library(corrplot)
library(stats)
library(marginaleffects)
library(mice)
library(lattice)
library(ggplot2)
library(data.table)
```


```{r}
#options(na.action = "na.exclude")

# variables to include:
c = c("Chlorophyll","logchla", "Regions", "Date","Year", "Month",  "season","SAC","lagSAC",
       "EXPORT","logEXPORT","OUT","lagOUT", "DissAmmonia",
       "lagDissAmmonia", "TotPhos","lagTotPhos","DissNitrateNitrite", 
       "lagDissNitrateNitrite","Temperature","lagTemperature","Secchi", 
       "lagSecchi","Conductivity","lagConductivity","season_num", "DissOrthophos", "lagDissOrthophos", "clam_step","logDissNitrateNitrite","logDissAmmonia","logDissOrthophos","logConductivity","logSAC","logOUT")#, "Biomass","Grazing_rate","Filtration_rate", "SJR","lagSJR", "Index","lagIndex")

df <- read.csv("Data/regional_integrated_dataset3.csv" )

df$Date = as.Date(df$Date)

# 1979 is the start of the study period.
df <- subset(df,df$Date >= "1979-01-01")
df <- subset(df,df$Date <= "2022-12-31")

# Changing categorical variables to factors
# need to change regions to factor farther down so dataframe splitting by region works properly
#df$Regions = as.factor(df$Regions)
df$season = as.factor(df$season)
df$season_num = as.numeric(df$season)


#Replacing clam variables with a step function (pre/post potamocorbula introduction to the Bay in 1987)
df <- df%>%mutate(clam_step=if_else(Year>1987, "post clam", "pre clam")) 
df$clam_step <- as.factor(df$clam_step)

# log transforming chl-a to achieve normal distribution of response variable.
df$logchla <- log(df$Chlorophyll)

# log transforming some variables to better visualize
df$logDissNitrateNitrite <- log(df$DissNitrateNitrite)
df$logDissAmmonia <- log(df$DissAmmonia)
df$logDissOrthophos <- log(df$DissOrthophos)
df$logConductivity <- log(df$Conductivity)
df$logSAC <- log(df$SAC)
df$logOUT <- log(df$OUT)
df$logEXPORT <- log(df$EXPORT)

# Dropping NAs from the dataset, only for dredging.

 df <- subset(df, select = c) #%>%
    #drop_na()

```


# Impute

We are using Multiple Imputations by Chained Equations (MICE) to impute missing data.

- Transforming the data first and then imputing is the preferred strategy.
(https://019b2da8-edfb-a262-61be-7973c056d9ae.share.connect.posit.cloud/mi-fitting.html#ref-von_hippel_2009)

- See this link for a graphic of the imputation process: https://stackoverflow.com/questions/54075836/multiple-imputed-datasets-pooling-results

- Multiple imputed versions of the dataset are created, then each imputed dataset is analysed with the GAM, then results from each of the GAMs are "pooled" together.

- Using Predictive Mean Matching (pmm)

- Pooling median p-value (MPV) is useful for analyzing pooled GAMs 
      - "take the median of all imputed within-imputation GAM p-values as the pooled p-value measuring the strength of evidence for the association." https://pmc.ncbi.nlm.nih.gov/articles/PMC9123297/ 

- Need to Regionally separate data before imputing, then impute each dataset, and then combine into imputed datasets.

Questions:
  - what are the special considerations for using MICE with GAMs?
    
## See Missing Data 

```{r}
ini <- mice(df, maxit=0)
ini$nmis
```

```{r}
md.pattern(df)
```

## Separate Dataset Regionally

```{r}
#we need to remove the flow variables so we don't have duplicated values for each region
df_flow <- df[,c("Date","Regions","SAC", "OUT",  "EXPORT", "logEXPORT","logOUT","lagOUT","logSAC","lagSAC")]

df_no_flow <- df[,!(names(df) %in% c("SAC", "logSAC", "OUT", "logOUT", "EXPORT", "logEXPORT","lagOUT", "lagSAC"))]

df_Central <- df_no_flow %>% filter(Regions == "Central")

r <- unique(df_no_flow$Regions)
region_list <- list()
for (region in r){
  df_regional <- df_no_flow %>% filter(Regions == region)
  region <- gsub(" ","",region)
  region_list[[region]] <- df_regional
}
```

## Impute

Impute all of the regional datasets 5 times

```{r include=FALSE}

# This creates a list of mids objects
regional_imp <- list()
for (region_name in as.character(names(region_list))){
  current_df <- region_list[[region_name]]
  predictors <- quickpred(current_df)
  imp <- mice(current_df,seed = 123, print=F, predictorMatrix = predictors)
  new_name <- paste0(region_name, "_imp")
  regional_imp[[new_name]]<- imp
}
```

## inspect convergence regionally {.tabset}

The plot shows the mean and the standard deviation of the imputed values only. Generally want to see the streams intermingle and be free of any trends at the later iterations. (https://www.gerkovink.com/miceVignettes/Convergence_pooling/Convergence_and_pooling.html)

### Central

```{r}
#check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$Central_imp, maxit=35, print=F)
plot(imp40)
```
Check imputation values

```{r}
stripplot(regional_imp$Central_imp)
```

### Confluence

```{r}
#check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$Confluence_imp, maxit=35, print=F)
plot(imp40)
```

Check imputation values

```{r}
stripplot(regional_imp$Confluence_imp)
```

### North

```{r}
#check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$NorthDelta_imp, maxit=35, print=F)
plot(imp40)
```
Check imputation values

```{r}
stripplot(regional_imp$NorthDelta_imp)
```
### South

```{r}
#check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$South_imp, maxit=35, print=F)
plot(imp40)
```

Check imputation values

```{r}
stripplot(regional_imp$South_imp)
```

### Suisun Bay

```{r}
#check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$SuisunBay_imp, maxit=35, print=F)
plot(imp40)
```

Check imputation values

```{r}
stripplot(regional_imp$SuisunBay_imp)
```


## Make Combined imputed datasets

```{r}
# Create imputed datasets for each region
regional_imp_dfs <- list()
action_seq <- as.list(seq(1,5))
for (region_imp in as.character(names(regional_imp))){
      current_imp <- regional_imp[[region_imp]]
      for (i in action_seq){
        df_new <- complete(current_imp,action = i, include=FALSE)
        df_new <- left_join( df_new,df_flow, by=c("Date","Regions"))
        new_name <- paste0(region_imp, "_df", i)
        regional_imp_dfs[[new_name]]<- df_new}}

```


## Add Lag terms to the imputed datasets

```{r}

df_list_lag <- lapply(regional_imp_dfs, function(y) {
#Step 1: summarize by season 
wq_r_sum_lagseason <- y %>%
  group_by(Regions,Year,season) %>%
  summarize(across(where(is.numeric),\(x) mean(x, na.rm = TRUE)), .groups="drop_last")

#Step 2: Add a column for the following season.
#This seems counter-intuitive because the lag should be the prior season,
#but this is a merge column: it will merge back with the prior season
#in the original table. 
wq_r_sum_lagseason <- wq_r_sum_lagseason %>%
  select(!c(lagOUT,lagSAC,lagTemperature,lagDissAmmonia,lagDissNitrateNitrite, lagTotPhos,lagConductivity,lagDissOrthophos))%>%
  rename_with(~paste0("lag",.x)) %>%
  rename(Year = "lagYear",season = "lagseason",Regions="lagRegions") %>% 
  mutate(lagseason = case_when(
    season == "Spring" ~ "Summer",
    season == "Summer" ~ "Autumn",
    season == "Autumn" ~ "Winter",
    season == "Winter" ~ "Spring" 
  )) %>%
  mutate(lagseasonyear = case_when(
    season == "Autumn" ~ Year+1,
    !(season == "Autumn") ~ Year
  )) %>%
  mutate(lagseasonyear = paste0(lagseason,lagseasonyear)) %>%
  ungroup %>% 
  select(lagseasonyear, Regions, laglogDissAmmonia,laglogDissNitrateNitrite,laglogConductivity,
         laglogSAC,laglogOUT, laglogDissOrthophos, laglogEXPORT)


y <- y %>% 
  mutate(seasonyear = paste0(season,year(Date))) %>%
  merge(wq_r_sum_lagseason,by.x=c("Regions","seasonyear"),by.y=c("Regions","lagseasonyear"),all.x=T) %>%
  select(-seasonyear)%>%
  relocate(Date)
})

```

## Combine the imputed datasets

Now we have to combine the 25 regional datasets into 5 datasets that include all regions.

```{r}
### df1
df1 <- distinct(rbind(df_list_lag$Central_imp_df1, df_list_lag$Confluence_imp_df1,
             df_list_lag$NorthDelta_imp_df1, df_list_lag$South_imp_df1,df_list_lag$SuisunBay_imp_df1)) %>%
  # z-score as the last step before the GAM
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

### df2
df2 <- distinct(rbind(df_list_lag$Central_imp_df2, df_list_lag$Confluence_imp_df2,
             df_list_lag$NorthDelta_imp_df2, df_list_lag$South_imp_df2,df_list_lag$SuisunBay_imp_df2))%>%
 # z-score 
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

### df3

df3 <- distinct(rbind(df_list_lag$Central_imp_df3, df_list_lag$Confluence_imp_df3,
             df_list_lag$NorthDelta_imp_df3, df_list_lag$South_imp_df3,df_list_lag$SuisunBay_imp_df3))%>%
  # z-score 
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

###df4
df4 <- distinct(rbind(df_list_lag$Central_imp_df4, df_list_lag$Confluence_imp_df4,
             df_list_lag$NorthDelta_imp_df4, df_list_lag$South_imp_df4,df_list_lag$SuisunBay_imp_df4))%>%
  # z-score 
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))


###df5

df5 <- distinct(rbind(df_list_lag$Central_imp_df5, df_list_lag$Confluence_imp_df5,
             df_list_lag$NorthDelta_imp_df5, df_list_lag$South_imp_df5,df_list_lag$SuisunBay_imp_df5))%>%
  # z- score
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

# Make regions a factor
df1$Regions <- as.factor(df1$Regions)
df2$Regions <- as.factor(df2$Regions)
df3$Regions <- as.factor(df3$Regions)
df4$Regions <- as.factor(df4$Regions)
df5$Regions <- as.factor(df5$Regions)

```


# Fit models with Imputed datasets {.tabset}

## m3.po3

```{r}

m3.po4.imp1 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + (OUT*season) + clam_step + s(OUT) + s(lagOUT) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df1, method = 'ML', select=TRUE)

m3.po4.imp2 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + (OUT*season) + clam_step + s(OUT) + s(lagOUT) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df2, method = 'ML', select=TRUE)

m3.po4.imp3 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + (OUT*season) + clam_step + s(OUT) + s(lagOUT) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df3, method = 'ML', select=TRUE)

m3.po4.imp4 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + (OUT*season) + clam_step + s(OUT) + s(lagOUT) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df4, method = 'ML', select=TRUE)

m3.po4.imp5 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) +  s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + (OUT*season) + clam_step + s(OUT) + s(lagOUT) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df5, method = 'ML', select=TRUE)
```

fit again with log transformed terms

```{r}

m3.po4.imp1.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + (logOUT*season) + clam_step + s(logOUT) + s(laglogOUT) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df1, method = 'ML', select=TRUE)

m3.po4.imp2.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) +  s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + (logOUT*season) + clam_step + s(logOUT) + s(laglogOUT) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df2, method = 'ML', select=TRUE)

m3.po4.imp3.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + (logOUT*season) + clam_step + s(logOUT) + s(laglogOUT) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df3, method = 'ML', select=TRUE)

m3.po4.imp4.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite)  + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + (logOUT*season) + clam_step + s(logOUT) + s(laglogOUT) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df4, method = 'ML', select=TRUE)

m3.po4.imp5.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite)  + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + (logOUT*season) + clam_step + s(logOUT) + s(laglogOUT) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df5, method = 'ML', select=TRUE)
```

## m4.po4

```{r}
# fit m4.po4 with all 5 imputed datasets

#removing SAC here because it was shrunk to 0. Should SAC*season also be removed?

m4.po4.imp1 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df1, method = 'ML', select=TRUE)

m4.po4.imp2 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df2, method = 'ML', select=TRUE)

m4.po4.imp3 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df3, method = 'ML', select=TRUE)

m4.po4.imp4 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df4, method = 'ML', select=TRUE)

m4.po4.imp5 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) +  s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df5, method = 'ML', select=TRUE)
```

fit again with log transformed terms

```{r}

m4.po4.imp1.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + (logSAC*season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df1, method = 'ML', select=TRUE)

m4.po4.imp2.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) +  s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + (logSAC*season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df2, method = 'ML', select=TRUE)

m4.po4.imp3.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + (logSAC*season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df3, method = 'ML', select=TRUE)

m4.po4.imp4.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite)  + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + (logSAC*season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df4, method = 'ML', select=TRUE)

m4.po4.imp5.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite)  + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + (logSAC*season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(laglogDissOrthophos), data=df5, method = 'ML', select=TRUE)
```


## m6.po4

```{r}
m6.po4.imp1 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(EXPORT)+ s(lagDissOrthophos),data= df1, method = 'ML', select=TRUE)

m6.po4.imp2 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite)  + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(EXPORT)+ s(lagDissOrthophos),data= df2, method = 'ML', select=TRUE)

m6.po4.imp3 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite)  + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(EXPORT)+ s(lagDissOrthophos),data= df3, method = 'ML', select=TRUE)

m6.po4.imp4 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite)  + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(EXPORT)+ s(lagDissOrthophos),data= df4, method = 'ML', select=TRUE)

m6.po4.imp5 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite)  + s(lagDissNitrateNitrite) + s(Temperature) + s(Conductivity) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(EXPORT)+ s(lagDissOrthophos),data= df5, method = 'ML', select=TRUE)

```

fit again with log transformations


m6.po4

```{r}
m6.po4.imp1.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + (logSAC*season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(EXPORT)+ s(laglogDissOrthophos),data= df1, method = 'ML', select=TRUE)

m6.po4.imp2.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + (logSAC*season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(EXPORT)+ s(laglogDissOrthophos),data= df2, method = 'ML', select=TRUE)

m6.po4.imp3.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite)+ s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + (logSAC*season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(EXPORT)+ s(laglogDissOrthophos),data= df3, method = 'ML', select=TRUE)

m6.po4.imp4.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + (logSAC*season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(EXPORT)+ s(laglogDissOrthophos),data= df4, method = 'ML', select=TRUE)

m6.po4.imp5.log <- gam((logchla) ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(laglogDissNitrateNitrite) + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + (logSAC*season) + clam_step + s(laglogSAC) + s(laglogDissAmmonia)+ s(logDissOrthophos)+ s(EXPORT)+ s(laglogDissOrthophos),data= df5, method = 'ML', select=TRUE)

```

## Compare AIC/BIC

To see the comparison of AIC/BIC of all the models in the model selection process, see Model_Selection.Rmd


```{r}
aic_all<- AIC(m3.po4.imp1,m3.po4.imp2,m3.po4.imp3,m3.po4.imp4,m3.po4.imp5, m4.po4.imp1,m4.po4.imp2,m4.po4.imp3,m4.po4.imp4,m4.po4.imp5, m6.po4.imp1,m6.po4.imp2,m6.po4.imp3,m6.po4.imp4,m6.po4.imp5)
```
```{r}
bic_all <- BIC(m3.po4.imp1,m3.po4.imp2,m3.po4.imp3,m3.po4.imp4,m3.po4.imp5, m4.po4.imp1,m4.po4.imp2,m4.po4.imp3,m4.po4.imp4,m4.po4.imp5, m6.po4.imp1,m6.po4.imp2,m6.po4.imp3,m6.po4.imp4,m6.po4.imp5)
```


This table ranks the models by their AIC and shows the difference in AIC/BIC values of each model from the lowest AIC/BIC of all models. The code is taken from here: https://interagencyecologicalprogram.github.io/ND-FASTR/rtm_chl_models_flow_weekly_avg.html#Purpose 

m6.po4 and m4.po4 look to be performing the best

```{r}
# AIC values
aic_all_t <- aic_all %>%
  as_tibble(rownames = "Model") %>%
  mutate(Model_Number = 1:15, .before = Model)

# BIC values
bic_all_t <- bic_all %>%
  as_tibble(rownames = "Model")

# Combine AIC and BIC and calculate differences from lowest value
aic_bic_all <-
  left_join(aic_all_t, bic_all_t, by = join_by(Model, df)) %>%
  mutate(across(c(AIC, BIC), ~ .x - min(.x), .names = "{.col}_delta")) %>%
  select(starts_with("Model"), df, starts_with("AIC"), starts_with("BIC"))

# Sort by AIC
aic_bic_all %>% arrange(AIC)

```

# Plot Predictions {.tabset}

## m3.po4 plotted predictions {.tabset}

```{r}
model_list_m3 = list(m3.po4.imp1,m3.po4.imp2,m3.po4.imp3,m3.po4.imp4,m3.po4.imp5)
```

### DissAmmonia

```{r}

lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "DissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### DissNitrateNitrite

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "DissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissNitrateNitrite

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "lagDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### Temperature

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "Temperature", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Conductivity

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "Conductivity", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### season

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "season", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Regions

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "Regions", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### clam_step

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "clam_step", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### OUT

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "OUT", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagOUT

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "lagOUT", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissAmmonia

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "lagDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### DissOrthophos

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "DissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissOrthophos

```{r}
lapply(model_list_m3, function(mod) {
    plot_predictions(mod, condition = "lagDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

## Log m3.po4 plotted predictions {.tabset}

```{r}
model_list_m3_log = list(m3.po4.imp1.log,m3.po4.imp2.log,m3.po4.imp3.log,m3.po4.imp4.log,m3.po4.imp5.log)
```

### logDissAmmonia

```{r}

lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "logDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logDissNitrateNitrite

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "logDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogDissNitrateNitrite

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logConductivity

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "logConductivity", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logOUT

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "logOUT", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogOUT

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "laglogOUT", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### loglagDissAmmonia

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logDissOrthophos

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "logDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogDissOrthophos

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Regions

```{r}
lapply(model_list_m3_log, function(mod) {
    plot_predictions(mod, condition = "Regions", points = 0.5) +
      theme_classic(base_size = 12)
})
```

## m4.po4 plotted predictions {.tabset}

```{r}
model_list_m4 = list(m4.po4.imp1,m4.po4.imp2,m4.po4.imp3,m4.po4.imp4,m4.po4.imp5)
```

### DissAmmonia

```{r}

lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "DissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### DissNitrateNitrite

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "DissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```
### lagDissNitrateNitrite

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "lagDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Temperature

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "Temperature", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Conductivity

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "Conductivity", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### season

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "season", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Regions

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "Regions", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### clam_step

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "clam_step", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### lagSAC

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "lagSAC", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissAmmonia

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "lagDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### DissOrthophos

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "DissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissOrthophos

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "lagDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

## Log m4.po4 plotted predictions {.tabset}

```{r}
model_list_m4_log = list(m4.po4.imp1.log,m4.po4.imp2.log,m4.po4.imp3.log,m4.po4.imp4.log,m4.po4.imp5.log)
```

### logDissAmmonia

```{r}

lapply(model_list_m4_log, function(mod) {
    plot_predictions(mod, condition = "logDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logDissNitrateNitrite

```{r}
lapply(model_list_m4_log, function(mod) {
    plot_predictions(mod, condition = "logDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```
### laglogDissNitrateNitrite

```{r}
lapply(model_list_m4_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logConductivity

```{r}
lapply(model_list_m4_log, function(mod) {
    plot_predictions(mod, condition = "logConductivity", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogSAC

```{r}
lapply(model_list_m4_log, function(mod) {
    plot_predictions(mod, condition = "laglogSAC", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### loglagDissAmmonia

```{r}
lapply(model_list_m4_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logDissOrthophos

```{r}
lapply(model_list_m4_log, function(mod) {
    plot_predictions(mod, condition = "logDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogDissOrthophos

```{r}
lapply(model_list_m4_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

## m6.po4 plotted predicitons {.tabset}

```{r}
model_list_m6 = list(m6.po4.imp1,m6.po4.imp2,m6.po4.imp3,m6.po4.imp4,m6.po4.imp5)
```

### DissAmmonia

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "DissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### DissNitrateNitrite

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "DissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissNitrateNitrite

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "lagDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Temperature

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "Temperature", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Conductivity

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "Conductivity", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### season

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "season", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Regions

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "Regions", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### clam_step

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "clam_step", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagSAC

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "lagSAC", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### EXPORT

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "EXPORT", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissAmmonia

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "lagDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### DissOrthophos

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "DissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissOrthophos

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "lagDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

## Log m6.po4 plotted predicitons {.tabset}

```{r}
model_list_m6_log = list(m6.po4.imp1.log,m6.po4.imp2.log,m6.po4.imp3.log,m6.po4.imp4.log,m6.po4.imp5.log)
```


### logDissAmmonia

```{r}

lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = "logDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

###logDissNitrateNitrite

```{r}

lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = "logDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogDissNitrateNitrite

```{r}

lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logConductivity

```{r}

lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = "logConductivity", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogSAC

```{r}

lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = "laglogSAC", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogDissAmmonia

```{r}

lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### logDissOrthophos

```{r}

lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = "logDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### laglogDissOrthophos

```{r}

lapply(model_list_m6_log, function(mod) {
    plot_predictions(mod, condition = "laglogDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

# Plot Marginal Effects

Only plotting the marginal effects of the first model

## m3.po4 marginal effects

Plotting marginal effects for first imputed dataset

```{r}
c <- c("logDissAmmonia",
          "laglogDissAmmonia",
          "logDissNitrateNitrite",
          "laglogDissNitrateNitrite",
          "Temperature",
          "logConductivity",
          "logOUT",
          "laglogOUT")

lapply(c, function(x) {
  plot_slopes(m3.po4.imp1.log, variables = x, condition = x)
})

```

## m4.po4 marginal effects

Plotting marginal effects for first imputed dataset

```{r}
c <- c("logDissAmmonia",
          "laglogDissAmmonia",
          "logDissNitrateNitrite",
          "laglogDissNitrateNitrite",
          "Temperature",
          "logConductivity",
          "laglogSAC")

lapply(c, function(x) {
  plot_slopes(m4.po4.imp1.log, variables = x, condition = x)
})

```


## m6.po4 marginal effects

Plotting marginal effects for first imputed dataset

```{r}
c <- c("logDissAmmonia",
          "laglogDissAmmonia",
          "logDissNitrateNitrite",
          "laglogDissNitrateNitrite",
          "Temperature",
          "logConductivity",
          "laglogSAC",
           "EXPORT")

lapply(c, function(x) {
  plot_slopes(m6.po4.imp1.log, variables = x, condition = x)
})

```

# Results {.tabset}

For information about how to pool results to numerically evaluate relationships described in a GAM fitted with MI datasets, see this paper about the MPV rule:
https://pmc.ncbi.nlm.nih.gov/articles/PMC9123297/#Sec1

## m3.po4  {.tabset}

### GAM check

```{r}
all_diagnostics <- sapply(model_list_m3_log, function(x) gam.check(x))
```

### Summary

```{r}
summary(m3.po4.imp1.log)
```

## m4.po4 {.tabset}

### GAM check

```{r}
all_diagnostics <- sapply(model_list_m4_log, function(x) gam.check(x))
```

### Summary

```{r}
summary(m4.po4.imp1.log)
```

## m6.po4 {.tabset}

### GAM check

```{r}
all_diagnostics <- sapply(model_list_m6_log, function(x) gam.check(x))
```

### Summary

```{r}
summary(m6.po4.imp1.log)
```

# Resources
https://gsp.humboldt.edu/olm/R/05_03_GAM.html

https://www.geeksforgeeks.org/r-language/generalized-additive-models-using-r/

GIF^(1/(2DF)) >2 - moderate collinearity, GIF>3.16 = high collinearity
https://019b2da8-edfb-a262-61be-7973c056d9ae.share.connect.posit.cloud/mlr-collinearity.html

GAM interpretation:
https://ecogambler.netlify.app/blog/interpreting-gams/

NA exclude
https://stats.stackexchange.com/questions/492955/should-i-use-na-omit-or-na-exclude-in-a-linear-model-in-r 
