---
title:
  "Global GAM SHORT"
author:
  "Ryan Hankins"
output: 
  html_document:
        toc: TRUE
        toc_float: TRUE
        code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, tidy = "styler")
```
# Background on this RMD file

This file continues the work in Global_GAM_RH.RMD. Here, I remove outliers and change nutrient terms to linear. I also only work with the top imputed model, and do not repeat any of the imputation diagnostic procedures. 

# Purpose

The purpose of these models is to understand the effects of environmental drivers and antecedent conditions on Delta-wide phytoplankton biomass (chl-a). 

# Prepare Data

```{r, include=FALSE}

library(mgcv)
library(tidyverse)
library(marginaleffects)
library(MuMIn)
library(corrplot)
library(stats)
library(mice)
library(lattice)
library(ggplot2)
library(data.table)
```


```{r}

# variables to include in the analysis:
c = c("Chlorophyll","logchla", "Regions", "Date","Year", "Month",  "season","SAC","lagSAC",
       "EXPORT","logEXPORT","OUT","lagOUT", "DissAmmonia",
       "lagDissAmmonia", "TotPhos","lagTotPhos","DissNitrateNitrite", 
       "lagDissNitrateNitrite","Temperature","lagTemperature","Secchi", 
       "lagSecchi","Conductivity","lagConductivity","season_num", "DissOrthophos", "lagDissOrthophos", "clam_step","logDissNitrateNitrite","logDissAmmonia","logDissOrthophos","logConductivity","logSAC","logOUT")#, "Biomass","Grazing_rate","Filtration_rate", "SJR","lagSJR", "Index","lagIndex")

df <- read.csv("Data/regional_integrated_dataset3.csv" )
df$Date = as.Date(df$Date)

# 1979 is the start of the study period.
df <- subset(df,df$Date >= "1979-01-01")
# 2022 is the end of the study period
df <- subset(df,df$Date <= "2022-12-31")

# Changing categorical variables to factors
# need to change regions to factor farther down so dataframe splitting by region works properly
#df$Regions = as.factor(df$Regions)
df$season = as.factor(df$season)
df$season_num = as.numeric(df$season)


# Replacing clam variables with a step function (pre/post potamocorbula introduction to the Bay in 1987)
df <- df%>%mutate(clam_step=if_else(Year>1987, "post clam", "pre clam")) 
df$clam_step <- as.factor(df$clam_step)

# log transforming chl-a to achieve normal distribution of response variable.
df$logchla <- log(df$Chlorophyll)

# log transforming some variables to better visualize
df$logDissNitrateNitrite <- log(df$DissNitrateNitrite)
df$logDissAmmonia <- log(df$DissAmmonia)
df$logDissOrthophos <- log(df$DissOrthophos)
df$logConductivity <- log(df$Conductivity)
df$logSAC <- log(df$SAC)
df$logOUT <- log(df$OUT)
df$logEXPORT <- log(df$EXPORT)


df <- subset(df, select = c)

```


## Remove outliers {.tabset}

### Investigating problem variables

```{r}
histogram(df$logDissAmmonia)
boxplot(df$logDissAmmonia)
histogram(df$logDissOrthophos)
boxplot(df$logDissOrthophos)
histogram(df$logDissNitrateNitrite)
boxplot(df$logDissNitrateNitrite)
``` 

### Remove outliers based on z-score

removing outliers based on z-score, assuming the log transformations of the targeted variables have a normal distribution.

-3 - 3 is a commonly used range beyond which values are considered outliers.
https://statsandr.com/blog/outliers-detection-in-r/#z-scores 


```{r}

clean_df_z <- df 

clean_df_z$logDissAmmonia[abs(scale(clean_df_z$logDissAmmonia)) > 3.29] <- NA

clean_df_z$logDissNitrateNitrite[abs(scale(clean_df_z$logDissNitrateNitrite)) > 3.29] <- NA

clean_df_z$logDissOrthophos[abs(scale(clean_df_z$logDissOrthophos)) > 3.29] <- NA

outliers_z <- anti_join(df, clean_df_z)

print(paste("rows changed:",nrow(outliers_z)))

```


```{r}
histogram(clean_df_z$logDissAmmonia)
boxplot(clean_df_z$logDissAmmonia)

histogram(clean_df_z$logDissOrthophos)
boxplot(clean_df_z$logDissOrthophos)

histogram(clean_df_z$logDissNitrateNitrite)
boxplot(clean_df_z$logDissNitrateNitrite)

```

### remove based on IQR method

```{r}
#testing removing outliers based on IQR
#Define upper and lower bounds for variables

ub_lda <- quantile(df$logDissAmmonia, 0.75, na.rm=TRUE) + 1.5*IQR(df$logDissAmmonia, na.rm=TRUE)
ub_ldn <- quantile(df$logDissNitrateNitrite, 0.75, na.rm=TRUE) + 1.5*IQR(df$logDissNitrateNitrite, na.rm=TRUE)
ub_ldo <- quantile(df$logDissOrthophos, 0.75, na.rm=TRUE) + 1.5*IQR(df$logDissOrthophos, na.rm=TRUE)

lb_lda <- quantile(df$logDissAmmonia, 0.25, na.rm=TRUE) - 1.5*IQR(df$logDissAmmonia, na.rm=TRUE)
lb_ldn <- quantile(df$logDissNitrateNitrite, 0.25, na.rm=TRUE) - 1.5*IQR(df$logDissNitrateNitrite, na.rm=TRUE)
lb_ldo <- quantile(df$logDissOrthophos, 0.25, na.rm=TRUE) - 1.5*IQR(df$logDissOrthophos, na.rm=TRUE)

outliers_lda <-  df$logDissAmmonia[df$logDissAmmonia < lb_lda | df$logDissAmmonia > ub_lda]
outliers_lda <- outliers_lda[!is.na(outliers_lda)]

outliers_ldn <- df$logDissNitrateNitrite[df$logDissNitrateNitrite < lb_ldn | df$logDissNitrateNitrite > ub_ldn ]
outliers_ldn <- outliers_ldn[!is.na(outliers_ldn)]

outliers_ldo <- df$logDissOrthophos[df$logDissOrthophos < lb_ldo | df$logDissOrthophos > ub_ldo]
outliers_ldo <- outliers_ldo[!is.na(outliers_ldo)]

#########################################
# Remove outliers

clean_df_iqr <- df

clean_df_iqr$logDissAmmonia[clean_df_iqr$logDissAmmonia < lb_lda | clean_df_iqr$logDissAmmonia > ub_lda ] <- NA

clean_df_iqr$logDissNitrateNitrite[clean_df_iqr$logDissNitrateNitrite < lb_ldn | clean_df_iqr$logDissNitrateNitrite > ub_ldn ] <- NA

clean_df_iqr$logDissOrthophos[clean_df_iqr$logDissOrthophos < lb_ldo | clean_df_iqr$logDissOrthophos > ub_ldo ] <- NA


outliers_iqr <- anti_join(df, clean_df_iqr)

paste("rows changed:",nrow(outliers_iqr))
```

```{r}
histogram(clean_df_iqr$logDissAmmonia)
boxplot(clean_df_iqr$logDissAmmonia)

histogram(clean_df_iqr$logDissOrthophos)
boxplot(clean_df_iqr$logDissOrthophos)

histogram(clean_df_iqr$logDissNitrateNitrite)
boxplot(clean_df_iqr$logDissNitrateNitrite)

```

## Choose dataset

```{r}
# choose which outlier adjusted dataset to use:
# outliers removed by z-score:
df <- clean_df_z
# outliers removed by IQR
#df <- clean_df_iqr
```
## Separate dataset regionally

```{r}
# We need to remove the flow variables so we don't have duplicated values for each region

df_flow <- df[,c("Date","Regions","SAC", "OUT",  "EXPORT", "logEXPORT","logOUT","lagOUT","logSAC","lagSAC")]

df_no_flow <- df[,!(names(df) %in% c("SAC", "logSAC", "OUT", "logOUT", "EXPORT", "logEXPORT","lagOUT", "lagSAC"))]

r <- unique(df_no_flow$Regions)
region_list <- list()
for (region in r){
  df_regional <- df_no_flow %>% filter(Regions == region)
  region <- gsub(" ","",region)
  region_list[[region]] <- df_regional
}
```

## Impute

Impute all of the regional datasets 5 times

```{r include=FALSE}

# This creates a list of mids objects

regional_imp <- list()
for (region_name in as.character(names(region_list))){
  current_df <- region_list[[region_name]]
  predictors <- quickpred(current_df)
  imp <- mice(current_df,seed = 123, print=F, predictorMatrix = predictors)
  new_name <- paste0(region_name, "_imp")
  regional_imp[[new_name]]<- imp
}
```

## Make combined imputed datasets

```{r}
# Create imputed datasets for each region
regional_imp_dfs <- list()
action_seq <- as.list(seq(1,5))
for (region_imp in as.character(names(regional_imp))){
      current_imp <- regional_imp[[region_imp]]
      for (i in action_seq){
        df_new <- complete(current_imp,action = i, include=FALSE)
        df_new <- left_join( df_new,df_flow, by=c("Date","Regions"))
        new_name <- paste0(region_imp, "_df", i)
        regional_imp_dfs[[new_name]]<- df_new}}

```


## Add lag terms to the imputed datasets

```{r}
# This code is taken from clamdata_cleanup_AB_261125.R

df_list_lag <- lapply(regional_imp_dfs, function(y) {
#Step 1: summarize by season 
wq_r_sum_lagseason <- y %>%
  group_by(Regions,Year,season) %>%
  summarize(across(where(is.numeric),\(x) mean(x, na.rm = TRUE)), .groups="drop_last")

#Step 2: Add a column for the following season.
#This seems counter-intuitive because the lag should be the prior season,
#but this is a merge column: it will merge back with the prior season
#in the original table. 
wq_r_sum_lagseason <- wq_r_sum_lagseason %>%
  select(!c(lagOUT,lagSAC,lagTemperature,lagDissAmmonia,lagDissNitrateNitrite, lagTotPhos,lagConductivity,lagDissOrthophos))%>%
  rename_with(~paste0("lag",.x)) %>%
  rename(Year = "lagYear",season = "lagseason",Regions="lagRegions") %>% 
  mutate(lagseason = case_when(
    season == "Spring" ~ "Summer",
    season == "Summer" ~ "Autumn",
    season == "Autumn" ~ "Winter",
    season == "Winter" ~ "Spring" 
  )) %>%
  mutate(lagseasonyear = case_when(
    season == "Autumn" ~ Year+1,
    !(season == "Autumn") ~ Year
  )) %>%
  mutate(lagseasonyear = paste0(lagseason,lagseasonyear)) %>%
  ungroup %>% 
  select(lagseasonyear, Regions, laglogDissAmmonia,laglogDissNitrateNitrite,laglogConductivity,
         laglogSAC,laglogOUT, laglogDissOrthophos, laglogEXPORT)


y <- y %>% 
  mutate(seasonyear = paste0(season,year(Date))) %>%
  merge(wq_r_sum_lagseason,by.x=c("Regions","seasonyear"),by.y=c("Regions","lagseasonyear"),all.x=T) %>%
  select(-seasonyear)%>%
  relocate(Date)
})

```

## Combine the imputed datasets

Now we have to combine the 25 regional datasets into 5 datasets that include all regions.This is also where we are z-scoring the datasets for better comparison of the predictors' relative effects.

```{r}
### df1
df1 <- distinct(rbind(df_list_lag$Central_imp_df1, df_list_lag$Confluence_imp_df1,
             df_list_lag$NorthDelta_imp_df1, df_list_lag$South_imp_df1,df_list_lag$SuisunBay_imp_df1)) %>%
  # z-score as the last step before the GAM
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

### df2
df2 <- distinct(rbind(df_list_lag$Central_imp_df2, df_list_lag$Confluence_imp_df2,
             df_list_lag$NorthDelta_imp_df2, df_list_lag$South_imp_df2,df_list_lag$SuisunBay_imp_df2))%>%
 # z-score 
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

### df3

df3 <- distinct(rbind(df_list_lag$Central_imp_df3, df_list_lag$Confluence_imp_df3,
             df_list_lag$NorthDelta_imp_df3, df_list_lag$South_imp_df3,df_list_lag$SuisunBay_imp_df3))%>%
  # z-score 
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

### df4
df4 <- distinct(rbind(df_list_lag$Central_imp_df4, df_list_lag$Confluence_imp_df4,
             df_list_lag$NorthDelta_imp_df4, df_list_lag$South_imp_df4,df_list_lag$SuisunBay_imp_df4))%>%
  # z-score 
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))


### df5

df5 <- distinct(rbind(df_list_lag$Central_imp_df5, df_list_lag$Confluence_imp_df5,
             df_list_lag$NorthDelta_imp_df5, df_list_lag$South_imp_df5,df_list_lag$SuisunBay_imp_df5))%>%
  # z- score
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

# Make regions a factor
df1$Regions <- as.factor(df1$Regions)
df2$Regions <- as.factor(df2$Regions)
df3$Regions <- as.factor(df3$Regions)
df4$Regions <- as.factor(df4$Regions)
df5$Regions <- as.factor(df5$Regions)

```

# Fit models

This time, I am fitting the top performing models with linear terms for the nutrients.

```{r}
m3.po4.imp1.log <- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + laglogDissNitrateNitrite + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logOUT, by = season) + clam_step + s(logOUT) + s(laglogOUT) + laglogDissAmmonia+ logDissOrthophos+ laglogDissOrthophos, data=df1, method = 'ML', select=TRUE)

m4.po4.imp1.log <- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + laglogDissNitrateNitrite + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logSAC, by =season) + clam_step + s(laglogSAC) + laglogDissAmmonia + logDissOrthophos + laglogDissOrthophos, data=df1, method = 'ML', select=TRUE)

m6.po4.imp1.log <- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + laglogDissNitrateNitrite + s(Temperature) + s(logConductivity) + s(Regions, bs='re') + season + s(logSAC, by =season) + clam_step + s(laglogSAC) + laglogDissAmmonia+ logDissOrthophos+ s(EXPORT)+ laglogDissOrthophos,data= df1, method = 'ML', select=TRUE)
```


# Check AIC/BIC

```{r echo=FALSE}

aic_all<- AIC(m3.po4.imp1.log, m4.po4.imp1.log, m6.po4.imp1.log)

bic_all <- BIC(m3.po4.imp1.log, m4.po4.imp1.log, m6.po4.imp1.log)
```


This table ranks the models by their AIC and shows the difference in AIC/BIC values of each model from the lowest AIC/BIC of all models. The code is taken from here: https://interagencyecologicalprogram.github.io/ND-FASTR/rtm_chl_models_flow_weekly_avg.html#Purpose 


```{r}
# AIC values
aic_all_t <- aic_all %>%
  as_tibble(rownames = "Model")

# BIC values
bic_all_t <- bic_all %>%
  as_tibble(rownames = "Model")

# Combine AIC and BIC and calculate differences from lowest value
aic_bic_all <-
  left_join(aic_all_t, bic_all_t, by = join_by(Model, df)) %>%
  mutate(across(c(AIC, BIC), ~ .x - min(.x), .names = "{.col}_delta")) %>%
  select(starts_with("Model"), df, starts_with("AIC"), starts_with("BIC"))

# Sort by AIC
aic_bic_all %>% arrange(AIC)

# These are performing better with AIC/BIC than the GAMs without the linear terms and the outliers removed
```

# M3 Results {.tabset}

## GAM Check

```{r}
gam.check(m3.po4.imp1.log)
```

## Summary

```{r}
summary(m3.po4.imp1.log)
```

## Plot Predictions

```{r echo=FALSE}
var_list = list("logDissAmmonia","logDissNitrateNitrite","laglogDissNitrateNitrite","logConductivity", "logOUT","laglogOUT", "laglogDissAmmonia","logDissOrthophos","laglogDissOrthophos","Regions")
```

```{r}

lapply(var_list, function(var) {
    plot_predictions(m3.po4.imp1.log, condition = var, points = 0.5) +
      theme_classic(base_size = 12)
})
```


```{r}
 plot_predictions(m3.po4.imp1.log, condition = c("logOUT","season"), points = 0.5) +
    theme_classic(base_size = 12)
```

# M4 Results {.tabset}

## GAM Check

```{r}
gam.check(m4.po4.imp1.log)
```

## Summary

```{r}
summary(m4.po4.imp1.log)
```


## Plot Predictions

```{r echo=FALSE}
var_list = list("logDissAmmonia", "laglogDissAmmonia", "logDissNitrateNitrite","laglogDissNitrateNitrite",  "logDissOrthophos","laglogDissOrthophos", "logConductivity", "laglogSAC","Regions")
```

```{r}

lapply(var_list, function(var) {
    plot_predictions(m4.po4.imp1.log, condition = var, points = 0.5) +
      theme_classic(base_size = 12)
})
``` 

```{r}
plot_predictions(m4.po4.imp1.log, condition = c("logSAC","season"), points = 0.5) +
    theme_classic(base_size = 12)
```

# M6 Results {.tabset}

## GAM Check

```{r}
gam.check(m6.po4.imp1.log)
```

## Summary

```{r}
summary(m6.po4.imp1.log)
```



## Plot Predictions

```{r echo = FALSE}
var_list = list("logDissAmmonia", "laglogDissAmmonia", "logDissNitrateNitrite","laglogDissNitrateNitrite",  "logDissOrthophos","laglogDissOrthophos", "logConductivity","EXPORT", "laglogSAC","Regions")
```

```{r}

lapply(var_list, function(var) {
    plot_predictions(m6.po4.imp1.log, condition = var, points = 0.5) +
      theme_classic(base_size = 12)
})
```


```{r}
plot_predictions(m6.po4.imp1.log, condition = c("logSAC","season"), points = 0.5) +
    theme_classic(base_size = 12)
```

