---
title:
  "Global GAM 2"
author:
  "Ryan Hankins"
output: 
  html_document:
        toc: TRUE
        toc_float: TRUE
        code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, tidy = "styler")
```
# Background on this RMD file

This file continues the work in Global_GAM_RH.RMD. Here, I remove outliers, change the nutrient terms to linear, remove season as a term, check and remove the random effect of region, and change the seasonallag to a 3-month lag.
# Purpose

The purpose of these models is to understand the effects of environmental drivers and antecedent conditions on Delta-wide phytoplankton biomass (chl-a). 

# Prepare Data

```{r, include=FALSE}

library(mgcv)
library(tidyverse)
library(marginaleffects)
library(MuMIn)
library(corrplot)
library(stats)
library(mice)
library(lattice)
library(ggplot2)
library(data.table)
library(lmerTest)
library(lubridate)
```


```{r}

# variables to include in the analysis:
c = c("Chlorophyll","logchla", "Regions", "Date","Year", "Month",  "season","SAC",
       "EXPORT","logEXPORT","OUT", "DissAmmonia",
        "TotPhos","DissNitrateNitrite", 
       "Temperature","lagTemperature","Secchi", 
       "Conductivity","season_num", "DissOrthophos",  "clam_step","logDissNitrateNitrite","logDissAmmonia","logDissOrthophos","logConductivity","logSAC","logOUT")#, "Biomass","Grazing_rate","Filtration_rate", "SJR","lagSJR", "Index","lagIndex")

df <- read.csv("Data/regional_integrated_dataset3.csv" )
df$Date = as.Date(df$Date)

# 1979 is the start of the study period.
df <- subset(df,df$Date >= "1979-01-01")
# 2022 is the end of the study period
df <- subset(df,df$Date <= "2022-12-31")

# Changing categorical variables to factors
# need to change regions to factor farther down so dataframe splitting by region works properly
#df$Regions = as.factor(df$Regions)
df$season = as.factor(df$season)
df$season_num = as.numeric(df$season)


# Replacing clam variables with a step function (pre/post potamocorbula introduction to the Bay in 1987)
df <- df%>%mutate(clam_step=if_else(Year>1987, "post clam", "pre clam")) 
df$clam_step <- as.factor(df$clam_step)

# log transforming chl-a to achieve normal distribution of response variable.
df$logchla <- log(df$Chlorophyll)

# log transforming some variables to better visualize
df$logDissNitrateNitrite <- log(df$DissNitrateNitrite)
df$logDissAmmonia <- log(df$DissAmmonia)
df$logDissOrthophos <- log(df$DissOrthophos)
df$logConductivity <- log(df$Conductivity)
df$logSAC <- log(df$SAC)
df$logOUT <- log(df$OUT)
df$logEXPORT <- log(df$EXPORT)

df <- subset(df, select = c)

# Make sure all dates are present for lag variables later on.
df %>% complete(Date = seq(min(Date), max(Date), by = "1 month"))

```


## Remove outliers {.tabset}

### Investigating problem variables

```{r}
histogram(df$logDissAmmonia)
boxplot(df$logDissAmmonia)
histogram(df$logDissOrthophos)
boxplot(df$logDissOrthophos)
histogram(df$logDissNitrateNitrite)
boxplot(df$logDissNitrateNitrite)
``` 

### Remove outliers based on z-score

removing outliers based on z-score, assuming the log transformations of the targeted variables have a normal distribution.

-3 - 3 is a commonly used range beyond which values are considered outliers.
https://statsandr.com/blog/outliers-detection-in-r/#z-scores 


```{r}

clean_df_z <- df 

clean_df_z$logDissAmmonia[abs(scale(clean_df_z$logDissAmmonia)) > 3.29] <- NA

clean_df_z$logDissNitrateNitrite[abs(scale(clean_df_z$logDissNitrateNitrite)) > 3.29] <- NA

clean_df_z$logDissOrthophos[abs(scale(clean_df_z$logDissOrthophos)) > 3.29] <- NA

outliers_z <- anti_join(df, clean_df_z)

df <- clean_df_z

print(paste("rows changed:",nrow(outliers_z)))

```


```{r}
histogram(clean_df_z$logDissAmmonia)
boxplot(clean_df_z$logDissAmmonia)

histogram(clean_df_z$logDissOrthophos)
boxplot(clean_df_z$logDissOrthophos)

histogram(clean_df_z$logDissNitrateNitrite)
boxplot(clean_df_z$logDissNitrateNitrite)

```

### remove based on IQR method

```{r}
# #testing removing outliers based on IQR
# #Define upper and lower bounds for variables
# 
# ub_lda <- quantile(df$logDissAmmonia, 0.75, na.rm=TRUE) + 1.5*IQR(df$logDissAmmonia, na.rm=TRUE)
# ub_ldn <- quantile(df$logDissNitrateNitrite, 0.75, na.rm=TRUE) + 1.5*IQR(df$logDissNitrateNitrite, na.rm=TRUE)
# ub_ldo <- quantile(df$logDissOrthophos, 0.75, na.rm=TRUE) + 1.5*IQR(df$logDissOrthophos, na.rm=TRUE)
# 
# lb_lda <- quantile(df$logDissAmmonia, 0.25, na.rm=TRUE) - 1.5*IQR(df$logDissAmmonia, na.rm=TRUE)
# lb_ldn <- quantile(df$logDissNitrateNitrite, 0.25, na.rm=TRUE) - 1.5*IQR(df$logDissNitrateNitrite, na.rm=TRUE)
# lb_ldo <- quantile(df$logDissOrthophos, 0.25, na.rm=TRUE) - 1.5*IQR(df$logDissOrthophos, na.rm=TRUE)
# 
# outliers_lda <-  df$logDissAmmonia[df$logDissAmmonia < lb_lda | df$logDissAmmonia > ub_lda]
# outliers_lda <- outliers_lda[!is.na(outliers_lda)]
# 
# outliers_ldn <- df$logDissNitrateNitrite[df$logDissNitrateNitrite < lb_ldn | df$logDissNitrateNitrite > ub_ldn ]
# outliers_ldn <- outliers_ldn[!is.na(outliers_ldn)]
# 
# outliers_ldo <- df$logDissOrthophos[df$logDissOrthophos < lb_ldo | df$logDissOrthophos > ub_ldo]
# outliers_ldo <- outliers_ldo[!is.na(outliers_ldo)]
# 
# #########################################
# # Remove outliers
# 
# clean_df_iqr <- df
# 
# clean_df_iqr$logDissAmmonia[clean_df_iqr$logDissAmmonia < lb_lda | clean_df_iqr$logDissAmmonia > ub_lda ] <- NA
# 
# clean_df_iqr$logDissNitrateNitrite[clean_df_iqr$logDissNitrateNitrite < lb_ldn | clean_df_iqr$logDissNitrateNitrite > ub_ldn ] <- NA
# 
# clean_df_iqr$logDissOrthophos[clean_df_iqr$logDissOrthophos < lb_ldo | clean_df_iqr$logDissOrthophos > ub_ldo ] <- NA
# 
# 
# outliers_iqr <- anti_join(df, clean_df_iqr)
# 
# paste("rows changed:",nrow(outliers_iqr))
```

```{r}
# histogram(clean_df_iqr$logDissAmmonia)
# boxplot(clean_df_iqr$logDissAmmonia)
# 
# histogram(clean_df_iqr$logDissOrthophos)
# boxplot(clean_df_iqr$logDissOrthophos)
# 
# histogram(clean_df_iqr$logDissNitrateNitrite)
# boxplot(clean_df_iqr$logDissNitrateNitrite)

```

## Separate dataset regionally

```{r}
# We need to remove the flow variables so we don't have duplicated values for each region

df_flow <- df[,c("Date","Regions","SAC", "OUT",  "EXPORT", "logEXPORT","logOUT","logSAC")]

df_no_flow <- df[,!(names(df) %in% c("SAC", "logSAC", "OUT", "logOUT", "EXPORT", "logEXPORT"))]

r <- unique(df_no_flow$Regions)
region_list <- list()
for (region in r){
  df_regional <- df_no_flow %>% filter(Regions == region)
  region <- gsub(" ","",region)
  region_list[[region]] <- df_regional
}
```

## Impute

Impute all of the regional datasets 5 times

```{r include=FALSE}

# This creates a list of mids objects

regional_imp <- list()
for (region_name in as.character(names(region_list))){
  current_df <- region_list[[region_name]]
  predictors <- quickpred(current_df)
  imp <- mice(current_df,seed = 123, print=F, predictorMatrix = predictors)
  new_name <- paste0(region_name, "_imp")
  regional_imp[[new_name]]<- imp
}
```

## Make combined imputed datasets

```{r}
# Create imputed datasets for each region
regional_imp_dfs <- list()
action_seq <- as.list(seq(1,5))
for (region_imp in as.character(names(regional_imp))){
      current_imp <- regional_imp[[region_imp]]
      for (i in action_seq){
        df_new <- complete(current_imp,action = i, include=FALSE)
        df_new <- left_join( df_new,df_flow, by=c("Date","Regions"))
        new_name <- paste0(region_imp, "_df", i)
        regional_imp_dfs[[new_name]]<- df_new}}

```


## Add lag terms to the imputed datasets

```{r}
# This code is taken from clamdata_cleanup_AB_261125.R
# 
# df_list_lag <- lapply(regional_imp_dfs, function(y) {
# #Step 1: summarize by season 
# wq_r_sum_lagseason <- y %>%
#   group_by(Regions,Year,season) %>%
#   summarize(across(where(is.numeric),\(x) mean(x, na.rm = TRUE)), .groups="drop_last")
# 
# #Step 2: Add a column for the following season.
# #This seems counter-intuitive because the lag should be the prior season,
# #but this is a merge column: it will merge back with the prior season
# #in the original table. 
# wq_r_sum_lagseason <- wq_r_sum_lagseason %>%
#   select(!c(lagOUT,lagSAC,lagTemperature,lagDissAmmonia,lagDissNitrateNitrite, lagTotPhos,lagConductivity,lagDissOrthophos))%>%
#   rename_with(~paste0("lag",.x)) %>%
#   rename(Year = "lagYear",season = "lagseason",Regions="lagRegions") %>% 
#   mutate(lagseason = case_when(
#     season == "Spring" ~ "Summer",
#     season == "Summer" ~ "Autumn",
#     season == "Autumn" ~ "Winter",
#     season == "Winter" ~ "Spring" 
#   )) %>%
#   mutate(lagseasonyear = case_when(
#     season == "Autumn" ~ Year+1,
#     !(season == "Autumn") ~ Year
#   )) %>%
#   mutate(lagseasonyear = paste0(lagseason,lagseasonyear)) %>%
#   ungroup %>% 
#   select(lagseasonyear, Regions, laglogDissAmmonia,laglogDissNitrateNitrite,laglogConductivity,
#          laglogSAC,laglogOUT, laglogDissOrthophos, laglogEXPORT)
# 
# 
# y <- y %>% 
#   mutate(seasonyear = paste0(season,year(Date))) %>%
#   merge(wq_r_sum_lagseason,by.x=c("Regions","seasonyear"),by.y=c("Regions","lagseasonyear"),all.x=T) %>%
#   select(-seasonyear)%>%
#   relocate(Date)
# })

```

```{r}
#  Create the 3 month lag
df_list_lag <- lapply(regional_imp_dfs, function(y){

  lagged_df <- y %>%
    # Group by region so completion happens independently per region
    group_by(Regions) %>%
    
    # Sort chronologically within each group
    arrange(Regions, Date) %>%
    
    # Lag all numeric columns by 3 rows
    mutate(across(where(is.numeric), ~ lag(.x, n = 3), .names = "lag3_{col}")) %>%

    ungroup()
})

```

## Inspect convergence regionally {.tabset}

The plot shows the mean and the standard deviation of the imputed values only. Generally, we want to see the streams intermingle and be free of any trends at the later iterations. (https://www.gerkovink.com/miceVignettes/Convergence_pooling/Convergence_and_pooling.html)

### Central

```{r}
# Check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$Central_imp, maxit=35, print=F)
plot(imp40)
```


### Confluence

```{r}
#check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$Confluence_imp, maxit=35, print=F)
plot(imp40)
```


### North

```{r}
#check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$NorthDelta_imp, maxit=35, print=F)
plot(imp40)
```

### South

```{r}
#check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$South_imp, maxit=35, print=F)
plot(imp40)
```


### Suisun Bay

```{r}
#check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$SuisunBay_imp, maxit=35, print=F)
plot(imp40)
```



## Combine the imputed datasets

Now we have to combine the 25 regional datasets into 5 datasets that include all regions.This is also where we are z-scoring the datasets for better comparison of the predictors' relative effects.

```{r}
### df1
df1 <- distinct(rbind(df_list_lag$Central_imp_df1, df_list_lag$Confluence_imp_df1,
             df_list_lag$NorthDelta_imp_df1, df_list_lag$South_imp_df1,df_list_lag$SuisunBay_imp_df1)) %>%
  # z-score as the last step before the GAM
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

### df2
df2 <- distinct(rbind(df_list_lag$Central_imp_df2, df_list_lag$Confluence_imp_df2,
             df_list_lag$NorthDelta_imp_df2, df_list_lag$South_imp_df2,df_list_lag$SuisunBay_imp_df2))%>%
 # z-score 
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

### df3

df3 <- distinct(rbind(df_list_lag$Central_imp_df3, df_list_lag$Confluence_imp_df3,
             df_list_lag$NorthDelta_imp_df3, df_list_lag$South_imp_df3,df_list_lag$SuisunBay_imp_df3))%>%
  # z-score 
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

### df4
df4 <- distinct(rbind(df_list_lag$Central_imp_df4, df_list_lag$Confluence_imp_df4,
             df_list_lag$NorthDelta_imp_df4, df_list_lag$South_imp_df4,df_list_lag$SuisunBay_imp_df4))%>%
  # z-score 
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))


### df5

df5 <- distinct(rbind(df_list_lag$Central_imp_df5, df_list_lag$Confluence_imp_df5,
             df_list_lag$NorthDelta_imp_df5, df_list_lag$South_imp_df5,df_list_lag$SuisunBay_imp_df5))%>%
  # z- score
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

# Make regions a factor
df1$Regions <- as.factor(df1$Regions)
df2$Regions <- as.factor(df2$Regions)
df3$Regions <- as.factor(df3$Regions)
df4$Regions <- as.factor(df4$Regions)
df5$Regions <- as.factor(df5$Regions)

```

# Check Regional Effects

Using a linear mixed effects model to check the regional random effect

```{r}
m3.1.lme <- lmer((logchla) ~ logDissAmmonia + logDissNitrateNitrite + lag3_logDissNitrateNitrite + Temperature + logConductivity + (1|Regions) + logOUT + clam_step + logOUT + lag3_logOUT + lag3_logDissAmmonia+ logDissOrthophos+ lag3_logDissOrthophos, data=df1)
```

```{r}
summary(m3.1.lme)
```

# Fit models

This time, I am fitting the top performing models with linear terms for the nutrients, removing season, the random effect of regions, and changing the seasonal lag to a 3-month lag.

```{r}
m3.1 <- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + lag3_logDissNitrateNitrite + s(Temperature) + s(logConductivity)  + s(logOUT) + clam_step + s(logOUT) + s(lag3_logOUT) + lag3_logDissAmmonia+ logDissOrthophos+ lag3_logDissOrthophos, data=df1, method = 'ML', select=TRUE)

m3.2 <- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + lag3_logDissNitrateNitrite + s(Temperature) + s(logConductivity)  + s(logOUT) + clam_step + s(logOUT) + s(lag3_logOUT) + lag3_logDissAmmonia+ logDissOrthophos+ lag3_logDissOrthophos, data=df2, method = 'ML', select=TRUE)

m3.3 <- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + lag3_logDissNitrateNitrite + s(Temperature) + s(logConductivity) + s(logOUT) + clam_step + s(logOUT) + s(lag3_logOUT) + lag3_logDissAmmonia+ logDissOrthophos+ lag3_logDissOrthophos, data=df3, method = 'ML', select=TRUE)

m3.4 <- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + lag3_logDissNitrateNitrite + s(Temperature) + s(logConductivity) + s(logOUT) + clam_step + s(logOUT) + s(lag3_logOUT) + lag3_logDissAmmonia+ logDissOrthophos+ lag3_logDissOrthophos, data=df4, method = 'ML', select=TRUE)

m3.5 <- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + lag3_logDissNitrateNitrite + s(Temperature) + s(logConductivity)  + s(logOUT) + clam_step + s(logOUT) + s(lag3_logOUT) + lag3_logDissAmmonia+ logDissOrthophos+ lag3_logDissOrthophos, data=df5, method = 'ML', select=TRUE)
```

```{r}
m4.1 <- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + lag3_logDissNitrateNitrite + s(Temperature) + s(logConductivity)+ s(logSAC) + clam_step + s(lag3_logSAC) + lag3_logDissAmmonia + logDissOrthophos + lag3_logDissOrthophos, data=df1, method = 'ML', select=TRUE)

m4.2<- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + lag3_logDissNitrateNitrite + s(Temperature) + s(logConductivity) + s(logSAC) + clam_step + s(lag3_logSAC) + lag3_logDissAmmonia + logDissOrthophos + lag3_logDissOrthophos, data=df2, method = 'ML', select=TRUE)

m4.3 <- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + lag3_logDissNitrateNitrite + s(Temperature) + s(logConductivity)  + s(logSAC) + clam_step + s(lag3_logSAC) + lag3_logDissAmmonia + logDissOrthophos + lag3_logDissOrthophos, data=df3, method = 'ML', select=TRUE)

m4.4 <- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + lag3_logDissNitrateNitrite + s(Temperature) + s(logConductivity)  + s(logSAC) + clam_step + s(lag3_logSAC) + lag3_logDissAmmonia + logDissOrthophos + lag3_logDissOrthophos, data=df4, method = 'ML', select=TRUE)

m4.5 <- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + lag3_logDissNitrateNitrite + s(Temperature) + s(logConductivity)  + s(logSAC) + clam_step + s(lag3_logSAC) + lag3_logDissAmmonia + logDissOrthophos + lag3_logDissOrthophos, data=df5, method = 'ML', select=TRUE)
```

```{r}
m6.1 <- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + lag3_logDissNitrateNitrite + s(Temperature) + s(logConductivity) + s(logSAC) + clam_step + s(lag3_logSAC) + lag3_logDissAmmonia+ logDissOrthophos+ s(EXPORT)+ lag3_logDissOrthophos,data= df1, method = 'ML', select=TRUE)

m6.2 <- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + lag3_logDissNitrateNitrite + s(Temperature) + s(logConductivity) + s(logSAC) + clam_step + s(lag3_logSAC) + lag3_logDissAmmonia+ logDissOrthophos+ s(EXPORT)+ lag3_logDissOrthophos,data= df2, method = 'ML', select=TRUE)

m6.3<- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + lag3_logDissNitrateNitrite + s(Temperature) + s(logConductivity)  + s(logSAC) + clam_step + s(lag3_logSAC) + lag3_logDissAmmonia+ logDissOrthophos+ s(EXPORT)+ lag3_logDissOrthophos,data= df3, method = 'ML', select=TRUE)

m6.4 <- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + lag3_logDissNitrateNitrite + s(Temperature) + s(logConductivity)  + s(logSAC) + clam_step + s(lag3_logSAC) + lag3_logDissAmmonia+ logDissOrthophos+ s(EXPORT)+ lag3_logDissOrthophos,data= df4, method = 'ML', select=TRUE)

m6.5 <- gam((logchla) ~ logDissAmmonia + logDissNitrateNitrite + lag3_logDissNitrateNitrite + s(Temperature) + s(logConductivity)  + s(logSAC) + clam_step + s(lag3_logSAC) + lag3_logDissAmmonia+ logDissOrthophos+ s(EXPORT)+ lag3_logDissOrthophos,data= df5, method = 'ML', select=TRUE)
```


# Check AIC/BIC

```{r echo=FALSE}

aic_all<- AIC(m3.1, m3.2, m3.3, m3.4, m3.5, m4.1, m4.2, m4.3, m4.4, m4.5, m6.1, m6.2, m6.3, m6.4, m6.5)

bic_all <- BIC(m3.1, m3.2, m3.3, m3.4, m3.5, m4.1, m4.2, m4.3, m4.4, m4.5, m6.1, m6.2, m6.3, m6.4, m6.5)
```


This table ranks the models by their AIC and shows the difference in AIC/BIC values of each model from the lowest AIC/BIC of all models. The code is taken from here: https://interagencyecologicalprogram.github.io/ND-FASTR/rtm_chl_models_flow_weekly_avg.html#Purpose 


```{r}
# AIC values
aic_all_t <- aic_all %>%
  as_tibble(rownames = "Model")

# BIC values
bic_all_t <- bic_all %>%
  as_tibble(rownames = "Model")

# Combine AIC and BIC and calculate differences from lowest value
aic_bic_all <-
  left_join(aic_all_t, bic_all_t, by = join_by(Model, df)) %>%
  mutate(across(c(AIC, BIC), ~ .x - min(.x), .names = "{.col}_delta")) %>%
  select(starts_with("Model"), df, starts_with("AIC"), starts_with("BIC"))

# Sort by AIC
aic_bic_all %>% arrange(AIC)

```

# M3 Results {.tabset}

## GAM Check

```{r}
gam.check(m3.1)
```

## Summary

```{r}
summary(m3.1)
```

## Plot Predictions

```{r echo=FALSE}
var_list = list("logDissAmmonia","logDissNitrateNitrite","lag3_logDissNitrateNitrite","logConductivity", "logOUT","lag3_logOUT", "lag3_logDissAmmonia","logDissOrthophos","lag3_logDissOrthophos", "logOUT", "clam_step")
```

```{r}

lapply(var_list, function(var) {
    plot_predictions(m3.1, condition = var, points = 0.5) +
      theme_classic(base_size = 12)
})
```

## Plot Marginal Effects

```{r}
lapply(var_list, function(x) {
  plot_slopes(m3.1, variables = x, condition = x)
})
```

# M4 Results {.tabset}

## GAM Check

```{r}
gam.check(m4.1)
```

## Summary

```{r}
summary(m4.1)
```


## Plot Predictions

```{r echo=FALSE}
var_list = list("logDissAmmonia", "lag3_logDissAmmonia", "logDissNitrateNitrite","lag3_logDissNitrateNitrite",  "logDissOrthophos","lag3_logDissOrthophos", "logConductivity", "lag3_logSAC", "clam_step")
```

```{r}

lapply(var_list, function(var) {
    plot_predictions(m4.1, condition = var, points = 0.5) +
      theme_classic(base_size = 12)
})
``` 

## Plot Marginal Effects

```{r}
lapply(var_list, function(x) {
  plot_slopes(m4.1, variables = x, condition = x)
})
```

# M6 Results {.tabset}

## GAM Check

```{r}
gam.check(m6.1)
```

## Summary

```{r}
summary(m6.1)
```

## Plot Predictions

```{r echo = FALSE}
var_list = list("logDissAmmonia", "lag3_logDissAmmonia", "logDissNitrateNitrite","lag3_logDissNitrateNitrite",  "logDissOrthophos","lag3_logDissOrthophos", "logConductivity","EXPORT", "lag3_logSAC","logSAC", "clam_step")
```

```{r}

lapply(var_list, function(var) {
    plot_predictions(m6.1, condition = var, points = 0.5) +
      theme_classic(base_size = 12)
})
```

## Plot Marginal Effects

```{r}
lapply(var_list, function(x) {
  plot_slopes(m6.1, variables = x, condition = x)
})
```

