---
title:
  "Global GAM"
author:
  "Ryan Hankins"
output: 
  html_document:
        toc: TRUE
        toc_float: TRUE
        code_folding: show
        df_print: paged 
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, tidy = "styler")
```



# Prepare Data

```{r, include=FALSE}

library(mgcv)
library(tidyverse)
library(marginaleffects)
library(MuMIn)
library(car)
library(corrplot)
library(stats)
library(marginaleffects)
library(mice)
library(lattice)
library(missForest)
```


```{r}
#options(na.action = "na.exclude")

# variables to include:
c = c( "logchla", "Regions", "Date","Year", "Month",  "season","SAC","lagSAC",
       "EXPORT","OUT","lagOUT", "SJR", "lagSJR", "Index","lagIndex","DissAmmonia",
       "lagDissAmmonia", "TotPhos","lagTotPhos","DissNitrateNitrite", 
       "lagDissNitrateNitrite","Temperature","lagTemperature","Secchi", 
       "lagSecchi","Conductivity",  "lagConductivity","season_num", "DissOrthophos", "lagDissOrthophos", "clam_step")#, #"Biomass","Grazing_rate",#"Filtration_rate")


df <- read.csv("Data/regional_integrated_dataset3.csv" )

df$Date = as.Date(df$Date)

# 1979 is the start of the study period.
df <- subset(df,df$Date >= "1979-01-01")

# Changing categorical variables to factors
df$Regions = as.factor(df$Regions)
df$season = as.factor(df$season)
df$season_num = as.numeric(df$season)


#Replacing clam variables with a step function (pre/post potamocorbula introduction to the Bay in 1987)
df <- df%>%mutate(clam_step=if_else(Year>1987, "post clam", "pre clam")) 
df$clam_step <- as.factor(df$clam_step)

# log transforming chl-a to achieve normal distribution of response variable.
df$logchla <- log(df$Chlorophyll)

# other option is using na.exclude on gams
# Dropping NAs from the dataset. 
 df <- subset(df, select = c) #%>%
    #drop_na()

# z-score standardizing the dataset for more logical comparison of predictors. This facilitates comparison of effect sizes between variables.

df <- df %>% mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))
df
```


# Model Selection

## Choose Models With Best Combination of Flow Variables 

```{r}
#options(na.action = "na.fail")
# OUT (no EXPORT, no step)
# m1 <- gam((logchla) ~ DissAmmonia +
#                DissNitrateNitrite +
#                Temperature +
#                Conductivity +
#                OUT +
#                s(Regions, bs='re') +
#                season +
#                (OUT*season) +
#                Grazing_rate +
#                lagOUT +
#                lagDissAmmonia +
#                lagDissNitrateNitrite,
#              data = df,
#              method = 'ML'
#              )
# 
# # # SAC (no EXPORT, no step)
# m2 <- gam((logchla) ~ DissAmmonia +
#                DissNitrateNitrite +
#                Temperature +
#                Conductivity +
#                SAC +
#                s(Regions, bs='re') +
#                season +
#                (SAC*season) +
#                Grazing_rate +
#                lagSAC +
#                lagDissAmmonia +
#                lagDissNitrateNitrite,
#              data = df,
#              method = 'ML'
#              )
```



```{r}

# # OUT + Step (no EXPORT)
# m3 <- gam((logchla) ~ DissAmmonia +
#                DissNitrateNitrite +
#                Temperature +
#                Conductivity +
#                OUT +
#                s(Regions, bs='re') +
#                season +
#                (OUT*season) +
#                clam_step +
#                lagOUT +
#                lagDissAmmonia +
#                lagDissNitrateNitrite,
#              data = df,
#              method = 'ML',
#              na.action = na.exclude)
# 
# # SAC + Step (no EXPORT)
# m4 <- gam((logchla) ~ DissAmmonia +
#                DissNitrateNitrite +
#                Temperature +
#                Conductivity +
#                SAC + 
#                s(Regions, bs='re') +
#                season +
#                (SAC*season) +
#                clam_step +
#                lagSAC +
#                lagDissAmmonia,
#               # lagDissNitrateNitrite, # removing based on dredge
#              data = df,
#              method = 'ML',
#              na.action = na.exclude)
# 
# # OUT + EXPORT + Step
# # m5 <- gam((logchla) ~ DissAmmonia +
# #                DissNitrateNitrite +
# #                Temperature +
# #                Conductivity +
# #                OUT + 
# #                EXPORT +
# #                s(Regions, bs='re') +
# #                season +
# #                (OUT*season) +
# #                clam_step +
# #                lagOUT +
# #                lagDissAmmonia +
# #                lagDissNitrateNitrite,
# #              data = df,
# #              method = 'ML'
# #              )
# 
# #SAC+ EXPORT + Step (best performing combo)
# m6 <- gam((logchla) ~ DissAmmonia +
#                DissNitrateNitrite +
#                Temperature +
#                Conductivity +
#                SAC + 
#                EXPORT +
#                s(Regions, bs='re') +
#                season +
#                # (SAC*season) + # removing based on dredge
#                clam_step +
#                lagSAC +
#                lagDissAmmonia,
#                #lagDissNitrateNitrite, # removing based on dredge
#              data = df,
#              method = 'ML',
#              na.action = na.exclude)

# m6 but with Season as a numeric smooth term with bs=cc (cyclic cubic spline)
# m7 <- gam((logchla) ~ DissAmmonia +
#                DissNitrateNitrite +
#                Temperature +
#                Conductivity +
#                SAC +
#                EXPORT +
#                s(Regions, bs='re') +
#                s(season_num, bs='cc', k=4) +
#                (SAC*season) +
#                clam_step +
#                lagSAC +
#                lagDissAmmonia +
#                lagDissNitrateNitrite,
#              data = df,
#              method = 'ML'
#              )
# SJR + OUT + EXPORT
# m8 <- gam((logchla) ~ DissAmmonia +
#                DissNitrateNitrite +
#                Temperature +
#                Conductivity +
#                SJR + 
#                EXPORT +
#                s(Regions, bs='re') +
#                s(season_num, bs='cc', k=4) +
#                (SJR*season) +
#                clam_step +
#                lagSJR +
#                lagDissAmmonia +
#                lagDissNitrateNitrite,
#              data = df,
#              method = 'ML'
#              )
# SJR + OUT (no EXPORT)
# m9 <- gam((logchla) ~ DissAmmonia +
#                DissNitrateNitrite +
#                Temperature +
#                Conductivity +
#                SJR + 
#                s(Regions, bs='re') +
#                s(season_num, bs='cc', k=4) +
#                (SJR*season) +
#                clam_step +
#                lagSJR +
#                lagDissAmmonia +
#                lagDissNitrateNitrite,
#              data = df,
#              method = 'ML'
#              )

```
## Compare with AIC
```{r}
#AIC(m3, m4,m6)
```

```{r}
#BIC(m3, m4, m6)
```

## Decisions/observations:
- remove month random effect
- keep season, try cyclic cubic spline bs="cc", set k=4 (m7)
  - season cyclic cubic spline removed after seeing it is insignificant in the models
- add region as a random effect
- models performed better with clam data as a step function

  
# Reduce variables in top models: {.tabset}
Dredging the top models to determine which variables to remove. 

## m1 
```{r}
# OUT (no EXPORT, no step)
# df <- subset(df, select = c)%>%
#    drop_na()
# options(na.action = "na.fail")
#m.1.dredge <- dredge(m1)
#m.1.dredge

```
## m2 
```{r}
# # SAC (no EXPORT, no step)
# m.2.dredge <- dredge(m2)
# head(m.2.dredge, 3)
```
## m.3 - active
```{r}
# OUT + Step (no EXPORT)
#m.3.dredge <- dredge(m3)
# head(m.3.dredge, 3)
```
Global model call: gam(formula = (logchla) ~ DissAmmonia + DissNitrateNitrite + Temperature + Conductivity + OUT + s(Regions, bs = "re") +  season + (OUT * season) + clam_step + lagOUT + lagDissAmmonia + lagDissNitrateNitrite, data = df, method = "ML")
---
Model selection table 
       (Int) clm_stp      Cnd      DsA     DNN     lDA      lDN      OUT s(Rgn,"re")
4032 -0.3008       + -0.06939 -0.03112 0.07865 0.10820 -0.03881 -0.08777           +
4000 -0.2931       + -0.07331 -0.02951 0.07022 0.10270          -0.08830           +
3996 -0.2982       + -0.07408          0.06258 0.09478          -0.08383           +
     ssn    Tmp OUT:ssn df    logLik   AICc delta weight
4032   + 0.3048       + 19 -1795.546 3631.5  0.00  0.391
4000   + 0.3031       + 18 -1796.595 3631.6  0.05  0.382
3996   + 0.3102       + 17 -1798.133 3632.6  1.09  0.227
Models ranked by AICc(x) 

## m.4 - active
```{r}
# SAC + Step (no EXPORT)
# m.4.dredge <- dredge(m4)
# head(m.4.dredge, 3)
```
Global model call: gam(formula = (logchla) ~ DissAmmonia + DissNitrateNitrite + 
    Temperature + Conductivity + SAC + s(Regions, bs = "re") + 
    season + (SAC * season) + clam_step + lagSAC + lagDissAmmonia + 
    lagDissNitrateNitrite, data = df, method = "ML")
---
Model selection table 
       (Int) clm_stp      Cnd      DsA     DNN    lDA      lDN s(Rgn,"re")      SAC
4000 -0.2851       + -0.07562 -0.03230 0.06620 0.1078                    + -0.05475
1984 -0.3065       + -0.07300 -0.03363 0.07382 0.1098 -0.04205           + -0.06866
4032 -0.2917       + -0.07216 -0.03369 0.07389 0.1126 -0.03481           + -0.05380
     ssn    Tmp SAC:ssn df    logLik   AICc delta weight
4000   + 0.2971       + 18 -1796.554 3631.5  0.00  0.362
1984   + 0.2817         16 -1798.675 3631.7  0.17  0.332
4032   + 0.2988       + 19 -1795.696 3631.8  0.33  0.306
Models ranked by AICc(x) 

## m5 
```{r}
# OUT + EXPORT + Step
#m.5.dredge <- dredge(m5)
#m.5.dredge
```
## m.6 -active
```{r}

# SAC+ EXPORT + Step (best performing combo)
# m.6.dredge <- dredge(m6)
# head(m.6.dredge, 3)
```
Global model call: gam(formula = (logchla) ~ DissAmmonia + DissNitrateNitrite + 
    Temperature + Conductivity + SAC + EXPORT + s(Regions, bs = "re") + 
    season + (SAC * season) + clam_step + lagSAC + lagDissAmmonia + 
    lagDissNitrateNitrite, data = df, method = "ML")
---
Model selection table 
       (Int) clm_stp      Cnd      DsA     DNN      EXP    lDA      lDN     lSA
4032 -0.2813       + -0.07495 -0.02849 0.07064 -0.04236 0.1118          0.04141
4096 -0.2891       + -0.07215 -0.02976 0.07762 -0.04208 0.1160 -0.03312 0.03720
8000 -0.2828       + -0.07825 -0.02835 0.06788 -0.03971 0.1106                 
     s(Rgn,"re")      SAC ssn    Tmp SAC:ssn df    logLik   AICc delta weight
4032           + -0.07462   + 0.2805         17 -1795.397 3627.1  0.00  0.398
4096           + -0.07332   + 0.2824         18 -1794.624 3627.6  0.50  0.310
8000           + -0.05055   + 0.2955       + 19 -1793.662 3627.8  0.61  0.293
Models ranked by AICc(x) 

## m7
```{r}
# m6 but with as a numeric smooth term with bs=cc (cyclic cubic spline)
# m.7.dredge <- dredge(m7)
# head(m.7.dredge, 3)
```
## m8
```{r}
# SJR+EXPORT+ Step
# m.8.dredge <- dredge(m8)
# head(m.7.dredge, 3)
```
# Compare top models with AIC/BIC
```{r}
# #top.m1 <- get.models(m.1.dredge, 1)[[1]]
# #top.m2 <- get.models(m.2.dredge, 1)[[1]]
# top.m3 <- get.models(m.3.dredge, 1)[[1]]
# top.m4 <- get.models(m.4.dredge, 1)[[1]]
# #top.m5 <- get.models(m.5.dredge, 1)[[1]]
# top.m6 <- get.models(m.6.dredge, 1)[[1]]
#top.m7 <- get.models(m.7.dredge, 1)[[1]]
#top.m8 <- get.models(m.8.dredge, 1)[[1]]
```
compare the top dredged models with AIC and BIC
```{r}
# AIC(top.m3, top.m4, top.m6)
```

```{r}
# BIC(top.m3, top.m4, top.m6)
```

# Remove Season {.tabset}
The interaction with Sac:season was not significant in m6 but was in m3 and m4
```{r}
# OUT + Step (no EXPORT)
# m3.no.season <- gam((logchla) ~ DissAmmonia +
#                DissNitrateNitrite +
#                Temperature +
#                Conductivity +
#                OUT +
#                s(Regions, bs='re') +
#                clam_step +
#                lagOUT +
#                lagDissAmmonia +
#                lagDissNitrateNitrite,
#              data = df,
#              method = 'ML',
#              na.action = na.exclude)
# 
# # SAC + Step (no EXPORT)
# m4.no.season <- gam((logchla) ~ DissAmmonia +
#                DissNitrateNitrite +
#                Temperature +
#                Conductivity +
#                SAC + 
#                s(Regions, bs='re') +
#                clam_step +
#                lagSAC +
#                lagDissAmmonia,
#               # lagDissNitrateNitrite,
#              data = df,
#              method = 'ML',
#              na.action = na.exclude)
# 
# #SAC+ EXPORT + Step (best performing combo)
# m6.no.season <- gam((logchla) ~ DissAmmonia +
#                DissNitrateNitrite +
#                Temperature +
#                Conductivity +
#                SAC + 
#                EXPORT +
#                s(Regions, bs='re') +
#                clam_step +
#                lagSAC +
#                lagDissAmmonia,
#                #lagDissNitrateNitrite,
#              data = df,
#              method = 'ML',
#             na.action = na.exclude )

```

## Check AIC of models so far

```{r}
#AIC(m3.no.season, m4.no.season, m6.no.season, m3, m4, m6)
```
```{r}
#BIC(m3.no.season, m4.no.season, m6.no.season, m3, m4, m6)
```

## m3 no season = active
```{r}
# m3.no.season.dredge <- dredge(m3.no.season)
# head(m3.no.season.dredge, 3)
```
Global model call: gam(formula = (logchla) ~ DissAmmonia + DissNitrateNitrite + 
    Temperature + Conductivity + OUT + s(Regions, bs = "re") + 
    clam_step + lagOUT + lagDissAmmonia + lagDissNitrateNitrite, 
    data = df, method = "ML")
---
Model selection table 
       (Int) clm_stp     Cnd      DsA     DNN    lDA     lDN     lOU      OUT
1024 -0.0662       + -0.1052 -0.04578 0.06000 0.1497 0.09286 0.08020 -0.05812
1016 -0.0630       + -0.1061 -0.03797         0.1462 0.10920 0.07587 -0.06411
1020 -0.0666       + -0.1081          0.04648 0.1376 0.10250 0.08302 -0.05307
     s(Rgn,"re")    Tmp df    logLik   AICc delta weight
1024           + 0.2986 14 -1847.929 3726.1  0.00  0.763
1016           + 0.2853 13 -1850.386 3729.0  2.88  0.181
1020           + 0.3087 13 -1851.550 3731.3  5.21  0.056
Models ranked by AICc(x) 

## m4 no season - active
```{r}
# m4.no.season.dredge <- dredge(m4.no.season)
# head(m4.no.season.dredge, 3)
```

Global model call: gam(formula = (logchla) ~ DissAmmonia + DissNitrateNitrite + 
    Temperature + Conductivity + SAC + s(Regions, bs = "re") + 
    clam_step + lagSAC + lagDissAmmonia + lagDissNitrateNitrite, 
    data = df, method = "ML")
---
Model selection table 
        (Int) clm_stp     Cnd      DsA     DNN    lDA     lDN     lSA s(Rgn,"re")
1024 -0.06646       + -0.1095 -0.05020 0.06015 0.1537 0.09403 0.08867           +
1016 -0.06300       + -0.1112 -0.04238         0.1503 0.11010 0.08414           +
1020 -0.06710       + -0.1121          0.04537 0.1402 0.10460 0.08995           +
          SAC    Tmp df    logLik   AICc delta weight
1024 -0.07314 0.2914 14 -1846.089 3722.4  0.00  0.788
1016 -0.07904 0.2780 13 -1848.559 3725.3  2.90  0.185
1020 -0.06614 0.3034 13 -1850.452 3729.1  6.69  0.028
Models ranked by AICc(x) 

## m6 no season - active
```{r}
# m6.no.season.dredge <- dredge(m6.no.season)
# head(m6.no.season.dredge, 3)
```
Global model call: gam(formula = (logchla) ~ DissAmmonia + DissNitrateNitrite + 
    Temperature + Conductivity + SAC + EXPORT + s(Regions, bs = "re") + 
    clam_step + lagSAC + lagDissAmmonia + lagDissNitrateNitrite, 
    data = df, method = "ML")
---
Model selection table 
        (Int) clm_stp     Cnd      DsA     DNN      EXP    lDA     lDN     lSA
2048 -0.06292       + -0.1091 -0.03903 0.07505 -0.07952 0.1537 0.06009 0.09201
1984 -0.06246       + -0.1069 -0.04251 0.08893 -0.08610 0.1671         0.09169
2044 -0.06323       + -0.1111          0.06448 -0.08353 0.1433 0.06652 0.09316
     s(Rgn,"re")      SAC    Tmp df    logLik   AICc delta weight
2048           + -0.05762 0.3034 15 -1830.625 3693.5  0.00  0.620
1984           + -0.05493 0.3059 14 -1832.536 3695.3  1.78  0.255
2044           + -0.05147 0.3133 14 -1833.244 3696.7  3.21  0.125
Models ranked by AICc(x) 

# Add Dissolved Orthophosphate
```{r}
# # OUT (no EXPORT)
# m3.po4 <- gam((logchla) ~ DissAmmonia +
#                DissNitrateNitrite +
#                Temperature +
#                Conductivity +
#                OUT +
#                s(Regions, bs='re') +
#                season +
#                (OUT*season) +
#                clam_step +
#                lagOUT +
#                lagDissAmmonia +
#                lagDissNitrateNitrite+
#                DissOrthophos +
#                lagDissOrthophos,
#              data = df,
#              method = 'ML'
#              )
# 
# # SAC  (no EXPORT)
# m4.po4 <- gam((logchla) ~ DissAmmonia +
#                DissNitrateNitrite +
#                Temperature +
#                Conductivity +
#                SAC + 
#                s(Regions, bs='re') +
#                season +
#                (SAC*season) +
#                clam_step +
#                lagSAC +
#                lagDissAmmonia +
#               # lagDissNitrateNitrite+
#                DissOrthophos +
#                lagDissOrthophos,
#              data = df,
#              method = 'ML'
#              )
# 
# #SAC+ EXPORT
# m6.po4 <- gam((logchla) ~ DissAmmonia +
#                DissNitrateNitrite +
#                Temperature +
#                Conductivity +
#                SAC + 
#                EXPORT +
#                s(Regions, bs='re') +
#                season +
#                (SAC*season) +
#                clam_step +
#                lagSAC +
#                lagDissAmmonia +
#              #  lagDissNitrateNitrite+
#                DissOrthophos +
#                lagDissOrthophos,
#              data = df,
#              method = 'ML'
#              )
# # OUT (no EXPORT)
# m3.no.season.po4 <- gam((logchla) ~ DissAmmonia +
#                DissNitrateNitrite +
#                Temperature +
#                Conductivity +
#                OUT +
#                s(Regions, bs='re') +
#                clam_step +
#                lagOUT +
#                lagDissAmmonia +
#                lagDissNitrateNitrite+
#                DissOrthophos +
#                lagDissOrthophos,
#              data = df,
#              method = 'ML'
#              )
# 
# # SAC + Step (no EXPORT)
# m4.no.season.po4 <- gam((logchla) ~ DissAmmonia +
#                DissNitrateNitrite +
#                Temperature +
#                Conductivity +
#                SAC + 
#                s(Regions, bs='re') +
#                clam_step +
#                lagSAC +
#                lagDissAmmonia +
#              #  lagDissNitrateNitrite+
#                DissOrthophos +
#                lagDissOrthophos,
#              data = df,
#              method = 'ML'
#              )
# 
# #SAC+ EXPORT + Step (best performing combo)
# m6.no.season.po4 <- gam((logchla) ~ DissAmmonia +
#                DissNitrateNitrite +
#                Temperature +
#                Conductivity +
#                SAC + 
#                EXPORT +
#                s(Regions, bs='re') +
#                clam_step +
#                lagSAC +
#                lagDissAmmonia +
#              #  lagDissNitrateNitrite +
#                DissOrthophos +
#                lagDissOrthophos,
#              data = df,
#              method = 'ML'
#              )

```

# AIC/BIC of models before smoothing predictors
```{r}
#AIC(m3,m4,m6,m3.no.season,m4.no.season,m6.no.season,m3.po4,m4.po4,m6.po4,m3.no.season.po4,m4.no.season.po4,m6.no.season.po4)
```

```{r}
#BIC(m3,m4,m6,m3.no.season,m4.no.season,m6.no.season,m3.po4,m4.po4,m6.po4,m3.no.season.po4,m4.no.season.po4,m6.no.season.po4)
```

# Shrink top models: {.tabset}
- EDF close to 0 it is effectively removed from the model
- linear terms have EDF close to 1

## m3

Notes:
s(OUT) has an EDF close to 0

```{r}
# OUT + Step (no EXPORT)
m3.shrink <- gam((logchla) ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity) +
               s(OUT) +
               s(Regions, bs='re') +
               season +
               (OUT*season) +
               clam_step +
               s(lagOUT) +
               s(lagDissAmmonia) +
               s(lagDissNitrateNitrite),
             data = df,
             method = 'ML',
             select=TRUE,
             na.action=na.omit)
gam.check(m3.shrink)

```
## m4

Notes:
S(SAC) has an EDF close to 0
raised conductivity k because EDF close to ref.df
```{r}
# SAC + Step (no EXPORT)
m4.shrink <- gam((logchla) ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity, k=12) +
               s(SAC) + 
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               s(lagSAC) +
               s(lagDissAmmonia),
               #s(lagDissNitrateNitrite), # removed based on dredge
             data = df,
             method = 'ML',
             select=TRUE)
summary(m4.shrink)
```
s(Sac) insignificant

## m6
```{r}

m6.shrink <- gam((logchla) ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity, k=12) +
               s(SAC) + 
               s(EXPORT) +
               s(Regions, bs='re') +
              # s(SAC*season) + #removed based on dredge
               season +
               clam_step +
               s(lagSAC) +
               s(lagDissAmmonia),
              # s(lagDissNitrateNitrite), # removed based on dredge
             data = df,
             method = 'ML', 
              select=TRUE
             )


summary(m6.shrink)
```
s(SAC) linear

## m3.no.season
```{r}
m3.no.season.shrink <- gam((logchla) ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity) +
               s(OUT) +
               s(Regions, bs='re') +
               clam_step +
               s(lagOUT) +
               s(lagDissAmmonia) +
               s(lagDissNitrateNitrite),
             data = df,
             method = 'ML',
             select=TRUE)
summary(m3.no.season.shrink)
```

## m4.no.season
```{r}
# SAC + Step (no EXPORT)
m4.no.season.shrink <- gam((logchla) ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity, k=12) +
               s(SAC) + 
               s(Regions, bs='re') +
               clam_step +
               s(lagSAC) +
               s(lagDissAmmonia), #+
              # s(lagDissNitrateNitrite),
             data = df,
             method = 'ML',
             select=TRUE)
summary(m4.no.season.shrink)
```

## m6.no.season

Notes:
EDF of s(EXPORT) is close to 1

```{r}
# SAC + EXPORT 
m6.no.season.shrink <- gam((logchla) ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity, k=12) +
               s(SAC) + 
               s(EXPORT) +
               s(Regions, bs='re') +
               clam_step +
               s(lagSAC) +
               s(lagDissAmmonia),# +
              # s(lagDissNitrateNitrite),
             data = df,
             method = 'ML',
             select=TRUE)
summary(m6.no.season.shrink)
```

## m3.po4
```{r}
# OUT  (no EXPORT)
m3.po4.shrink <- gam((logchla) ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity) +
               s(OUT) +
               s(Regions, bs='re') +
               season +
               (OUT*season) +
               clam_step +
               s(lagOUT) +
               s(lagDissAmmonia) +
               s(lagDissNitrateNitrite)+
               s(DissOrthophos)+
               s(lagDissOrthophos),
             data = df,
             method = 'ML',
             select=TRUE)
summary(m3.po4.shrink)
```
## m4.po4

Notes:
EDF of s(SAC) is close to 0

```{r}
# SAC (no EXPORT)
m4.po4.shrink <- gam((logchla) ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity, k=12) +
               s(SAC) + 
               s(Regions, bs='re') +
               season +
               (SAC*season) +
               clam_step +
               s(lagSAC) +
               s(lagDissAmmonia)+
             #  s(lagDissNitrateNitrite) +
               s(DissOrthophos)+
               s(lagDissOrthophos),
             data = df,
             method = 'ML',
             select=TRUE)
summary(m4.po4.shrink)
```


## m6.po4

EDF of s(SAC) is close

```{r}

m6.po4.shrink <- gam((logchla) ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity, k=12) +
               s(SAC) + 
               s(EXPORT) +
               s(Regions, bs='re') +
               season +
               (SAC:season) +
               clam_step +
               s(lagSAC) +
               s(lagDissAmmonia)+
             #  s(lagDissNitrateNitrite)+
               s(DissOrthophos)+
               s(lagDissOrthophos),
             data = df,
             method = 'ML', 
              select=TRUE
             )


summary(m6.po4.shrink)
```

## m3.no.season.po4
```{r}
m3.no.season.po4.shrink <- gam((logchla) ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity) +
               s(OUT) +
               s(Regions, bs='re') +
               clam_step +
               s(lagOUT) +
               s(lagDissAmmonia) +
               s(lagDissNitrateNitrite)+
               s(DissOrthophos)+
               s(lagDissOrthophos),
             data = df,
             method = 'ML',
             select=TRUE)
summary(m3.no.season.po4.shrink)
```

## m4.no.season.po4
```{r}
# SAC + Step (no EXPORT)
m4.no.season.po4.shrink <- gam((logchla) ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity, k=12) +
               s(SAC) + 
               s(Regions, bs='re') +
               clam_step +
               s(lagSAC) +
               s(lagDissAmmonia)+
             #  s(lagDissNitrateNitrite)+
               s(DissOrthophos)+
               s(lagDissOrthophos),
             data = df,
             method = 'ML',
             select=TRUE)
summary(m4.no.season.po4.shrink)
```

## m6.no.season.po4

EDF of S(EXPORT) close to 1

```{r}

m6.no.season.po4.shrink <- gam((logchla) ~ s(DissAmmonia) +
               s(DissNitrateNitrite) +
               s(Temperature) +
               s(Conductivity, k=12) +
               s(SAC) + 
               s(EXPORT) +
               s(Regions, bs='re') +
               clam_step +
               s(lagSAC) +
               s(lagDissAmmonia) +
             #  s(lagDissNitrateNitrite)+
               s(DissOrthophos)+
               s(lagDissOrthophos),
             data = df,
             method = 'ML',
             select=TRUE)
summary(m6.no.season.shrink)
```
# Closer look at models 

## m3 {.tabset}

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity)  +  s(OUT) +  s(Regions, bs='re') + season + (OUT*season) +
  clam_step + s(lagOUT) +s(lagDissAmmonia) + s(lagDissNitrateNitrite)

### concurvity

<0.3 - low, >0.5 - some concern, >0.8 - very high

Observations:
lagDNN x DNN = 0.5

```{r}
concurvity(m3.shrink)
```
Print predictor combinations with worst-case index of concurvity >0.5
```{r}
w<-concurvity(m3.shrink, full =FALSE)
wdf<- as.data.frame(w$worst)
i <- which(wdf > 0.5, arr.ind = TRUE)
r <- data.frame(
  Row = rownames(wdf)[i[, 1]],
  Column = colnames(wdf)[i[, 2]],
  Value = wdf[i]
)
r <- r[r$Row != r$Column, ]
r
```

### VIF


```{r}

m.l<- lm(formula = logchla ~ OUT  + clam_step,
            data=df)
vif(m.l)
```


### plotted predictions

```{r class.source = 'fold-hide'}

c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "OUT",
          "season",
          "Regions", 
          "clam_step",
          "lagOUT",
          "lagDissNitrateNitrite",
          "lagDissAmmonia")

lapply(c, function(x) {
  plot_predictions(m3.shrink, x , points = .5) +
    theme_classic(base_size = 12)
})

```

### marginal effects

The marginal effects in this plot can be interpreted as measuring the change in Y that is associated with a small increase in X, for different baseline values of X.

```{r class.source = 'fold-hide'}
c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "OUT",
          "season",
          "Regions", 
          "clam_step",
          "lagOUT",
          "lagDissAmmonia",
          "lagDissNitrateNitrite")

lapply(c, function(x) {
  plot_slopes(m3.shrink, variables = x, condition = x)
})

```

### diagnostics

```{r}
gam.check(m3.shrink)
```

## m4 {.tabset}

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity, k=12) + s(SAC) + 
               s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)

### concurvity

<0.3 - low, >0.5 - some concern, >0.8 - very high


```{r}
concurvity(m4.shrink)
```
Print predictor combinations with worst-case index of concurvity >0.5
```{r}
w <- concurvity(m4.shrink, full =FALSE)
wdf<- as.data.frame(w$worst)
i <- which(wdf > 0.5, arr.ind = TRUE)
r <- data.frame(
  Row = rownames(wdf)[i[, 1]],
  Column = colnames(wdf)[i[, 2]],
  Value = wdf[i]
)
r <- r[r$Row != r$Column, ]
r
```

### VIF

```{r}

m.l<- lm(formula = logchla ~ SAC + clam_step,
            data=df)
vif(m.l)
```

### plotted predictions

```{r class.source = 'fold-hide'}

c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "SAC",
          "season",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia")

lapply(c, function(x) {
  plot_predictions(m4.shrink, x , points = .5) +
    theme_classic(base_size = 12)
})

```

### marginal effects

The marginal effects in this plot can be interpreted as measuring the change in Y that is associated with a small increase in X, for different baseline values of X.

```{r class.source = 'fold-hide'}
c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "SAC",
          "season",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia")

lapply(c, function(x) {
  plot_slopes(m4.shrink, variables = x, condition = x)
})

```

### diagnostics

```{r}
gam.check(m4.shrink)
```

## m6 {.tabset}

m6: SAC + EXPORT

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity, k=12) + s(SAC) + 
               s(EXPORT) + s(Regions, bs='re') + season + clam_step + 
               s(lagSAC) + s(lagDissAmmonia) 

### concurvity

<0.3 - low, >0.5 - some concern, >0.8 - very high

observations:
EXPORT x SAC = 0.45

```{r}
concurvity(m6.shrink)
```
Print predictor combinations with worst-case index of concurvity >0.5
```{r}
w<-concurvity(m6.shrink, full =FALSE)
wdf<- as.data.frame(w$worst)
i <- which(wdf > 0.5, arr.ind = TRUE)
r <- data.frame(
  Row = rownames(wdf)[i[, 1]],
  Column = colnames(wdf)[i[, 2]],
  Value = wdf[i]
)
r <- r[r$Row != r$Column, ]
r
```

### VIF


```{r}

m.l<- lm(formula = logchla ~ SAC+ EXPORT + clam_step,
            data=df)
vif(m.l)
```

### plotted predicitons

```{r class.source = 'fold-hide'}
# Plot data with model fit
c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "SAC",
          "EXPORT",
          "season",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia")

lapply(c, function(x) {
  plot_predictions(m6.shrink, condition = x , points = .5) +
    theme_classic(base_size = 12)
})

```

### marginal effects


The marginal effects in this plot can be interpreted as measuring the change in Y that is associated with a small increase in X, for different baseline values of X.

```{r class.source = 'fold-hide'}

c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "SAC",
          "EXPORT",
          "season",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia")


lapply(c, function(x) {
  plot_slopes(m6.shrink, condition = x , variables = x) +
    theme_classic(base_size = 12)
})

```

### diagnostics

```{r}
gam.check(m6.shrink)
```

## m3.no.season {.tabset}

m3.no.season: OUT, no season or OUT:season

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity)
            + s(OUT) + s(Regions, bs='re') + clam_step + s(lagOUT) + s(lagDissAmmonia) + s(lagDissNitrateNitrite)

### concurvity

<0.3 - low, >0.5 - some concern, >0.8 - very high

```{r}
concurvity(m3.no.season.shrink)
```
Print predictor combinations with worst-case index of concurvity >0.5
```{r}
w<-concurvity(m3.no.season.shrink, full =FALSE)
wdf<- as.data.frame(w$worst)
i <- which(wdf > 0.5, arr.ind = TRUE)
r <- data.frame(
  Row = rownames(wdf)[i[, 1]],
  Column = colnames(wdf)[i[, 2]],
  Value = wdf[i]
)
r <- r[r$Row != r$Column, ]
r
```


### plotted predictions

```{r}

c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "OUT",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia")

lapply(c, function(x) {
  plot_predictions(m3.no.season.shrink, x , points = .5) +
    theme_classic(base_size = 12)
})

```

### marginal effects


The marginal effects in this plot can be interpreted as measuring the change in Y that is associated with a small increase in X, for different baseline values of X.

```{r}


c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "OUT",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia")


lapply(c, function(x) {
  plot_slopes(m3.no.season.shrink, variables = x , condition = x) +
    theme_classic(base_size = 12)
})

```

### diagnostics

```{r}
gam.check(m3.no.season.shrink)
```


## m4.no.season {.tabset}

m4.no.season: SAC, no season or SAC:season

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) +
              s(Conductivity, k=12) + s(SAC)+s(Regions, bs='re') + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(lagDissNitrateNitrite)

### concurvity

<0.3 - low, >0.5 - some concern, >0.8 - very high

```{r}
concurvity(m4.no.season.shrink)
```
Print predictor combinations with worst-case index of concurvity >0.5
```{r}
w<- concurvity(m4.no.season.shrink, full =FALSE)
wdf<- as.data.frame(w$worst)
i <- which(wdf > 0.5, arr.ind = TRUE)
r <- data.frame(
  Row = rownames(wdf)[i[, 1]],
  Column = colnames(wdf)[i[, 2]],
  Value = wdf[i]
)
r <- r[r$Row != r$Column, ]
r
```


### plotted predictions

```{r}

c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "SAC",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia",
          "lagDissNitrateNitrite")

lapply(c, function(x) {
  plot_predictions(m4.no.season.shrink, x , points = .5) +
    theme_classic(base_size = 12)
})

```

### marginal effects


The marginal effects in this plot can be interpreted as measuring the change in Y that is associated with a small increase in X, for different baseline values of X.

```{r}

c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "SAC",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia",
          "lagDissNitrateNitrite")

lapply(c, function(x) {
  plot_slopes(m4.no.season.shrink, condition = x , variables=x) +
    theme_classic(base_size = 12)
})

```

### diagnostics

```{r}
gam.check(m4.no.season.shrink)
```


## m6.no.season {.tabset}

m6.no.season: SAC + EXPORT, no season or SAC:season

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) +  s(Temperature) + s(Conductivity, k=12) +
               s(SAC) +  s(EXPORT) + s(Regions, bs='re') + clam_step + s(lagSAC) + s(lagDissAmmonia) + s(lagDissNitrateNitrite)

### concurvity

<0.3 - low, >0.5 - some concern, >0.8 - very high

```{r}
concurvity(m6.no.season.shrink)
```

Print predictor combinations with worst-case index of concurvity >0.5
```{r}
w<-concurvity(m6.no.season.shrink, full =FALSE)
wdf<- as.data.frame(w$worst)
i <- which(wdf > 0.5, arr.ind = TRUE)
r <- data.frame(
  Row = rownames(wdf)[i[, 1]],
  Column = colnames(wdf)[i[, 2]],
  Value = wdf[i]
)
r <- r[r$Row != r$Column, ]
r
```

### VIF


```{r}

m.l<- lm(formula = logchla ~ SAC+ EXPORT + clam_step,
            data=df)
vif(m.l)
```

### plotted predictions

```{r}
# Plot data with model fit
c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "SAC",
          "EXPORT",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia")

lapply(c, function(x) {
  plot_predictions(m6.no.season.shrink, x , points = .5) +
    theme_classic(base_size = 12)
})

```

### marginal effects

The marginal effects in this plot can be interpreted as measuring the change in Y that is associated with a small increase in X, for different baseline values of X.

```{r}

c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "SAC",
          "EXPORT",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia")

lapply(c, function(x) {
  plot_slopes(m6.no.season.shrink, variables = x , condition = x) +
    theme_classic(base_size = 12)
})

```


### diagnostics

```{r}
gam.check(m6.no.season.shrink)
```


## m3.po4 {.tabset}

m3.po4: OUT + PO4 + lagPO4

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) +s(Temperature) + s(Conductivity) + s(OUT) +
               s(Regions, bs='re') + season +(OUT*season) + clam_step + s(lagOUT) + s(lagDissAmmonia) +
               s(lagDissNitrateNitrite) + s(DissOrthophos) + s(lagDissOrthophos)

### concurvity

<0.3 - low, >0.5 - some concern, >0.8 - very high

Observations:
lagDNN x DNN = 0.5

```{r}
concurvity(m3.po4.shrink)
```

Print predictor combinations with worst-case index of concurvity >0.5
```{r}
w<-concurvity(m3.po4.shrink, full =FALSE)
wdf<- as.data.frame(w$worst)
i <- which(wdf > 0.5, arr.ind = TRUE)
r <- data.frame(
  Row = rownames(wdf)[i[, 1]],
  Column = colnames(wdf)[i[, 2]],
  Value = wdf[i]
)
r <- r[r$Row != r$Column, ]
r
```


### plotted predictions

```{r class.source = 'fold-hide'}
# Plot data with model fit
c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "OUT",
          "season",
          "Regions", 
          "clam_step",
          "lagOUT",
          "lagDissAmmonia",
          "DissOrthophos",
          "lagDissOrthophos")

lapply(c, function(x) {
  plot_predictions(m3.po4.shrink, x , points = .5) +
    theme_classic(base_size = 12)
})

```

### marginal effects

The marginal effects in this plot can be interpreted as measuring the change in Y that is associated with a small increase in X, for different baseline values of X.

```{r class.source = 'fold-hide'}
# Plot data with model fit
c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "OUT",
          "season",
          "Regions", 
          "clam_step",
          "lagOUT",
          "lagDissAmmonia",
          "DissOrthophos",
          "lagDissOrthophos")

lapply(c, function(x) {
  plot_slopes(m3.po4.shrink, variables = x, condition = x)
})

```

### diagnostics

```{r}
gam.check(m3.po4.shrink)
```

## m4.po4 {.tabset}

m4.po4: SAC + PO4 + lagPO4

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) +  s(Conductivity, k=12) +
               s(SAC) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) +
               s(lagDissAmmonia)  + s(DissOrthophos) + s(lagDissOrthophos)

### concurvity

<0.3 - low, >0.5 - some concern, >0.8 - very high


```{r}
concurvity(m4.po4.shrink)
```

```{r}
w<-concurvity(m4.po4.shrink, full =FALSE)
wdf<- as.data.frame(w$worst)
i <- which(wdf > 0.5, arr.ind = TRUE)
r <- data.frame(
  Row = rownames(wdf)[i[, 1]],
  Column = colnames(wdf)[i[, 2]],
  Value = wdf[i]
)
r <- r[r$Row != r$Column, ]
r
```

### plotted predictions

```{r class.source = 'fold-hide'}
# Plot data with model fit
c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "lagDissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "SAC",
          "season",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia",
          "DissOrthophos",
          "lagDissOrthophos")

lapply(c, function(x) {
  plot_predictions(m4.po4.shrink, x , points = .5) +
    theme_classic(base_size = 12)
})

```

### marginal effects

The marginal effects in this plot can be interpreted as measuring the change in Y that is associated with a small increase in X, for different baseline values of X.

```{r class.source = 'fold-hide'}
c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "SAC",
          "season",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia",
          "DissOrthophos",
          "lagDissOrthophos")

lapply(c, function(x) {
  plot_slopes(m4.po4.shrink, variables = x, condition = x)
})

```

### diagnostics

```{r}
gam.check(m4.po4.shrink)
```

## m6.po4 {.tabset}

m6.po4: SAC + EXPORT + PO4 + lagPO4

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity, k=12) +
               s(SAC) + s(EXPORT) + s(Regions, bs='re') +  season +  clam_step + s(lagSAC) +
               s(lagDissAmmonia) + s(DissOrthophos)+s(lagDissOrthophos)

### concurvity

<0.3 - low, >0.5 - some concern, >0.8 - very high

observations:
EXPORT x SAC = 0.45

```{r}
concurvity(m6.po4.shrink)
```
Print predictor combinations with worst-case index of concurvity >0.5
```{r}
w<-concurvity(m6.po4.shrink, full =FALSE)
wdf<- as.data.frame(w$worst)
i <- which(wdf > 0.5, arr.ind = TRUE)
r <- data.frame(
  Row = rownames(wdf)[i[, 1]],
  Column = colnames(wdf)[i[, 2]],
  Value = wdf[i]
)
r <- r[r$Row != r$Column, ]
r
```

### plotted predicitons

```{r class.source = 'fold-hide'}
# Plot data with model fit
c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "SAC",
          "EXPORT",
          "season",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia",
          "DissOrthophos",
          "lagDissOrthophos")

lapply(c, function(x) {
  plot_predictions(m6.po4.shrink, condition = x , points = .5) +
    theme_classic(base_size = 12)
})

```

### marginal effects


The marginal effects in this plot can be interpreted as measuring the change in Y that is associated with a small increase in X, for different baseline values of X.

```{r class.source = 'fold-hide'}

c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "SAC",
          "EXPORT",
          "season",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia",
          "lagDissNitrateNitrite",
          "DissOrthophos",
          "lagDissOrthophos")

lapply(c, function(x) {
  plot_slopes(m6.po4.shrink, condition = x , variables = x) +
    theme_classic(base_size = 12)
})

```

### diagnostics

```{r}
gam.check(m6.po4.shrink)
```

## m3.no.season.po4 {.tabset}

m3.no.season.po4: OUT + PO4 + lagPO4, no season or OUT:season

- gam((logchla) ~ s(DissAmmonia) +  s(DissNitrateNitrite) + s(Temperature) +
               s(Conductivity) + s(OUT) + s(Regions, bs='re') + clam_step +
               s(lagOUT) + s(lagDissAmmonia) + s(lagDissNitrateNitrite)+ s(DissOrthophos) + s(lagDissOrthophos)

### concurvity

<0.3 - low, >0.5 - some concern, >0.8 - very high

```{r}
concurvity(m3.no.season.po4.shrink)
```
Print predictor combinations with worst-case index of concurvity >0.5
```{r}
w<- concurvity(m3.no.season.po4.shrink, full =FALSE)
wdf<- as.data.frame(w$worst)
i <- which(wdf > 0.5, arr.ind = TRUE)
r <- data.frame(
  Row = rownames(wdf)[i[, 1]],
  Column = colnames(wdf)[i[, 2]],
  Value = wdf[i]
)
r <- r[r$Row != r$Column, ]
r
```


### plotted predictions

```{r}
# Plot data with model fit
c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "OUT",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia",
          "lagDissNitrateNitrite",
          "DissOrthophos",
          "lagDissOrthophos")

lapply(c, function(x) {
  plot_predictions(m3.no.season.po4.shrink, x , points = .5) +
    theme_classic(base_size = 12)
})

```

### marginal effects


The marginal effects in this plot can be interpreted as measuring the change in Y that is associated with a small increase in X, for different baseline values of X.

```{r}

c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "OUT",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia",
          "lagDissNitrateNitrite",
          "DissOrthophos",
          "lagDissOrthophos")

lapply(c, function(x) {
  plot_slopes(m3.no.season.po4.shrink, variables = x , condition = x) +
    theme_classic(base_size = 12)
})

```

### diagnostics

```{r}
gam.check(m3.no.season.po4.shrink)
```


## m4.no.season.po4 {.tabset}

m4.no.season.po4: SAC + PO4 + lagPO4, no season or SAC:season

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) +
               s(Conductivity, k=12) + s(SAC) +  s(Regions, bs='re') + clam_step + s(lagSAC) +
               s(lagDissAmmonia) + s(DissOrthophos)+ s(lagDissOrthophos)

### concurvity

<0.3 - low, >0.5 - some concern, >0.8 - very high

```{r}
concurvity(m4.no.season.po4.shrink)
```
Print predictor combinations with worst-case index of concurvity >0.5
```{r}
w<- concurvity(m4.no.season.po4.shrink, full =FALSE)
wdf<- as.data.frame(w$worst)
i <- which(wdf > 0.5, arr.ind = TRUE)
r <- data.frame(
  Row = rownames(wdf)[i[, 1]],
  Column = colnames(wdf)[i[, 2]],
  Value = wdf[i]
)
r <- r[r$Row != r$Column, ]
r
```


### plotted predictions

```{r}

c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "SAC",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia",
          "DissOrthophos",
          "lagDissOrthophos")

lapply(c, function(x) {
  plot_predictions(m4.no.season.po4.shrink, x , points = .5) +
    theme_classic(base_size = 12)
})

```

### marginal effects


The marginal effects in this plot can be interpreted as measuring the change in Y that is associated with a small increase in X, for different baseline values of X.

```{r}

c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "SAC",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia",
          "lagDissNitrateNitrite",
          "DissOrthophos",
          "lagDissOrthophos")

lapply(c, function(x) {
  plot_slopes(m4.no.season.po4.shrink, condition = x , variables=x) +
    theme_classic(base_size = 12)
})

```

### diagnostics

```{r}
gam.check(m4.no.season.po4.shrink)
```


## m6.no.season.po4 {.tabset}

m6.no.season.po4: SAC + EXPORT + PO4 + lagPO4, no season or SAC:season

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) +
               s(Conductivity, k=12) + s(SAC) + s(EXPORT) + s(Regions, bs='re') +
               clam_step + s(lagSAC) + s(lagDissAmmonia) + 
               s(DissOrthophos)+ s(lagDissOrthophos)

### concurvity

<0.3 - low, >0.5 - some concern, >0.8 - very high

```{r}
concurvity(m6.no.season.po4.shrink)
```
Print predictor combinations with worst-case index of concurvity >0.5
```{r}
w<-concurvity(m6.no.season.po4.shrink, full =FALSE)
wdf<- as.data.frame(w$worst)
i <- which(wdf > 0.5, arr.ind = TRUE)
r <- data.frame(
  Row = rownames(wdf)[i[, 1]],
  Column = colnames(wdf)[i[, 2]],
  Value = wdf[i]
)
r <- r[r$Row != r$Column, ]
r
```


### plotted predictions

```{r}
# Plot data with model fit
c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "SAC",
          "EXPORT",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia",
          "DissOrthophos",
          "lagDissOrthophos")

lapply(c, function(x) {
  plot_predictions(m6.no.season.shrink, x , points = .5) +
    theme_classic(base_size = 12)
})

```

### marginal effects

The marginal effects in this plot can be interpreted as measuring the change in Y that is associated with a small increase in X, for different baseline values of X.

```{r}
# Plot data with model fit
c <- c("DissAmmonia",
          "DissNitrateNitrite",
          "Temperature",
          "Conductivity",
          "SAC",
          "EXPORT",
          "Regions", 
          "clam_step",
          "lagSAC",
          "lagDissAmmonia",
          "lagDissNitrateNitrite",
          "DissOrthophos",
          "lagDissOrthophos")

lapply(c, function(x) {
  plot_slopes(m6.no.season.po4.shrink, variables = x , condition = x) +
    theme_classic(base_size = 12)
})

```


### diagnostics

```{r}
gam.check(m6.no.season.po4.shrink)
```

# Compare all the models with AIC/BIC

As a reminder, here are the models we are working with. Next to the model name I have written the main difference with the other models.


m3: OUT

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity)  +  s(OUT) +  s(Regions, bs='re') + season + (OUT*season) +
  clam_step + s(lagOUT) +s(lagDissAmmonia) + s(lagDissNitrateNitrite)

m4: SAC

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity, k=12) + s(SAC) + 
               s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)

m6: SAC + EXPORT

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity, k=12) + s(SAC) + 
               s(EXPORT) + s(Regions, bs='re') + season + clam_step + 
               s(lagSAC) + s(lagDissAmmonia)

m3.no.season: OUT, no season or OUT:season

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity)
            + s(OUT) + s(Regions, bs='re') + clam_step + s(lagOUT) + s(lagDissAmmonia) + s(lagDissNitrateNitrite)

m4.no.season: SAC, no season or SAC:season

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) +
              s(Conductivity, k=12) + s(SAC)+s(Regions, bs='re') + clam_step + s(lagSAC) + s(lagDissAmmonia) + s(lagDissNitrateNitrite)

m6.no.season: SAC + EXPORT, no season or SAC:season

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) +  s(Temperature) + s(Conductivity, k=12) +
               s(SAC) +  s(EXPORT) + s(Regions, bs='re') + clam_step + s(lagSAC) + s(lagDissAmmonia) +
               s(lagDissNitrateNitrite)
               
m3.po4: OUT + PO4 + lagPO4

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) +s(Temperature) + s(Conductivity) + s(OUT) +
               s(Regions, bs='re') + season +(OUT*season) + clam_step + s(lagOUT) + s(lagDissAmmonia) +
               s(lagDissNitrateNitrite) + s(DissOrthophos) + s(lagDissOrthophos)
               
           
m4.po4: SAC + PO4 + lagPO4

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) +  s(Conductivity, k=12) +
               s(SAC) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) +
               s(lagDissAmmonia) + s(lagDissNitrateNitrite) + s(DissOrthophos) + s(lagDissOrthophos)

m6.po4: SAC + EXPORT + PO4 + lagPO4

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity, k=12) +
               s(SAC) + s(EXPORT) + s(Regions, bs='re') +  season +  clam_step + s(lagSAC) +
               s(lagDissAmmonia)+ s(lagDissNitrateNitrite) + s(DissOrthophos)+s(lagDissOrthophos)

m3.no.season.po4: OUT + PO4 + lagPO4, no season or OUT:season

- gam((logchla) ~ s(DissAmmonia) +  s(DissNitrateNitrite) + s(Temperature) +
               s(Conductivity) + s(OUT) + s(Regions, bs='re') + clam_step +
               s(lagOUT) + s(lagDissAmmonia) + s(lagDissNitrateNitrite)+ s(DissOrthophos) + s(lagDissOrthophos)
               
m4.no.season.po4: SAC + PO4 + lagPO4, no season or SAC:season

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) +
               s(Conductivity, k=12) + s(SAC) +  s(Regions, bs='re') + clam_step + s(lagSAC) +
               s(lagDissAmmonia)+ s(lagDissNitrateNitrite) + s(DissOrthophos)+ s(lagDissOrthophos)
  
m6.no.season.po4: SAC + EXPORT + PO4 + lagPO4, no season or SAC:season

- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) +
               s(Conductivity, k=12) + s(SAC) + s(EXPORT) + s(Regions, bs='re') +
               clam_step + s(lagSAC) + s(lagDissAmmonia) + s(lagDissNitrateNitrite)+
               s(DissOrthophos)+ s(lagDissOrthophos)
        
## AIC
```{r}
aic_all<- AIC(m3.shrink,m4.shrink,m6.shrink,m3.no.season.shrink,m4.no.season.shrink,m6.no.season.shrink,m3.po4.shrink,m4.po4.shrink,m6.po4.shrink,m3.no.season.po4.shrink,m4.no.season.po4.shrink,m6.no.season.po4.shrink)
aic_all
```
## BIC
```{r}
bic_all <- BIC(m3.shrink,m4.shrink,m6.shrink,m3.no.season.shrink,m4.no.season.shrink,m6.no.season.shrink,m3.po4.shrink,m4.po4.shrink,m6.po4.shrink,m3.no.season.po4.shrink,m4.no.season.po4.shrink,m6.no.season.po4.shrink)
bic_all
```

## Determine top models

This table ranks the models by their AIC and shows the difference in AIC/BIC values of each model from the lowest AIC/BIC of all models. The code is taken from here: https://interagencyecologicalprogram.github.io/ND-FASTR/rtm_chl_models_flow_weekly_avg.html#Purpose 

m6.po4 and m4.po4 look to be performing the best

```{r}
# AIC values
aic_all_t <- aic_all %>%
  as_tibble(rownames = "Model") %>%
  mutate(Model_Number = 1:12, .before = Model)

# BIC values
bic_all_t <- bic_all %>%
  as_tibble(rownames = "Model")

# Combine AIC and BIC and calculate differences from lowest value
aic_bic_all <-
  left_join(aic_all_t, bic_all_t, by = join_by(Model, df)) %>%
  mutate(across(c(AIC, BIC), ~ .x - min(.x), .names = "{.col}_delta")) %>%
  select(starts_with("Model"), df, starts_with("AIC"), starts_with("BIC"))

# Sort by AIC
aic_bic_all %>% arrange(AIC)

```

# Impute

We are considering Multiple Imputations by Chained Equations (MICE) to impute missing data.

- Transforming the data first and then imputing is the preferred strategy.
(https://019b2da8-edfb-a262-61be-7973c056d9ae.share.connect.posit.cloud/mi-fitting.html#ref-von_hippel_2009)

- see this link for a graphic of the imputation process: https://stackoverflow.com/questions/54075836/multiple-imputed-datasets-pooling-results

- multiple imputed versions of the dataset are created, then each imputed dataset is analysed with the GAM, then results from each of the GAMs are "pooled" together.

- using Predictive Mean Matching (pmm)

Questions:
  - should we use random forest imputation instead of MICE? It is more straightforward but less robust.
  
  - what are the special considerations for using MICE with GAMs?
    - number of knots must be the same? (https://stats.stackexchange.com/questions/662448/gam-model-inference-following-mice
-imputation)

  - pooling median p-value (MPV) is useful for analyzing pooled GAMs 
      - "take the median of all imputed within-imputation GAM p-values as the pooled p-value measuring the strength of evidence for the association." https://pmc.ncbi.nlm.nih.gov/articles/PMC9123297/ 
    

```{r}
ini <- mice(df, maxit=0)
ini$nmis
```
```{r}
md.pattern(df)
```
    

# Attempt MICE imputation {.tabset}


## Impute

imputing with predictive mean matching (pmm)
```{r}
imp <- mice(df,seed = 124, print=F)
class(imp)
```

```{r}
imp$method
```

## inspect convergence

The plot shows the mean and the standard deviation of the imputed values only. Generally want to see the streams intermingle and be free of any trends at the later iterations. (https://www.gerkovink.com/miceVignettes/Convergence_pooling/Convergence_and_pooling.html)

```{r}
plot(imp)
```

```{r}
#check if convergence is maintained over more iterations.
imp40 <- mice.mids(imp, maxit=35, print=F)
plot(imp40)
```



## check imputation values

```{r}
stripplot(imp)
```

# Fit GAM with imputed datasets {.tabset}

```{r}
# create imputed datasets
df1<- complete(imp, action=1, include=FALSE)
df2<- complete(imp, action=2, include=FALSE)
df3<- complete(imp, action=3, include=FALSE)
df4<- complete(imp, action=4, include=FALSE)
df5<- complete(imp, action=5, include=FALSE)
```

```{r}
# fit m4.po4 with all 5 imputed datasets

#removing SAC here because it was shrunk to 0. Should SAC*season also be removed?

m4.po4.imp1 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity, k=12) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df1, method = 'ML', select=TRUE)

m4.po4.imp2 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity, k=12) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df2, method = 'ML', select=TRUE)

m4.po4.imp3 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity, k=12) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df3, method = 'ML', select=TRUE)

m4.po4.imp4 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity, k=12) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df4, method = 'ML', select=TRUE)

m4.po4.imp5 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity, k=12) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(lagDissOrthophos), data=df5, method = 'ML', select=TRUE)
```


```{r}
m6.po4.imp1 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity, k=12) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(EXPORT)+ s(lagDissOrthophos),data= df1, method = 'ML', select=TRUE)

m6.po4.imp2 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity, k=12) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(EXPORT)+ s(lagDissOrthophos),data= df2, method = 'ML', select=TRUE)

m6.po4.imp3 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity, k=12) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(EXPORT)+ s(lagDissOrthophos),data= df3, method = 'ML', select=TRUE)

m6.po4.imp4 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity, k=12) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(EXPORT)+ s(lagDissOrthophos),data= df4, method = 'ML', select=TRUE)

m6.po4.imp5 <- gam((logchla) ~ s(DissAmmonia) + s(DissNitrateNitrite) + s(Temperature) + s(Conductivity, k=12) + s(Regions, bs='re') + season + (SAC*season) + clam_step + s(lagSAC) + s(lagDissAmmonia)+ s(DissOrthophos)+ s(EXPORT)+ s(lagDissOrthophos),data= df5, method = 'ML', select=TRUE)

```

## pooled median p-values

### m4.po4

```{r}
model_list_m4 = list(m4.po4.imp1,m4.po4.imp2,m4.po4.imp3,m4.po4.imp4,m4.po4.imp5)

p_matrix <- sapply(model_list_m4, function(x) summary(x)$s.table[, 4])


median_p_per_term <- apply(p_matrix, 1, median)


print(median_p_per_term)
```

### m6.po4

```{r}
model_list_m6 = list(m6.po4.imp1,m6.po4.imp2,m6.po4.imp3,m6.po4.imp4,m6.po4.imp5)

p_matrix <- sapply(model_list_m6, function(x) summary(x)$s.table[, 4])


median_p_per_term <- apply(p_matrix, 1, median)


print(median_p_per_term)
```
## m4.po4 plotted predictions {.tabset}

### DissAmmonia

```{r}

lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "DissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### DissNitrateNitrite

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "DissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Temperature

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "Temperature", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Conductivity

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "Conductivity", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### season

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "season", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Regions

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "Regions", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### clam_step

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "clam_step", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### lagSAC

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "lagSAC", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissAmmonia

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "lagDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### DissOrthophos

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "DissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissOrthophos

```{r}
lapply(model_list_m4, function(mod) {
    plot_predictions(mod, condition = "lagDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

## m6.po4 plotted predicitons {.tabset}

### DissAmmonia

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "DissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### DissNitrateNitrite

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "DissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissNitrateNitrite

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "lagDissNitrateNitrite", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Temperature

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "Temperature", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Conductivity

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "Conductivity", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### season

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "season", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### Regions

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "Regions", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### clam_step

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "clam_step", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagSAC

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "lagSAC", points = 0.5) +
      theme_classic(base_size = 12)
})
```


### EXPORT

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "EXPORT", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissAmmonia

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "lagDissAmmonia", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### DissOrthophos

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "DissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```

### lagDissOrthophos

```{r}

lapply(model_list_m6, function(mod) {
    plot_predictions(mod, condition = "lagDissOrthophos", points = 0.5) +
      theme_classic(base_size = 12)
})
```


# Data exploration  

- outputs can be found in presentations folder, logchla exploration

```{r class.source = 'fold-hide'}
# hist(df$Chlorophyll)
# sqrt_chla <- sqrt(df$Chlorophyll)
# hist(sqrt_chla)
# df$log_chla <- log(df$Chlorophyll)
# hist(df$logchla)
# log10_chla <- log10(df$Chlorophyll)
# hist(log10_chla)
# 
# 
# log seems to have the best fit.
# 
# # biplots
# 
# #scatter{r}
# lapply(c, function(x) {
#   ggplot(df, mapping =aes(x = .data[[x]], y = logchla, color= Regions)) +
#     ylab("log(Chlorphyll a)")+
#     geom_point()+
#     facet_wrap(~season, ncol=1)+
#     theme_classic()})
# # 
# # # linear regression
# lapply(c, function(x) {
#   ggplot(df, mapping =aes(x = .data[[x]], y = logchla, color= Regions)) +
#     ylab("log(Chlorphyll a)")+
#     geom_smooth(method="lm", se=FALSE)+
#     facet_wrap(~season, ncol=1)+
#     theme_classic()})
# # 
# # 
# # ggplot(df, mapping =aes(x = Date, y=log_chla, color=Regions))+
# #   ylab("log(Chlorphyll a)")+
# #   geom_line()+
# #   scale_x_date(date_breaks= "10 years")+
# #   facet_wrap(~season, ncol=1)+
# #   theme_classic()
# # 
# # ggplot(df, mapping =aes(x = Date, y=log_chla))+
# #   ylab("log(Chlorphyll a)")+
# #   geom_line()+
# #   scale_x_date(date_breaks= "10 years")+
# #   facet_wrap(~season, ncol=1)+
# #   theme_classic()
# # 
# # # linear regression
# # lapply(c, function(x) {
# #   ggplot(df, mapping =aes(x = .data[[x]], y = logchla, color=season)) +
# #     ylab("log(Chlorphyll a)")+
# #     geom_smooth(method="lm", se=FALSE)+
# #     geom_point()+
# #     theme_classic()})
# 
#Corr plot
df_numeric = df2 %>% select(where(is.numeric))
df_numeric = df_numeric %>% drop_na()
m = cor(df_numeric)
# check syntax to remove NAs for individual predictors
corrplot(m,
         method = "color",
         type = "upper",
         addCoef.col = "black",
         number.cex = 0.5,
         tl.cex = 0.7,
         tl.col = "black",
         col = colorRampPalette(c("blue", "white", "red"))(200)
)
# 
# #######################################
# # Min Max normalized data exploration #
# #######################################
# 
# 
# df_minmax <- df %>% mutate(across(!c(Regions,season), ~as.vector(scale(.,center=min(.), scale = max(.)-min(.)))))
# 
# 
# # linear regression
# lapply(c, function(x) {
#   ggplot(df_minmax, mapping =aes(x = .data[[x]], y = logchla, color=season)) +
#     ylab("log(Chlorphyll a)")+
#     geom_smooth(method="lm", se=FALSE)+
#     geom_point()+
#     theme_classic()})
# 
# #######################################
# # z-score standardized data exploration #
# #######################################
# 
# 
# df_std <- df %>% mutate(across(!c(Regions,season), ~as.vector(scale(.))))
# 
# 
# # linear regression
lapply(c, function(x) {
  ggplot(df2, mapping =aes(x = .data[[x]], y = logchla, color=season)) +
    ylab("log(Chlorphyll a)")+
    geom_smooth(method="lm", se=FALSE)+
    geom_point()+
    theme_classic()})
lapply(c, function(x) {
  ggplot(df2, mapping =aes(x = .data[[x]], y = logchla)) +
    ylab("log(Chlorphyll a)")+
    geom_smooth(method="lm", se=TRUE)+
    geom_point()+
    theme_classic()})
# 
# #############################
# # Box Plot for Step Change  #
# #############################
# 
# 
ggplot(df2, mapping =aes(x = clam_step, y = logchla)) +
  ylab("log(Chlorphyll a)")+
  geom_boxplot()+
  theme_classic()

```


# Resources
https://gsp.humboldt.edu/olm/R/05_03_GAM.html

https://www.geeksforgeeks.org/r-language/generalized-additive-models-using-r/

GIF^(1/(2DF)) >2 - moderate collinearity, GIF>3.16 = high collinearity
https://019b2da8-edfb-a262-61be-7973c056d9ae.share.connect.posit.cloud/mlr-collinearity.html

GAM interpretation:
https://ecogambler.netlify.app/blog/interpreting-gams/

NA exclude
https://stats.stackexchange.com/questions/492955/should-i-use-na-omit-or-na-exclude-in-a-linear-model-in-r 
