---
title:
  "final Summer GAMs"
author:
  "Ryan Hankins"
output: 
  rmdformats::html_clean:
    lightbox: true
    thumbnails: false
---

# Prepare Data {.tabset}

```{r, include=FALSE}
library(tidyverse)
library(corrplot)
library(DT)
library(mgcv)
library(tidyverse)
library(marginaleffects)
library(stats)
library(mice)
library(ggplot2)
```



```{r}

# variables to include in the analysis:

df <- read.csv("Data/regional_integrated_dataset3.csv" )

df$Date = as.Date(df$Date)

# 1979 is the start of the study period.
df <- subset(df,df$Date >= "1979-01-01")
# 2022 is the end of the study period
df <- subset(df,df$Date <= "2022-12-31")

# Changing categorical variables to factors
# need to change regions to factor farther down so dataframe splitting by region works properly
df$Regions = as.factor(df$Regions)
df$season = as.factor(df$season)
df$season_num = as.numeric(df$season)

# log transforming chl-a to achieve normal distribution of response variable.
df$logchla <- log(df$Chlorophyll)

# log transforming some variables to better visualize
df$logDissNitrateNitrite <- log(df$DissNitrateNitrite)
df$logDissAmmonia <- log(df$DissAmmonia)
df$logDissOrthophos <- log(df$DissOrthophos)
df$logConductivity <- log(df$Conductivity)
df$logSAC <- log(df$SAC)
df$logOUT <- log(df$OUT)
df$logSJR <- log(df$SJR)

#Replacing clam variables with a step function (pre/post potamocorbula introduction to the Bay in 1987)
df <- df%>%mutate(clam_step=if_else(Year>1987, "post clam", "pre clam")) 
df$clam_step <- as.factor(df$clam_step)
df$Chlorophyll <- as.numeric(df$Chlorophyll)
```


## make the final dataset

```{r}
c = c("Chlorophyll","logchla", "Regions", "Date","Year", "Month", "season","SAC",
       "EXPORT","logEXPORT","OUT", "DissAmmonia", "DissNitrateNitrite", 
       "Temperature","Secchi", "Conductivity","season_num", "DissOrthophos","clam_step","logDissNitrateNitrite","logDissAmmonia","logDissOrthophos","logConductivity","logSAC","logOUT")

smr <- df %>% filter(season == "Summer")%>%subset(select=c)
```



## Separate dataset regionally

```{r}
# We need to remove the flow variables so we don't have duplicated values for each region

df_flow <- smr[,c("Date","Regions","SAC", "OUT",  "EXPORT", "logEXPORT","logOUT","logSAC")]

df_no_flow <- smr[,!(names(df) %in% c("SAC", "logSAC", "OUT", "logOUT", "EXPORT", "logEXPORT"))]

r <- unique(df_no_flow$Regions)
region_list <- list()
for (region in r){
  df_regional <- df_no_flow %>% filter(Regions == region)
  region <- gsub(" ","",region)
  region_list[[region]] <- df_regional
}
```

## Impute

Impute all of the regional datasets 5 times

```{r include=FALSE}

# This creates a list of mids objects

regional_imp <- list()
for (region_name in as.character(names(region_list))){
  current_df <- region_list[[region_name]]
  predictors <- quickpred(current_df)
  imp <- mice(current_df,seed = 123, print=F, predictorMatrix = predictors)
  new_name <- paste0(region_name, "_imp")
  regional_imp[[new_name]]<- imp
}
```

## Inspect convergence regionally {.tabset}

The plot shows the mean and the standard deviation of the imputed values only. Generally, we want to see the streams intermingle and be free of any trends at the later iterations. (https://www.gerkovink.com/miceVignettes/Convergence_pooling/Convergence_and_pooling.html)

### Central

```{r}
# Check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$Central_imp, maxit=35, print=F)
plot(imp40)
```


### Confluence

```{r}
#check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$Confluence_imp, maxit=35, print=F)
plot(imp40)
```


### North

```{r}
#check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$NorthDelta_imp, maxit=35, print=F)
plot(imp40)
```

### South

```{r}
#check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$South_imp, maxit=35, print=F)
plot(imp40)
```


### Suisun Bay

```{r}
#check if convergence is maintained over many iterations.
imp40 <- mice.mids(regional_imp$SuisunBay_imp, maxit=35, print=F)
plot(imp40)
```


## Make combined imputed datasets

```{r}
# Create imputed datasets for each region
regional_imp_dfs <- list()
action_seq <- as.list(seq(1,5))
for (region_imp in as.character(names(regional_imp))){
      current_imp <- regional_imp[[region_imp]]
      for (i in action_seq){
        df_new <- complete(current_imp,action = i, include=FALSE)
        df_new <- left_join( df_new,df_flow, by=c("Date","Regions"))
        new_name <- paste0(region_imp, "_df", i)
        regional_imp_dfs[[new_name]]<- df_new}}

```


## Add lag terms to the imputed datasets

```{r}
# This code is taken from clamdata_cleanup_AB_261125.R

df_list_lag <- lapply(regional_imp_dfs, function(y) {
#Step 1: summarize by season 
wq_r_sum_lagseason <- y %>%
  group_by(Regions,Year,season) %>%
  summarize(across(where(is.numeric),\(x) mean(x, na.rm = TRUE)), .groups="drop_last")

#Step 2: Add a column for the following season.
#This seems counter-intuitive because the lag should be the prior season,
#but this is a merge column: it will merge back with the prior season
#in the original table. 
wq_r_sum_lagseason <- wq_r_sum_lagseason %>%
  select(!c(lagOUT,lagSAC,lagTemperature,lagDissAmmonia,lagDissNitrateNitrite, lagTotPhos,lagConductivity,lagDissOrthophos))%>%
  rename_with(~paste0("lag",.x)) %>%
  rename(Year = "lagYear",season = "lagseason",Regions="lagRegions") %>% 
  mutate(lagseason = case_when(
    season == "Spring" ~ "Summer",
    season == "Summer" ~ "Autumn",
    season == "Autumn" ~ "Winter",
    season == "Winter" ~ "Spring" 
  )) %>%
  mutate(lagseasonyear = case_when(
    season == "Autumn" ~ Year+1,
    !(season == "Autumn") ~ Year
  )) %>%
  mutate(lagseasonyear = paste0(lagseason,lagseasonyear)) %>%
  ungroup %>% 
  select(lagseasonyear, Regions, laglogDissAmmonia,laglogDissNitrateNitrite,laglogConductivity,
         laglogSAC,laglogOUT, laglogDissOrthophos, laglogEXPORT)


y <- y %>% 
  mutate(seasonyear = paste0(season,year(Date))) %>%
  merge(wq_r_sum_lagseason,by.x=c("Regions","seasonyear"),by.y=c("Regions","lagseasonyear"),all.x=T) %>%
  select(-seasonyear)%>%
  relocate(Date)
})

```

## Combine the imputed datasets

Now we have to combine the 25 regional datasets into 5 datasets that include all regions.This is also where we are z-scoring the datasets for better comparison of the predictors' relative effects.

```{r}
### df1
df1 <- distinct(rbind(df_list_lag$Central_imp_df1, df_list_lag$Confluence_imp_df1,
             df_list_lag$NorthDelta_imp_df1, df_list_lag$South_imp_df1,df_list_lag$SuisunBay_imp_df1)) %>%
  # z-score as the last step before the GAM
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

### df2
df2 <- distinct(rbind(df_list_lag$Central_imp_df2, df_list_lag$Confluence_imp_df2,
             df_list_lag$NorthDelta_imp_df2, df_list_lag$South_imp_df2,df_list_lag$SuisunBay_imp_df2))%>%
 # z-score 
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

### df3

df3 <- distinct(rbind(df_list_lag$Central_imp_df3, df_list_lag$Confluence_imp_df3,
             df_list_lag$NorthDelta_imp_df3, df_list_lag$South_imp_df3,df_list_lag$SuisunBay_imp_df3))%>%
  # z-score 
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

### df4
df4 <- distinct(rbind(df_list_lag$Central_imp_df4, df_list_lag$Confluence_imp_df4,
             df_list_lag$NorthDelta_imp_df4, df_list_lag$South_imp_df4,df_list_lag$SuisunBay_imp_df4))%>%
  # z-score 
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))


### df5

df5 <- distinct(rbind(df_list_lag$Central_imp_df5, df_list_lag$Confluence_imp_df5,
             df_list_lag$NorthDelta_imp_df5, df_list_lag$South_imp_df5,df_list_lag$SuisunBay_imp_df5))%>%
  # z- score
  mutate(across(!c(Regions,season,Date,Month, Year, season_num, clam_step), ~as.vector(scale(.))))

# Make regions a factor
df1$Regions <- as.factor(df1$Regions)
df2$Regions <- as.factor(df2$Regions)
df3$Regions <- as.factor(df3$Regions)
df4$Regions <- as.factor(df4$Regions)
df5$Regions <- as.factor(df5$Regions)

```

