---
title:
  "Reduce Summer GAMs"
author:
  "Ryan Hankins"
output: 
  rmdformats::html_clean:
    lightbox: true
    thumbnails: false
---
# Starting models:

Summer Model 

Model 1: Summer Chla ~ s(logNH4) + s(logPO4) + s(log(NO2+NO3)) + s(temperature) + s(logconductivity) + Sac Valley index + previous year(Sac Valley index) + s(Region:logSAC) + s(Region:logSJ)+ s(Region:Exports) + Pre/post-clam + s(Region:lag(logSAC)) + s(Region:lag(logSJ)) + lag(logNH4) + s(lag(PO4)) + s(lag(logNO2+NO3)) + (1|Month)  

Model 2: Summer Chla ~ s(logNH4) + s(logPO4) + s(log(NO2+NO3)) + s(temperature) + s(logconductivity) + Sac Valley index + previous year(Sac Valley index) + Pre/post-clam + s(lag(logNDOI)) + s(Region:log(NDOI)) + lag(logNH4) + s(lag(PO4)) + s(lag(logNO2+NO3)) + (1|Month)  

# Prepare Data {.tabset}

```{r, include=FALSE}
library(tidyverse)
library(corrplot)
library(DT)
library(mgcv)
library(tidyverse)
library(marginaleffects)
library(stats)
library(mice)
library(ggplot2)
```



```{r}

# variables to include in the analysis:

df <- read.csv("Data/regional_integrated_dataset3.csv" )

df$Date = as.Date(df$Date)

# 1979 is the start of the study period.
df <- subset(df,df$Date >= "1979-01-01")
# 2022 is the end of the study period
df <- subset(df,df$Date <= "2022-12-31")

# Changing categorical variables to factors
# need to change regions to factor farther down so dataframe splitting by region works properly
df$Regions = as.factor(df$Regions)
df$season = as.factor(df$season)
df$season_num = as.numeric(df$season)

# log transforming chl-a to achieve normal distribution of response variable.
df$logchla <- log(df$Chlorophyll)

# log transforming some variables to better visualize
df$logDissNitrateNitrite <- log(df$DissNitrateNitrite)
df$logDissAmmonia <- log(df$DissAmmonia)
df$logDissOrthophos <- log(df$DissOrthophos)
df$logConductivity <- log(df$Conductivity)
df$logSAC <- log(df$SAC)
df$logOUT <- log(df$OUT)
df$logSJR <- log(df$SJR)

#Replacing clam variables with a step function (pre/post potamocorbula introduction to the Bay in 1987)
df <- df%>%mutate(clam_step=if_else(Year>1987, "post clam", "pre clam")) 
df$clam_step <- as.factor(df$clam_step)
df$Chlorophyll <- as.numeric(df$Chlorophyll)
```

## add lag terms
```{r}
# Create monthly lag terms
df<- df %>% arrange(Year, Month) %>% mutate(across(c(DissAmmonia,logDissAmmonia,DissOrthophos,logDissOrthophos,DissNitrateNitrite,logDissNitrateNitrite,Temperature, Conductivity,logConductivity, SAC, logSAC, SJR, logSJR, EXPORT, OUT, logOUT),
                                       ~ lag(.x, n = 1), 
                                       .names = "lag{.col}_mo"))

# Create seasonal lag terms for the log terms
# This code is taken from clamdata_cleanup_AB_261125.R
#Step 1: summarize by season 
wq_r_sum_lagseason <- df %>%
  group_by(Regions,Year,season) %>%
  summarize(across(where(is.numeric),\(x) mean(x, na.rm = TRUE)), .groups="drop_last")

#Step 2: Add a column for the following season.
#This seems counter-intuitive because the lag should be the prior season,
#but this is a merge column: it will merge back with the prior season
#in the original table. 
wq_r_sum_lagseason <- wq_r_sum_lagseason %>%
  select(!c(lagOUT,lagSAC,lagTemperature,lagDissAmmonia,lagDissNitrateNitrite, lagTotPhos,lagConductivity,lagDissOrthophos))%>%
  rename_with(~paste0("lag",.x)) %>%
  rename(Year = "lagYear",season = "lagseason", Regions="lagRegions") %>% 
  mutate(lagseason = case_when(
    season == "Spring" ~ "Summer",
    season == "Summer" ~ "Autumn",
    season == "Autumn" ~ "Winter",
    season == "Winter" ~ "Spring" 
  )) %>%
  mutate(lagseasonyear = case_when(
    season == "Autumn" ~ Year+1,
    !(season == "Autumn") ~ Year
  )) %>%
  mutate(lagseasonyear = paste0(lagseason,lagseasonyear)) %>%
  ungroup %>% 
  select(lagseasonyear, Regions, laglogDissAmmonia,laglogDissNitrateNitrite, laglogDissOrthophos, laglogConductivity,
         laglogSAC,laglogOUT, laglogSJR, laglogDissOrthophos, lagEXPORT)


df <- df %>% 
  mutate(seasonyear = paste0(season,year(Date))) %>%
  merge(wq_r_sum_lagseason,by.x=c("Regions","seasonyear"),by.y=c("Regions","lagseasonyear"),all.x=T) %>%
  select(-seasonyear)%>%
  relocate(Date)

```



## make the final dataset

```{r}
c = c("logchla", "season", "Regions", "Date","Year", "Month","SAC","lagSAC", "lagSAC_mo", "logSAC", "laglogSAC","laglogSAC_mo", "EXPORT", "lagEXPORT", "lagEXPORT_mo", "OUT","lagOUT","lagOUT_mo", "logOUT", "laglogOUT", "laglogOUT_mo", "SJR", "lagSJR","lagSJR_mo","logSJR", "laglogSJR", "laglogSJR_mo", "Index","lagIndex","DissAmmonia","lagDissAmmonia","lagDissAmmonia_mo","logDissAmmonia","laglogDissAmmonia","laglogDissAmmonia_mo","DissOrthophos","lagDissOrthophos","logDissOrthophos","laglogDissOrthophos","laglogDissOrthophos_mo","DissNitrateNitrite","lagDissNitrateNitrite","logDissNitrateNitrite","laglogDissNitrateNitrite","laglogDissNitrateNitrite_mo","Temperature","lagTemperature","lagTemperature_mo","Conductivity",  "lagConductivity","lagConductivity_mo","logConductivity","laglogConductivity_mo","clam_step")


smr <- df %>% filter(season == "Summer")%>%subset(select=c)
```


# Reduce Summer Model

## smr.1.test {.tabset}
We are using the complete cases data set here.

Model 1: Summer Chla ~ s(logNH4) + s(logPO4) + s(log(NO2+NO3)) + s(temperature) + s(logconductivity) + Sac Valley index + previous year(Sac Valley index) + s(Region:logSAC) + s(Region:logSJ)+ s(Region:Exports) + Pre/post-clam + s(Region:lag(logSAC)) + s(Region:lag(logSJ)) + lag(logNH4) + s(lag(PO4)) + s(lag(logNO2+NO3)) + (1|Month)

```{r cache = TRUE}
smr.1.test <- gam(logchla ~ s(logDissAmmonia) + s(logDissOrthophos) + s(logDissNitrateNitrite) + s(Temperature) + s(logConductivity, by=Regions) + s(Index) + s(lagIndex) + s(logSAC, by=Regions) + s(logSJR, by=Regions) + s(EXPORT, by=Regions) + s(laglogSAC_mo, by=Regions) + s(laglogSJR_mo, by=Regions) + clam_step + s(laglogDissAmmonia_mo) + s(laglogDissOrthophos_mo)+ s(laglogDissNitrateNitrite_mo) + s(Month, bs="re"), data=smr, method = 'ML', select=TRUE)
```
### summary 

```{r}
summary(smr.1.test)
```


### GAM Check
```{r}
gam.check(smr.1.test)
```
LogDissOrthophos, Index, lagIndex, laglogDissNitrateNitrite_mo, Month

### Plot Predictions

```{r}
c = c("logDissAmmonia","laglogDissAmmonia_mo","logDissOrthophos","laglogDissOrthophos_mo","logDissNitrateNitrite","laglogDissNitrateNitrite_mo","Temperature","lagTemperature_mo","laglogConductivity_mo","clam_step", "Month")
lapply(c, function(x) {
  plot_predictions(smr.1.test, x , points = .5) +
    theme_classic(base_size = 12)
})

```
Terms with regional interactions:

```{r}
c = c("logSAC", "laglogSAC_mo", "EXPORT","lagEXPORT_mo", "logSJR", "laglogSJR_mo", "Index","lagIndex","logConductivity")
lapply(c, function(x) {
    plot_predictions(smr.1.test, condition = c(x,"Regions"), points = 0.5) +
      theme_classic(base_size = 12)
})
```

## smr.2.test {.tabset}

Model 2: Summer Chla ~ s(logNH4) + s(logPO4) + s(log(NO2+NO3)) + s(temperature) + s(logconductivity) + Sac Valley index + previous year(Sac Valley index) + Pre/post-clam + s(lag(logNDOI)) + s(Region:log(NDOI)) + lag(logNH4) + s(lag(PO4)) + s(lag(logNO2+NO3)) + (1|Month) 

```{r cache=TRUE}

smr.2.test <- gam(logchla ~ s(logDissAmmonia) + s(logDissOrthophos) + s(logDissNitrateNitrite) + s(Temperature) + s(logConductivity, by=Regions) + s(Index) + s(lagIndex) + clam_step + s(logOUT) + s(logOUT, by = Regions) + s(laglogDissAmmonia_mo) + s(laglogDissOrthophos_mo)+ s(laglogDissNitrateNitrite_mo) + s(Month, bs="re"), data=smr, method = 'ML', select=TRUE)
```

### summary 

```{r}
summary(smr.2.test)
```

### GAM Check
```{r}
gam.check(smr.2.test)
```
logDissAmmonia

### Plot predicitons

```{r}
c = c( "logDissAmmonia","laglogDissAmmonia","logDissOrthophos","laglogDissOrthophos_mo","logDissNitrateNitrite","laglogDissNitrateNitrite_mo","Temperature","lagTemperature_mo","laglogConductivity_mo","clam_step", "Month")
lapply(c, function(x) {
  plot_predictions(smr.2.test, x , points = .5) +
    theme_classic(base_size = 12)
})

```
Terms with regional interactions:

```{r}
c = c("logOUT",  "laglogOUT_mo", "Index","lagIndex","logConductivity")
lapply(c, function(x) {
    plot_predictions(smr.1.test, condition = c(x,"Regions"), points = 0.5) +
      theme_classic(base_size = 12)
})
```
## AIC/BIC
notes: smr_1_test performs better based on both AIC and BIC

AIC
```{r}
AIC(smr.1.test,smr.2.test)
```
BIC
```{r}
BIC(smr.1.test, smr.2.test)
```

# Run reduced summer models


## smr.1 {.tabset}

running smr.1 without LogDissOrthophos, Index, lagIndex, laglogDissNitrateNitrite_mo
retaining Month random effect

```{r cache = TRUE}
smr.1 <- gam(logchla ~ s(logDissAmmonia) + s(logDissNitrateNitrite) + s(Temperature) + s(logConductivity, by=Regions) + s(logSAC, by=Regions) + s(logSJR, by=Regions) + s(EXPORT, by=Regions) + s(laglogSAC_mo, by=Regions) + s(laglogSJR_mo, by=Regions) + clam_step + s(laglogDissAmmonia_mo) + s(laglogDissOrthophos_mo) + s(Month, bs="re"), data=smr, method = 'ML', select=TRUE)
```


### GAM Check
```{r}
gam.check(smr.1)
```
### Summary
```{r}
summary(smr.1)
```

## smr.2 {.tabset}

running smr.2 without logDissAmmonia

```{r cache = TRUE}


smr.2 <- gam(logchla ~ s(logDissOrthophos) + s(logDissNitrateNitrite) + s(Temperature) + s(logConductivity, by=Regions) + s(Index) + s(lagIndex) + clam_step + s(logOUT) + s(logOUT, by = Regions) + s(laglogDissAmmonia_mo) + s(laglogDissOrthophos_mo)+ s(laglogDissNitrateNitrite_mo) + s(Month, bs="re"), data=smr, method = 'ML', select=TRUE)
```

### GAM Check
```{r}
gam.check(smr.2)
```

### Summary
```{r}
summary(smr.2)
```

## Compare AIC/BIC of all summer models 

```{r}
AIC(smr.1.test,smr.1,smr.2.test, smr.2)
```

```{r}
BIC(smr.1.test,smr.1,smr.2.test, smr.2)
```

