---
title:
  "Seasonal GAM Data Exploration"
author:
  "Ryan Hankins"
output: 
  rmdformats::html_clean:
    lightbox: true
    thumbnails: false
    gallery: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, tidy = "styler")
```
# Models
Spring Model 

Model 1: Spring Chla ~ s(logNH4) + s(logPO4) + s(log(NO2+NO3)) + s(temperature) + s(logconductivity) + Sac Valley index + previous year(Sac Valley index) + s(Region:logSAC) + s(Region:logSJ)+ s(Region:logExports) + Pre/post-clam + s(Region:lag(logSAC)) + s(Region:lag(logSJ)) + lag(logNH4) + s(lag(PO4)) + s(lag(logNO2+NO3)) + (1|Month)  

Model 2: Spring Chla ~ s(logNH4) + s(logPO4) + s(log(NO2+NO3)) + s(temperature) + s(logconductivity) + Sac Valley index + previous year(Sac Valley index) + Pre/post-clam + s(lag(logNDOI)) + s(Region:log(NDOI)) + lag(logNH4) + s(lag(PO4)) + s(lag(logNO2+NO3)) + (1|Month)  

Summer Model 

Model 1: Summer Chla ~ s(logNH4) + s(logPO4) + s(log(NO2+NO3)) + s(temperature) + s(logconductivity) + Sac Valley index + previous year(Sac Valley index) + s(Region:logSAC) + s(Region:logSJ)+ s(Region:logExports) + Pre/post-clam + s(Region:lag(logSAC)) + s(Region:lag(logSJ)) + lag(logNH4) + s(lag(PO4)) + s(lag(logNO2+NO3)) + (1|Month)  

Model 2: Summer Chla ~ s(logNH4) + s(logPO4) + s(log(NO2+NO3)) + s(temperature) + s(logconductivity) + Sac Valley index + previous year(Sac Valley index) + Pre/post-clam + s(lag(logNDOI)) + s(Region:log(NDOI)) + lag(logNH4) + s(lag(PO4)) + s(lag(logNO2+NO3)) + (1|Month)  

# Prepare Data {.tabset}

```{r, include=FALSE}
library(tidyverse)
library(corrplot)
library(DT)
```


```{r}

# variables to include in the analysis:

df <- read.csv("Data/regional_integrated_dataset3.csv" )

df$Date = as.Date(df$Date)

# 1979 is the start of the study period.
df <- subset(df,df$Date >= "1979-01-01")
# 2022 is the end of the study period
df <- subset(df,df$Date <= "2022-12-31")

# Changing categorical variables to factors
# need to change regions to factor farther down so dataframe splitting by region works properly
#df$Regions = as.factor(df$Regions)
df$season = as.factor(df$season)
df$season_num = as.numeric(df$season)

# log transforming chl-a to achieve normal distribution of response variable.
df$logchla <- log(df$Chlorophyll)

# log transforming some variables to better visualize
df$logDissNitrateNitrite <- log(df$DissNitrateNitrite)
df$logDissAmmonia <- log(df$DissAmmonia)
df$logDissOrthophos <- log(df$DissOrthophos)
df$logConductivity <- log(df$Conductivity)
df$logSAC <- log(df$SAC)
df$logOUT <- log(df$OUT)
df$logSJR <- log(df$SJR)

#Replacing clam variables with a step function (pre/post potamocorbula introduction to the Bay in 1987)
df <- df%>%mutate(clam_step=if_else(Year>1987, "post clam", "pre clam")) 
df$clam_step <- as.factor(df$clam_step)
df$Chlorophyll <- as.numeric(df$Chlorophyll)
```

## add lag terms
```{r}
# Create monthly lag terms
df<- df %>% arrange(Year, Month) %>% mutate(across(c(DissAmmonia,logDissAmmonia,DissOrthophos,logDissOrthophos,DissNitrateNitrite,logDissNitrateNitrite,Temperature, Conductivity,logConductivity, SAC, logSAC, SJR, logSJR, EXPORT, OUT, logOUT, Index),
                                       ~ lag(.x, n = 1), 
                                       .names = "lag{.col}_mo"))

# Create seasonal lag terms for the log terms
# This code is taken from clamdata_cleanup_AB_261125.R
#Step 1: summarize by season 
wq_r_sum_lagseason <- df %>%
  group_by(Regions,Year,season) %>%
  summarize(across(where(is.numeric),\(x) mean(x, na.rm = TRUE)), .groups="drop_last")

#Step 2: Add a column for the following season.
#This seems counter-intuitive because the lag should be the prior season,
#but this is a merge column: it will merge back with the prior season
#in the original table. 
wq_r_sum_lagseason <- wq_r_sum_lagseason %>%
  select(!c(lagOUT,lagSAC,lagTemperature,lagDissAmmonia,lagDissNitrateNitrite, lagTotPhos,lagConductivity,lagDissOrthophos))%>%
  rename_with(~paste0("lag",.x)) %>%
  rename(Year = "lagYear",season = "lagseason", Regions="lagRegions") %>% 
  mutate(lagseason = case_when(
    season == "Spring" ~ "Summer",
    season == "Summer" ~ "Autumn",
    season == "Autumn" ~ "Winter",
    season == "Winter" ~ "Spring" 
  )) %>%
  mutate(lagseasonyear = case_when(
    season == "Autumn" ~ Year+1,
    !(season == "Autumn") ~ Year
  )) %>%
  mutate(lagseasonyear = paste0(lagseason,lagseasonyear)) %>%
  ungroup %>% 
  select(lagseasonyear, Regions, laglogDissAmmonia,laglogDissNitrateNitrite, laglogDissOrthophos, laglogConductivity,
         laglogSAC,laglogOUT, laglogSJR, laglogDissOrthophos, lagEXPORT)


df <- df %>% 
  mutate(seasonyear = paste0(season,year(Date))) %>%
  merge(wq_r_sum_lagseason,by.x=c("Regions","seasonyear"),by.y=c("Regions","lagseasonyear"),all.x=T) %>%
  select(-seasonyear)%>%
  relocate(Date)

```

## make the final dataset

```{r}
c = c("logchla", "season", "Regions", "Date","Year", "Month","SAC","lagSAC", "lagSAC_mo", "logSAC", "laglogSAC","laglogSAC_mo", "EXPORT", "lagEXPORT", "lagEXPORT_mo", "OUT","lagOUT","lagOUT_mo", "logOUT", "laglogOUT", "laglogOUT_mo", "SJR", "lagSJR","lagSJR_mo","logSJR", "laglogSJR", "laglogSJR_mo", "Index","lagIndex","DissAmmonia","lagDissAmmonia","lagDissAmmonia_mo","logDissAmmonia","laglogDissAmmonia","laglogDissAmmonia_mo","DissOrthophos","lagDissOrthophos","logDissOrthophos","laglogDissOrthophos","laglogDissOrthophos_mo","DissNitrateNitrite","lagDissNitrateNitrite","logDissNitrateNitrite","laglogDissNitrateNitrite","laglogDissNitrateNitrite_mo","Temperature","lagTemperature","lagTemperature_mo","Conductivity",  "lagConductivity","lagConductivity_mo","logConductivity","laglogConductivity_mo","clam_step")

df <- df %>% subset(select=c)

smr <- df %>% filter(season == "Summer") 
smr_z <- smr %>% mutate(across(!c(Regions,season,Date,Month, Year, clam_step), ~as.vector(scale(.))))
spr <- df %>% filter(season == "Spring") %>% subset(select = c)
spr_z <- spr %>% mutate(across(!c(Regions,, season, Date,Month, Year,clam_step), ~as.vector(scale(.))))

```


# Summer {.tabset}


## correlation matrix

```{r}

smr_numeric = smr_z %>% select(where(is.numeric))
smr_numeric = smr_numeric %>% drop_na()
m = cor(smr_numeric)

m_df <- data.frame(m)
rmarkdown::paged_table(m_df)
```

```{r  out.width = "3000px"}
corrplot(m,
         method = "color",
         type = "upper",
         addCoef.col = "black",
         number.cex = 0.25,
         number.digits=2,
         tl.cex =0.45,
         tl.col = "black")
```

## bivariate plots (z-scored)

```{r}

lapply(c, function(x) {
  ggplot(smr_z, mapping =aes(x = .data[[x]], y = logchla)) +
    ylab("log(Chlorphyll a)")+
    geom_smooth(method="lm", se=FALSE)+
    geom_point()+
    theme_classic()})
```

## bivariate plots by region (z-scored)

```{r}

lapply(c, function(x) {
  ggplot(smr_z, mapping =aes(x = .data[[x]], y = logchla, color = Regions)) +
    ylab("log(Chlorphyll a)")+
    geom_smooth(method="lm", se=FALSE)+
    geom_point()+
    theme_classic()})
```




## bivariate plots (non-standardized data)


```{r}

lapply(c, function(x) {
  ggplot(smr, mapping =aes(x = .data[[x]], y = logchla)) +
    ylab("log(Chlorphyll a)")+
    geom_smooth(method="lm", se=FALSE)+
    geom_point()+
    theme_classic()})

```



# Spring {.tabset}


## correlation matrix

```{r}

spr_numeric = spr_z %>% select(where(is.numeric))
spr_numeric = spr_numeric %>% drop_na()
m = cor(spr_numeric)

m_df <- data.frame(m)
rmarkdown::paged_table(m_df)
```

```{r out.width = "3000px"}

corrplot(m,
         method = "color",
         type = "upper",
         addCoef.col = "black",
         number.cex = 0.25,
         number.digits=2,
         tl.cex =0.45,
         tl.col = "black")
```


## bivariate plots (z-scored)

```{r}

# linear regression
lapply(c, function(x) {
  ggplot(spr_z, mapping =aes(x = .data[[x]], y = logchla)) +
    ylab("log(Chlorphyll a)")+
    geom_smooth(method="lm", se=FALSE)+
    geom_point()+
    theme_classic()})
```

## bivariate plots by region (z-scored)

```{r}

# linear regression
lapply(c, function(x) {
  ggplot(spr_z, mapping =aes(x = .data[[x]], y = logchla, color = Regions)) +
    ylab("log(Chlorphyll a)")+
    geom_smooth(method="lm", se=FALSE)+
    geom_point()+
    theme_classic()})
```


## bivariate plots (non-standardized data)

```{r}

# linear regression
lapply(c, function(x) {
  ggplot(spr, mapping =aes(x = .data[[x]], y = logchla)) +
    ylab("log(Chlorphyll a)")+
    geom_smooth(method="lm", se=FALSE)+
    geom_point()+
    theme_classic()})
```

