---
title: "Tuolumne Stock-Recruit Analysis"
author: "Laura"
date: "2024-12-18"
output: html_document
---

Setup
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = FALSE)
```

Read in packages
```{r Read in Packages, include=FALSE}
library(ggplot2)
library(plyr)
library(dplyr)
library(splines)
library(tidyverse)
library(caret)
library(ggrepel)
library(here)
library(FSA)
library(nlstools)
library(corrplot)
theme_set(theme_classic())
```

File Location
```{r File Location, include=FALSE}
data_input <- here("~/Sac_Delta_analyses/stock_recruit/tuolumne-stock-recruit-habitat-files")
figures<- here("~/Sac_Delta_analyses/stock_recruit/figures")  
waterford_rst<-read_csv(here("Waterford_RST_2022.csv"))
```

Read-in and manipulate data
```{r Read-in and manipulate data, include=FALSE}
#join covariate data, log values for habitat and flow, and calc. z-scores
tuol_flow<-read_csv(here(data_input, "flow_covariates_tuolumne.csv"))%>%select(-c(gage_id,source))%>%mutate(log_value=log(value))%>%pivot_wider(names_from=covariate, values_from=c(value, log_value)) #log flow because of diminishing returns of increasing flow (as in Munsch et al. 2020)

tuol_temp<-read_csv(here(data_input, "temperature_covariates_tuolumne.csv"))%>%na.omit()%>%mutate(log_value=log(value))%>%pivot_wider(names_from=covariate, values_from=c(value, log_value))

tuol_habitat<-read_csv(here(data_input, "habitat_covariates_tuolumne.csv"))%>%select(-stream)%>%mutate(log_value=log(value+1))%>%pivot_wider(names_from=c(habitat_type, covariate), values_from=c(value, log_value)) 
waterford<-waterford_rst%>%left_join(tuol_flow,by=c("Outmigration.year"="year"))%>%left_join(tuol_temp,by=c("Outmigration.year"="year"))%>%left_join(tuol_habitat, by=c("Outmigration.year"="year"))%>%filter(Outmigration.year!=2020)%>%select(-c("covariate_type.x", "covariate_type.y"))%>%pivot_longer(cols=c(-Outmigration.year, -spawner_abundance, -Waterford_abundance, -Grayson_abundance), names_to="Covariate_name", values_to="Covariate_value") #remove 2020 due to COVID sampling gaps
```

Explore Relationships Among Juvenile Abundance and Covariates
```{r Explore Relationships Among Juvenile Abundance and Covariates}
ggplot(waterford, aes(x=Covariate_value, y=(log(Waterford_abundance))))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw()

ggplot(waterford, aes(x=Covariate_value, y=(log(Grayson_abundance))))+
  geom_point()+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw()
```

Examine collinearity among covariates
```{r Examine collinearity among covariates, include=FALSE}
#filter covariates
covariates_list<-c("log_value_spawning_mean_spawn_acres", "log_value_spawning_mean_wetted_acres", "log_value_in_channel_mean_ic_rearing_acres", "log_value_in_channel_mean_wetted_acres", "log_value_total_rearing_mean_total_rearing_acres", "log_value_total_rearing_mean_wetted_acres", "log_value_floodplain_mean_acres_inundated_wua", "log_value_floodplain_mean_acres_inundated_wetted_area","log_value_mean_temp_oct_nov","log_value_mean_temp_may","log_value_mean_jan_may_flow")
waterford_reduced<-waterford%>%filter(Covariate_name %in% covariates_list)%>%na.omit()

waterford_mat<-waterford_reduced%>%pivot_wider(names_from=Covariate_name, values_from=Covariate_value)%>%select(-c(Outmigration.year, spawner_abundance, Waterford_abundance, Grayson_abundance))%>%na.omit()
m=cor(waterford_mat)
colnames(m)<-c("Jan-May flow", "Oct-Nov temp", "Spawning wetted", "Spawning WUA", "Floodplain wetted","Floodplain WUA", "In-channel wetted", "In-channel WUA", "Total rearing wetted", "Total rearing WUA", "May temp") 
rownames(m)<-c("Jan-May flow", "Oct-Nov temp", "Spawning wetted", "Spawning WUA", "Floodplain wetted","Floodplain WUA", "In-channel wetted", "In-channel WUA", "Total rearing wetted", "Total rearing WUA", "May temp")
png(filename="~/Sac_Delta_analyses/stock_recruit/figures/Tuolumne_corrplot.png", width=800, height=600)
corrplot(m, method = "number", type = "upper") # show only upper side
dev.off()
```

```{r print corrplot}
corrplot(m, method = "number", type = "upper") 
```

Linear Waterford models (upstream RST) single environmental covariate
```{r Linear Waterford models with single environmental covariate (upstream RST)}
linear_sr_waterford<-function(covariate){
  {lm(log(Waterford_abundance)~log(spawner_abundance) + Covariate_value -1, data=filter(waterford_reduced, covariate==Covariate_name), na.action=na.omit)
  }
}

linear_sr_waterford_models<-map(set_names(unique(waterford_reduced$Covariate_name)), linear_sr_waterford)
linear_sr_models_resids<-map_dfr(linear_sr_waterford_models, function(x) resid(x)%>%as_tibble(), .id = "Covariate_name")%>%
  dplyr::rename(resid=value)

ggplot(linear_sr_models_resids, aes(x=resid))+
  geom_histogram(bins=15)+
  facet_wrap(~Covariate_name, scales = "free")+
  theme_bw()                               

map(linear_sr_waterford_models, summary)
map(linear_sr_waterford_models, AIC)
map(linear_sr_waterford_models, coefficients)
```

Linear Waterford models with Oct-Nov temp and spring covariates
```{r Linear Waterford models with Oct_nov temp + spring covariates}

waterford_reduced<-waterford_reduced%>%mutate(covariate_season=if_else(Covariate_name%in%c("log_value_mean_temp_oct_nov", "log_value_spawning_mean_wetted_acres", "log_value_spawning_mean_spawn_acres"), "fall", "spring")) #dive into two data frames and then recombine
waterford_reduced_fall<-waterford_reduced%>%filter(covariate_season=="fall")%>%rename(c(Fall_covariate_name=Covariate_name, Fall_covariate_value=Covariate_value))%>%select(-covariate_season)
waterford_reduced_spring<-waterford_reduced%>%filter(covariate_season=="spring")%>%rename(c(Spring_covariate_name=Covariate_name, Spring_covariate_value=Covariate_value))%>%select(-covariate_season)
waterford_reduced_fall_spring<-full_join(waterford_reduced_fall, waterford_reduced_spring)

#run with temperature as the fall covariate value
waterford_reduced_fall_spring_temp<-waterford_reduced_fall_spring%>%filter(Fall_covariate_name=="log_value_mean_temp_oct_nov")
linear_sr_waterford_fall_spring_temp<-function(covariate){
  {lm(log(Waterford_abundance)~log(spawner_abundance) + Fall_covariate_value + Spring_covariate_value -1, data=filter(waterford_reduced_fall_spring_temp, covariate==Spring_covariate_name), na.action=na.omit)
  }
}

linear_sr_waterford_models_fall_spring_temp<-map(set_names(unique(waterford_reduced_fall_spring_temp$Spring_covariate_name)), linear_sr_waterford_fall_spring_temp)
linear_sr_models_resids_fall_spring_temp<-map_dfr(linear_sr_waterford_models_fall_spring_temp, function(x) resid(x)%>%as_tibble(), .id = "Spring_covariate_name")%>%
  dplyr::rename(resid=value)

ggplot(linear_sr_models_resids_fall_spring_temp, aes(x=resid))+
  geom_histogram(bins=15)+
  facet_wrap(~Spring_covariate_name, scales = "free")+
  theme_bw()                               

map(linear_sr_waterford_models_fall_spring_temp, summary)
map(linear_sr_waterford_models_fall_spring_temp, AIC)
linear_sr_coefficients_fall_spring_temp<-map(linear_sr_waterford_models_fall_spring_temp, coefficients)
```

Linear Waterford models with Spawning WUA and spring covariates
```{r Linear Waterford models with Spawning WUA and spring covariates}

waterford_reduced_fall_spring_spawn_wua<-waterford_reduced_fall_spring%>%filter(Fall_covariate_name=="log_value_spawning_mean_spawn_acres")
linear_sr_waterford_fall_spring_spawn_wua<-function(covariate){
  {lm(log(Waterford_abundance)~log(spawner_abundance) + Fall_covariate_value + Spring_covariate_value -1, data=filter(waterford_reduced_fall_spring_spawn_wua, covariate==Spring_covariate_name), na.action=na.omit)
  }
}

linear_sr_waterford_models_fall_spring_spawn_wua<-map(set_names(unique(waterford_reduced_fall_spring_spawn_wua$Spring_covariate_name)), linear_sr_waterford_fall_spring_spawn_wua)
linear_sr_models_resids_fall_spring_spawn_wua<-map_dfr(linear_sr_waterford_models_fall_spring_spawn_wua, function(x) resid(x)%>%as_tibble(), .id = "Spring_covariate_name")%>%
  dplyr::rename(resid=value)

ggplot(linear_sr_models_resids_fall_spring_spawn_wua, aes(x=resid))+
  geom_histogram(bins=15)+
  facet_wrap(~Spring_covariate_name, scales = "free")+
  theme_bw()                               

map(linear_sr_waterford_models_fall_spring_spawn_wua, summary)
map(linear_sr_waterford_models_fall_spring_spawn_wua, AIC)
linear_sr_coefficients_fall_spring_spawn_wua<-map(linear_sr_waterford_models_fall_spring_spawn_wua, coefficients)
```

Linear Waterford model with spawning WUA and Oct-Nov temp
```{r Linear Waterford model with spawning WUA and Oct-Nov temp}

waterford_wide<-waterford_reduced%>%pivot_wider(names_from= Covariate_name, values_from = Covariate_value)
waterford_fall_lm2<-lm(log(Waterford_abundance)~log(spawner_abundance) + log_value_mean_temp_oct_nov + log_value_spawning_mean_spawn_acres -1, data=waterford_wide, na.action=na.omit)
AIC(waterford_fall_lm2)
summary(waterford_fall_lm2)
```

Waterford Ricker Models: single environmental covariate
```{r Waterford Ricker Models with single environmental covariate}
r<-srFuns(type="Ricker", param=3)
rs<-srStarts(Waterford_abundance~spawner_abundance, data=waterford_reduced, type="Ricker")
unlist(rs)

#run Ricker model with no covariates
r_starters_list <- list(a = rs$a,
                        b = rs$b)

waterford_ricker <- nls(log(Waterford_abundance) ~
               log((a * spawner_abundance) * exp(-b * spawner_abundance)),
             data = waterford_reduced,
             start = r_starters_list)
summary(waterford_ricker)
AIC(waterford_ricker)
plot(waterford_ricker)

r_starters_list2 <- list(a = rs$a,
                        b = rs$b, f=0.2)


#flow covariate model- refit on the waterford_reduced dataset
waterford_ricker_flow <- nls(log(Waterford_abundance) ~
               log((a * spawner_abundance) * exp(-b * spawner_abundance))+ f*log_value_mean_jan_may_flow,
             data = waterford_wide, start = c(a = rs$a, b = rs$b, f=1), lower = c(0.1, 0.0000001, 0.1), algorithm = "port")
summary(waterford_ricker_flow)
AIC(waterford_ricker_flow)

#spring temperature covariate model
waterford_ricker_may_temp <- nls(log(Waterford_abundance) ~
               log((a * spawner_abundance) * exp(-b * spawner_abundance))+ f*log_value_mean_temp_may,
             data = waterford_wide, start = c(a = rs$a, b = rs$b, f=-1), lower = c(0.1, 0.0000001, -5), algorithm = "port", nls.control(maxiter=150,minFactor = 0.00000000000001))
summary(waterford_ricker_may_temp)
AIC(waterford_ricker_may_temp)

#fall temperature covariate model
waterford_ricker_oct_temp <- nls(log(Waterford_abundance) ~
               log((a * spawner_abundance) * exp(-b * spawner_abundance))+ f*log_value_mean_temp_oct_nov,
             data = waterford_wide, start = c(a = rs$a, b = rs$b, f=-1), lower = c(0.1, 0.0000001, -5), algorithm = "port", nls.control(maxiter=150,minFactor = 0.00000000000001))
summary(waterford_ricker_oct_temp)
AIC(waterford_ricker_oct_temp)

#spawning wetted area covariate model
waterford_ricker_spawning_wetted <- nls(log(Waterford_abundance) ~
               log((a * spawner_abundance) * exp(-b * spawner_abundance))+ f*log_value_spawning_mean_wetted_acres,
             data = waterford_wide, start = c(a = rs$a, b = rs$b, f=1), lower = c(0.1, 0.0000001, 0), algorithm = "port", nls.control(maxiter=150,minFactor = 0.00000000000001))
summary(waterford_ricker_spawning_wetted)
AIC(waterford_ricker_spawning_wetted)

#spawning wua covariate model
waterford_ricker_spawning_wua <- nls(log(Waterford_abundance) ~
               log((a * spawner_abundance) * exp(-b * spawner_abundance))+ f*log_value_spawning_mean_spawn_acres,
             data = waterford_wide, start = c(a = rs$a, b = rs$b, f=2), lower = c(0.1, 0.0000001, 0), algorithm = "port", nls.control(maxiter=150,minFactor = 0.00000000000001))
summary(waterford_ricker_spawning_wua)
AIC(waterford_ricker_spawning_wua)

#in-channel wetted acres
waterford_ricker_in_channel_wetted <- nls(log(Waterford_abundance) ~
               log((a * spawner_abundance) * exp(-b * spawner_abundance))+ f*log_value_in_channel_mean_wetted_acres,
             data = waterford_wide, start = c(a = rs$a, b = rs$b, f=1), lower = c(0.1, 0.0000001, 0), algorithm = "port", nls.control(maxiter=150,minFactor = 0.00000000000001))
summary(waterford_ricker_in_channel_wetted)
AIC(waterford_ricker_in_channel_wetted)

#in-channel wua
waterford_ricker_in_channel_wua <- nls(log(Waterford_abundance) ~
               log((a * spawner_abundance) * exp(-b * spawner_abundance))+ f*log_value_in_channel_mean_ic_rearing_acres,
             data = waterford_wide, start = c(a = rs$a, b = rs$b, f=1), lower = c(0.1, 0.0000001, 0), algorithm = "port", nls.control(maxiter=150,minFactor = 0.00000000000001))
summary(waterford_ricker_in_channel_wua)
AIC(waterford_ricker_in_channel_wua)

#total rearing wetted
waterford_ricker_total_rearing_wetted <- nls(log(Waterford_abundance) ~
               log((a * spawner_abundance) * exp(-b * spawner_abundance))+ f*log_value_total_rearing_mean_wetted_acres,
             data = waterford_wide, start = c(a = rs$a, b = rs$b, f=1), lower = c(0.1, 0.0000001, 0), algorithm = "port", nls.control(maxiter=150,minFactor = 0.00000000000001))
summary(waterford_ricker_total_rearing_wetted)
AIC(waterford_ricker_total_rearing_wetted)

#total rearing wua
waterford_ricker_total_rearing_wua <- nls(log(Waterford_abundance) ~
               log((a * spawner_abundance) * exp(-b * spawner_abundance))+ f*log_value_total_rearing_mean_total_rearing_acres,
             data = waterford_wide, start = c(a = rs$a, b = rs$b, f=2), lower = c(0.1, 0.0000001, 0), algorithm = "port", nls.control(maxiter=150,minFactor = 0.00000000000001))
summary(waterford_ricker_total_rearing_wua)
AIC(waterford_ricker_total_rearing_wua)

#floodplain rearing wetted
waterford_ricker_floodplain_wetted <- nls(log(Waterford_abundance) ~
               log((a * spawner_abundance) * exp(-b * spawner_abundance))+ f*log_value_floodplain_mean_acres_inundated_wetted_area,
             data = waterford_wide, start = c(a = rs$a, b = rs$b, f=0.3), lower = c(0.1, 0.0000001, 0), algorithm = "port", nls.control(maxiter=150,minFactor = 0.00000000000001))
summary(waterford_ricker_floodplain_wetted)
AIC(waterford_ricker_floodplain_wetted)

#floodplain rearing wua
waterford_ricker_floodplain_wua <- nls(log(Waterford_abundance) ~
               log((a * spawner_abundance) * exp(-b * spawner_abundance))+ f*log_value_floodplain_mean_acres_inundated_wua,
             data = waterford_wide, start = c(a = rs$a, b = rs$b, f=0.3), lower = c(0.1, 0.0000001, 0), algorithm = "port", nls.control(maxiter=150,minFactor = 0.00000000000001))
summary(waterford_ricker_floodplain_wua)
AIC(waterford_ricker_floodplain_wua)
```
