# Load libraries/data

```{r}
library(vegan)
library(tidyverse)
library(lubridate)
library(compositions)
```

Using base dataset (integrated_dataset_2008-2022.Rdata)

```{r}
# load data
load(here::here('Data/integrated_dataset_2008-2024.Rdata'))

# remove entirely duplicated rows
dat <- distinct(dat)

# remove basically duplicated rows (mvi_class and modeled class are NA)
dat <- dat %>%
  filter(!(is.na(mvi_class) & is.na(modeled_prob_harmful)))

# remove spaces in Algal Group names
dat <- dat %>%
  mutate(AlgalGroup = str_remove_all(AlgalGroup, '\\s+'))

# remove LSZs
dat <- dat %>%
  filter(!(Station %in% c('LSZ2','LSZ6','LSZ2-SJR','LSZ6-SJR')))

# use only Count Method == 'field'
dat <- dat %>%
  filter(!(Survey %in% 'DWR-EMP' & CountMethodTaxa %in% c('transect','halfchamber')))

df_test <- dat %>% select(Year, Month, Region, Station, SAC)
```

# Basic Checks

## Unique values

```{r}
lapply(dat[sapply(dat, \(x) is.character(x) || is.factor(x))], unique)
```

Notice issues with Labs/Surveys:

- USGS-BP and DWR-EMP don't have overlapping, comparable metadata according to PESP, but can check for relevant stations

- Lab's aren't comparable, but can check in context of presence/absence data specifically

# Lab Checks

## Check lab timelines

```{r}
df_lab_plot <- dat %>%
  filter(!is.na(Lab), !is.na(Year), !is.na(Month)) %>%
  distinct(Year, Month, Lab) %>%
  group_by(Year, Month) %>%
  summarize(
    Lab = paste(sort(unique(Lab)), collapse = ' + '),
    .groups = 'drop'
  ) %>%
  mutate(
    Year = factor(Year, levels = sort(unique(Year), decreasing = TRUE)),
    Month = factor(Month, levels = 1:12)
  )

ggplot(df_lab_plot, aes(x = Month, y = Year, fill = Lab)) +
  geom_tile(color = 'blue') +
  scale_fill_manual(
    values = c(
      'BSA' = 'lightblue',
      'EcoAnalysts' = 'orange',
      'Dick Dufford' = 'purple',
      'BSA + EcoAnalysts' = 'green',
      'BSA + Dick Dufford' = 'pink',
      'Dick Dufford + EcoAnalysts' = 'yellow',
      'BSA + Dick Dufford + EcoAnalysts' = 'gray70'
    ),
    drop = FALSE
  ) +
  labs(
    x = 'Month',
    y = 'Year',
    fill = 'Lab'
  ) +
  theme_minimal()
```

## Check which stations are which labs

```{r}
dat %>%
  distinct(Station, Lab, Date) %>%
  count(Station, Lab, name = 'n_events') %>%
  pivot_wider(
    names_from = Lab,
    values_from = n_events,
    values_fill = 0
  ) %>%
  arrange(Station)
```

Subset Data

```{r}
lab_checks <- dat %>%
  select(
    Survey:Cells_per_mL,
    tot_cells,
    modeled_prob_harmful
  )

# check duplicates
dup_check <- lab_checks %>%
  filter(!is.na(Cells_per_mL)) %>%
  count(Survey, Station, Lab, Year, Month, AlgalGroup) %>%
  filter(n > 1)

dup_check

# remove Dick Dufford
lab_checks <- lab_checks %>%
  filter(Lab != 'Dick Dufford')
```

Consolidate 'Other' Taxa

```{r}
pa_cols_o <- c(
  'Cyanobacteria',
  'CentricDiatoms',
  'PennateDiatoms',
  'Cryptophytes',
  'GreenAlgae',
  'Other'
)

core_cols <- c(
  'Survey',
  'Station',
  'Lab',
  'Year',
  'Month',
  'tot_cells'
)

df_wide_pa <- lab_checks %>%
  filter(!is.na(Cells_per_mL), !is.na(AlgalGroup)) %>%
  mutate(
    AlgalGroup = if_else(AlgalGroup %in% pa_cols_o, AlgalGroup, 'Other')
  )


df_wide_pa <- df_wide_pa %>%
  group_by(Survey, Station, Lab, Year, Month, AlgalGroup, tot_cells) %>%
  summarize(
    Cells_per_mL = sum(Cells_per_mL, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  pivot_wider(
    id_cols = c(Survey, Station, Lab, Year, Month, tot_cells),
    names_from = AlgalGroup,
    values_from = Cells_per_mL,
    values_fill = 0
  ) %>%
  mutate(across(all_of(pa_cols_o), ~ as.integer(.x > 0)))
```

## PCoA w/ Other

```{r}
df_pcoa_o <- df_wide_pa %>%
  mutate(
    across(all_of(pa_cols_o), ~ replace_na(as.integer(.x), 0L)),
    n_present = rowSums(across(all_of(pa_cols_o)))
  ) %>%
  filter(n_present > 0)

pa_mat_o <- df_pcoa_o %>%
  select(all_of(pa_cols_o)) %>%
  as.data.frame()

d_jac_o <- vegdist(
  pa_mat_o,
  method = 'jaccard',
  binary = TRUE
)

pcoa_o <- wcmdscale(
  d_jac_o,
  k = 2,
  eig = TRUE
)

scores_pcoa_o <- as.data.frame(pcoa_o$points)
colnames(scores_pcoa_o)[1:2] <- c('PCoA1', 'PCoA2')

scores_pcoa_o <- scores_pcoa_o %>%
  select(PCoA1, PCoA2) %>%
  bind_cols(
    df_pcoa_o %>%
      select(all_of(core_cols), all_of(pa_cols_o))
  )
```

```{r}
eig_o <- pcoa_o$eig

axis_var <- tibble(
  axis = paste0('PCoA', seq_along(eig_o)),
  eigenvalue = eig_o
) %>%
  filter(eigenvalue > 0) %>%
  mutate(percent = 100 * eigenvalue / sum(eigenvalue))

p1_o <- round(axis_var$percent[1], 1)
p2_o <- round(axis_var$percent[2], 1)

scores_count_o <- scores_pcoa_o %>%
  group_by(Lab, across(all_of(pa_cols_o))) %>%
  summarize(
    PCoA1 = mean(PCoA1),
    PCoA2 = mean(PCoA2),
    n = n(),
    .groups = 'drop'
  ) %>%
  group_by(Lab) %>%
  mutate(
    n_lab = sum(n),
    prop = n / n_lab
  ) %>%
  ungroup()

ggplot(scores_count_o, aes(PCoA1, PCoA2, color = Lab, size = prop)) +
  geom_point(alpha = 0.9, shape = 21, fill = NA, stroke = 1.2, aes(color = Lab)) +
  scale_size_area(max_size = 8, labels = scales::percent) +
  coord_fixed() +
  labs(
    title = 'Presence/absence of algal groups (w Other)',
    x = paste0('PCoA1 (', p1_o, '%)'),
    y = paste0('PCoA2 (', p2_o, '%)'),
    color = 'Lab',
    size = '% of lab samples'
  ) +
  theme_minimal()
```

## PCOA w/o Other

```{r}
pa_cols_wo <- c(
  'Cyanobacteria',
  'CentricDiatoms',
  'PennateDiatoms',
  'Cryptophytes',
  'GreenAlgae'
)

df_pcoa_wo <- df_wide_pa %>%
  mutate(
    across(all_of(pa_cols_wo), ~ replace_na(as.integer(.x), 0L)),
    n_present = rowSums(across(all_of(pa_cols_wo)))
  ) %>%
  filter(n_present > 0)

pa_mat_wo <- df_pcoa_wo %>%
  select(all_of(pa_cols_wo)) %>%
  as.data.frame()

d_jac_wo <- vegdist(
  pa_mat_wo,
  method = 'jaccard',
  binary = TRUE
)

pcoa_wo <- wcmdscale(
  d_jac_wo,
  k = 2,
  eig = TRUE
)

scores_pcoa_wo <- as.data.frame(pcoa_wo$points)
colnames(scores_pcoa_wo)[1:2] <- c('PCoA1', 'PCoA2')

scores_pcoa_wo <- scores_pcoa_wo %>%
  select(PCoA1, PCoA2) %>%
  bind_cols(
    df_pcoa_wo %>%
      select(all_of(core_cols), all_of(pa_cols_wo))
  )
```

```{r}
eig_wo <- pcoa_wo$eig

axis_var <- tibble(
  axis = paste0('PCoA', seq_along(eig_wo)),
  eigenvalue = eig_wo
) %>%
  filter(eigenvalue > 0) %>%
  mutate(percent = 100 * eigenvalue / sum(eigenvalue))

p1_wo <- round(axis_var$percent[1], 1)
p2_wo <- round(axis_var$percent[2], 1)

scores_count_wo <- scores_pcoa_wo %>%
  group_by(Lab, across(all_of(pa_cols_wo))) %>%
  summarize(
    PCoA1 = mean(PCoA1),
    PCoA2 = mean(PCoA2),
    n = n(),
    .groups = 'drop'
  ) %>%
  group_by(Lab) %>%
  mutate(
    n_lab = sum(n),
    prop = n / n_lab
  ) %>%
  ungroup()

ggplot(scores_count_wo, aes(PCoA1, PCoA2, color = Lab, size = prop)) +
  geom_point(alpha = 0.9, shape = 21, fill = NA, stroke = 1.2, aes(color = Lab)) +
  scale_size_area(max_size = 8, labels = scales::percent) +
  coord_fixed() +
  labs(
    title = 'Presence/absence of algal groups (w/o Other)',
    x = paste0('PCoA1 (', p1_wo, '%)'),
    y = paste0('PCoA2 (', p2_wo, '%)'),
    color = 'Lab',
    size = '% of lab samples'
  ) +
  theme_minimal()
```

## Detection frequency

```{r}
df_wide_pa %>%
  group_by(Lab) %>%
  summarize(
    n_samples = n(),
    across(all_of(pa_cols_o), ~ mean(.x, na.rm = TRUE)),
    .groups = 'drop'
  ) %>%
  pivot_longer(
    cols = all_of(pa_cols_o),
    names_to = 'Group',
    values_to = 'prop_present'
  ) %>%
  ggplot(aes(Group, prop_present, fill = Lab)) +
  geom_col(position = 'dodge') +
  labs(
    title = 'Detection frequency',
    x = 'Algal group',
    y = 'samps w/ group',
    fill = 'Lab'
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Number of samples per lab

```{r}
df_wide_pa %>%
  count(Lab, name = 'n_rows')
```

Different labs seem to matter

# Survey Checks

Filtering down to BSA data, do further method differences matter?

## Check method differences

```{r}
dat %>%
  filter(Lab == 'BSA') %>%
  group_by(Station) %>%
  summarize(
    SampleMethod = paste(sort(unique(SampleMethod)), collapse = ', '),
    CountMethodSample = paste(sort(unique(CountMethodSample)), collapse = ', '),
    Magnification = paste(sort(unique(Magnification)), collapse = ', '),
    .groups = 'drop'
  ) %>%
  arrange(Station)
```

## PCA Plots

Subset data

```{r}
surv_checks <- dat %>%
  select(
    Survey:AlgalGroup,
    relative_cell_density,
    tot_cells,
    mvi_class
  ) %>%
  filter(Lab == 'BSA')

# check no duplicates per event; looks good
event_check <- surv_checks %>%
  filter(!is.na(relative_cell_density), !is.na(AlgalGroup)) %>%
  group_by(Survey, Station, Lab, Year, Month) %>%
  summarize(
    n_tot_cells = n_distinct(tot_cells, na.rm = FALSE),
    tot_cells_values = paste(unique(tot_cells), collapse = ', '),
    n_mvi_class = n_distinct(mvi_class, na.rm = FALSE),
    mvi_class_values = paste(unique(mvi_class), collapse = ', '),
    .groups = 'drop'
  ) %>%
  filter(n_tot_cells > 1 | n_mvi_class > 1)
```

Pivot data

```{r}
pa_cols_o <- c(
  'Cyanobacteria',
  'CentricDiatoms',
  'PennateDiatoms',
  'Cryptophytes',
  'GreenAlgae',
  'Other'
)

df_wide_count <- surv_checks %>%
  filter(!is.na(relative_cell_density), !is.na(AlgalGroup)) %>%
  mutate(
    AlgalGroup = if_else(AlgalGroup %in% pa_cols_o, AlgalGroup, 'Other')
  ) %>%
  group_by(Survey, Station, Lab, Year, Month, AlgalGroup, tot_cells, mvi_class) %>%
  summarize(
    relative_cell_density = sum(relative_cell_density, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  pivot_wider(
    id_cols = c(Survey, Station, Lab, Year, Month, tot_cells, mvi_class),
    names_from = AlgalGroup,
    values_from = relative_cell_density,
    values_fill = 0
  )
```

```{r}
comp_mat <- df_wide_count %>%
  select(all_of(pa_cols_o)) %>%
  as.matrix()

comp_mat <- comp_mat[rowSums(comp_mat) > 0, ]

comp_closed <- comp_mat / rowSums(comp_mat)

comp_closed[comp_closed == 0] <- 0.000001
comp_closed <- comp_closed / rowSums(comp_closed)

comp_clr <- clr(comp_closed)

pca <- prcomp(comp_clr, center = TRUE, scale. = FALSE)

summary(pca)

round(pca$rotation, 3)
```

```{r}
scores <- as.data.frame(pca$x)
scores$mvi_class <- df_wide_count$mvi_class
scores$survey <- df_wide_count$Survey

ggplot(scores, aes(PC1, PC2, color = survey)) +
  geom_point(alpha = 0.6) +
  labs(x = paste0('PC1 (', round(100*summary(pca)$importance[2,1],1), '%)'),
       y = paste0('PC2 (', round(100*summary(pca)$importance[2,2],1), '%)')) +
  theme_minimal()
```

```{r}
scores <- as.data.frame(pca$x)
scores$mvi_class <- df_wide_count$mvi_class
scores$survey <- df_wide_count$Survey

ggplot(scores, aes(PC1, PC2, color = mvi_class)) +
  geom_point(alpha = 0.6) +
  labs(x = paste0('PC1 (', round(100*summary(pca)$importance[2,1],1), '%)'),
       y = paste0('PC2 (', round(100*summary(pca)$importance[2,2],1), '%)')) +
  theme_minimal()
```

```{r}
ggplot(scores, aes(PC1, PC2)) +
  geom_density_2d(aes(color = survey), linewidth = 0.4) +
  labs(x = paste0('PC1 (', round(100*summary(pca)$importance[2,1],1), '%)'),
       y = paste0('PC2 (', round(100*summary(pca)$importance[2,2],1), '%)')) +
  theme_minimal()
```

```{r}
det_freq <- df_wide_count %>%
  group_by(Survey) %>%
  summarize(
    n_samples = n(),
    across(all_of(pa_cols_o), ~ mean(.x > 0, na.rm = TRUE) * 100),
    .groups = 'drop'
  ) %>%
  pivot_longer(
    cols = all_of(pa_cols_o),
    names_to = 'Group',
    values_to = 'detection_freq'
  )
```

## Bar Graph

```{r}
ggplot(det_freq, aes(Group, detection_freq, fill = Survey)) +
  geom_col(position = 'dodge') +
  labs(
    title = 'Detection frequency',
    x = 'Algal group',
    y = 'Samples with group present (%)',
    fill = 'Survey'
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Might be real, but might just be different sites. Feel okay keeping as-is.

# Format Export Data

Format/subset data for further exploratory and data analysis.

Starting from `dat` to be careful. Data saved in 'Short Term Analysis/st_data'

## Bottom and other random data

Remove bottom data, since it's not used. Also sample depth, 89% NA

```{r}
df_export <- dat %>%
  select(-contains('_Bottom')) %>%
  select(-c(Sample_depth_surface, Sample_depth_nutr_surface))
```

## Lab Metadata

Can remove (figured it out here) *except* for Lab

```{r}
df_export <- df_export %>%
  select(-c(CountMethodSample, CountMethodTaxa, Magnification, SampleMethod, DepthType))
```

## Dufford Lab

Fully overlaps with BSA, can remove

```{r}
df_export <- df_export %>%
  filter(Lab != 'Dick Dufford')

unique(df_export$Lab)
```

## Clams

Only a few sites have clams, so used monthly regional averages. Subset out the original data.

Also, make it more clear the monthly rates are regional averages.

```{r}
df_export <- df_export %>%
  rename(monreg_biomass = monthly_biomass,
         monreg_graze_rate = monthly_grazing_rate,
         monreg_filt_rate = monthly_filtration_rate)

colnames(df_export)
```

## Cell Density

We've decided to use `cells` as the unit for density. Removing Unit/Bio columns

```{r}
df_export <- df_export %>%
  select(-c(Biovolume_per_mL, Units_per_mL, tot_bio, tot_units, relative_biovolume, relative_unit_density))
```

Aside from AlgalGroup, relative_cell_density, and Cells_per_mL, every value should be the same per event (Station/Date/Lab). Check this.

```{r}
event_cols <- c('Station', 'Date', 'Lab')

vary_cols <- c(
  'AlgalGroup',
  'relative_cell_density',
  'Cells_per_mL'
)

check_cols <- setdiff(names(df_export), c(event_cols, vary_cols))

event_check <- df_export %>%
  group_by(across(all_of(event_cols))) %>%
  summarize(
    across(
      all_of(check_cols),
      ~ n_distinct(.x, na.rm = FALSE)
    ),
    .groups = 'drop'
  ) %>%
  pivot_longer(
    cols = all_of(check_cols),
    names_to = 'column',
    values_to = 'n_distinct'
  ) %>%
  filter(n_distinct > 1)

event_check
```

Since this is true, we can pivot wider.

## Check Number of Events

First, check the number of distinct events; this is the number of rows the dataframe should have in the end.

```{r}
event_cols <- c('Station', 'Date', 'Lab')

df_events <- df_export %>%
  distinct(across(all_of(event_cols)))

n_events <- nrow(df_events)

n_events
```

## Check No Phyto Data

First, though, we check which rows (if any) don't have corresponding phyto data.

Using Cell Density, it appears all rows have some data (this is not true for Unit/Biovolume Density).

```{r}
df_export <- df_export %>%
  rename(
    RelCellDensity = relative_cell_density,
    AbsCellDensity = Cells_per_mL
  )
```

```{r}
event_cols <- c('Station', 'Date', 'Lab')

df_export %>%
  group_by(across(all_of(event_cols))) %>%
  summarize(
    n_rows = n(),
    n_nonmissing = sum(!is.na(RelCellDensity)),
    n_positive = sum(RelCellDensity > 0, na.rm = TRUE),
    all_na = all(is.na(RelCellDensity) | RelCellDensity == 0),
    .groups = 'drop'
  ) %>%
  filter(all_na)
```

## Pivot Algal Groups

Consolidating less relevant algal groups into 'other' category.

```{r}
alg_cols <- c(
  'Cyanobacteria',
  'CentricDiatoms',
  'PennateDiatoms',
  'Cryptophytes',
  'GreenAlgae',
  'Other'
)

other_groups <- df_export %>%
  filter(!is.na(AlgalGroup), !AlgalGroup %in% alg_cols) %>%
  distinct(OtherGroup = AlgalGroup) %>%
  arrange(OtherGroup)

other_groups
```

```{r}
# define distinct events
id_cols <- setdiff(
  names(df_export),
  c('AlgalGroup', 'RelCellDensity', 'AbsCellDensity')
)

df_events <- df_export %>%
  distinct(across(all_of(id_cols)))

# populate for events w/ phyto data
df_alg_wide <- df_export %>%
  filter(!is.na(AlgalGroup)) %>%
  mutate(
    AlgalGroup = if_else(AlgalGroup %in% alg_cols, AlgalGroup, 'Other')
  ) %>%
  group_by(across(all_of(id_cols)), AlgalGroup) %>%
  summarize(
    RelCellDensity = if (all(is.na(RelCellDensity))) NA_real_ else sum(RelCellDensity, na.rm = TRUE),
    AbsCellDensity = if (all(is.na(AbsCellDensity))) NA_real_ else sum(AbsCellDensity, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  pivot_wider(
    id_cols = all_of(id_cols),
    names_from = AlgalGroup,
    values_from = c(RelCellDensity, AbsCellDensity),
    values_fill = 0,
    names_glue = '{.value}_{AlgalGroup}'
  )

# join w/ events w/o phyto data
df_full_export <- df_events %>%
  left_join(df_alg_wide, by = id_cols)
```

## Check Rows

Row total should equal the number of events (2098).

```{r}
event_cols <- c('Station', 'Date', 'Lab')

df_events <- df_full_export %>%
  distinct(across(all_of(event_cols)))

n_events <- nrow(df_events)

n_events
```

## Export 1: Phyto Metadata

For further analysis, subset by phytoplankton metadata columns

```{r}
df_phyto_meta <- df_full_export %>%
  select(Survey:Microcystis, RelCellDensity_CentricDiatoms:AbsCellDensity_PennateDiatoms)
```

## Export 2: No Phyto Metadata

For analyses without phytoplankton metadata.

```{r}
df_cells <- df_full_export %>%
  select(Survey:tot_cells, modeled_class, mvi_class:AbsCellDensity_PennateDiatoms)
```

## Export

```{r}
wq <- wq %>% ungroup()

save(
  df_phyto_meta,
  df_cells,
  wq,
  file = here::here('Short Term Analysis/st_data/01_afterlab.Rda')
)
```
