# Load libraries/data

```{r}
library(tidyverse)
library(lubridate)
```

Using `01_afterlab.Rda`

```{r}
# load data
load(here::here('Short Term Analysis/st_data/01_afterlab.Rda'))
```

# Format data

## Subset WQ

Subset "wq" to comparable data based on the formatted "dat" data (now df_cells).

Add columns for "YearGroup" (2008-2014, 2015-2022) and Season (OnSeason/OffSeason), referring to HAB Season.

```{r}
wq_sub <- wq %>%
  ungroup() %>%
  semi_join(
    df_cells %>%
      distinct(Station, Date),
    by = c("Station", "Date")
  )

wq_sub <- wq_sub %>%
  mutate(
    Year = as.integer(format(Date, '%Y')),
    YearGroup = case_when(
      Year %in% 2008:2014 ~ '08-14',
      Year %in% 2015:2022 ~ '15-22'
    ),
    Season = if_else(
      as.integer(format(Date, '%m')) %in% 5:10,
      'OnSeason',
      'OffSeason'
    )
  ) %>%
  filter(!is.na(YearGroup))
```

## Add Phyto Groups

Add columns for "YearGroup" (2008-2014, 2015-2022) and Season (OnSeason/OffSeason), referring to HAB Season.

Also add other groupings for display purposes.

```{r}
df_phyto_meta <- df_phyto_meta %>%
  mutate(
    Year = as.integer(format(Date, '%Y')),
    YearGroup = case_when(
      Year %in% 2008:2014 ~ '08-14',
      Year %in% 2015:2022 ~ '15-22'
    ),
    Season = if_else(
      as.integer(format(Date, '%m')) %in% 5:10,
      'OnSeason',
      'OffSeason'
    ),
    Group = paste(YearGroup, Season, sep = ': '),
    MVI_bin = Microcystis,
    Ptox = round(ptox),
    MVI_Ptox = paste(MVI_bin, Ptox, sep = ' : ')
  ) %>%
  filter(!is.na(YearGroup))
```

# Checks

## MVI (categorical)

Uses wq data; using subsetted data for better comparison.

```{r}
wq_sub %>%
  mutate(
    MVI = round(Microcystis)
  ) %>%
  count(MVI, Season) %>%
  pivot_wider(
    names_from = Season,
    values_from = n,
    values_fill = 0
  )
```

```{r}
wq_sub %>%
  mutate(
    Group = paste(YearGroup, Season, sep = ': '),
    MVI = round(Microcystis)
  ) %>%
  count(MVI, Group) %>%
  pivot_wider(
    names_from = Group,
    values_from = n,
    values_fill = 0
  )
```

## MVI (binary)

```{r}
df_phyto_meta %>%
  count(MVI_bin, Group) %>%
  pivot_wider(
    names_from = Group,
    values_from = n,
    values_fill = 0
  )
```

## MVI:Ptox

```{r}
df_phyto_meta %>%
  count(MVI_Ptox, Group) %>%
  pivot_wider(
    names_from = Group,
    values_from = n,
    values_fill = 0
  )
```

## Bloom

*Bloom*: binary value based on whether Chla is \>1 SD above the mean within the given region

```{r}
df_phyto_meta %>%
  count(bloom, Group) %>%
  pivot_wider(
    names_from = Group,
    values_from = n,
    values_fill = 0
  )
```

## Bloom:Class

*Bloom*: binary value based on whether Chla is \>1 SD above the mean within the given region

*Class*: HAB/BAB classification from `HAB_BAB_class` variable

- NA: no bloom? very few (only 6) of these

```{r}
df_phyto_meta %>%
  count(Bloom_Class, Group) %>%
  pivot_wider(
    names_from = Group,
    values_from = n,
    values_fill = 0
  )
```

## Harmful Class

Based off Bloom:Class

- HAB: Class == HAB; Bloom irrelevant

- BAB: Class == BAB; Bloom == Yes

- No bloom: Class == BAB; Bloom == No

```{r}
df_phyto_meta %>%
  count(mvi_class, Group) %>%
  pivot_wider(
    names_from = Group,
    values_from = n,
    values_fill = 0
  )
```

# Check Ptox

## MVI:Ptox:Chla

```{r}
df_phyto_meta %>%
  mutate(
    Chla = case_when(
      bloom == 'Yes' ~ 1,
      bloom == 'No' ~ 0
    ),
    MVI_Ptox_Chla = paste(MVI_bin, Ptox, Chla, sep = ' : ')
  ) %>%
  filter(!is.na(YearGroup)) %>%
  group_by(MVI_Ptox_Chla, Group) %>%
  summarise(
    num_HAB = sum(pred_class == 'HAB', na.rm = TRUE),
    num_total = sum(pred_class %in% c('HAB', 'BAB'), na.rm = TRUE),
    pct_HAB = round(100 * num_HAB / num_total, 2),
    pct_num = sprintf('%.1f%% (%d)', pct_HAB, num_HAB),
    .groups = 'drop'
  ) %>%
  pivot_wider(
    id_cols = MVI_Ptox_Chla,
    names_from = Group,
    values_from = pct_num
  ) %>%
  select(MVI_Ptox_Chla, matches(': OnSeason$'))
```

## "Confusion Matrix"

How often does MVI vs modeling (pred_class) disagree? What role does ptox play when they do?

```{r}
df_phyto_meta %>%
  filter(MVI_bin %in% c(0, 1)) %>%
  mutate(pred_ptox = paste(pred_class, Ptox, sep = ' : ')) %>%
  count(MVI_bin, pred_ptox) %>%
  pivot_wider(
    names_from = pred_ptox,
    values_from = n,
    values_fill = 0
  )
```

```{r}
df_phyto_meta %>%
  filter(MVI_bin %in% c(0, 1)) %>%
  mutate(
    pred_class2 = case_when(
      pred_class %in% c("BAB", "NA") ~ "NotHAB",
      is.na(pred_class) ~ "NotHAB",
      TRUE ~ pred_class
    ),
    pred_ptox = paste(pred_class2, Ptox, sep = " : ")
  ) %>%
  count(MVI_bin, pred_ptox) %>%
  pivot_wider(
    names_from = pred_ptox,
    values_from = n,
    values_fill = 0
  )
```

# Modeled

```{r}
df_phyto_meta %>%
  count(modeled, Group) %>%
  pivot_wider(
    names_from = Group,
    values_from = n,
    values_fill = 0
  )
```

## % Modeled (All)

```{r}
df_phyto_meta %>%
  group_by(YearGroup) %>%
  summarise(
    n_total = n(),
    n_modeled = sum(modeled == 'Yes', na.rm = TRUE),
    pct_modeled = round(100 * mean(modeled == 'Yes', na.rm = TRUE),2),
    .groups = 'drop'
  )
```

## % Modeled (by Season)

```{r}
df_phyto_meta %>%
  group_by(Group) %>%
  summarise(
    n_total = n(),
    n_modeled = sum(modeled == 'Yes', na.rm = TRUE),
    pct_modeled = round(100 * mean(modeled == 'Yes', na.rm = TRUE),2),
    .groups = 'drop'
  )
```
