After a closer look at the phytoplankton data, decided to work with subsets. We will create all our subsets here.

"dat": has community and WQ data and a HAB/BAB classification
- but if they had an MVI but not community data, not in dat


# Load libraries/data

```{r}
library(tidyverse)
library(lubridate)
```

Using `01_afterlab.Rda`

```{r}
# load data
load(here::here('Short Term Analysis/st_data/01_afterlab.Rda'))
```

# Clean Data

Remove data with no Bloom Classification at all (only 8 rows)

```{r}
df_cells <- df_cells %>%
  filter(!is.na(mvi_class))
```

Rename "NA" to "No Bloom".

```{r}
df_cells <- df_cells %>%
  mutate(
    modeled_class = case_when(
      is.na(modeled_class) ~ 'No bloom',
      TRUE ~ modeled_class
    )
  )
```

Add `YearMonth` column (YYYY-MM-01).

```{r}
df_cells <- df_cells %>%
  mutate(
    YearMonth = as.Date(sprintf('%s-%02d-01', Year, Month))
  ) %>%
  relocate(YearMonth, .before = Date)
```

Combine Turbidities

```{r}
# check no overlap
df_cells %>%
  summarize(n_overlap = sum(!is.na(TurbidityNTU) & !is.na(TurbidityFNU)))

# combine
df_cells <- df_cells %>%
  mutate(Turbidity = coalesce(TurbidityNTU, TurbidityFNU)) %>%
  select(-TurbidityNTU, -TurbidityFNU) %>%
  relocate(Turbidity, .after = Conductivity)
```

Remove analytes with \> 70% NA

```{r}
df_cells <- df_cells %>%
  select(-c(DissBromide, TON, TotChloride, DissolvedOxygenPercent))
```

Remove colinear analytes

```{r}
# Secchi ~ Turbidity; light availability

df_cells <- df_cells %>%
  select(-c(Secchi))
```

# Save Data

```{r}
save(df_cells, file = here::here('Short Term Analysis/st_data/03_cleandata.Rda'))
```

# Subset: MVI only

Filter to year >= 2015 (when MVI started)

```{r}
df_mvi <- df_cells %>%
  filter(Year >= 2015 & Lab == 'BSA')
```

```{r}
save(df_mvi, file = here::here('Short Term Analysis/st_data/03_cleandata.Rda'))
```

# Full: Environmental Data

Use all environmental data, regardless if there's corresponding phytoplankton data or not (with cell density, this is not an issue).

```{r}
df_full <- df_cells %>%
  select(-modeled)

lapply(df_full[sapply(df_full, \(x) is.character(x) || is.factor(x))], unique)
```
