---
execute:
  echo: false
---

Exploratory data analysis for using all data.

TODO: fix using 'modeled'; ignores off season
- 2015 is when MVI starts

prioritize running 2015-present with BSA, MVI data
- can run others after for exploratory/comparison purposes
- **eg. EA (2008-2014) modeled data**
- eg. same BSA time period w/ modeled data
  - would need to be clear about the second set of community models (since circular)
    - can do 'w/o ptox'
    
    
TODO: threshold analysis (with multinomial)
- 

**QUESTION: What was done with \< RL values??**

# Load Libraries/Data

```{r}
library(tidyverse)
library(lubridate)
library(mice)
library(NADA)

conflicted::conflict_prefer('filter', 'dplyr')
```

Using `03_cleanlab.Rda`

```{r}
# load data
load(here::here('Short Term Analysis/st_data/03_cleandata.Rda'))

# set regions variable
regions <- unique(df_mvi$Region)

# remove salinity
df_mvi <- df_mvi %>%
  select(-Salinity)

# set analytes variable
analytes <- df_mvi %>%
  select(Chlorophyll:monreg_filt_rate) %>%
  names()
```

Note that this data does not subset for rows with phyto data; ie. this is all data with a Bloom Classification, regardless if it has corresponding phyto data or not (though, with cell density, all do).

# Unique Check

```{r}
lapply(df_mvi[sapply(df_mvi, \(x) is.character(x) || is.factor(x))], unique)
```
## Index Notes

Note: Index is numeric index, not categorical. Adding in categorical for later.


```{r}
unique(df_mvi$Index)
```

...also, is Index for SAC or SJV? Does that matter for flow variables? Confirmed there's only one index per Water Year

```{r}
df_mvi <- df_mvi %>%
  mutate(WaterYear = if_else(as.integer(as.character(Month)) >= 10,
                             as.integer(as.character(Year)) + 1,
                             as.integer(as.character(Year))))

df_mvi %>%
  group_by(WaterYear) %>%
  summarize(n_index = n_distinct(Index, na.rm = TRUE), .groups = 'drop') %>%
  filter(n_index > 1)
```
Also some NAs, due to missing 2023/2024.

```{r}
df_mvi %>%
  filter(is.na(Index)) %>%
  select(Station, Year) %>%
  distinct() %>%
  arrange(Year, Station)
```

```{r}
df_mvi <- df_mvi %>%
  mutate(
    Index = case_when(
      Year == 2023 ~ 8.45,
      Year == 2024 ~ 8.98,
      TRUE         ~ Index
    ),
    Index_Cat = case_when(
      is.na(Index)  ~ NA_character_,
      Index >= 9.2  ~ 'W',
      Index >  7.8  ~ 'AN',
      Index >  6.5  ~ 'BN',
      Index >  5.4  ~ 'D',
      TRUE          ~ 'C'
    ),
    Index_Cat = factor(Index_Cat, levels = c('C', 'D', 'BN', 'AN', 'W'))
  )
```

# Station/Region Check

## 2015-2024

Per region, which stations have any data per event (Month+Year)?

```{r}
for (r in regions) {

  plot_df <- df_mvi %>%
    filter(Region == r) %>%
    mutate(YearMonth = as.Date(YearMonth)) %>%
    count(Station, YearMonth, name = 'n_rows') %>%
    complete(
      Station,
      YearMonth = seq.Date(
        min(YearMonth, na.rm = TRUE),
        max(YearMonth, na.rm = TRUE),
        by = 'month'
      ),
      fill = list(n_rows = 0)
    ) %>%
    mutate(row_exists = n_rows > 0)

  p <- plot_df %>%
    ggplot(aes(x = YearMonth, y = Station, fill = row_exists)) +
    geom_tile(color = 'blue') +
    scale_fill_manual(
      values = c('FALSE' = 'red', 'TRUE' = 'lightblue')
    ) +
    labs(
      title = r,
      x = NULL,
      y = 'Station',
      fill = 'Row exists'
    ) +
    theme_minimal()

  print(p)
}
```



```{r}
for (r in regions) {
  first_yearmonth_tbl <- df_mvi %>%
    filter(Region == r) %>%
    mutate(YearMonth = as.Date(YearMonth)) %>%
    group_by(Station) %>%
    summarize(
      first_YearMonth = min(YearMonth, na.rm = TRUE),
      .groups = 'drop'
    ) %>%
    arrange(first_YearMonth, Station)
  
  print(first_yearmonth_tbl)
}
```

**Central**: D28A, MD10A, and P8 can be used (possibly C9, starts about 1 year later though)

**Confluence**: D26 and D19 only (D22, D16, D12 start a bit later; D24 the same time as C9; NZ068 about 6 months after that).

**Suisun Bay**: D8, D7, D6, D4 (D10 starts same time as NZ068).

### Recommendation: Start either 06/01/2016 (exclude two stations) or 05/01/2017 (include all). Only D24 should be completely removed.

TODO: check when 'summer/spring' start.

### Remove stations without enough data

```{r}
df_mvi <- df_mvi %>%
  filter(!(Station %in% c('D24')))
```

### Subset date

TODO: decide

# Central Region

## Data Coverage

```{r}
df_central_all <- df_mvi %>%
  filter(Region == 'Central') %>%
  filter(Station %in% c('D28A', 'MD10A', 'P8'))
```

```{r}
for (a in analytes) {
  p <- df_central_all %>%
    mutate(missing = is.na(.data[[a]])) %>%
    ggplot(aes(x = YearMonth, y = Station, fill = missing)) +
      geom_tile(color = 'blue') +
      scale_fill_manual(values = c('FALSE' = 'lightblue', 'TRUE' = 'red')) +
      labs(title = paste0(a, ' - Central'),
           x = NULL, y = 'Station', fill = 'Missing') +
      theme_minimal()
  print(p)
}

```

### Conclusions

TODO: talk to besty about clam data

DissAmmonia and VSS have large gaps...

But, there should be data...

```{r}
df_central_all %>%
  summarize(
    n = n(),
    across(all_of(analytes), ~ sum(is.na(.x)))
  ) %>%
  pivot_longer(-n, names_to = 'analyte', values_to = 'n_na') %>%
  mutate(pct_na = round(100 * n_na / n, 1)) %>%
  arrange(desc(n_na))
```


## ISSUES

### Flow and clams

Missing 2023/2024 data

### Nutrients

Non-detects are being read in as 'NA'.

Need to properly account for censored data. Easiest option: uniform imputation between close-to-0 and reporting limit (should really be done multiple times and pooled, but the impact is likely small, check with sensitivity check; can do for the final analysis; also applies to other imputed values).

For now, will assume all NAs are < RL and that Ammonia's RL was always 0.05. (most properly account for in HABvBAB file (eg. include Detect column), but unable to fully run code).

# Impute Nutrients (< RL)

TODO: rn just DissAmmonia, should add in other relevant nutrients

## Check sensitivity

TODO: right now, assumes same RL for a given nutrient

```{r}
censor_check <- function(x, rl = 0.05, fl = 0.01, n_draws = 200) {

  nd  <- is.na(x)
  obs <- ifelse(nd, rl, x)

  draws <- replicate(n_draws,
    { v <- replace(x, nd, runif(sum(nd), fl, rl)); c(mean(v), sd(v)) })

  ros_vals <- as.data.frame(ros(obs, nd))$modeled
  km       <- cenfit(obs, nd)

  tibble(
    method = c('zero', 'half_rl', 'at_rl', 'uniform', 'ros', 'kaplan_meier'),
    mean = c(mean(replace(x, nd, 0)),
               mean(replace(x, nd, rl / 2)),
               mean(replace(x, nd, rl)),
               mean(draws[1, ]),
               mean(ros_vals),
               as.numeric(mean(km)['mean'])),
    q1 = c(NA, NA, NA, quantile(draws[1, ], 0.025, names = FALSE), NA, NA),
    q4 = c(NA, NA, NA, quantile(draws[1, ], 0.975, names = FALSE), NA, NA),
    pct_nd = round(100 * mean(nd), 1)
  ) %>%
    mutate(across(c(mean, q1, q4), ~ round(.x, 3)))
}
```

```{r}
rls <- c(DissAmmonia = 0.05, DissNitrateNitrite = 0.05, DissOrthophos = 0.05)

sens_all <- imap_dfr(rls, function(rl, a) {
  df_mvi %>%
    group_by(Station) %>%
    group_modify(~ censor_check(.x[[a]], rl = rl, fl = rl/5)) %>%
    ungroup() %>%
    mutate(analyte = a, .before = 1)
})

sens_all
```

## Impute

Define all nutrients with < RL values

```{r}
nutrients <- c(
  'DissAmmonia',
  'DissNitrateNitrite',
  'TotPhos',
  'DissOrthophos',
  'DON',
  'TKN',
  'DissCalcium',
  'DissChloride',
  'DissSilica',
  'TOC',
  'DOC',
  'TDS',
  'TSS',
  'VSS',
  'Chlorophyll',
  'Pheophytin'
)
```

Impute < RL values

TODO: rn, assumes all have 0.05 RL for entire time, and that all missing data is < RL data. THIS IS NOT TRUE; MUST BE FIXED.

```{r}
# custom mice method: uniform draw between close-to-zero and RL
mice.impute.unifrl <- function(y, ry, x, wy = NULL, fl = 0.01, rl = 0.05, ...) {
  if (is.null(wy)) wy <- !ry
  runif(sum(wy), fl, rl)
}

# no need to split by station since not conditioned on anything
meth <- make.method(df_mvi)
meth[] <- ''
meth[nutrients] <- 'unifrl'

pred <- make.predictorMatrix(df_mvi)
pred[] <- 0

imp <- mice(df_mvi, method = meth, predictorMatrix = pred,
            m = 10, maxit = 1, printFlag = FALSE,
            fl = 0.01, rl = 0.05)
```

```{r}
complete(imp, 'long') %>%
  group_by(.imp, Station) %>%
  summarize(mean_amm = mean(DissAmmonia), .groups = 'drop') %>%
  group_by(Station) %>%
  summarize(
    mean = mean(mean_amm),
    q1 = quantile(mean_amm, 0.025),
    q4 = quantile(mean_amm, 0.975),
    .groups = 'drop'
  ) %>%
  mutate(across(c(mean, q1, q4), ~ round(.x, 3)))
```

Check density curves

```{r}
# long_imp <- complete(imp, 'long')
# 
# nutrients <- c('DissAmmonia','DissOrthophos')
# 
# for (nut in nutrients) {
# 
#   d_imp <- long_imp %>% select(.imp, Station, value = all_of(nut))
#   d_obs <- df_mvi   %>% select(Station, value = all_of(nut)) %>% filter(!is.na(value))
# 
#   n_lab <- df_mvi %>%
#     group_by(Station) %>%
#     summarize(
#       lab = paste0(first(Station), ' (', sum(!is.na(.data[[nut]])), 'd / ',
#                    sum(is.na(.data[[nut]])), 'nd)'),
#       .groups = 'drop'
#     )
# 
#   p <- ggplot(d_imp %>% left_join(n_lab, by = 'Station'),
#               aes(value, group = .imp)) +
#     geom_density(colour = 'red', linewidth = 0.2) +
#     geom_density(data = d_obs %>% left_join(n_lab, by = 'Station'),
#                  aes(value), inherit.aes = FALSE,
#                  colour = 'steelblue', linewidth = 0.8) +
#     facet_wrap(~ lab, scales = 'free') +
#     labs(title = nut, x = nut, y = 'Density') +
#     theme_minimal()
# 
#   print(p)
# }
```

## Add imputed vals to df

```{r}
# pick one chain
df_mvi <- complete(imp, 1)
```

# Impute Nutrients (truly missing)

TODO: add in after this info is known

# Impute Clams

TODO: Right now, running on subset without 2023/2024 (no data)

```{r}
df_mvisub <- df_mvi %>%
  filter(!Year %in% c(2023, 2024))
```

## Sensitivity check

TODO

## Impute

```{r}
system_vars <- c('monreg_biomass', 'monreg_graze_rate', 'monreg_filt_rate')

predictor_vars <- c(
  'Year',
  'Month',
  'Region',
  'Temperature',
  'Conductivity'
)
```

```{r}
df_mice <- df_mvisub %>%
  mutate(
    Year   = factor(Year),
    Month  = factor(Month),
    Region = factor(Region)
  ) %>%
  group_by(Year, Month, Region) %>%
  summarize(
    Temperature  = mean(Temperature, na.rm = TRUE),
    Conductivity = mean(Conductivity, na.rm = TRUE),
    across(all_of(system_vars), first),
    .groups = 'drop'
  )

meth <- make.method(df_mice)
meth[] <- ''
meth[system_vars] <- 'pmm'

pred <- make.predictorMatrix(df_mice)
pred[,] <- 0
pred[system_vars, c(predictor_vars, system_vars)] <- 1
diag(pred) <- 0

imp <- mice(
  df_mice,
  m       = 20,
  maxit   = 20,
  method  = meth,
  predictorMatrix = pred,
  seed    = 42
)
```

```{r}
plot(imp)
```

```{r}
densityplot(imp)
stripplot(imp, pch = 20, cex = 0.7)
```

## Add back in

### TEMP FIX

Subsetting all of mvi to 2022

```{r}
df_mvi <- df_mvi %>%
  filter(!c(Year %in% c(2023, 2024)))
```


```{r}
# pick one chain
df_mvi <- df_mvi %>%
  mutate(
    Year   = factor(Year),
    Month  = factor(Month),
    Region = factor(Region)
  ) %>%
  select(-all_of(system_vars)) %>%
  left_join(complete(imp, 1) %>% select(Year, Month, Region, all_of(system_vars)),
            by = c("Year", "Month", "Region"))
```

# Export

```{r}
save(df_mvi, file = here::here('Short Term Analysis/st_data/04_imputedata.Rda'))
```
